### 2017-02-11 manuscript.rmd
added manuscript files

---
bibliography: manuscript.bib
csl: systematic-biology.csl
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
---

```{r preliminaries, include=FALSE}
	# Load packages

	# General
	library( tidyverse )
	library( stringr )
	library( forcats )
  library( magrittr )
	library( digest )
	library( knitr )
	library( jsonlite )
	library( parallel )

	# Formatting
	library( xtable )

	# Graphics
	library( cowplot )
	library( ggtree )
	library( GGally )
	library( fields ) # tim.colors
	library( seriation )
	library( gridExtra )
	library( RGraphics )

	# Biological
	library( edgeR )
	library( treeio )
	library( hutan )	# https://github.com/caseywdunn/hutan
	library( ape )
	library( agalmar )	# https://github.com/caseywdunn/agalmar
	library( picante )  # cor.table
	library( phytools ) # phylosig

	# Configure knitr, see http://yihui.name/knitr/options
	opts_knit$set( progress = TRUE, verbose = TRUE )

	# Set system computational parameters
	cores = detectCores() - 1
	if ( cores < 1 ) {
		cores = 1
	}
	set.seed( 23456 )

	# Set paths to input data
	json_name = "agalma/agalma.json"

```

```{r load_data, include=FALSE, cache=TRUE}

  # Load and parse phylogenies and expression data from the 
  # json file produced by agalma

	# Eventually species tree will be parsed from json, for now specify temporary tree
	# agalma.data = jsonlite::fromJSON( json_name )
	# species_tree = read.tree( text=agalma.data$speciestree )

	# For now, specify the tree as plain text
	species_tree_text = "(((((((((((((((Nanomia_bijuga:0.120563,Agalma_elegans:0.0575132)100:0.0339417,Lychnagalma_utricularia:0.0749548)100:0.0140092,(Cordagalma_sp_:0.117222,Forskalia_asymmetrica:0.0739343)100:0.0113754)100:0.0159061,(Resomia_ornicephala:0.036642,Frillagalma_vityazi:0.0604811)100:0.0190101)100:0.024866,(((Physonect_sp_:0.0317839,Marrus_claudanielis:0.0284231)100:0.0125343,Stephalia_dilata:0.0841317)82:0.00517707,Erenna_richardi:0.0414121)100:0.0124079)100:0.0161254,((((Chelophyes_appendiculata:0.0574339,Abylopsis_tetragona:0.0562432)100:0.11489,(Chuniphyes_multidentata:0.0345263,Kephyes_ovata:0.0306657)100:0.0273795)100:0.0225022,Lilyopsis_fluoracantha:0.197884)100:0.018739,((Prayidae_D27SS7:0.0600685,Prayidae_D27D2:0.0464896)100:0.00909199,Hippopodius_hippopus:0.100508)100:0.011385)100:0.0822577)100:0.0232853,((Bargmannia_elongata:0.0175209,Bargmannia_amoena:0.0201366)100:0.0187736,Bargmannia_lata:0.031329)100:0.0887499)100:0.0300558,((Apolemia_sp_:0.0291459,Apolemia_lanosa:0.0332208)95:0.00837843,Apolemia_white:0.0422333)100:0.0841266)100:0.0313749,(Physalia_physalis:0.082276,Rhizophysa_filiformis:0.105414)100:0.0750894)100:0.133951,(Hydractinia_symbiolongicarpus:0.100907,Podocoryna_carnea:0.246708)100:0.104625)99:0.0219002,Clytia_hemisphaerica:0.331138)97:0.0219604,(Hydra_magnipapillata:0.268083,Ectopleura_larynx:0.273479)100:0.11387)100:0.0827444,Aegina_citrea:0.408585)100:0.148851,(Atolla_vanhoeffeni:0.282409,Alatina_alata:0.328817):0.058436)100:0.419544,(Aiptasia_pallida:0.199088,Nematostella_vectensis:0.188379)100:0);"

	species_tree = read.tree( text=species_tree_text )

```

```{r pre_process_species_trees, include=FALSE, cache=TRUE}

	# Repartition length of edges descended from root so they are non-zero
	species_tree = hutan::slide_root_edges( species_tree )

	# species_tree = ape::root( species_tree, c( "Nematostella_vectensis", "Aiptasia_pallida" ), resolve.root=TRUE )
	# species_tree = ape::unroot(species_tree)

	stopifnot( species_tree$Nnode == length( species_tree$node.label ) )

	species_tree = ape::ladderize( species_tree )

	# Update species names
	species_tree$tip.label = sub( '_', ' ', species_tree$tip.label )

	# Make ultrametric tree
	# By default, root has a depth of 1
	species_ultrametric = chronos( species_tree, lambda=1, model="correlated", quiet=TRUE )
	class( species_ultrametric ) = "phylo"

```

# Siphonophore Phylogeny

Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Cat Munro^1^, Freya Goetz^1,5^, Phil Pugh, Steven H.D. Haddock^4^, Casey W. Dunn^1^*

^1^ Department of Ecology and Evolutionary Biology, Brown University, Providence, RI, USA

^2^ Current address: Department of Molecular & Cellular Biology, University of California at Davis, Davis, CA, USA

^3^ Current address: Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA, USA

^4^ Brown Data Science Practice, Brown University, Brown University, Providence, RI, USA

^5^ Current address: Smithsonian, Washington DC, USA

^4^ Monterey Bay Aquarium Research Institute, Moss Landing, CA, USA

\* Corresponding author, casey_dunn@brown.edu

## Abstract

## Introduction

Siphonophores are...

## Methods

This manuscript is an executable document computed directly from the data, providing an explicit and reproducible description of all findings. All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017). The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")`.

#### Collecting

Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in supplementary table 1. Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water scuba dives. *Chelophyes appendiculata* and *Hippopodius hippopus* specimen were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 04/13/11. Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, and at the United States National Museum (Smithsonian Institution), Washington, DC. Accession numbers are given in supplementary table X. In cases where physical vouchers were unavailable we provide photographs to document species identity (table x).

#### mRNA isolation and RNAseq library preparation

When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table 1). mRNA was extracted directly from tissue using Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011) or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006). Extractions were performed according to the manufacturer's instruction. In case of very small total RNA quantities, only a single round of bead purification was performed (supplementary table 1). Resulting higher rRNA read counts were dealt with further downstream in the bioinformatics workflow. Libraries were prepped for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep. Master Mix Set (NEB, #E6110S). We collected long read paired end Illumina data for *de novo* transcriptome assembly. Summary statistics for expression libaries are given in Table 1.

#### Analysis

Sequence assembly, annotation, phylogenetic analysis, and derivation of expression counts were conducted with the tool Agalma [@Dunn:2013kw], v0.XXX. Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.

## Results and Discussion

### Sample collecting and sequencing

```{r summary_table, results='asis', cache=TRUE, echo=FALSE, comment=NA, warning=FALSE}

```
Table 1 (at the end of file): Summary statistics for libraries.

### Species phylogeny

```{r species_phylogram, cache=TRUE, echo=FALSE, comment=NA, warning=FALSE, fig.cap="Phylogram of siphonophore relationships. Node labels indicate bootstrap support percent, unnumbered nodes have 100% support. The image was rendered with ggtree [@Yu:2016fo]" }

	siphonophora_node = ape::mrca( species_tree )[ "Physalia physalis", "Agalma elegans" ]
	cystonectae_node = ape::mrca( species_tree )[ "Physalia physalis", "Rhizophysa filiformis" ]
	calycophora_node = ape::mrca( species_tree )[ "Kephyes ovata", "Hippopodius hippopus" ]
	euphysonectae_node = ape::mrca( species_tree )[ "Agalma elegans", "Erenna richardi" ]

	support = as.numeric( species_tree$node.label )
	support[ support > 99 ] = NA

	# pad tip values
	support = c( rep( NA, length( species_tree$tip.label ) ), support )

	p = ggtree( species_tree ) + 
			geom_text( aes( label=support ), vjust=-.5, hjust=1, size=2.5, col="darkslategray4" ) +
			geom_tiplab( size=3 ) +
			xlim( NA, 4 ) + 
			geom_cladelabel( node=cystonectae_node, label="Cystonectae", align=FALSE, offset=1, color='red' ) +
			geom_cladelabel( node=calycophora_node, label="Calycophora", align=FALSE, offset=1, color='blue' ) +
			geom_strip(
				which( species_tree$tip.label=="Agalma elegans" ), 
				which( species_tree$tip.label=="Apolemia white" ), 
				barsize=0.5, 
				color='orange', 
				offset=1, 
				label='"Physonectae"'
			) +
			geom_strip( 
				which( species_tree$tip.label=="Hydractinia symbiolongicarpus" ), 
				which( species_tree$tip.label=="Nematostella vectensis" ), 
				barsize=0.5, 
				color='gray', 
				offset=1, 
				label='"Outgroups"'
			)

	flip( p, euphysonectae_node, calycophora_node )

```

The phylogenetic relationships between siphonophores species presented here is based on XXX genes from XXX siphonophore species, including new data presented here for XXX species. These findings are entirely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy]. Cystonectae is the sister group to the remaining siphonophores, and Calycophorae is nested within the paraphyletic "Physonectae". In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this XXX-gene transcriptome analysis. These findings include XXX.

## Conclusion

## Acknowledgements

This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057. Data transfer was supported by NSF RII-C2 EPS-1005789. Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island. We also thank the MBARI crews and ROV pilots for collection of the specimens.

## Supplementary Information

### Supplementary Analyses

### Software versions

This manuscript was computed on `r format( Sys.time(), "%a %b %d %X %Y" )` with the following R package versions.

```{r session_summary, echo=FALSE, comment=NA}
	sessionInfo()
```

## References

### 2017-02-11 character_coding/CharacterEvolution.Rmd
Started R Markdown document for character coding

---
title: "## Evolutionary History of the Main Biological Characters of Siphonophores"
author: "Alejandro Damian Serrano"
date: "2/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Packages BASIC
library('magrittr')
library('tidyverse')
library('stringr')

#Packages BIO-PHYLO
library('vegan')
library("ape")
library("ggtree")
library("phangorn")
library("phytools")
library("picante")
library("geiger")
library("phylobase")
library("fields")
library("adephylo")
library("geomorph")

#Working directory
setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")

#Load data
read.csv('main_characters.csv', header = T, sep = ',') -> cdata
rownames(cdata) = cdata$Species

#Load tree
tree = read.tree('ultimatephylo2017.tre')
tree$tip.label = str_replace_all(tree$tip.label,'_',' ')
tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
ultratree <- chronos(tree)

#Temporarily prune data
cdata %<>% filter(rownames(.) %in% tree$tip.label)
rownames(cdata) = cdata$Species
cdata %<>% .[-1]

#Named vectors for variables
sex = cdata$Sex
names(sex) = rownames(cdata)

nectosome = cdata$Nectosome
names(nectosome) = rownames(cdata)

necto_pos = cdata$Nectosome.position
names(necto_pos) = rownames(cdata)

necto_types = cdata$Nectophores
names(necto_types) = rownames(cdata)

palpons = cdata$Palpons
names(palpons) = rownames(cdata)

tentilla = cdata$Tentilla
names(tentilla) = rownames(cdata)

pneuma = cdata$Pneumatophore
names(pneuma) = rownames(cdata)

vrange = cdata$Vertical.range
names(vrange) = rownames(cdata)

bracts = cdata$Bract.types
names(bracts) = rownames(cdata)

#Depth pruned data
dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
dpruned_tree <- drop.tip(ultratree, which(!(tree$tip.label %in% rownames(dpruned_data))))
dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]

depth_median = dpruned_data$Depth.Median
names(depth_median) = rownames(dpruned_data)

depth_mean = dpruned_data$Depth.Mean
names(depth_mean) = rownames(dpruned_data)

depth_min = dpruned_data$Depth.Min
names(depth_min) = rownames(dpruned_data)

depth_max = dpruned_data$Depth.Max
names(depth_max) = rownames(dpruned_data)

depth_sd = dpruned_data$Depth.StdDev
names(depth_sd) = rownames(dpruned_data)

depth_range = depth_max - depth_min

abundance = dpruned_data$Depth.Count
names(abundance) = rownames(dpruned_data)

```

##The Tree:
```{r, echo=F}
par(ask=F)
par(mfrow=c(1,1))
plotTree(ultratree, lwd = 4)

```

Siphonohores...

## Stochastic mapping plots

###Sex distribution:

```{r, echo=F}
#Simmap SEX
plotTree(ultratree, lwd = 4)
make.simmap(ultratree, sex, nsim = 100) -> sex_sim
sex_sim %>% plotSimmap(lwd = 4, add = T)
colors = c("black", "red")
names(colors) = c("Dioeicous", "Monoeicous")
nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(sex_sim)

```

### 2017-02-11 character_coding/character_evolution.R

@@ -64,6 +64,31 @@ names(vrange) = rownames(cdata)
 bracts = cdata$Bract.types
 names(bracts) = rownames(cdata)
 
+#Depth pruned data
+dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
+dpruned_tree <- drop.tip(ultratree, which(!(tree$tip.label %in% rownames(dpruned_data))))
+dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
+
+depth_median = dpruned_data$Depth.Median
+names(depth_median) = rownames(dpruned_data)
+
+depth_mean = dpruned_data$Depth.Mean
+names(depth_mean) = rownames(dpruned_data)
+
+depth_min = dpruned_data$Depth.Min
+names(depth_min) = rownames(dpruned_data)
+
+depth_max = dpruned_data$Depth.Max
+names(depth_max) = rownames(dpruned_data)
+
+depth_sd = dpruned_data$Depth.StdDev
+names(depth_sd) = rownames(dpruned_data)
+
+depth_range = depth_max - depth_min
+
+abundance = dpruned_data$Depth.Count
+names(abundance) = rownames(dpruned_data)
+
 ### Plots ###
 par(ask=F)
 par(mfrow=c(1,1))
@@ -180,31 +205,6 @@ add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=
 #Contmap bract types
 contMap(ultratree, bracts)
 <Return>
-
-#Contmap depth
-dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
-dpruned_tree <- drop.tip(ultratree, which(!(tree$tip.label %in% rownames(dpruned_data))))
-dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
-
-depth_median = dpruned_data$Depth.Median
-names(depth_median) = rownames(dpruned_data)
-
-depth_mean = dpruned_data$Depth.Mean
-names(depth_mean) = rownames(dpruned_data)
-
-depth_min = dpruned_data$Depth.Min
-names(depth_min) = rownames(dpruned_data)
-
-depth_max = dpruned_data$Depth.Max
-names(depth_max) = rownames(dpruned_data)
-
-depth_sd = dpruned_data$Depth.StdDev
-names(depth_sd) = rownames(dpruned_data)
-
[…]

### 2017-02-12 character_coding/CharacterEvolution.Rmd
added more simmaps

@@
 
+```{r, include=F}
+#Make simmap objects
+make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+
+```
+
 ###Sex distribution:
@@
 plotTree(ultratree, lwd = 4)
-make.simmap(ultratree, sex, nsim = 100) -> sex_sim
 sex_sim %>% plotSimmap(lwd = 4, add = T)
@@
 ```
+
+###Presence of palpons:
+```{r, echo=F}
+#Simmap PALPONS
+plotTree(ultratree, lwd = 4)
+make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+palpons_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(palpons_sim)
+```
+
+###Presence of tentilla:
+```{r, echo=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+tentilla_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim)
+```
+
+###Presence of Nectophores:
+```{r, echo=F}
+
+```
+
+###Position of Nectosome:
+```{r, echo=F}
+
+```
+
+###Nectophore number and types:
+
+###Presence of Pneumatophore:
+```{r, echo=F}
+
+```
+
+###Ocean Depth Zone:
+```{r, echo=F}
+
+```

### 2017-02-13 character_coding/CharacterEvolution.Rmd
separated make.simmaps from plotSimmaps

--- 2017-02-11 character_coding/CharacterEvolution.Rmd
@@
 
+```{r, echo=F}
+#Make simmap objects
+make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
+
+```
+
 ###Sex distribution:
@@
 ```
+
+###Presence of palpons:
+```{r, echo=F}
+#Simmap PALPONS
+plotTree(ultratree, lwd = 4)
+palpons_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(palpons_sim)
+```
+
+###Presence of tentilla:
+```{r, echo=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+tentilla_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim)
+```
+
+###Presence of Nectophores:
+```{r, echo=F}
+#Simmap NECTOSOME
+plotTree(ultratree, lwd = 4)
+nectosome_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(nectosome_sim)
+```
+
+###Position of Nectosome:
+```{r, echo=F}
+#SIMMAP NECTOSOME POSITION
+plotTree(ultratree, lwd = 4)
+necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green")
+names(colors) = c("Dorsal", "No nectosome", "Ventral")
+nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Nectophore number and types:
+```{r, echo=F}
+#Simmap NECTOPHORE TYPES
+plotTree(ultratree, lwd = 4)
+necto_types_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green", "blue")
+names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
+nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Presence of Pneumatophore:
+```{r, echo=F}
+#Simmap PNEUMATOPHORE
+plotTree(ultratree, lwd = 4)
+pneuma_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(pneuma_sim)
+```
+
+###Number of Bract types:
+```{r}
+#Simmap bract-types
+plotTree(ultratree, lwd = 4)
+bracts_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green3", "blue")
+names(colors) = c("No bracts", "1", "2", "4")
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Ocean Depth Zones:
+```{r, echo=F}
+#Simmap Vertical.Range
+plotTree(ultratree, lwd = 4)
+vrange_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "cyan", "blue", "green", "red")
+names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
+nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```

### 2017-02-13 character_coding/CharacterEvolution.Rmd
added continous variable reconstructionplots and phylosig analyses

--- 2017-02-11 character_coding/CharacterEvolution.Rmd
@@
 ##The Tree:
-```{r, echo=F}
+```{r ultratree, echo=F}
 par(ask=F)
@@
 
+```{r simmap_making, echo=F}
+#Make simmap objects
+make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
+
+```
+
+When we reconstruct the evolutionary history of sex distribution in siphonophore species, we obtain the following results:
+```{r sex_results} 
+make.simmap(ultratree, sex, nsim = 100)
+```
+
 ###Sex distribution:
 
-```{r, echo=F}
+```{r sex, echo=F}
 #Simmap SEX
 plotTree(ultratree, lwd = 4)
-make.simmap(ultratree, sex, nsim = 100) -> sex_sim
 sex_sim %>% plotSimmap(lwd = 4, add = T)
@@
 ```
+
+###Presence of palpons:
+```{r palpons, echo=F}
+#Simmap PALPONS
+plotTree(ultratree, lwd = 4)
+palpons_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(palpons_sim)
+```
+
+###Presence of tentilla:
+```{r tentilla, echo=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+tentilla_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim)
+```
+
+###Presence of Nectophores:
+```{r nectosome, echo=F}
+#Simmap NECTOSOME
+plotTree(ultratree, lwd = 4)
+nectosome_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(nectosome_sim)
+```
+
+###Position of Nectosome:
+```{r nectosome_position, echo=F}
+#SIMMAP NECTOSOME POSITION
+plotTree(ultratree, lwd = 4)
+necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green")
+names(colors) = c("Dorsal", "No nectosome", "Ventral")
+nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Nectophore number and types:
+```{r nectophore_number_types, echo=F}
+#Simmap NECTOPHORE TYPES
+plotTree(ultratree, lwd = 4)
+necto_types_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green", "blue")
+names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
+nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Presence of Pneumatophore:
+```{r pneumatophore, echo=F}
+#Simmap PNEUMATOPHORE
+plotTree(ultratree, lwd = 4)
+pneuma_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(pneuma_sim)
+```
+
+###Number of Bract types:
+```{r bract_types}
+#Simmap bract-types
+plotTree(ultratree, lwd = 4)
+bracts_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green3", "blue")
+names(colors) = c("No bracts", "1", "2", "4")
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###Ocean Depth Zones:
+```{r depth_zones, echo=F}
+#Simmap Vertical.Range
+plotTree(ultratree, lwd = 4)
+vrange_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "cyan", "blue", "green", "red")
+names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
+nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+##Continuous traits: Brownian motion reconstructions
+
+```{r cont_bract_types, echo=F}
+contMap(ultratree, bracts)
+```
+
+###Median depth reconstruction:
+```{r median_cont, echo=F}
+depth_value = depth_median
+depthmap <- contMap(dpruned_tree, depth_value) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+
+###Bathymetrics phylomorphospace
+```{r depth_phylomorpho, echo=F}
+par(oma=rep(4,4))
+phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=0.8, label = "horizontal")
+
+#with time from root...
+evotime = fastBM(dpruned_tree,nsim=2)
+evotime_map <- contMap(dpruned_tree,evotime[,1])
+evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
+phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=0.8, label = "horizontal")
+#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
+```
+
+###Abundance of each taxon in the Monterey Bay sampling site
+```{r counts_bars, echo=F}
+par(oma=rep(1,4))
+log_abundance = log(abundance)
+log_abundance["Physalia physalis"] = 0
+plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
+```
+
+###Uncertainty inclusive traitgram for depth range
+```{r uncertainty_depth, echo=F}
+contTrait <- depth_range
+A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
+paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
+trans<-as.character(floor(0:50/2))
+trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
+# plot
+for(i in 0:50){
+  p<-i/length(trans)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
+}
+phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
+```
+
+###Bathymetrics scattergram array
+```{r depth_scattergram, echo=F}
+fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
+```
+
+###Distribution of sex across siphonophore taxa
+```{r ggtree_sex, echo=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = sex)
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Sex", reverse=T)) 
+```
+
+###Overview of main binary traits
+```{r ggtree_bin_heatmap, echo=F}
+p <- ggtree(ultratree)
+p <- p + geom_tiplab(size=3)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
+rownames(f_cdata) = rownames(cdata)
+names(f_cdata)[4] = "Float"
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
+```
+
+##Phylogenetic signal in the character data
+###Agnostic branch length tree generation
+```{r flat_tree}
+flat_tree = tree
+flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
+```
+
+###Phylogenetic signal in bract number
+```{r phylosig_bracts}
+phylosignal(bracts, ultratree)
+physignal(bracts, ultratree)
+```
+
+###Phylogenetic signal in Nectosome, Palpons, Tentilla, and Pneumetophore
+```{r phylosig_binaries, echo=F}
+multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+```
+
+###Phylogenetic signal in VARS bathymetrics
+```{r phylosig_depth, echo=F}
+physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+physignal(depth_range, dpruned_tree)
+multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+```
+
+###Phylogenetic PCA for bathymetrics
+```{r pPCA_depth, echo=F}
+phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+```
+
+###Phylogenetic ANOVA for depth
+Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
+```{r phylANOVA_depth, echo=F}
+phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
+```

### 2017-02-13 character_coding/CharacterEvolution.Rmd
removed messages and warning console outputs from the knitted html

--- 2017-02-12 character_coding/CharacterEvolution.Rmd
@@
 ##The Tree:
-```{r, echo=F}
+```{r ultratree, echo=F, message=F, warning=F}
 par(ask=F)
@@
 
-```{r, include=F}
+```{r simmap_making, echo=F, message=F, warning=F}
 #Make simmap objects
 make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
 
+```
+
+When we reconstruct the evolutionary history of sex distribution in siphonophore species, we obtain the following results:
+```{r sex_results} 
+make.simmap(ultratree, sex, nsim = 100)
 ```
@@
 
-```{r, echo=F}
+```{r sex, echo=F, message=F, warning=F}
 #Simmap SEX
@@
 ###Presence of palpons:
-```{r, echo=F}
+```{r palpons, echo=F, message=F, warning=F}
 #Simmap PALPONS
 plotTree(ultratree, lwd = 4)
-make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
 palpons_sim %>% plotSimmap(lwd = 4, add = T)
@@
 ###Presence of tentilla:
-```{r, echo=F}
+```{r tentilla, echo=F, message=F, warning=F}
 #Simmap TENTILLA
 plotTree(ultratree, lwd = 4)
-make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
 tentilla_sim %>% plotSimmap(lwd = 4, add = T)
@@
 ###Presence of Nectophores:
-```{r, echo=F}
-
+```{r nectosome, echo=F, message=F, warning=F}
+#Simmap NECTOSOME
+plotTree(ultratree, lwd = 4)
+nectosome_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(nectosome_sim)
 ```
@@
 ###Position of Nectosome:
-```{r, echo=F}
-
+```{r nectosome_position, echo=F, message=F, warning=F}
+#SIMMAP NECTOSOME POSITION
+plotTree(ultratree, lwd = 4)
+necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green")
+names(colors) = c("Dorsal", "No nectosome", "Ventral")
+nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 ```
@@
 ###Nectophore number and types:
+```{r nectophore_number_types, echo=F, message=F, warning=F}
+#Simmap NECTOPHORE TYPES
+plotTree(ultratree, lwd = 4)
+necto_types_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green", "blue")
+names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
+nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
 
 ###Presence of Pneumatophore:
-```{r, echo=F}
-
+```{r pneumatophore, echo=F, message=F, warning=F}
+#Simmap PNEUMATOPHORE
+plotTree(ultratree, lwd = 4)
+pneuma_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(pneuma_sim)
 ```
 
-###Ocean Depth Zone:
-```{r, echo=F}
+###Number of Bract types:
+```{r bract_types}
+#Simmap bract-types
+plotTree(ultratree, lwd = 4)
+bracts_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green3", "blue")
+names(colors) = c("No bracts", "1", "2", "4")
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
 
+###Ocean Depth Zones:
+```{r depth_zones, echo=F, message=F, warning=F}
+#Simmap Vertical.Range
+plotTree(ultratree, lwd = 4)
+vrange_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "cyan", "blue", "green", "red")
+names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
+nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 ```
+
+##Continuous traits: Brownian motion reconstructions
+
+```{r cont_bract_types, echo=F, message=F, warning=F}
+contMap(ultratree, bracts)
+```
+
+###Median depth reconstruction:
+```{r median_cont, echo=F, message=F, warning=F}
+depth_value = depth_median
+depthmap <- contMap(dpruned_tree, depth_value) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+
+###Bathymetrics phylomorphospace
+```{r depth_phylomorpho, echo=F, message=F, warning=F}
+par(oma=rep(4,4))
+phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=0.8, label = "horizontal")
+
+#with time from root...
+evotime = fastBM(dpruned_tree,nsim=2)
+evotime_map <- contMap(dpruned_tree,evotime[,1])
+evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
+phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=0.8, label = "horizontal")
+#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
+```
+
+###Abundance of each taxon in the Monterey Bay sampling site
+```{r counts_bars, echo=F, message=F, warning=F}
+par(oma=rep(1,4))
+log_abundance = log(abundance)
+log_abundance["Physalia physalis"] = 0
+plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
+```
+
+###Uncertainty inclusive traitgram for depth range
+```{r uncertainty_depth, echo=F, message=F, warning=F}
+contTrait <- depth_range
+A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
+paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
+trans<-as.character(floor(0:50/2))
+trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
+# plot
+for(i in 0:50){
+  p<-i/length(trans)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
+}
+phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
+```
+
+###Bathymetrics scattergram array
+```{r depth_scattergram, echo=F, message=F, warning=F}
+fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
+```
+
+###Distribution of sex across siphonophore taxa
+```{r ggtree_sex, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = sex)
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Sex", reverse=T)) 
+```
+
+###Overview of main binary traits
+```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
+p <- ggtree(ultratree)
+p <- p + geom_tiplab(size=3)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
+rownames(f_cdata) = rownames(cdata)
+names(f_cdata)[4] = "Float"
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
+```
+
+##Phylogenetic signal in the character data
+###Agnostic branch length tree generation
+```{r flat_tree}
+flat_tree = tree
+flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
+```
+
+###Phylogenetic signal in bract number
+```{r phylosig_bracts}
+phylosignal(bracts, ultratree)
+physignal(bracts, ultratree)
+```
+
+###Phylogenetic signal in Nectosome, Palpons, Tentilla, and Pneumetophore
+```{r phylosig_binaries, echo=F, message=F, warning=F}
+multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+```
+
+###Phylogenetic signal in VARS bathymetrics
+```{r phylosig_depth, echo=F, message=F, warning=F}
+physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+physignal(depth_range, dpruned_tree)
+multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+```
+
+###Phylogenetic PCA for bathymetrics
+```{r pPCA_depth, echo=F, message=F, warning=F}
+phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+```
+
+###Phylogenetic ANOVA for depth
+Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
+```{r phylANOVA_depth, echo=F, message=F, warning=F}
+phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
+phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
+```

### 2017-02-14 character_coding/CharacterEvolution.Rmd

@@
 
-When we reconstruct the evolutionary history of sex distribution in siphonophore species, we obtain the following results:
-```{r sex_results} 
-make.simmap(ultratree, sex, nsim = 100)
-```
+When we reconstruct the evolutionary history of different traits in siphonophore species, we obtain the following results:
 
@@
 ###Number of Bract types:
-```{r bract_types}
+```{r bract_types, echo=F, message=F, warning=F}
 #Simmap bract-types
@@
 ##Continuous traits: Brownian motion reconstructions
-
+###Number of bract types reconstruction:
 ```{r cont_bract_types, echo=F, message=F, warning=F}
@@
 depth_value = depth_median
-depthmap <- contMap(dpruned_tree, depth_value) %>% setMap(colors=c("cyan","blue","black"))
+depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
 plot(depthmap)
@@
 phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=0.8, label = "horizontal")
+```
 
-#with time from root...
+###Bathymetrics phylomorphospace with time from root as color
+```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
 evotime = fastBM(dpruned_tree,nsim=2)
-evotime_map <- contMap(dpruned_tree,evotime[,1])
+evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
 evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
@@
 physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
-physignal(depth_range, dpruned_tree)
 multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)

### 2017-02-14 character_coding/CharacterEvolution.Rmd
"fixed graphical issues"

@@
 
-## Stochastic mapping plots
-
+##Stochastic trait mapping:
 ```{r simmap_making, echo=F, message=F, warning=F}
 #Make simmap objects
+print("SIMMAP Sex Distribution")
 make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+print("SIMMAP Palpons")
 make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+print("SIMMAP Tentilla")
 make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+print("SIMMAP Nectosome")
 make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+print("SIMMAP Nectosome Position")
 make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+print("SIMMAP Nectophore Types")
 make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+print("SIMMAP Pneumatophore")
 make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+print("SIMMAP Bract Types")
 make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+print("SIMMAP Habitat")
 make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
@@
 
-###Sex distribution:
-
-```{r sex, echo=F, message=F, warning=F}
+###SIMMAP Sex distribution:
+```{r sex, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #Simmap SEX
@@
 colors = c("black", "red")
-names(colors) = c("Dioeicous", "Monoeicous")
+names(colors) = c("Dioecious", "Monoecious")
 nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 densityMap(sex_sim)
-
 ```
 
-###Presence of palpons:
-```{r palpons, echo=F, message=F, warning=F}
+Posterior probabilities of states (1 Dioeic)
+
+###SIMMAP Presence of palpons:
+```{r palpons, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #Simmap PALPONS
@@
 
-###Presence of tentilla:
-```{r tentilla, echo=F, message=F, warning=F}
+###SIMMAP Presence of tentilla:
+```{r tentilla, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #Simmap TENTILLA
@@
 
-###Presence of Nectophores:
-```{r nectosome, echo=F, message=F, warning=F}
+###SIMMAP Presence of Nectophores:
+```{r nectosome, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #Simmap NECTOSOME
@@
 
-###Position of Nectosome:
-```{r nectosome_position, echo=F, message=F, warning=F}
+###SIMMAP Position of Nectosome:
+```{r nectosome_position, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #SIMMAP NECTOSOME POSITION
@@
 
-###Nectophore number and types:
-```{r nectophore_number_types, echo=F, message=F, warning=F}
+###SIMMAP Nectophore number and types:
+```{r nectophore_number_types, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
 #Simmap NECTOPHORE TYPES
@@
 
-###Presence of Pneumatophore:
-```{r pneumatophore, echo=F, message=F, warning=F}
+###SIMMAP Presence of Pneumatophore:
+```{r pneumatophore, echo=F, message=F, warning=F, fig.height=7, fig.width=12}
 #Simmap PNEUMATOPHORE
@@
 
-###Number of Bract types:
-```{r bract_types, echo=F, message=F, warning=F}
+###SIMMAP Number of Bract types:
+```{r bract_types, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
 #Simmap bract-types
 plotTree(ultratree, lwd = 4)
-bracts_sim %>% plotSimmap(lwd = 4, add = T)
+bracts_sim %>% plotSimmap(lwd = 4, add = T) -> bract_sim_plot
+#plot(bract_sim_plot[2])
 colors = c("black", "red", "green3", "blue")
@@
 
-###Ocean Depth Zones:
-```{r depth_zones, echo=F, message=F, warning=F}
+###SIMMAP Habitat:
+```{r depth_zones, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
 #Simmap Vertical.Range
@@
 ###Bathymetrics phylomorphospace
-```{r depth_phylomorpho, echo=F, message=F, warning=F}
-par(oma=rep(4,4))
+```{r depth_phylomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
+par(oma=rep(3,4))
 phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=0.8, label = "horizontal")
@@
 ###Bathymetrics phylomorphospace with time from root as color
-```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
+```{r depth_phylochronomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
 evotime = fastBM(dpruned_tree,nsim=2)
@@
 ###Uncertainty inclusive traitgram for depth range
-```{r uncertainty_depth, echo=F, message=F, warning=F}
+```{r uncertainty_depth, echo=F, fig.height=14, fig.width=14, message=F, warning=F}
 contTrait <- depth_range
@@
 
-###Overview of main binary traits
-```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
+###Overview grid of main binary traits
+```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, fig.height=10, fig.width=18}
 p <- ggtree(ultratree)
@@
 ##Phylogenetic signal in the character data
-###Agnostic branch length tree generation
+Agnostic branch length tree generation:
 ```{r flat_tree}
@@
 
+```{r phylosig_bracts_declare, include=F, echo=F, message=F, warning=F}
+physig_bracts_simp = phylosignal(bracts, ultratree)
+physig_bracts_mult = physignal(bracts, ultratree)
+```
 ###Phylogenetic signal in bract number
-```{r phylosig_bracts}
-phylosignal(bracts, ultratree)
-physignal(bracts, ultratree)
+```{r phylosig_bracts, echo=F}
+physig_bracts_simp
+physig_bracts_mult
 ```
 
-###Phylogenetic signal in Nectosome, Palpons, Tentilla, and Pneumetophore
+```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
+binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+```
+###Phylogenetic signal in Binary Traits
 ```{r phylosig_binaries, echo=F, message=F, warning=F}
-multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+binaries_phylosig
 ```
 
+```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
+depth_physig = physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+depth_multisig = multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+```
 ###Phylogenetic signal in VARS bathymetrics
 ```{r phylosig_depth, echo=F, message=F, warning=F}
-physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
-multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+depth_physig
+depth_multisig
 ```
@@
 ###Phylogenetic PCA for bathymetrics
-```{r pPCA_depth, echo=F, message=F, warning=F}
+```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
 phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
@@
 ###Phylogenetic ANOVA for depth
+```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
+z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
+z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
+z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
+z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
+z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
+z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
+z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
+```
 Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
 ```{r phylANOVA_depth, echo=F, message=F, warning=F}
-phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
-phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
+print("Depth vs Sex Distribution")
+z_sex
+print("Depth vs Nectosome")
+z_necto
+print("Depth vs Nectophore Types and Number")
+z_nectotype
+print("Depth vs Nectosome Position")
+z_nectopos
+print("Depth vs Palpons")
+z_palp
+print("Depth vs Tentilla")
+z_tentilla
+print("Depth vs Pneumatophore")
+z_float
 ```

### 2017-02-14 character_coding/CharacterEvolution.Rmd
fixed spelling of monoecious/dioecious

@@
 ---
-title: "## Evolutionary History of the Main Biological Characters of Siphonophores"
+title: "## Evolutionary History of Siphonophore Biology"
 author: "Alejandro Damian Serrano"
-date: "2/11/2017"
 output: html_document
@@
 
-Siphonohores...
+Siphonohores have evolved a fascinating diversity of morphological features, zooid types, life history traits, and habitats. Here we explore the evolutionary history of some of these features.
 
@@
 
-Posterior probabilities of states (1 Dioeic)
+Posterior probabilities of states (0 Dioecious, 1 Monoecious).
 
@@
 
+Posterior probabilities of states (1 Present, 0 Absent).
+
 ###SIMMAP Presence of tentilla:
@@
 
+Posterior probabilities of states (1 Present, 0 Absent).
+
+Physalia physalis may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to Apolemia.
+
 ###SIMMAP Presence of Nectophores:
@@
 ```
+
+Posterior probabilities of states (1 Present, 0 Absent).
 
@@
 
+Posterior probabilities of states (1 Present, 0 Absent).
+
+It is not clear whether or not the stem group of siphonophores had a pneumatophore.
+
 ###SIMMAP Number of Bract types:
@@
 ```
+Trait = Number of Bract Types
 
@@
 ###Bathymetrics scattergram array
-```{r depth_scattergram, echo=F, message=F, warning=F}
+```{r depth_scattergram, echo=F, fig.height=11, fig.width=15, message=F, warning=F}
 fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))

### 2017-02-14 character_coding/CharacterEvolution.Rmd
fixed character ordering in ggtree plots

@@
 par(oma=rep(3,4))
-phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=0.8, label = "horizontal")
+phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
 ```
@@
 evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
-phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=0.8, label = "horizontal")
+phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
 #add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
@@
 exp_tree$edge.length = 10*ultratree$edge.length
-dd = data.frame(taxa  = tree$tip.label, trait = sex)
+dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
 p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
 p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
-p+theme(legend.position="right") + guides(color=guide_legend(title="Sex", reverse=T)) 
+p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
+```
+
+###Pneumatophore presence across siphonophore taxa
+```{r ggtree_pneuma, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(pneuma[match(tree$tip.label, names(pneuma))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Pneumatophore", reverse=T)) 
+```
+
+###Tentilla presence across siphonophore taxa
+```{r ggtree_tentillq, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
+```
+
+###Palpon presence across siphonophore taxa
+```{r ggtree_palpon, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
 ```

### 2017-02-15 manuscript.rmd
added author
added sam to author list
added character coding to manuscript.rmd supplementary content

@@
 	library( phytools ) # phylosig
+  library( vegan )
+  library( ggtree )
+  library( phangorn )
+  library( geiger )
+  library( phylobase )
+  library( fields )
+  library( adephylo )
+  library( geomorph ) #physig
 
@@
 
-Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Cat Munro^1^, Freya Goetz^1,5^, Phil Pugh, Steven H.D. Haddock^4^, Casey W. Dunn^1^*
+Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Cat Munro^1^, Alejandro Damian Serrano^1^, Samuel H Church^1, 5^, Freya Goetz^1,5^, Phil Pugh, Steven H.D. Haddock^4^, Casey W. Dunn^1^*
 
@@
 ^4^ Brown Data Science Practice, Brown University, Brown University, Providence, RI, USA
+
+^5^ Harvard
 
@@
 
+## Character Evolution
+
+Siphonohores have evolved a fascinating diversity of morphological features, zooid types, life history traits, and habitats. Here we explore the evolutionary history of some of these features.
+
 ## Conclusion
@@
 
+```{r char_evo_setup, include=F, echo=F}
+#Working directory
+setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")
+
+#Load data
+read.csv('main_characters.csv', header = T, sep = ',') -> cdata
+rownames(cdata) = cdata$Species
+
+#Load tree
+tree = species_ultrametric
+tree$tip.label = str_replace_all(tree$tip.label,'_',' ')
+tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
+ultratree <- chronos(tree)
+
+#Temporarily prune data
+cdata %<>% filter(rownames(.) %in% tree$tip.label)
+rownames(cdata) = cdata$Species
+cdata %<>% .[-1]
+
+#Named vectors for variables
+sex = cdata$Sex
+names(sex) = rownames(cdata)
+
+nectosome = cdata$Nectosome
+names(nectosome) = rownames(cdata)
+
+necto_pos = cdata$Nectosome.position
+names(necto_pos) = rownames(cdata)
+
+necto_types = cdata$Nectophores
+names(necto_types) = rownames(cdata)
+
+palpons = cdata$Palpons
+names(palpons) = rownames(cdata)
+
+tentilla = cdata$Tentilla
+names(tentilla) = rownames(cdata)
+
+pneuma = cdata$Pneumatophore
+names(pneuma) = rownames(cdata)
+
+vrange = cdata$Vertical.range
+names(vrange) = rownames(cdata)
+
+bracts = cdata$Bract.types
+names(bracts) = rownames(cdata)
+
+#Depth pruned data
+dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
+dpruned_tree <- drop.tip(ultratree, which(!(tree$tip.label %in% rownames(dpruned_data))))
+dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
+
+depth_median = dpruned_data$Depth.Median
+names(depth_median) = rownames(dpruned_data)
+
+depth_mean = dpruned_data$Depth.Mean
+names(depth_mean) = rownames(dpruned_data)
+
+depth_min = dpruned_data$Depth.Min
+names(depth_min) = rownames(dpruned_data)
+
+depth_max = dpruned_data$Depth.Max
+names(depth_max) = rownames(dpruned_data)
+
+depth_sd = dpruned_data$Depth.StdDev
+names(depth_sd) = rownames(dpruned_data)
+
+depth_range = depth_max - depth_min
+
+abundance = dpruned_data$Depth.Count
+names(abundance) = rownames(dpruned_data)
+
+```
+
+The Tree:
+```{r ultratree, echo=F, message=F, warning=F}
+par(ask=F)
+par(mfrow=c(1,1))
+plotTree(ultratree, lwd = 4)
+
+```
+
+Stochastic trait mapping:
+```{r simmap_making, echo=F, message=F, warning=F}
+#Make simmap objects
+print("SIMMAP Sex Distribution")
+make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+print("SIMMAP Palpons")
+make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+print("SIMMAP Tentilla")
+make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+print("SIMMAP Nectosome")
+make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+print("SIMMAP Nectosome Position")
+make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+print("SIMMAP Nectophore Types")
+make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+print("SIMMAP Pneumatophore")
+make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+print("SIMMAP Bract Types")
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+print("SIMMAP Habitat")
+make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
+
+```
+
+When we reconstruct the evolutionary history of different traits in siphonophore species, we obtain the following results:
+
+SIMMAP Sex distribution:
+```{r sex, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap SEX
+plotTree(ultratree, lwd = 4)
+sex_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Dioecious", "Monoecious")
+nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(sex_sim)
+```
+
+Posterior probabilities of states (0 Dioecious, 1 Monoecious).
+
+SIMMAP Presence of palpons:
+```{r palpons, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap PALPONS
+plotTree(ultratree, lwd = 4)
+palpons_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(palpons_sim)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
+
+SIMMAP Presence of tentilla:
+```{r tentilla, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+tentilla_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
+
+Physalia physalis may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to Apolemia.
+
+SIMMAP Presence of Nectophores:
+```{r nectosome, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap NECTOSOME
+plotTree(ultratree, lwd = 4)
+nectosome_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(nectosome_sim)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
+
+SIMMAP Position of Nectosome:
+```{r nectosome_position, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#SIMMAP NECTOSOME POSITION
+plotTree(ultratree, lwd = 4)
+necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green")
+names(colors) = c("Dorsal", "No nectosome", "Ventral")
+nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+SIMMAP Nectophore number and types:
+```{r nectophore_number_types, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap NECTOPHORE TYPES
+plotTree(ultratree, lwd = 4)
+necto_types_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red", "green", "blue")
+names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
+nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+SIMMAP Presence of Pneumatophore:
+```{r pneumatophore, echo=F, message=F, warning=F, fig.height=7, fig.width=12}
+#Simmap PNEUMATOPHORE
+plotTree(ultratree, lwd = 4)
+pneuma_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(pneuma_sim)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
+
+It is not clear whether or not the stem group of siphonophores had a pneumatophore.
+
+SIMMAP Number of Bract types:
+```{r bract_types, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+#Simmap bract-types
+plotTree(ultratree, lwd = 4)
+bracts_sim %>% plotSimmap(lwd = 4, add = T) -> bract_sim_plot
+#plot(bract_sim_plot[2])
+colors = c("black", "red", "green3", "blue")
+names(colors) = c("No bracts", "1", "2", "4")
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+SIMMAP Habitat:
+```{r depth_zones, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+#Simmap Vertical.Range
+plotTree(ultratree, lwd = 4)
+vrange_sim %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "cyan", "blue", "green", "red")
+names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
+nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+Continuous traits: Brownian motion reconstructions
+Number of bract types reconstruction:
+```{r cont_bract_types, echo=F, message=F, warning=F}
+contMap(ultratree, bracts)
+```
+Trait = Number of Bract Types
+
+Median depth reconstruction:
+```{r median_cont, echo=F, message=F, warning=F}
+depth_value = depth_median
+depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+
+Bathymetrics phylomorphospace
+```{r depth_phylomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
+par(oma=rep(3,4))
+phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
+```
+
+Bathymetrics phylomorphospace with time from root as color
+```{r depth_phylochronomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
+evotime = fastBM(dpruned_tree,nsim=2)
+evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
+evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
+phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
+#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
+```
+
+Abundance of each taxon in the Monterey Bay sampling site
+```{r counts_bars, echo=F, message=F, warning=F}
+par(oma=rep(1,4))
+log_abundance = log(abundance)
+log_abundance["Physalia physalis"] = 0
+plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
+```
+
+Uncertainty inclusive traitgram for depth range
+```{r uncertainty_depth, echo=F, fig.height=14, fig.width=14, message=F, warning=F}
+contTrait <- depth_range
+A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
+paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
+trans<-as.character(floor(0:50/2))
+trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
+# plot
+for(i in 0:50){
+  p<-i/length(trans)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
+  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
+}
+phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
+```
+
+Bathymetrics scattergram array
+```{r depth_scattergram, echo=F, fig.height=11, fig.width=15, message=F, warning=F}
+fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
+```
+
+Distribution of sex across siphonophore taxa
+```{r ggtree_sex, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
+```
+
+Pneumatophore presence across siphonophore taxa
+```{r ggtree_pneuma, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(pneuma[match(tree$tip.label, names(pneuma))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Pneumatophore", reverse=T)) 
+```
+
+Tentilla presence across siphonophore taxa
+```{r ggtree_tentillq, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
+```
+
+Palpon presence across siphonophore taxa
+```{r ggtree_palpon, echo=F, message=F, warning=F}
+exp_tree = ultratree
+exp_tree$edge.length = 10*ultratree$edge.length
+dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
+p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
+p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
+p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
+```
+
+Overview grid of main binary traits
+```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, fig.height=10, fig.width=18}
+p <- ggtree(ultratree)
+p <- p + geom_tiplab(size=3)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
+rownames(f_cdata) = rownames(cdata)
+names(f_cdata)[4] = "Float"
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
+```
+
+Phylogenetic signal in the character data
+Agnostic branch length tree generation:
+```{r flat_tree}
+flat_tree = tree
+flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
+```
+
+```{r phylosig_bracts_declare, include=F, echo=F, message=F, warning=F}
+physig_bracts_simp = phylosignal(bracts, ultratree)
+physig_bracts_mult = physignal(bracts, ultratree)
+```
+Phylogenetic signal in bract number
+```{r phylosig_bracts, echo=F}
+physig_bracts_simp
+physig_bracts_mult
+```
+
+```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
+binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+```
+Phylogenetic signal in Binary Traits
+```{r phylosig_binaries, echo=F, message=F, warning=F}
+binaries_phylosig
+```
+
+```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
+depth_physig = physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+depth_multisig = multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+```
+Phylogenetic signal in VARS bathymetrics
+```{r phylosig_depth, echo=F, message=F, warning=F}
+depth_physig
+depth_multisig
+```
+
+Phylogenetic PCA for bathymetrics
+```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
+phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+```
+
+Phylogenetic ANOVA for depth
+```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
+z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
+z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
+z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
+z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
+z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
+z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
+z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
+```
+Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
+```{r phylANOVA_depth, echo=F, message=F, warning=F}
+print("Depth vs Sex Distribution")
+z_sex
+print("Depth vs Nectosome")
+z_necto
+print("Depth vs Nectophore Types and Number")
+z_nectotype
+print("Depth vs Nectosome Position")
+z_nectopos
+print("Depth vs Palpons")
+z_palp
+print("Depth vs Tentilla")
+z_tentilla
+print("Depth vs Pneumatophore")
+z_float
+```
+
 ### Software versions

### 2017-02-21 character_coding/CharacterEvolution.Rmd
added pGLS.Ives for dummy coded sex

@@
 ```
+
+###PGLS
+```{r pgls_depth_sex, echo=F, message=F, warning=F}
+dp_sex <- dpruned_data$Sex
+dp_sex = as.numeric(dp_sex)
+names(dp_sex) = names(depth_median)
+dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
+#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
+#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
+pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
+gls(dp_sex ~ depth_median)
+```

### 2017-02-21 character_coding/CharacterEvolution.Rmd
included phylolm alternative and knitted

--- 2017-02-14 character_coding/CharacterEvolution.Rmd
@@
 library("geomorph")
+library("phylolm")
 
@@
 ```
+
+###PGLS
+```{r pgls_depth_sex, echo=F, message=F, warning=F}
+dp_sex <- dpruned_data$Sex
+dp_sex = as.numeric(dp_sex)
+names(dp_sex) = names(depth_median)
+dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
+#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
+#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
+pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
+phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
+gls(dp_sex ~ depth_median)
+```

### 2017-02-23 manuscript.rmd
added pgls to manuscript

@@
   library( geomorph ) #physig
+  library( phylolm )
 
@@
 
+PGLS - Sex distribution
+```{r pgls_depth_sex, echo=F, message=F, warning=F}
+dp_sex <- dpruned_data$Sex
+dp_sex = as.numeric(dp_sex)
+names(dp_sex) = names(depth_median)
+dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
+#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
+#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
+pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
+phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
+gls(dp_sex ~ depth_median)
+```
+
 ### Software versions

### 2017-02-23 manuscript.rmd
started writing the evolution of monoecy

@@
 Siphonohores have evolved a fascinating diversity of morphological features, zooid types, life history traits, and habitats. Here we explore the evolutionary history of some of these features.
+
+In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters for which their evolutionary history was left equivocal, due to unresolved relationships between physonects.
+With our current cladistic resolution, we were able to determine the following results:
+
+###Evolution of Monoecy
+
+Haeckel XXX noticed for the first time that some siphonophores were monoeicous and others were dioiecious.
+Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
+Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
+There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the Marrus-Erenna clade.
 

### 2017-02-28 manuscript.rmd
Added subheaders to the character evolution section
fleshed out character evolution section - tentilla and habitat use

@@
 Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
-There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the Marrus-Erenna clade.
+There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the *Marrus-Erenna* clade.
 
-## Conclusion
+###The Evolution of Zooid Types
+
+###The Gain and Loss of Tentilla
+
+The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla.
+Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
+Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
+The only two siphonophore genera lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX).
+Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses.
+
+A key issue here is how we code for absence of tentilla, especially in the case of *Physalia physalis*.
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D bands which contain all active and functionally arranged nematocysts used by the organism for prey capture.
+We suspect these may be reduced tentilla.
+So far we do not have enough tissue-specific gene expression and histological data to accurately resolve this question.
+However, when we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change radically (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*.
+
+###The Evolution of Vertical Habitat Use
+
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (ref , *Bargmannia sp* 3888m VARS unpublished).
+While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny suggests the siphonophore MRCA was planktonic, as most extant taxa are.
+Some interesting questions arise from these facts, including 1) what was the bathymetric niche of the siphonophore MRCA, and 2) how did siphonophore's vertical habitat use of the water columns evolve along the phylogeny.
+
+##Discussion
+
+## Conclusions
 

### 2017-02-28 character_coding/CharacterEvolution.Rmd
added Physalia-modified tentilla reconstruction analyses

@@
 
+tentilla2 = tentilla
+tentilla2['Physalia physalis'] = 1
+
 pneuma = cdata$Pneumatophore
@@
 make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+print("SIMMAP Tentilla2")
+make.simmap(ultratree, tentilla2, nsim = 100) -> tentilla_sim2
 print("SIMMAP Nectosome")
@@
 Physalia physalis may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to Apolemia.
+
+SIMMAP Presence of tentilla - *Physalia* corrected:
+```{r tentilla2, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+tentilla_sim2 %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim2)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
 

### 2017-02-28 manuscript.rmd
added Physalia-modified tentilla reconstruction analyses

@@
 Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses.
+However, this reconstruction bears some uncertainty on the state of the MRCA, suggesting a ~30% probability of 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.
 
-A key issue here is how we code for absence of tentilla, especially in the case of *Physalia physalis*.
+A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*.
 The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D bands which contain all active and functionally arranged nematocysts used by the organism for prey capture.
@@
 However, when we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change radically (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*.
+The main implication of this scenario is the unambiguous tentilla-bearing state of the MRCA.
 
@@
 Some interesting questions arise from these facts, including 1) what was the bathymetric niche of the siphonophore MRCA, and 2) how did siphonophore's vertical habitat use of the water columns evolve along the phylogeny.
+Our results indicate a mesopelagic MRCA, with several convergent transition events to epipelagic and bathypelagic waters.
+There was only a single transition to benthic lifestyle on the stem of *Rhodaliidae*.
 
 ##Discussion
+
+The strong phylogenetic signal in the characters traditinally used for taxonomic diagnostics is a positive indicator of the applicability and unambiguity of these characters.
 
@@
 names(tentilla) = rownames(cdata)
+
+tentilla2 = tentilla
+tentilla2['Physalia physalis'] = 1 #In the scenario where we consider Physalias knobs reduced not lost tentilla
 
@@
 
-Physalia physalis may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to Apolemia.
+*Physalia physalis* may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to *Apolemia*.
+
+SIMMAP Presence of tentilla - *Physalia* corrected:
+```{r tentilla2, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+#Simmap TENTILLA
+plotTree(ultratree, lwd = 4)
+tentilla_sim2 %>% plotSimmap(lwd = 4, add = T)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(tentilla_sim2)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
 

### 2017-03-10 character_coding/CharacterEvolution.Rmd
added FactoMineR PCA plots

@@
 library("phylolm")
+library("FactoMineR")
+library("factoextra")
 
@@
 ###Phylogenetic PCA for bathymetrics
-```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
+```{r pPCA_depth, echo=F, fig.height=10, fig.width=19, message=F, warning=F}
 phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+
+dpruned_data[,c(13,14,15,17,18)] %>% PCA(graph=F) -> Zpca
+Zpca %>% fviz_screeplot(addlabels=T)
+Zpca %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
+Zpca %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
+
+phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") -> ppcaZ
+prcomp( dpruned_data[,c(13,14,15,17,18)]) -> rpcaZ
+rpcaZ$rotation = ppcaZ$L
+rpcaZ$x = ppcaZ$S
+rpcaZ$scale = TRUE
+
+rpcaZ %>% fviz_screeplot(addlabels=T)
+rpcaZ %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
+rpcaZ %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
+
+rpcaZ %>% fviz_pca_ind(repulse = T)
+
 ```

### 2017-03-10 manuscript.rmd
added FactoMineR PCA plots

@@
 	library( RGraphics )
+  library( factoextra )
+  library( FactoMineR )
 
@@
 phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+
+dpruned_data[,c(13,14,15,17,18)] %>% PCA(graph=F) -> Zpca
+Zpca %>% fviz_screeplot(addlabels=T)
+Zpca %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
+Zpca %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
+
+phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") -> ppcaZ
+prcomp( dpruned_data[,c(13,14,15,17,18)]) -> rpcaZ
+rpcaZ$rotation = ppcaZ$L
+rpcaZ$x = ppcaZ$S
+rpcaZ$scale = TRUE
+
+rpcaZ %>% fviz_screeplot(addlabels=T)
+rpcaZ %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
+rpcaZ %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
+
+rpcaZ %>% fviz_pca_ind(repulse = T)
 ```

### 2017-04-26 manuscript.rmd
fleshed out ZooidTypeEvolution section

@@
 
-In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters for which their evolutionary history was left equivocal, due to unresolved relationships between physonects.
+In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters left with equivocal evolutionary histories, due to unresolved relationships between physonects.
 With our current cladistic resolution, we were able to determine the following results:
@@
 
-Haeckel XXX noticed for the first time that some siphonophores were monoeicous and others were dioiecious.
+[Citation needed] noticed for the first time that some siphonophores were monoeicous and others were dioiecious.
 Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
@@
 ###The Evolution of Zooid Types
+
+One of the most striking aspects of siphonophore biology is their diversity of unique zooid types.
+Other colonial cnidarians (such as Hydractinia) and some bryozoans (example) have been found to have up to X different zooid types [Citation1 , Citation 2].
+The siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores).
+Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
+Here we reconstruct the evolutionary origins ofthe different zooid types and subtypes on the present transcriptome tree.
+
+Nectophores are retained modified medusae Codonophora use for swimming.
+The nectosome is the region of the colony that develops from the nectosomal growth zone.
+Unlike the siphosomal growth zone, the nectosome does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
+If fact, with the exception of  *Physalia physalis* (which grows small nectophores near the gonodendra), siphonophore nectophores are exclusively found on the nectosome.
+It is likely that the MRCA of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
+We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora.
+The nectosome probably arose as a duplication of the siphosome, followed by functional specialization in propelling the colony.
+Figure X shows the origin of the dorsal nectosome in the Agalmatidae (sensu stricto) from an ancestral ventral nectosome.
+
+All Codonophora have a nectosome, but the number and subtypes of nectophores present varies greatly between species.
+As shown in Figure X, most Codonophora presents the ancestral nectosome with multiple nectophores of the same subtype.
+However, Calycophorans evolved a different system with just 2 nectophores of one type.
+This shift may be associated with the loss of the pneumatophore.
+Not all Calycophorans remained with this arrangement.
+The Hippopodidae returned to bearing multiple identical nectophores, many of which are inactive and serve functions of defense (like a shell to retract in) and buoyancy.
+As in the rest of Calycophorans, the Hippopodids only use 2 nectophores to propel the colony.
+Another interesting shift occurs in the branch leading to Diphyomorpha, where the 2 nectophores specialize into 2 subtypes, associated with a shift into a vertically aligned position and pointed bell shapes.
+The 2 types function together in a coupled hydrodynamic system that allows very fast escape responses (Mackie 1964).
+
+Bracts also present a complex evolutionary history of subtypes (Figure X).
+Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
+As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
+The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and independently evolved into 2 subtypes in *Erenna* and *Nanomia*.
+During the branch leading to Forskalidae, bract subtypes evolved from 1 to 4 morphologically distinct forms.
+Bracts are important for protection of the delicate zooids and to help maintain neutral buoyancy.
+Some calycophorans are able to actively regulate the Na/NH4 balance to adjust their buoyancy (ref).
+
+The ancestral siphonophore certainly had a pneumatophore (Figure X).
+This zooid type helps the colony float, maintain its pose, and regulate geotaxis (Church, 2013).
+Despite its diverse biological functions, it is lost in Calycophorae and never gained again in that clade.
+Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.
+
+Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
+They were present in the MRCA of siphonophores (Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
+These taxa must have found other avenues to effectively circulate nutrients across the colony.
 

### 2017-06-28 manuscript.rmd
included Physalia tentilla GV comment

@@
 Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
-The only two siphonophore genera lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX).
+The only two siphonophore genera regarded as lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX).
 Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses.
@@
 A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D bands which contain all active and functionally arranged nematocysts used by the organism for prey capture.
-We suspect these may be reduced tentilla.
-So far we do not have enough tissue-specific gene expression and histological data to accurately resolve this question.
-However, when we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change radically (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*.
-The main implication of this scenario is the unambiguous tentilla-bearing state of the MRCA.
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D buttons which contain all active and functionally arranged nematocysts used by the organism for prey capture.
+Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
+Hessinger & Ford 1988 (in the Biology of Nematocysts) described these bands as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm.
+This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change to a more parsimonious scenario (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*.
+This result suggests an unambiguous tentilla-bearing state of the MRCA.
 

### 2017-10-10 manuscript.rmd
first pass at dockerfile

@@
   library( phylobase )
-  library( fields )
   library( adephylo )

### 2017-10-10 manuscript.rmd
minor tweeks

@@
 	library( ggtree )
-	library( GGally )
+#	library( GGally )
 	library( fields ) # tim.colors
@@
 	library( gridExtra )
-	library( RGraphics )
+#	library( RGraphics )
   library( factoextra )
@@
 	# Biological
-	library( edgeR )
+#	library( edgeR )
 	library( treeio )
@@
 	library( ape )
-	library( agalmar )	# https://github.com/caseywdunn/agalmar
+#	library( agalmar )	# https://github.com/caseywdunn/agalmar
 	library( picante )  # cor.table
@@
   library( phylobase )
+  library( fields )
   library( adephylo )
@@
 ```{r char_evo_setup, include=F, echo=F}
-#Working directory
-setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")
 
 #Load data
-read.csv('main_characters.csv', header = T, sep = ',') -> cdata
+read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
 rownames(cdata) = cdata$Species
@@
 ultratree <- chronos(tree)
+class( ultratree ) = "phylo"
 

### 2017-10-10 character_coding/CharacterEvolution.Rmd
changed index column sequence for depth physignal

@@
 ```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
-depth_physig = physignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
-depth_multisig = multiPhylosignal(dpruned_data[,c(13,14,15,17,18)], dpruned_tree)
+depth_physig = physignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
+depth_multisig = multiPhylosignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
 ```

### 2017-10-10 character_coding/CharacterEvolution.Rmd
changed index column sequence for depth PCAs

@@
 #Working directory
-setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")
+#setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")
+setwd("character_coding")
 
@@
 ```{r pPCA_depth, echo=F, fig.height=10, fig.width=19, message=F, warning=F}
-phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot
 
-dpruned_data[,c(13,14,15,17,18)] %>% PCA(graph=F) -> Zpca
+dpruned_data[,c(14,15,16,18,19)] %>% PCA(graph=F) -> Zpca
 Zpca %>% fviz_screeplot(addlabels=T)
@@
 
-phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") -> ppcaZ
-prcomp( dpruned_data[,c(13,14,15,17,18)]) -> rpcaZ
+phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") -> ppcaZ
+prcomp( dpruned_data[,c(14,15,16,18,19)]) -> rpcaZ
 rpcaZ$rotation = ppcaZ$L

### 2017-10-10 manuscript.rmd
changed index column sequence for depth PCAs

@@
 ```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
-phyl.pca(dpruned_tree, dpruned_data[,c(13,14,15,17,18)], mode="corr") %>% biplot
+phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot
 
-dpruned_data[,c(13,14,15,17,18)] %>% PCA(graph=F) -> Zpca
+dpruned_data[,c(14,15,16,18,19)] %>% PCA(graph=F) -> Zpca
 Zpca %>% fviz_screeplot(addlabels=T)

### 2017-10-10 manuscript.rmd
reexecuted pdf
simplified dependencies

@@
     fig_caption: yes
-    latex_engine: xelatex
 ---
@@
 	# Biological
-#	library( edgeR )
 	library( treeio )
@@
 	library( ape )
-#	library( agalmar )	# https://github.com/caseywdunn/agalmar
 	library( picante )  # cor.table

### 2017-10-10 manuscript.rmd
updated to new species tree

@@
 	# Biological
-	library( treeio )
 	library( hutan )	# https://github.com/caseywdunn/hutan
@@
 	# Set paths to input data
-	json_name = "agalma/agalma.json"
+	input_file_ml = "agalma/Outfiles/RAxML_bipartitions.alignment.fa"
 
@@
 
-  # Load and parse phylogenies and expression data from the
-  # json file produced by agalma
-
-	# Eventually species tree will be parsed from json, for now specify temporary tree
-	# agalma.data = jsonlite::fromJSON( json_name )
-	# species_tree = read.tree( text=agalma.data$speciestree )
-
-	# For now, specify the tree as plain text
-	species_tree_text = "(((((((((((((((Nanomia_bijuga:0.120563,Agalma_elegans:0.0575132)100:0.0339417,Lychnagalma_utricularia:0.0749548)100:0.0140092,(Cordagalma_sp_:0.117222,Forskalia_asymmetrica:0.0739343)100:0.0113754)100:0.0159061,(Resomia_ornicephala:0.036642,Frillagalma_vityazi:0.0604811)100:0.0190101)100:0.024866,(((Physonect_sp_:0.0317839,Marrus_claudanielis:0.0284231)100:0.0125343,Stephalia_dilata:0.0841317)82:0.00517707,Erenna_richardi:0.0414121)100:0.0124079)100:0.0161254,((((Chelophyes_appendiculata:0.0574339,Abylopsis_tetragona:0.0562432)100:0.11489,(Chuniphyes_multidentata:0.0345263,Kephyes_ovata:0.0306657)100:0.0273795)100:0.0225022,Lilyopsis_fluoracantha:0.197884)100:0.018739,((Prayidae_D27SS7:0.0600685,Prayidae_D27D2:0.0464896)100:0.00909199,Hippopodius_hippopus:0.100508)100:0.011385)100:0.0822577)100:0.0232853,((Bargmannia_elongata:0.0175209,Bargmannia_amoena:0.0201366)100:0.0187736,Bargmannia_lata:0.031329)100:0.0887499)100:0.0300558,((Apolemia_sp_:0.0291459,Apolemia_lanosa:0.0332208)95:0.00837843,Apolemia_white:0.0422333)100:0.0841266)100:0.0313749,(Physalia_physalis:0.082276,Rhizophysa_filiformis:0.105414)100:0.0750894)100:0.133951,(Hydractinia_symbiolongicarpus:0.100907,Podocoryna_carnea:0.246708)100:0.104625)99:0.0219002,Clytia_hemisphaerica:0.331138)97:0.0219604,(Hydra_magnipapillata:0.268083,Ectopleura_larynx:0.273479)100:0.11387)100:0.0827444,Aegina_citrea:0.408585)100:0.148851,(Atolla_vanhoeffeni:0.282409,Alatina_alata:0.328817):0.058436)100:0.419544,(Aiptasia_pallida:0.199088,Nematostella_vectensis:0.188379)100:0);"
-
-	species_tree = read.tree( text=species_tree_text )
+	species_tree = ape::read.tree( input_file_ml )
 
@@
 
+	species_tree = ape::ladderize( species_tree )
+
+	species_tree = 
+	  ape::root( 
+	    species_tree, 
+	    c( "Nematostella_vectensis", "Aiptasia_pallida" ), 
+	    resolve.root=TRUE 
+	   )
+
+  #medusozoa_node = getMRCA( species_tree, c("Alatina_alata", "Nanomia_bijuga") )
+
+  #species_tree = phytools::reroot( species_tree, medusozoa_node )
+
   # Repartition length of edges descended from root so they are non-zero
@@
 
-	# species_tree = ape::root( species_tree, c( "Nematostella_vectensis", "Aiptasia_pallida" ), resolve.root=TRUE )
 	# species_tree = ape::unroot(species_tree)
@@
 	stopifnot( species_tree$Nnode == length( species_tree$node.label ) )
-
-	species_tree = ape::ladderize( species_tree )
 

### 2017-10-11 manuscript.rmd
parses bootstrap and psoterior trees

@@
 	# Configure knitr, see http://yihui.name/knitr/options
-	opts_knit$set( progress = TRUE, verbose = TRUE )
+	opts_knit$set( 
+	  progress = TRUE, 
+	  verbose = TRUE, 
+	  include=FALSE, 
+	  cache=TRUE
+	 )
 
@@
 	# Set paths to input data
-	input_file_ml = "agalma/Outfiles/RAxML_bipartitions.alignment.fa"
+	tree_path = "agalma/Outfiles/"
+	input_file_ml = paste( tree_path, "RAxML_bipartitions.alignment.fa", sep="" )
+	input_file_bootstraps = paste( tree_path, "RAxML_bootstrap.alignment.fa", sep="" )
+
+	input_files_phylobayes = list.files( 
+	  path=tree_path, 
+	  pattern="chain\\d+_trimmed60_poisson.treelist" 
+	 )
+
+	# Analysis parameters
+	burn_in = 400
 
@@
 
-```{r load_data, include=FALSE, cache=TRUE}
+```{r load_data}
 
 	species_tree = ape::read.tree( input_file_ml )
+  bootstrap_trees = ape::read.tree( input_file_bootstraps )
+
+  phylobayes_trees = sapply(
+    input_files_phylobayes,
+    function( pb ) {
+      ape::read.tree( 
+        paste( tree_path, pb, sep = "" ) 
+      )[ -(1:burn_in)  ]
+    }
+  )
+
+  phylobayes_trees = unlist( phylobayes_tree,recursive=FALSE )
 
@@
 	   )
-
-  #medusozoa_node = getMRCA( species_tree, c("Alatina_alata", "Nanomia_bijuga") )
-
-  #species_tree = phytools::reroot( species_tree, medusozoa_node )
 

### 2017-10-12 manuscript.rmd
fixed var name
Added bootstraps and posteriors to species tree

@@
 	library( ggtree )
-#	library( GGally )
 	library( fields ) # tim.colors
@@
 	library( gridExtra )
-#	library( RGraphics )
   library( factoextra )
@@
 	opts_knit$set( 
-	  progress = TRUE, 
-	  verbose = TRUE, 
+	  progress=TRUE, 
+	  verbose=TRUE, 
 	  include=FALSE, 
-	  cache=TRUE
+	  cache=TRUE,
+	  echo=FALSE,
+	  message=FALSE
 	 )
@@
 
+```{r functions}
+
+  get_node_support = function(  tree, treeset, threshold=100 ){
+    # Returns a vector corresponding to nodes in tree, with percent of trees 
+    # in treeset that contain the corresponding bipartition
+    clades = prop.clades(tree, treeset) / length(treeset)
+    nodes = c(rep(NA,length(tree$tip.label)), round(clades*100)) # Offset with NA values for tips
+    nodes[ length(tree$tip.label) + 1 ] = NA # Set the root value to NA, not a robust way to find root but works in this case
+
+    nodes[ nodes >= threshold  ] = NA
+
+    return(nodes)
+  }
+
+  fix_names = function( tree ){
+    # Apply changes to any tip names
+    tree$tip.label = sub( '_', ' ', tree$tip.label )
+    return( tree )
+  }
+
+```
+
 ```{r load_data}
@@
 	species_tree = ape::read.tree( input_file_ml )
+  species_tree = fix_names( species_tree )
+
   bootstrap_trees = ape::read.tree( input_file_bootstraps )
@@
 
-  phylobayes_trees = unlist( phylobayes_tree,recursive=FALSE )
+  phylobayes_trees = unlist( phylobayes_trees,recursive=FALSE )
+  class( phylobayes_trees  ) = "multiPhylo"
+
+  bootstrap_trees = lapply( bootstrap_trees, fix_names )
+  phylobayes_trees = lapply( phylobayes_trees, fix_names )
 
@@
 
-```{r pre_process_species_trees, include=FALSE, cache=TRUE}
+```{r pre_process_species_trees}
 
@@
 	    species_tree, 
-	    c( "Nematostella_vectensis", "Aiptasia_pallida" ), 
+	    c( "Nematostella vectensis", "Aiptasia pallida" ), 
 	    resolve.root=TRUE 
@@
 
-	# Update species names
-	species_tree$tip.label = sub( '_', ' ', species_tree$tip.label )
-
 	# Make ultrametric tree
@@
 	class( species_ultrametric ) = "phylo"
+
+	# Update species names
+	species_ultrametric$tip.label = sub( '_', ' ', species_ultrametric$tip.label )
 
@@
 
-```{r summary_table, results='asis', cache=TRUE, echo=FALSE, comment=NA, warning=FALSE}
+```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
 
@@
 
-```{r species_phylogram, cache=TRUE, echo=FALSE, comment=NA, warning=FALSE, fig.cap="Phylogram of siphonophore relationships.
+```{r species_phylogram, echo=FALSE, comment=NA, warning=FALSE, fig.width=6.5, fig.height=7, fig.cap="Phylogram of siphonophore relationships.
 Node labels indicate bootstrap support percent, unnumbered nodes have 100% support.
@@
 
-	p = ggtree( species_tree ) +
-			geom_text( aes( label=support ), vjust=-.5, hjust=1, size=2.5, col="darkslategray4" ) +
+	p_phylogram = ggtree( species_tree ) +
+			# geom_text( aes( label=support ), vjust=-.5, hjust=1, size=2.5, col="darkslategray4" ) +
+	    geom_text(aes(label=get_node_support(species_tree, bootstrap_trees)), vjust=-.5, hjust=1, size=2.5, col="red") +
+      geom_text(aes(label=get_node_support(species_tree, phylobayes_trees)), vjust=1.5, hjust=1, size=2.5, col="blue") +
 			geom_tiplab( size=3 ) +
@@
 
-	flip( p, euphysonectae_node, calycophora_node )
+	flip( p_phylogram, euphysonectae_node, calycophora_node )
 

### 2017-10-13 manuscript.rmd
minor tweeks

@@
 	  progress=TRUE, 
-	  verbose=TRUE, 
-	  include=FALSE, 
+	  verbose=TRUE)
+
+	opts_chunk$set(
+	#  include=FALSE, 
 	  cache=TRUE,
@@
     # Apply changes to any tip names
-    tree$tip.label = sub( '_', ' ', tree$tip.label )
+    tree$tip.label = sub( '_', ' ', tree$tip.label ) # Change first underscore to space
+    tree$tip.label = sub( 'sp_', 'sp', tree$tip.label ) # Fix trailing _ in Physonect sp_
     return( tree )
@@
 	class( species_ultrametric ) = "phylo"
-
-	# Update species names
-	species_ultrametric$tip.label = sub( '_', ' ', species_ultrametric$tip.label )
 
@@
 
-The phylogenetic relationships between siphonophores species presented here is based on XXX genes from XXX siphonophore species, including new data presented here for XXX species.
+(XX This paragraph on sampling through Agalma analyses) The analyses presented here consider XXX siphonophore species and XXX outgroup species.
+This includes new data for XXX species.
+Summary stats on assemblies XX (Table XX).
+Matrix has XX genes, XX sites, and occupancy is XX.
+
+(XX This paragraph on summarizes phylogeny runs, apart from tree topology) Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
+We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
+This left `r length( phylobayes_trees )` trees in the posterior.
+XXConvergence...
+
 These findings are entirely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 
-```{r session_summary, echo=FALSE, comment=NA}
+```{r session_summary, echo=FALSE, include=TRUE, comment=NA}
 	sessionInfo()

### 2017-10-28 manuscript.rmd
added details to methods section

@@
 Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water scuba dives.
-*Chelophyes appendiculata* and *Hippopodius hippopus* specimen were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 04/13/11.
+*Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 04/13/11.
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, and at the United States National Museum (Smithsonian Institution), Washington, DC.
@@
 
-#### mRNA isolation and RNAseq library preparation
+#### Sequencing
 
 When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table 1).
-mRNA was extracted directly from tissue using Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011) or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
+mRNA was extracted directly from tissue using a variety of methods (supplementary table x): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Qymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
 Extractions were performed according to the manufacturer's instruction.
-In case of very small total RNA quantities, only a single round of bead purification was performed (supplementary table 1).
-Resulting higher rRNA read counts were dealt with further downstream in the bioinformatics workflow.
-Libraries were prepped for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep.
-Master Mix Set (NEB, #E6110S).
-We collected long read paired end Illumina data for *de novo* transcriptome assembly.
+Any resulting higher rRNA read counts were dealt with further downstream in the bioinformatics workflow.
+Libraries were prepared for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep Master Mix Set (NEB, #E6110S).
+We collected long read paired end Illumina data for *de novo* transcriptome assembly.In the case of large tissue inputs, libraries were sequenced separately for each tissue and subsequently were subsampled and pooled *in silico*.
+Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (supplementary table 1).
 Summary statistics for expression libaries are given in Table 1.
@@
 
-Sequence assembly, annotation, phylogenetic analysis, and derivation of expression counts were conducted with the tool Agalma [@Dunn:2013kw], v0.XXX.
-Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package.
-See Supplementary Information for R package version numbers.
+New data were analysed in conjunction with 31 publically available datasets.
+Sequence assembly, annotation, Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma [@Dunn:2013kw], v.
+1.00, and Bayesian Inference (BI) analyses were conducted using Phylobayes[@lartillot2009phylobayes] v.
+1.7a-mpi.
+Source code for all analysis steps, sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
+
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy.
+Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
+ML analyses were conducted using the WAG+Γ model of amino acid substitution, bootstrap values were estimated using 1000 replicates.
+
+Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.
 

### 2017-10-30 manuscript.rmd
updated phylogenetic methods section

@@
 
-New data were analysed in conjunction with 31 publically available datasets.
+New data were analysed in conjunction with 13 publically available datasets, with a total number of 43 species.
 Sequence assembly, annotation, Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma [@Dunn:2013kw], v.
@@
 
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy.
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids.
 Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
-ML analyses were conducted using the WAG+Γ model of amino acid substitution, bootstrap values were estimated using 1000 replicates.
+ML analyses were conducted on the unpartitioned supermatrix using the WAG+Γ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
+BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
+Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
+The CAT-GTR and CAT-poisson models did not converge after a long CPU time, and only the results from the CAT-poisson model are included here.
 
 Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.
+
+#### Hypothesis testing
 
@@
 
-(XX This paragraph on sampling through Agalma analyses) The analyses presented here consider XXX siphonophore species and XXX outgroup species.
+(XX This paragraph on sampling through Agalma analyses) The analyses presented here consider XXX siphonophore species and 8 outgroup species.
 This includes new data for XXX species.
@@
 Cystonectae is the sister group to the remaining siphonophores, and Calycophorae is nested within the paraphyletic "Physonectae".
-In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this XXX-gene transcriptome analysis.
+In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
 These findings include XXX.

### 2017-10-30 manuscript.rmd
removed error messages from simmap, knitted new ms

@@
 Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
-ML analyses were conducted on the unpartitioned supermatrix using the WAG+Γ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
+ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
 BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
@@
 print("SIMMAP Sex Distribution")
-make.simmap(ultratree, sex, nsim = 100) -> sex_sim
+make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
 print("SIMMAP Palpons")
-make.simmap(ultratree, palpons, nsim = 100) -> palpons_sim
+make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
 print("SIMMAP Tentilla")
-make.simmap(ultratree, tentilla, nsim = 100) -> tentilla_sim
+make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
 print("SIMMAP Tentilla2")
-make.simmap(ultratree, tentilla2, nsim = 100) -> tentilla_sim2
+make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
 print("SIMMAP Nectosome")
-make.simmap(ultratree, nectosome, nsim = 100) -> nectosome_sim
+make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
 print("SIMMAP Nectosome Position")
-make.simmap(ultratree, necto_pos, nsim = 100) -> necto_pos_sim
+make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
 print("SIMMAP Nectophore Types")
-make.simmap(ultratree, necto_types, nsim = 100) -> necto_types_sim
+make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
 print("SIMMAP Pneumatophore")
-make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
+make.simmap(ultratree, pneuma, nsim = 100, message=FALSE, pts=T) -> pneuma_sim
 print("SIMMAP Bract Types")
-make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+make.simmap(ultratree, bracts, nsim = 100, message=FALSE) -> bracts_sim
 print("SIMMAP Habitat")
-make.simmap(ultratree, vrange, nsim = 100) -> vrange_sim
+make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim
 
@@
 plotTree(ultratree, lwd = 4)
-sex_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Dioecious", "Monoecious")
+sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
 nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-palpons_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Absent", "Present")
+palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-tentilla_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Absent", "Present")
+tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-tentilla_sim2 %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Absent", "Present")
+tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-nectosome_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Absent", "Present")
+nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red", "green")
 names(colors) = c("Dorsal", "No nectosome", "Ventral")
+necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-necto_types_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red", "green", "blue")
 names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
+necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-pneuma_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "red")
 names(colors) = c("Absent", "Present")
+pneuma_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-bracts_sim %>% plotSimmap(lwd = 4, add = T) -> bract_sim_plot
-#plot(bract_sim_plot[2])
 colors = c("black", "red", "green3", "blue")
 names(colors) = c("No bracts", "1", "2", "4")
+bracts_sim %>% plotSimmap(lwd = 4, add = T,colors=colors) -> bract_sim_plot
+#plot(bract_sim_plot[2])
 nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
@@
 plotTree(ultratree, lwd = 4)
-vrange_sim %>% plotSimmap(lwd = 4, add = T)
 colors = c("black", "cyan", "blue", "green", "red")
 names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
+vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
 nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)

### 2017-10-30 manuscript.rmd
removed pneumatophore plot

@@
 
-Pneumatophore presence across siphonophore taxa
-```{r ggtree_pneuma, echo=F, message=F, warning=F}
-exp_tree = ultratree
-exp_tree$edge.length = 10*ultratree$edge.length
-dd = data.frame(taxa  = tree$tip.label, trait = as.factor(pneuma[match(tree$tip.label, names(pneuma))]))
-p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
-p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
-p+theme(legend.position="right") + guides(color=guide_legend(title="Pneumatophore", reverse=T)) 
-```
-
 Tentilla presence across siphonophore taxa

### 2017-10-30 manuscript.rmd
removed bract types analyses and plot

@@
 
-pneuma = cdata$Pneumatophore
-names(pneuma) = rownames(cdata)
-
 vrange = cdata$Vertical.range
 names(vrange) = rownames(cdata)
-
-bracts = cdata$Bract.types
-names(bracts) = rownames(cdata)
 
@@
 make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
-print("SIMMAP Pneumatophore")
-make.simmap(ultratree, pneuma, nsim = 100, message=FALSE, pts=T) -> pneuma_sim
-print("SIMMAP Bract Types")
-make.simmap(ultratree, bracts, nsim = 100, message=FALSE) -> bracts_sim
 print("SIMMAP Habitat")
@@
 
-SIMMAP Presence of Pneumatophore:
-```{r pneumatophore, echo=F, message=F, warning=F, fig.height=7, fig.width=12}
-#Simmap PNEUMATOPHORE
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-pneuma_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(pneuma_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(pneuma_sim)
-```
-
 Posterior probabilities of states (1 Present, 0 Absent).
@@
 It is not clear whether or not the stem group of siphonophores had a pneumatophore.
-
-SIMMAP Number of Bract types:
-```{r bract_types, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
-#Simmap bract-types
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red", "green3", "blue")
-names(colors) = c("No bracts", "1", "2", "4")
-bracts_sim %>% plotSimmap(lwd = 4, add = T,colors=colors) -> bract_sim_plot
-#plot(bract_sim_plot[2])
-nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-```
 
@@
 Continuous traits: Brownian motion reconstructions
-Number of bract types reconstruction:
-```{r cont_bract_types, echo=F, message=F, warning=F}
-contMap(ultratree, bracts)
-```
-Trait = Number of Bract Types
 
@@
 ```
-Phylogenetic signal in bract number
-```{r phylosig_bracts, echo=F}
-physig_bracts_simp
-physig_bracts_mult
-```
 

### 2017-10-30 manuscript.rmd
removed figure size constraints fromm simmap plots

@@
 SIMMAP Sex distribution:
-```{r sex, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r sex, echo=F, message=F, warning=F}
 #Simmap SEX
@@
 SIMMAP Presence of palpons:
-```{r palpons, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r palpons, echo=F, message=F, warning=F}
 #Simmap PALPONS
@@
 SIMMAP Presence of tentilla:
-```{r tentilla, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r tentilla, echo=F, message=F,warning=F}
 #Simmap TENTILLA
@@
 SIMMAP Presence of tentilla - *Physalia* corrected:
-```{r tentilla2, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r tentilla2, echo=F, message=F, warning=F}
 #Simmap TENTILLA
@@
 SIMMAP Presence of Nectophores:
-```{r nectosome, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r nectosome, echo=F, message=F, warning=F}
 #Simmap NECTOSOME
@@
 SIMMAP Position of Nectosome:
-```{r nectosome_position, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r nectosome_position, echo=F, message=F, warning=F}
 #SIMMAP NECTOSOME POSITION
@@
 SIMMAP Nectophore number and types:
-```{r nectophore_number_types, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
+```{r nectophore_number_types, echo=F, message=F, warning=F}
 #Simmap NECTOPHORE TYPES
@@
 SIMMAP Habitat:
-```{r depth_zones, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+```{r depth_zones, echo=F, message=F, warning=F}
 #Simmap Vertical.Range
@@
 Bathymetrics phylomorphospace
-```{r depth_phylomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
+```{r depth_phylomorpho, echo=F, message=F, warning=F}
 par(oma=rep(3,4))
@@
 Bathymetrics phylomorphospace with time from root as color
-```{r depth_phylochronomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
+```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
 evotime = fastBM(dpruned_tree,nsim=2)
@@
 Uncertainty inclusive traitgram for depth range
-```{r uncertainty_depth, echo=F, fig.height=14, fig.width=14, message=F, warning=F}
+```{r uncertainty_depth, echo=F, message=F, warning=F}
 contTrait <- depth_range
@@
 Bathymetrics scattergram array
-```{r depth_scattergram, echo=F, fig.height=11, fig.width=15, message=F, warning=F}
+```{r depth_scattergram, echo=F, message=F, warning=F}
 fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
@@
 Overview grid of main binary traits
-```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, fig.height=10, fig.width=18}
+```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
 p <- ggtree(ultratree)
@@
 Phylogenetic PCA for bathymetrics
-```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
+```{r pPCA_depth, echo=F, message=F, warning=F}
 phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot

### 2017-10-30 manuscript.rmd
removed phylosig analysis for bract type number
added mention to B. conifera lack of tentilla in manuscript

@@
 Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
-The only two siphonophore genera regarded as lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX).
+The only two siphonophore genera regarded as lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with most members bearing tentilla (*Rhizophysidae* with the exception of *Bathyphysa conifera* and *{Bargmannia, Diphyes}*) (FigXX).
 Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses.
@@
 
-```{r phylosig_bracts_declare, include=F, echo=F, message=F, warning=F}
-physig_bracts_simp = phylosignal(bracts, ultratree)
-physig_bracts_mult = physignal(bracts, ultratree)
-```
-
 ```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}

### 2017-10-30 manuscript.rmd
fixed double ii in Hippopodiidae

@@
 Not all Calycophorans remained with this arrangement.
-The Hippopodidae returned to bearing multiple identical nectophores, many of which are inactive and serve functions of defense (like a shell to retract in) and buoyancy.
-As in the rest of Calycophorans, the Hippopodids only use 2 nectophores to propel the colony.
+The Hippopodiidae returned to bearing multiple identical nectophores, many of which are inactive and serve functions of defense (like a shell to retract in) and buoyancy.
+As in the rest of Calycophorans, the Hippopiids only use 2 nectophores to propel the colony.
 Another interesting shift occurs in the branch leading to Diphyomorpha, where the 2 nectophores specialize into 2 subtypes, associated with a shift into a vertically aligned position and pointed bell shapes.
@@
 As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
-The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and independently evolved into 2 subtypes in *Erenna* and *Nanomia*.
+The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodiidae and independently evolved into 2 subtypes in *Erenna* and *Nanomia*.
 During the branch leading to Forskalidae, bract subtypes evolved from 1 to 4 morphologically distinct forms.

### 2017-10-30 manuscript.rmd
fixed - plots are now centered

--- 2017-10-30 manuscript.rmd
@@
 Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
-ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
+ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
 BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
@@
 
-The Tree:
+#####The Tree
+
 ```{r ultratree, echo=F, message=F, warning=F}
@@
 ```
-
-Stochastic trait mapping:
 ```{r simmap_making, echo=F, message=F, warning=F}
+#Stochastic trait mapping:
 #Make simmap objects
-print("SIMMAP Sex Distribution")
+#print("SIMMAP Sex Distribution")
 make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
-print("SIMMAP Palpons")
+#print("SIMMAP Palpons")
 make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
-print("SIMMAP Tentilla")
+#print("SIMMAP Tentilla")
 make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
-print("SIMMAP Tentilla2")
+#print("SIMMAP Tentilla2")
 make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
-print("SIMMAP Nectosome")
+#print("SIMMAP Nectosome")
 make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
-print("SIMMAP Nectosome Position")
+#print("SIMMAP Nectosome Position")
 make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
-print("SIMMAP Nectophore Types")
+#print("SIMMAP Nectophore Types")
 make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
-print("SIMMAP Habitat")
+#print("SIMMAP Habitat")
 make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim
@@
 
-SIMMAP Sex distribution:
+#####SIMMAP Sex distribution
 ```{r sex, echo=F, message=F, warning=F}
@@
 ```
-
 Posterior probabilities of states (0 Dioecious, 1 Monoecious).
 
-SIMMAP Presence of palpons:
+#####SIMMAP Presence of palpons
 ```{r palpons, echo=F, message=F, warning=F}
@@
 ```
-
 Posterior probabilities of states (1 Present, 0 Absent).
 
-SIMMAP Presence of tentilla:
+#####SIMMAP Presence of tentilla
 ```{r tentilla, echo=F, message=F,warning=F}
@@
 ```
-
 Posterior probabilities of states (1 Present, 0 Absent).
@@
 
-SIMMAP Presence of tentilla - *Physalia* corrected:
+#####SIMMAP Presence of tentilla - *Physalia* corrected
 ```{r tentilla2, echo=F, message=F, warning=F}
@@
 ```
-
 Posterior probabilities of states (1 Present, 0 Absent).
 
-SIMMAP Presence of Nectophores:
+#####SIMMAP Presence of Nectophores
 ```{r nectosome, echo=F, message=F, warning=F}
@@
 ```
-
 Posterior probabilities of states (1 Present, 0 Absent).
 
-SIMMAP Position of Nectosome:
+#####SIMMAP Position of Nectosome
 ```{r nectosome_position, echo=F, message=F, warning=F}
@@
 
-SIMMAP Nectophore number and types:
+#####SIMMAP Nectophore number and types
 ```{r nectophore_number_types, echo=F, message=F, warning=F}
@@
 ```
+Posterior probabilities of states (1 Present, 0 Absent).It is not clear whether or not the stem group of siphonophores had a pneumatophore.
 
-Posterior probabilities of states (1 Present, 0 Absent).
-
-It is not clear whether or not the stem group of siphonophores had a pneumatophore.
-
-SIMMAP Habitat:
+#####SIMMAP Habitat
 ```{r depth_zones, echo=F, message=F, warning=F}
@@
 
-Continuous traits: Brownian motion reconstructions
+####Continuous traits: Brownian motion reconstructions
 
-Median depth reconstruction:
+#####Median depth reconstruction:
 ```{r median_cont, echo=F, message=F, warning=F}
@@
 
-Bathymetrics phylomorphospace
+#####Bathymetrics phylomorphospace
 ```{r depth_phylomorpho, echo=F, message=F, warning=F}
@@
 
-Bathymetrics phylomorphospace with time from root as color
+#####Bathymetrics phylomorphospace with time from root as color
 ```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
@@
 
-Abundance of each taxon in the Monterey Bay sampling site
+####Abundance of each taxon in the Monterey Bay sampling site
 ```{r counts_bars, echo=F, message=F, warning=F}
@@
 
-Uncertainty inclusive traitgram for depth range
+###Uncertainty inclusive traitgram for depth range
 ```{r uncertainty_depth, echo=F, message=F, warning=F}
@@
 
-Bathymetrics scattergram array
+####Bathymetrics scattergram array
 ```{r depth_scattergram, echo=F, message=F, warning=F}
@@
 
-Distribution of sex across siphonophore taxa
+####Distribution of sex across siphonophore taxa
 ```{r ggtree_sex, echo=F, message=F, warning=F}
@@
 
-Tentilla presence across siphonophore taxa
+####Tentilla presence across siphonophore taxa
 ```{r ggtree_tentillq, echo=F, message=F, warning=F}
@@
 
-Palpon presence across siphonophore taxa
+####Palpon presence across siphonophore taxa
 ```{r ggtree_palpon, echo=F, message=F, warning=F}
@@
 
-Overview grid of main binary traits
+#####Overview grid of main binary traits
 ```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
@@
 
-Phylogenetic signal in the character data
-Agnostic branch length tree generation:
+####Phylogenetic signal in the character data
+
+#####Agnostic branch length tree generation:
 ```{r flat_tree}
@@
 ```
-Phylogenetic signal in Binary Traits
+#####Phylogenetic signal in Binary Traits
 ```{r phylosig_binaries, echo=F, message=F, warning=F}
@@
 ```
-Phylogenetic signal in VARS bathymetrics
+#####Phylogenetic signal in VARS bathymetrics
 ```{r phylosig_depth, echo=F, message=F, warning=F}
@@
 
-Phylogenetic PCA for bathymetrics
+#####Phylogenetic PCA for bathymetrics
 ```{r pPCA_depth, echo=F, message=F, warning=F}
@@
 
-Phylogenetic ANOVA for depth
+#####Phylogenetic ANOVA for depth
 ```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
@@
 Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
+
 ```{r phylANOVA_depth, echo=F, message=F, warning=F}
@@
 
-PGLS - Sex distribution
+#####PGLS - Sex distribution
 ```{r pgls_depth_sex, echo=F, message=F, warning=F}

### 2017-10-30 manuscript.rmd
added new library summary stats, xtable

@@
 
+	#summary statistics for libraries
+
+	summary_stats = read.table("Library_summary.csv", header=TRUE, sep=",")
+	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size standard deviation","Number of read pairs","Percent kept after rRNA removal","Percent kept after assembly","Adapter fails","Quality fails","Base Composition fails","Total assembled transcripts","Coding transcripts")
+	specimen_data = read.csv("Specimen_data.csv")
+names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
 ```
@@
 ```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
+
+xtable_summary<-xtable(summary_stats,auto=TRUE)
+align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
+print(xtable_summary, scalebox=0.7, include.rownames=FALSE)
 

### 2017-10-31 character_coding/character_evolution.R
ultratree <- ape::rotate(ultratree, 31) as suggested by @prp

@@ -19,7 +19,7 @@ library("adephylo")
 library("geomorph")
 
 #Working directory
-setwd("~/Dropbox/siphonophore_phylogeny_2017/character_coding")
+setwd("~/siphonophore_phylogeny_2017/character_coding")
 
 #Load data
 read.csv('main_characters.csv', header = T, sep = ',') -> cdata
@@ -92,6 +92,7 @@ names(abundance) = rownames(dpruned_data)
 ### Plots ###
 par(ask=F)
 par(mfrow=c(1,1))
+ultratree <- ape::rotate(ultratree, 31)
 
 plotTree(ultratree, lwd = 4)
 <Return>

### 2017-10-31 character_coding/CharacterEvolution.Rmd
adding bracts presence/absence

@@
 bracts = cdata$Bract.types
+bracts[which(bracts>0)] <- 1
 names(bracts) = rownames(cdata)
@@
 make.simmap(ultratree, pneuma, nsim = 100, pts=T) -> pneuma_sim
-print("SIMMAP Bract Types")
+print("SIMMAP Bracts")
 make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
@@
 It is not clear whether or not the stem group of siphonophores had a pneumatophore.
+
+###SIMMAP BRACTS:
+```{r bracts, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+#Simmap bracts
+plotTree(ultratree, lwd = 4)
+#plot(bract_sim_plot[2])
+colors = c("black", "red")
+names(colors) = c("Absent","Present")
+bracts_sim %>% plotSimmap(lwd = 4, add = T, colors=colors) -> bract_sim_plot
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
 

### 2017-10-31 manuscript.rmd
ultratree <- ape::rotate(ultratree, 31) as suggested by @prp
adding bracts presence/absence

@@
 par(mfrow=c(1,1))
+ultratree <- ape::rotate(ultratree, 31)
 plotTree(ultratree, lwd = 4)
@@
 make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
+#print("SIMMAP Bracts")
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
 #print("SIMMAP Habitat")
@@
 nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+```
+
+###SIMMAP BRACTS:
+```{r bracts, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+#Simmap bracts
+plotTree(ultratree, lwd = 4)
+#plot(bract_sim_plot[2])
+colors = c("black", "red")
+names(colors) = c("Absent","Present")
+bracts_sim %>% plotSimmap(lwd = 4, add = T, colors=colors) -> bract_sim_plot
+nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)

### 2017-10-31 character_coding/CharacterEvolution.Rmd
depth phylo-barplot figure code

@@
 
+###Depth barplot annotations figure
+```{r bathy_barplot, echo=F, message=F, warning=F}
+#Depth_data_VARS
+d <- lapply(runif(30, 0, 30), rnorm, n=100) #random data to be replaced
+ylim <- range(unlist(d))
+bx <- lapply(d, function(y) {
+  dd <- data.frame(y=y)
+  ggplot(dd, aes(x=1, y=y))+geom_boxplot() + ylim(ylim) + theme_inset()
+})
+names(bx) <- 1:30
+bx2 <- lapply(bx, function(g) g)
+p <- ggtree(dpruned_tree) + geom_tiplab()
+HJ <- (-1*p$data$x[1:length(p$data$x)])-1
+inset(p, bx2, width=3, height=.03, hjust= HJ) + xlim(0, 4) + coord_flip()
+
+```
+
 ```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}

### 2017-10-31 manuscript.rmd
included bracts 1/0 in tip annotation overview figure 4a

@@
 
+bracts = cdata$Bract.types
+bracts[which(bracts>0)] <- 1  #bracts changed to presence absence for tip annotation
+names(bracts) = rownames(cdata)
+cdata$Bract.types = bracts
+names(cdata)[8] = "Bracts"
+
 vrange = cdata$Vertical.range
@@
 
-#####Overview grid of main binary traits
+####Overview grid of main binary traits
 ```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
@@
 p <- p + geom_tiplab(size=3)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:8)]
 rownames(f_cdata) = rownames(cdata)
 names(f_cdata)[4] = "Float"
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1.6)
 ```

### 2017-10-31 manuscript.rmd
included eudoxia 1/0 in tip annotation overview figure 4a

@@
 p <- p + geom_tiplab(size=3)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:8)]
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:8,12)]
 rownames(f_cdata) = rownames(cdata)

### 2017-10-31 manuscript.rmd
included monoecy/dioecy in tip annotation overview figure 4a

--- 2017-10-31 manuscript.rmd
@@
 p <- ggtree(ultratree)
-p <- p + geom_tiplab(size=3)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:8)]
+p <- p + geom_tiplab(size=2.2)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
 rownames(f_cdata) = rownames(cdata)
-names(f_cdata)[4] = "Float"
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1.6)
+names(f_cdata)[5] = "Float"
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
 ```

### 2017-11-02 manuscript.rmd
small text modifications
fixed f_cdata row ordering

--- 2017-10-31 manuscript.rmd
@@
 
+Morphological character data were obtained from the literature, or from direct observation of available voucher material.
+
 Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.
@@
 ### Sample collecting and sequencing
-
-```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-
-xtable_summary<-xtable(summary_stats,auto=TRUE)
-align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
-print(xtable_summary, scalebox=0.7, include.rownames=FALSE)
-
-```
-Table 1 (at the end of file): Summary statistics for libraries.
 
@@
 ## Supplementary Information
+
+```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
+
+xtable_summary<-xtable(summary_stats,auto=TRUE)
+align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
+print(xtable_summary, scalebox=0.7, include.rownames=FALSE)
+
+```
+Table 1 (at the end of file): Summary statistics for libraries.
 
@@
 p <- ggtree(ultratree)
-p <- p + geom_tiplab(size=3)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:8)]
+p <- p + geom_tiplab(size=2.2)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
 rownames(f_cdata) = rownames(cdata)
-names(f_cdata)[4] = "Float"
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1.6)
+f_cdata[match(tree$tip.label,rownames(f_cdata)),]
+names(f_cdata)[5] = "Float"
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
 ```

### 2017-11-02 manuscript.rmd
phylobarplot ocde in manuscript

@@
 
+###Depth phylo-barplot
+```{r, bathy_barplot, echo=F, message=F, warning=F}
+d <- lapply(runif(27, 0, 20), rnorm, n=100) #Depth_data_VARS to substitute
+ylim <- range(unlist(d))
+bx <- lapply(d, function(y) {
+  dd <- data.frame(y=y)
+  ggplot(dd, aes(x=1, y=y))+geom_boxplot() + ylim(ylim) + theme_inset()
+})
+names(bx) <- 1:length(ultratree$tip.label)
+bx2 <- lapply(bx, function(g) g)
+p <- ggtree(ultratree) + geom_tiplab()
+HJ <- (-1*p$data$x[1:length(p$data$x)])-1
+inset(p, bx2, width=3, height=.03, hjust= HJ) + xlim(0, 4) + coord_flip()
+```
+
 ####Phylogenetic signal in the character data

### 2017-11-02 manuscript.rmd
reorganized wording in tentilla evolution section

@@
 Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
-Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
-The only two siphonophore genera regarded as lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with most members bearing tentilla (*Rhizophysidae* with the exception of *Bathyphysa conifera* and *{Bargmannia, Diphyes}*) (FigXX).
-Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses.
-However, this reconstruction bears some uncertainty on the state of the MRCA, suggesting a ~30% probability of 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.
+It is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
+The only siphonophores genera regarded as lacking tentilla are *Physalia physalis* and *Apolemia* spp., and *Bathyphysa conifera*.
+Since *B.
+sibogae* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
+When we reconstrauct the evolution of this character on the current phylogeny, we find that 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia*.
+However, this leaves a 30% support for a simple-tentacled MRCA followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.
 
 A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D buttons which contain all active and functionally arranged nematocysts used by the organism for prey capture.
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture.
 Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
-Hessinger & Ford 1988 (in the Biology of Nematocysts) described these bands as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm.
+Hessinger & Ford 1988 (in the Biology of Nematocysts) described *Physalia*'s buttons as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm.
 This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change to a more parsimonious scenario (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*.
-This result suggests an unambiguous tentilla-bearing state of the MRCA.
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Figure 5b).
 

### 2017-11-02 manuscript.rmd
added cliffhanger at the end of the tentilla evolution section
Fixed bract evolution text

@@
 
-Bracts also present a complex evolutionary history of subtypes (Figure X).
 Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
 As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
-The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and independently evolved into 2 subtypes in *Erenna* and *Nanomia*.
-During the branch leading to Forskalidae, bract subtypes evolved from 1 to 4 morphologically distinct forms.
+The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and in *Gymnopraia lapislazula*.
 Bracts are important for protection of the delicate zooids and to help maintain neutral buoyancy.
-Some calycophorans are able to actively regulate the Na/NH4 balance to adjust their buoyancy (ref).
+Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony Bidigare, R.
+R., & Biggs, D.
+C.
+(1980).
 
@@
 Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
-They were present in the MRCA of siphonophores (Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
+They were present in the MRCA of siphonophores (Supp Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
 These taxa must have found other avenues to effectively circulate nutrients across the colony.
@@
 When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Figure 5b).
+
+Siphonophore tentilla present an astounding diversity of sizes, shapes, colors, and nematocyst complements, and some have been observed to rapidly uncoil in contact with prey.
+Future research should explore the evolutionary history of these unique structures.
 

### 2017-11-02 manuscript.rmd
expanded MRCA on 1st mention

@@
 [Citation needed] noticed for the first time that some siphonophores were monoeicous and others were dioiecious.
-Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
+Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA (most recent common ancestor) as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
 Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
@@
 Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
-It is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles.
+It is still an open question whether the MRCA of Siphonophora had simple or branched tentacles.
 The only siphonophores genera regarded as lacking tentilla are *Physalia physalis* and *Apolemia* spp., and *Bathyphysa conifera*.

### 2017-11-02 manuscript.rmd
adjusted tone in some parts of character evolution text

@@
 In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters left with equivocal evolutionary histories, due to unresolved relationships between physonects.
-With our current cladistic resolution, we were able to determine the following results:
+With our current cladistic resolution, we were able to answer some of the questions left open:
 
@@
 Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
-Here we reconstruct the evolutionary origins ofthe different zooid types and subtypes on the present transcriptome tree.
+Here we reconstruct the evolutionary origins ofthe different zooid types on the present transcriptome tree.
 
-Nectophores are retained modified medusae Codonophora use for swimming.
+Nectophores are retained modified medusae that Codonophora use for coordinated colony-level swimming.
 The nectosome is the region of the colony that develops from the nectosomal growth zone.
@@
 If fact, with the exception of  *Physalia physalis* (which grows small nectophores near the gonodendra), siphonophore nectophores are exclusively found on the nectosome.
-It is likely that the MRCA of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
+It is possible that the MRCA of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
 We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora.
 The nectosome probably arose as a duplication of the siphosome, followed by functional specialization in propelling the colony.
-Figure X shows the origin of the dorsal nectosome in the Agalmatidae (sensu stricto) from an ancestral ventral nectosome.
+The nectosome has been lost within Codonophora in the genus *Athorybia*.
 
-All Codonophora have a nectosome, but the number and subtypes of nectophores present varies greatly between species.
+Following the colony development orientation framework presented in (citation), the nectosome can be located in a dorsal or a ventral position.
+Our ancestral reconstructions for this character (Supp figure) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae (sensu stricto) and the branch leading to the *Bargmannia* species.
+All Codonophora (with the exception of *Athorybia* species) have a nectosome, but the number and subtypes of nectophores present varies greatly between species.
 As shown in Figure X, most Codonophora presents the ancestral nectosome with multiple nectophores of the same subtype.
@@
 The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and in *Gymnopraia lapislazula*.
-Bracts are important for protection of the delicate zooids and to help maintain neutral buoyancy.
+Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy.
 Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony Bidigare, R.
@@
 
-The ancestral siphonophore certainly had a pneumatophore (Figure X).
-This zooid type helps the colony float, maintain its pose, and regulate geotaxis (Church, 2013).
-Despite its diverse biological functions, it is lost in Calycophorae and never gained again in that clade.
+The ancestral siphonophore certainly had a pneumatophore (Supp Figure X).
+This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
+Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  (Church, 2013) suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis.
+Despite its multiple biological functions, it was lost in the Calycophorae and never gained again in that clade.
 Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.
@@
 They were present in the MRCA of siphonophores (Supp Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
-These taxa must have found other avenues to effectively circulate nutrients across the colony.
+These taxa might have found other avenues to effectively circulate nutrients across the colony.
 

### 2017-11-03 manuscript.rmd
minor changes, new pdf (commented out ape::rotate(ultratree, 31) )

@@
 
-	summary_stats = read.table("Library_summary.csv", header=TRUE, sep=",")
+	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
 	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size standard deviation","Number of read pairs","Percent kept after rRNA removal","Percent kept after assembly","Adapter fails","Quality fails","Base Composition fails","Total assembled transcripts","Coding transcripts")
-	specimen_data = read.csv("Specimen_data.csv")
+	specimen_data = read.csv("supplementary_info/Specimen_data.csv")
 names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
@@
 
-Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Cat Munro^1^, Alejandro Damian Serrano^1^, Samuel H Church^1, 5^, Freya Goetz^1,5^, Phil Pugh, Steven H.D. Haddock^4^, Casey W. Dunn^1^*
+Catriona Munro^1^, Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Alejandro Damian Serrano^1,9^, Samuel H Church^1, 5^, Freya Goetz^1,6^, Phil Pugh^7^, Steven H.D.
+Haddock^8^, Casey W.
+Dunn^1,9^*
 
-^1^ Department of Ecology and Evolutionary Biology, Brown University, Providence, RI, USA
+^1^ Department of Ecology and Evolutionary Biology, Brown University, Providence, RI 02912, USA
 
-^2^ Current address: Department of Molecular & Cellular Biology, University of California at Davis, Davis, CA, USA
+^2^ Current address: Department of Molecular & Cellular Biology, University of California at Davis, Davis, CA 95616, USA
 
-^3^ Current address: Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA, USA
+^3^ Current address: Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA
 
-^4^ Brown Data Science Practice, Brown University, Brown University, Providence, RI, USA
+^4^ Brown Data Science Practice, Brown University, Brown University, Providence, RI 02912, USA
 
-^5^ Harvard
+^5^ Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, MA 02138, USA
 
-^5^ Current address: Smithsonian, Washington DC, USA
+^6^ Current address: Smithsonian Institution, National Museum of Natural History, Washington, DC 20560, USA
 
-^4^ Monterey Bay Aquarium Research Institute, Moss Landing, CA, USA
+^7^ National Oceanography Centre, Southampton, SO14 3ZH, UK
+
+^8^ Monterey Bay Aquarium Research Institute, Moss Landing, CA 95039, USA
+
+^9^ Current address: Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA
 

### 2017-11-03 manuscript.rmd
fixed issues with ape::rotate

@@
   species_tree = fix_names( species_tree )
+  species_tree_unladder =species_tree
 
@@
 #Load tree
-tree = species_ultrametric
-tree$tip.label = str_replace_all(tree$tip.label,'_',' ')
+tree = species_tree_unladder
 tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
+tree<-rotateNodes(tree,nodes=c(28,31))
 ultratree <- chronos(tree)
@@
 par(mfrow=c(1,1))
-ultratree <- ape::rotate(ultratree, 31)
 plotTree(ultratree, lwd = 4)

### 2017-11-03 manuscript.rmd
correction physical voucher location

@@
 *Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 04/13/11.
-Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, and at the United States National Museum (Smithsonian Institution), Washington, DC.
+Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
 Accession numbers are given in supplementary table X.

### 2017-11-04 manuscript.rmd
minor changes methods

@@
 When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table 1).
-mRNA was extracted directly from tissue using a variety of methods (supplementary table x): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Qymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
+mRNA was extracted directly from tissue using a variety of methods (supplementary table x): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
 Extractions were performed according to the manufacturer's instruction.
@@
 Libraries were prepared for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep Master Mix Set (NEB, #E6110S).
-We collected long read paired end Illumina data for *de novo* transcriptome assembly.In the case of large tissue inputs, libraries were sequenced separately for each tissue and subsequently were subsampled and pooled *in silico*.
+We collected long read paired end Illumina data for *de novo* transcriptome assembly.
+In the case of large tissue inputs, libraries were sequenced separately for each tissue, subsequently subsampled and pooled *in silico*.
 Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (supplementary table 1).

### 2017-11-07 manuscript.rmd
added commented code for tip-driven node rotation

@@
 tree<-rotateNodes(tree,nodes=c(28,31))
+#tree<-rotateNodes(tree,nodes=c(ape::mrca(tree)["Apolemia lanosa", "Agalma elegans"], ape::mrca(tree)["Abylopsis tetragona", "Hippopodius hippopus"]))
 ultratree <- chronos(tree)

### 2017-11-17 manuscript.rmd
added code to make barplot/violinplot into manuscript

@@
 
+######VARS data barplot code
+```{r vars_plot, echo=F, message=F, warning=F}
+#Get Raw Data
+raw_data = read.csv(file="characer_coding/trim_siphdepths.tsv", sep = "\t")[,c(4,3,5:11)]
+raw_data = raw_data[order(raw_data$ConceptName),]
+
+#Clean Raw Data
+names(raw_data)[1] = "Species"
+raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
+raw_data$Species[which(raw_data$Species == "Lilyopsis")] = "Lilyopsis fluoracantha"
+raw_data$Species[which(raw_data$Species == "Agalma")] = "Agalma elegans"
+raw_data$Species[which(raw_data$Species == "Frillagalma")] = "Frillagalma vityazi"
+raw_data$Species[which(raw_data$Species == "Stephalia")] = "Stephalia dilata"
+raw_data$Species[which(raw_data$Species == "Lychnagalma")] = "Lychnagalma utricularia"
+raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
+raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma ordinatum"
+raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
+raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
+
+#Adjust Data to tree structure
+raw_data[which(raw_data$Species %in% ultratree$tip.label),] -> intree_data
+tree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))] %>% length() -> missSPP
+missing_data_proxy <- raw_data[1:missSPP,]
+missing_data_proxy$Species = ultratree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))]
+missing_data_proxy$Depth = rep(20, missSPP)
+parched_data <- rbind(missing_data_proxy, raw_data[which(raw_data$Species %in% ultratree$tip.label),])
+parched_data = parched_data[which(!is.na(parched_data$Depth)),]
+parched_data$Species <- factor(parched_data$Species, levels = ultratree$tip.label)
+
+#Plots for FIGURE
+ggplot(parched_data, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
+ggplot(parched_data, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
+
+```
+
 ### Software versions

### 2017-11-20 manuscript.rmd
corrected typos, updated character driven rotate. Knits

@@
 tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
-tree<-rotateNodes(tree,nodes=c(28,31))
-#tree<-rotateNodes(tree,nodes=c(ape::mrca(tree)["Apolemia lanosa", "Agalma elegans"], ape::mrca(tree)["Abylopsis tetragona", "Hippopodius hippopus"]))
+#tree<-rotateNodes(tree,nodes=c(28,31))
+tree<-rotateNodes(tree,nodes=ape::mrca(tree)["Agalma elegans","Hippopodius hippopus"])
+
 ultratree <- chronos(tree)
@@
 #Get Raw Data
-raw_data = read.csv(file="characer_coding/trim_siphdepths.tsv", sep = "\t")[,c(4,3,5:11)]
+raw_data = read.csv(file="character_coding/trim_siphdepths.tsv", sep = "\t")[,c(4,3,5:11)]
 raw_data = raw_data[order(raw_data$ConceptName),]

### 2017-11-21 manuscript.rmd
removed random phylobarplot code

@@
 
-###Depth phylo-barplot
-```{r, bathy_barplot, echo=F, message=F, warning=F}
-d <- lapply(runif(27, 0, 20), rnorm, n=100) #Depth_data_VARS to substitute
-ylim <- range(unlist(d))
-bx <- lapply(d, function(y) {
-  dd <- data.frame(y=y)
-  ggplot(dd, aes(x=1, y=y))+geom_boxplot() + ylim(ylim) + theme_inset()
-})
-names(bx) <- 1:length(ultratree$tip.label)
-bx2 <- lapply(bx, function(g) g)
-p <- ggtree(ultratree) + geom_tiplab()
-HJ <- (-1*p$data$x[1:length(p$data$x)])-1
-inset(p, bx2, width=3, height=.03, hjust= HJ) + xlim(0, 4) + coord_flip()
-```
-
 ####Phylogenetic signal in the character data

### 2017-11-21 manuscript.rmd
fixed species misnamings from sp\s to sp

@@
 read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
+cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
 rownames(cdata) = cdata$Species

### 2017-11-21 manuscript.rmd
fixed sex state for Rudjakovia, fixed rotation of tree with all siph species

@@
 #tree<-rotateNodes(tree,nodes=c(28,31))
-tree<-rotateNodes(tree,nodes=ape::mrca(tree)["Agalma elegans","Hippopodius hippopus"])
 
@@
 class( ultratree ) = "phylo"
+nodes_to_rotate = c(ape::mrca(ultratree)["Physalia physalis","Kephyes ovata"], ape::mrca(ultratree)["Bargmannia lata","Hippopodius hippopus"])
+ultratree<-rotateNodes(ultratree,nodes=nodes_to_rotate)
 

### 2017-11-21 manuscript.rmd
spaced out text under simmap plots

@@
 ```
+
 Posterior probabilities of states (0 Dioecious, 1 Monoecious).
@@
 ```
+
 Posterior probabilities of states (1 Present, 0 Absent).
@@
 ```
+
 Posterior probabilities of states (1 Present, 0 Absent).It is not clear whether or not the stem group of siphonophores had a pneumatophore.

### 2017-11-21 manuscript.rmd
bracts PP density map

@@
 ###SIMMAP BRACTS:
-```{r bracts, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
+```{r bracts, echo=F, message=F, warning=F}
 #Simmap bracts
@@
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(bracts_sim)
 ```

### 2017-11-21 manuscript.rmd
fixed issues with dpruned_data matching species names

@@
 dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
-dpruned_tree <- drop.tip(ultratree, which(!(tree$tip.label %in% rownames(dpruned_data))))
+dpruned_tree <- drop.tip(ultratree, which(!(ultratree$tip.label %in% rownames(dpruned_data))))
 dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]

### 2017-11-21 manuscript.rmd
inverted depth axis in phenogram w/ uncertainty

@@
 ```{r uncertainty_depth, echo=F, message=F, warning=F}
-contTrait <- depth_range
+contTrait <- -depth_range
 A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)

### 2017-11-21 manuscript.rmd
ggtree colored tip label figures include=F

@@
 ####Distribution of sex across siphonophore taxa
-```{r ggtree_sex, echo=F, message=F, warning=F}
+```{r ggtree_sex, echo=F, message=F, warning=F, include=F}
 exp_tree = ultratree
@@
 ####Tentilla presence across siphonophore taxa
-```{r ggtree_tentillq, echo=F, message=F, warning=F}
+```{r ggtree_tentillq, echo=F, message=F, warning=F, include=F}
 exp_tree = ultratree
@@
 ####Palpon presence across siphonophore taxa
-```{r ggtree_palpon, echo=F, message=F, warning=F}
+```{r ggtree_palpon, echo=F, message=F, warning=F, include=F}
 exp_tree = ultratree

### 2017-11-21 manuscript.rmd
bract included in phylogenetic signal analysis of binary characters

@@
 ```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
-binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
+binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
 ```

### 2017-11-21 manuscript.rmd
started drafting intro

@@
 
-Siphonophores are...
+Siphonophores are a group of 188 valid species within Hydrozoa (Cnidaria), the vast majority of which are members of the plankton (fig.
+1).
+Siphonophores are colonial, and are composed of zooids that are each homologous to solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1988siphonophore; @dunn2006evolution].
+Siphonophores differ significantly from all other colonial hydrozoans in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
+A siphonophore colony arises from a single embryo, which forms a protozooid and a growth zone from which other genetically identical zooids bud asexually [@carre1969etude;@carre1991complete;@carre1967;@carre1995ordre].
+Each zooid arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig 1A, B & C) [@dunn2006evolution].
+
+Most siphonophores are planktonic, with the exception of one pleustonic species (*Physalia physalis*, Portuguese man of war) that lives at the interface of water and air, and a small clade of benthic siphonophores, the Rhodaliidae [@Totton1960;@pugh1983benthic].
 

### 2017-11-21 manuscript.rmd
rearranged supplementary info, removed some extras

@@
 
+```{r char_evo_setup, include=F, echo=F}
+#Load data
+read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
+cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
+rownames(cdata) = cdata$Species
+
+#Load tree
+tree = species_tree_unladder
+tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
+#tree<-rotateNodes(tree,nodes=c(28,31))
+
+ultratree <- chronos(tree)
+class( ultratree ) = "phylo"
+nodes_to_rotate = c(ape::mrca(ultratree)["Physalia physalis","Kephyes ovata"], ape::mrca(ultratree)["Bargmannia lata","Hippopodius hippopus"])
+ultratree<-rotateNodes(ultratree,nodes=nodes_to_rotate)
+
+#Temporarily prune data
+cdata %<>% filter(rownames(.) %in% tree$tip.label)
+rownames(cdata) = cdata$Species
+cdata %<>% .[-1]
+
+#Named vectors for variables
+sex = cdata$Sex
+names(sex) = rownames(cdata)
+
+nectosome = cdata$Nectosome
+names(nectosome) = rownames(cdata)
+
+necto_pos = cdata$Nectosome.position
+names(necto_pos) = rownames(cdata)
+
+necto_types = cdata$Nectophores
+names(necto_types) = rownames(cdata)
+
+palpons = cdata$Palpons
+names(palpons) = rownames(cdata)
+
+tentilla = cdata$Tentilla
+names(tentilla) = rownames(cdata)
+
+tentilla2 = tentilla
+tentilla2['Physalia physalis'] = 1 #In the scenario where we consider Physalias knobs reduced not lost tentilla
+
+bracts = cdata$Bract.types
+bracts[which(bracts>0)] <- 1  #bracts changed to presence absence for tip annotation
+names(bracts) = rownames(cdata)
+cdata$Bract.types = bracts
+names(cdata)[8] = "Bracts"
+
+vrange = cdata$Vertical.range
+names(vrange) = rownames(cdata)
+
+#Depth pruned data
+dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
+dpruned_tree <- drop.tip(ultratree, which(!(ultratree$tip.label %in% rownames(dpruned_data))))
+dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
+
+depth_median = dpruned_data$Depth.Median
+names(depth_median) = rownames(dpruned_data)
+
+depth_mean = dpruned_data$Depth.Mean
+names(depth_mean) = rownames(dpruned_data)
+
+depth_min = dpruned_data$Depth.Min
+names(depth_min) = rownames(dpruned_data)
+
+depth_max = dpruned_data$Depth.Max
+names(depth_max) = rownames(dpruned_data)
+
+depth_sd = dpruned_data$Depth.StdDev
+names(depth_sd) = rownames(dpruned_data)
+
+depth_range = depth_max - depth_min
+
+abundance = dpruned_data$Depth.Count
+names(abundance) = rownames(dpruned_data)
+
+```
+
+```{r simmap_making, echo=F, message=F, warning=F}
+#Stochastic trait mapping:
+#Make simmap objects
+#print("SIMMAP Sex Distribution")
+make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
+#print("SIMMAP Palpons")
+make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
+#print("SIMMAP Tentilla")
+make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
+#print("SIMMAP Tentilla2")
+make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
+#print("SIMMAP Nectosome")
+make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
+#print("SIMMAP Nectosome Position")
+make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+#print("SIMMAP Nectophore Types")
+make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
+#print("SIMMAP Bracts")
+make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+#print("SIMMAP Habitat")
+make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim
+
+```
+
+When we reconstruct the evolutionary history of different traits in siphonophore species, we obtain the following results:
+
+#####SIMMAP Sex distribution
+```{r sex, echo=F, message=F, warning=F}
+#Simmap SEX
+plotTree(ultratree, lwd = 4)
+colors = c("black", "red")
+names(colors) = c("Dioecious", "Monoecious")
+sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
+nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(sex_sim)
+```
+
+Posterior probabilities of states (0 Dioecious, 1 Monoecious).
+
+#####SIMMAP Presence of palpons
+```{r palpons, echo=F, message=F, warning=F}
+#Simmap PALPONS
+plotTree(ultratree, lwd = 4)
+colors = c("black", "red")
+names(colors) = c("Absent", "Present")
+palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(palpons_sim)
+```
+
+Posterior probabilities of states (1 Present, 0 Absent).
+
 Siphonohores have evolved a fascinating diversity of morphological features, zooid types, life history traits, and habitats. Here we explore the evolutionary history of some of these features.
@@
 
-### Supplementary Analyses
-
-```{r char_evo_setup, include=F, echo=F}
-
-#Load data
-read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
-cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
-rownames(cdata) = cdata$Species
-
-#Load tree
-tree = species_tree_unladder
-tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
-#tree<-rotateNodes(tree,nodes=c(28,31))
-
-ultratree <- chronos(tree)
-class( ultratree ) = "phylo"
-nodes_to_rotate = c(ape::mrca(ultratree)["Physalia physalis","Kephyes ovata"], ape::mrca(ultratree)["Bargmannia lata","Hippopodius hippopus"])
-ultratree<-rotateNodes(ultratree,nodes=nodes_to_rotate)
-
-#Temporarily prune data
-cdata %<>% filter(rownames(.) %in% tree$tip.label)
-rownames(cdata) = cdata$Species
-cdata %<>% .[-1]
-
-#Named vectors for variables
-sex = cdata$Sex
-names(sex) = rownames(cdata)
-
-nectosome = cdata$Nectosome
-names(nectosome) = rownames(cdata)
-
-necto_pos = cdata$Nectosome.position
-names(necto_pos) = rownames(cdata)
-
-necto_types = cdata$Nectophores
-names(necto_types) = rownames(cdata)
-
-palpons = cdata$Palpons
-names(palpons) = rownames(cdata)
-
-tentilla = cdata$Tentilla
-names(tentilla) = rownames(cdata)
-
-tentilla2 = tentilla
-tentilla2['Physalia physalis'] = 1 #In the scenario where we consider Physalias knobs reduced not lost tentilla
-
-bracts = cdata$Bract.types
-bracts[which(bracts>0)] <- 1  #bracts changed to presence absence for tip annotation
-names(bracts) = rownames(cdata)
-cdata$Bract.types = bracts
-names(cdata)[8] = "Bracts"
-
-vrange = cdata$Vertical.range
-names(vrange) = rownames(cdata)
-
-#Depth pruned data
-dpruned_data = cdata[which(!is.na(cdata$Depth.Median)), ]
-dpruned_tree <- drop.tip(ultratree, which(!(ultratree$tip.label %in% rownames(dpruned_data))))
-dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
-
-depth_median = dpruned_data$Depth.Median
-names(depth_median) = rownames(dpruned_data)
-
-depth_mean = dpruned_data$Depth.Mean
-names(depth_mean) = rownames(dpruned_data)
-
-depth_min = dpruned_data$Depth.Min
-names(depth_min) = rownames(dpruned_data)
-
-depth_max = dpruned_data$Depth.Max
-names(depth_max) = rownames(dpruned_data)
-
-depth_sd = dpruned_data$Depth.StdDev
-names(depth_sd) = rownames(dpruned_data)
-
-depth_range = depth_max - depth_min
-
-abundance = dpruned_data$Depth.Count
-names(abundance) = rownames(dpruned_data)
-
-```
-
-#####The Tree
-
-```{r ultratree, echo=F, message=F, warning=F}
-par(ask=F)
-par(mfrow=c(1,1))
-plotTree(ultratree, lwd = 4)
-
-```
-```{r simmap_making, echo=F, message=F, warning=F}
-#Stochastic trait mapping:
-#Make simmap objects
-#print("SIMMAP Sex Distribution")
-make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
-#print("SIMMAP Palpons")
-make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
-#print("SIMMAP Tentilla")
-make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
-#print("SIMMAP Tentilla2")
-make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
-#print("SIMMAP Nectosome")
-make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
-#print("SIMMAP Nectosome Position")
-make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
-#print("SIMMAP Nectophore Types")
-make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
-#print("SIMMAP Bracts")
-make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
-#print("SIMMAP Habitat")
-make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim
-
-```
-
-When we reconstruct the evolutionary history of different traits in siphonophore species, we obtain the following results:
-
-#####SIMMAP Sex distribution
-```{r sex, echo=F, message=F, warning=F}
-#Simmap SEX
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Dioecious", "Monoecious")
-sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
-nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(sex_sim)
-```
-
-Posterior probabilities of states (0 Dioecious, 1 Monoecious).
-
-#####SIMMAP Presence of palpons
-```{r palpons, echo=F, message=F, warning=F}
-#Simmap PALPONS
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(palpons_sim)
-```
-
-Posterior probabilities of states (1 Present, 0 Absent).
-
 #####SIMMAP Presence of tentilla
@@
 
-#####SIMMAP Position of Nectosome
-```{r nectosome_position, echo=F, message=F, warning=F}
-#SIMMAP NECTOSOME POSITION
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red", "green")
-names(colors) = c("Dorsal", "No nectosome", "Ventral")
-necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-```
-
 ###SIMMAP BRACTS:
@@
 
-#####SIMMAP Nectophore number and types
-```{r nectophore_number_types, echo=F, message=F, warning=F}
-#Simmap NECTOPHORE TYPES
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red", "green", "blue")
-names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
-necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-```
-
 Posterior probabilities of states (1 Present, 0 Absent).It is not clear whether or not the stem group of siphonophores had a pneumatophore.
-
-#####SIMMAP Habitat
-```{r depth_zones, echo=F, message=F, warning=F}
-#Simmap Vertical.Range
-plotTree(ultratree, lwd = 4)
-colors = c("black", "cyan", "blue", "green", "red")
-names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
-vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-```
-
-####Continuous traits: Brownian motion reconstructions
 
@@
 
-#####Bathymetrics phylomorphospace
-```{r depth_phylomorpho, echo=F, message=F, warning=F}
-par(oma=rep(3,4))
-phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
-```
-
-#####Bathymetrics phylomorphospace with time from root as color
-```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
-evotime = fastBM(dpruned_tree,nsim=2)
-evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
-evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
-phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
-#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
-```
-
-####Abundance of each taxon in the Monterey Bay sampling site
-```{r counts_bars, echo=F, message=F, warning=F}
-par(oma=rep(1,4))
-log_abundance = log(abundance)
-log_abundance["Physalia physalis"] = 0
-plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
-```
-
-###Uncertainty inclusive traitgram for depth range
-```{r uncertainty_depth, echo=F, message=F, warning=F}
-contTrait <- -depth_range
-A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
-paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
-trans<-as.character(floor(0:50/2))
-trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
-# plot
-for(i in 0:50){
-  p<-i/length(trans)
-  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
-  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
-}
-phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
-```
-
-####Bathymetrics scattergram array
-```{r depth_scattergram, echo=F, message=F, warning=F}
-fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
-```
-
-####Distribution of sex across siphonophore taxa
-```{r ggtree_sex, echo=F, message=F, warning=F, include=F}
-exp_tree = ultratree
-exp_tree$edge.length = 10*ultratree$edge.length
-dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
-p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
-p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
-p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
-```
-
-####Tentilla presence across siphonophore taxa
-```{r ggtree_tentillq, echo=F, message=F, warning=F, include=F}
-exp_tree = ultratree
-exp_tree$edge.length = 10*ultratree$edge.length
-dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
-p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
-p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
-p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
-```
-
-####Palpon presence across siphonophore taxa
-```{r ggtree_palpon, echo=F, message=F, warning=F, include=F}
-exp_tree = ultratree
-exp_tree$edge.length = 10*ultratree$edge.length
-dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
-p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
-p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
-p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
-```
-
 ####Overview grid of main binary traits
@@
 rownames(f_cdata) = rownames(cdata)
+f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
-names(f_cdata)[5] = "Float"
 gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
@@
 ####Phylogenetic signal in the character data
-
-#####Agnostic branch length tree generation:
-```{r flat_tree}
-flat_tree = tree
-flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
-```
-
 ```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
@@
 ```
+
 #####Phylogenetic signal in Binary Traits
 ```{r phylosig_binaries, echo=F, message=F, warning=F}
-binaries_phylosig
+binaries_phylosig # make it xtable
 ```
 
-```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
-depth_physig = physignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
-depth_multisig = multiPhylosignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
-```
 #####Phylogenetic signal in VARS bathymetrics
@@
 depth_multisig
-```
-
-#####Phylogenetic PCA for bathymetrics
-```{r pPCA_depth, echo=F, message=F, warning=F}
-phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot
-
-dpruned_data[,c(14,15,16,18,19)] %>% PCA(graph=F) -> Zpca
-Zpca %>% fviz_screeplot(addlabels=T)
-Zpca %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
-Zpca %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
-
-phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") -> ppcaZ
-prcomp( dpruned_data[,c(14,15,16,18,19)]) -> rpcaZ
-rpcaZ$rotation = ppcaZ$L
-rpcaZ$x = ppcaZ$S
-rpcaZ$scale = TRUE
-
-rpcaZ %>% fviz_screeplot(addlabels=T)
-rpcaZ %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
-rpcaZ %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
-
-rpcaZ %>% fviz_pca_ind(repulse = T)
-```
-
-#####Phylogenetic ANOVA for depth
-```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
-z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
-z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
-z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
-z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
-z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
-z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
-z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
-```
-Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
-
-```{r phylANOVA_depth, echo=F, message=F, warning=F}
-print("Depth vs Sex Distribution")
-z_sex
-print("Depth vs Nectosome")
-z_necto
-print("Depth vs Nectophore Types and Number")
-z_nectotype
-print("Depth vs Nectosome Position")
-z_nectopos
-print("Depth vs Palpons")
-z_palp
-print("Depth vs Tentilla")
-z_tentilla
-print("Depth vs Pneumatophore")
-z_float
-```
-
-#####PGLS - Sex distribution
-```{r pgls_depth_sex, echo=F, message=F, warning=F}
-dp_sex <- dpruned_data$Sex
-dp_sex = as.numeric(dp_sex)
-names(dp_sex) = names(depth_median)
-dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
-#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
-#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
-pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
-phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
-gls(dp_sex ~ depth_median)
 ```
@@
 names(raw_data)[1] = "Species"
+raw_data$Species = as.character(raw_data$Species)
 raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
@@
 raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
-raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma ordinatum"
+raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
 raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
 raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
+raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
+raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Thermopalia sp"
+raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Physonect sp"
 
@@
 
+#90% Cut-off of extrame values
+SPDepths = parched_data[,1:2]
+byspp_SPD = split(SPDepths, SPDepths$Species)
+for(i in 1:length(byspp_SPD)){
+  if(nrow(byspp_SPD[[i]])>200){
+#   byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 0.02*nrow(byspp_SPD[[i]]))
+    byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 200)
+  }
+}
+trimmed_SDdata = bind_rows(byspp_SPD)
+
+sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp", "Resomia ornicephala", "Frillagalma vityazi", "Physonect sp", "Marrus claudanielis", "Stephalia dilata", "Thermopalia sp", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Prayidae D27D2", "Prayidae D27SS7", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp", "Rhizophysa filiformis", "Physalia physalis")
+
+trimmed_SDdata$Species = as.character(trimmed_SDdata$Species)
+trimmed_SDdata = trimmed_SDdata[which(!is.na(trimmed_SDdata$Species)),]
+trimmed_SDdata$Species <- factor(trimmed_SDdata$Species, levels = sp_order)
+
+ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
+ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
+
 ```

### 2017-11-21 manuscript.rmd
updated bitex file. Knitted.

--- 2017-11-21 manuscript.rmd
@@
 A siphonophore colony arises from a single embryo, which forms a protozooid and a growth zone from which other genetically identical zooids bud asexually [@carre1969etude;@carre1991complete;@carre1967;@carre1995ordre].
-Each zooid arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig 1A, B & C) [@dunn2006evolution].
+Each zooid arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
+2) [@dunn2006evolution].
 

### 2017-11-21 manuscript.rmd
silenced simmap output for palpons and sex, placed upstream

--- 2017-11-21 manuscript.rmd
@@
 
-```{r simmap_making, echo=F, message=F, warning=F}
+```{r simmap_making, echo=F, message=F, warning=F, include=F}
 #Stochastic trait mapping:
@@
 
-When we reconstruct the evolutionary history of different traits in siphonophore species, we obtain the following results:
-
 #####SIMMAP Sex distribution
-```{r sex, echo=F, message=F, warning=F}
+```{r sex, echo=F, message=F, warning=F, include=F}
 #Simmap SEX
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Dioecious", "Monoecious")
-sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
-nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+#plotTree(ultratree, lwd = 4)
+#colors = c("black", "red")
+#names(colors) = c("Dioecious", "Monoecious")
+#sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
+#nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+#add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 densityMap(sex_sim)
@@
 
-Posterior probabilities of states (0 Dioecious, 1 Monoecious).
-
 #####SIMMAP Presence of palpons
-```{r palpons, echo=F, message=F, warning=F}
+```{r palpons, echo=F, message=F, warning=F, include=F}
 #Simmap PALPONS
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 densityMap(palpons_sim)
 ```
-
-Posterior probabilities of states (1 Present, 0 Absent).
 
@@
 
-Posterior probabilities of states (1 Present, 0 Absent).
-
-*Physalia physalis* may have reduced tentilla, which would indicate there was only one loss of tentilla at the branch leading to *Apolemia*.
-
 #####SIMMAP Presence of tentilla - *Physalia* corrected
@@
 
-#####Phylogenetic signal in VARS bathymetrics
-```{r phylosig_depth, echo=F, message=F, warning=F}
-depth_physig
-depth_multisig
-```
-
 ######VARS data barplot code

### 2017-11-21 manuscript.rmd
piping to pdf

@@
 #add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(sex_sim)
+densityMap(sex_sim) %>% pdf("density_sex_sim")
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(palpons_sim)
+densityMap(palpons_sim)%>% pdf("palpons_sim")
 ```

### 2017-11-21 manuscript.rmd
presence nectophores -> presence nectosome

@@
 
-#####SIMMAP Presence of Nectophores
+#####SIMMAP Presence of Nectosome
 ```{r nectosome, echo=F, message=F, warning=F}

### 2017-11-21 manuscript.rmd
silenced, moved up, outputted to pdf file -- VARS violin plot and Overview ggheatmap figure

@@
 
-#####SIMMAP Sex distribution
 ```{r sex, echo=F, message=F, warning=F, include=F}
@@
 
-#####SIMMAP Presence of palpons
 ```{r palpons, echo=F, message=F, warning=F, include=F}
@@
 densityMap(palpons_sim)%>% pdf("palpons_sim")
+```
+
+```{r vars_plot, echo=F, message=F, warning=F, include=F}
+#Get Raw Data
+raw_data = read.csv(file="character_coding/trim_siphdepths.tsv", sep = "\t")[,c(4,3,5:11)]
+raw_data = raw_data[order(raw_data$ConceptName),]
+
+#Clean Raw Data
+names(raw_data)[1] = "Species"
+raw_data$Species = as.character(raw_data$Species)
+raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
+raw_data$Species[which(raw_data$Species == "Lilyopsis")] = "Lilyopsis fluoracantha"
+raw_data$Species[which(raw_data$Species == "Agalma")] = "Agalma elegans"
+raw_data$Species[which(raw_data$Species == "Frillagalma")] = "Frillagalma vityazi"
+raw_data$Species[which(raw_data$Species == "Stephalia")] = "Stephalia dilata"
+raw_data$Species[which(raw_data$Species == "Lychnagalma")] = "Lychnagalma utricularia"
+raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
+raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
+raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
+raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
+raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
+raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Thermopalia sp"
+raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Physonect sp"
+
+#Adjust Data to tree structure
+raw_data[which(raw_data$Species %in% ultratree$tip.label),] -> intree_data
+tree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))] %>% length() -> missSPP
+missing_data_proxy <- raw_data[1:missSPP,]
+missing_data_proxy$Species = ultratree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))]
+missing_data_proxy$Depth = rep(20, missSPP)
+parched_data <- rbind(missing_data_proxy, raw_data[which(raw_data$Species %in% ultratree$tip.label),])
+parched_data = parched_data[which(!is.na(parched_data$Depth)),]
+parched_data$Species <- factor(parched_data$Species, levels = ultratree$tip.label)
+
+#90% Cut-off of extrame values
+SPDepths = parched_data[,1:2]
+byspp_SPD = split(SPDepths, SPDepths$Species)
+for(i in 1:length(byspp_SPD)){
+  if(nrow(byspp_SPD[[i]])>200){
+#   byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 0.02*nrow(byspp_SPD[[i]]))
+    byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 200)
+  }
+}
+trimmed_SDdata = bind_rows(byspp_SPD)
+
+sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp", "Resomia ornicephala", "Frillagalma vityazi", "Physonect sp", "Marrus claudanielis", "Stephalia dilata", "Thermopalia sp", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Prayidae D27D2", "Prayidae D27SS7", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp", "Rhizophysa filiformis", "Physalia physalis")
+
+trimmed_SDdata$Species = as.character(trimmed_SDdata$Species)
+trimmed_SDdata = trimmed_SDdata[which(!is.na(trimmed_SDdata$Species)),]
+trimmed_SDdata$Species <- factor(trimmed_SDdata$Species, levels = sp_order)
+
+#ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
+ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic() %>% pdf("trimmed_violinVars.pdf")
+
+```
+
+```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, include=F}
+p <- ggtree(ultratree)
+p <- p + geom_tiplab(size=2.2)
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
+rownames(f_cdata) = rownames(cdata)
+f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
+f_cdata[match(tree$tip.label,rownames(f_cdata)),]
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1) %>% pdf("overview_ggheatmap_chars.pdf")
 ```
@@
 
-Posterior probabilities of states (1 Present, 0 Absent).
-
 #####SIMMAP Presence of Nectosome
@@
 
-###SIMMAP BRACTS:
-```{r bracts, echo=F, message=F, warning=F}
-#Simmap bracts
-plotTree(ultratree, lwd = 4)
-#plot(bract_sim_plot[2])
-colors = c("black", "red")
-names(colors) = c("Absent","Present")
-bracts_sim %>% plotSimmap(lwd = 4, add = T, colors=colors) -> bract_sim_plot
-nodelabels(pie=(describe.simmap(bracts_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(bracts_sim)
-```
-
-Posterior probabilities of states (1 Present, 0 Absent).It is not clear whether or not the stem group of siphonophores had a pneumatophore.
-
 #####Median depth reconstruction:
@@
 plot(depthmap)
-```
-
-####Overview grid of main binary traits
-```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
-p <- ggtree(ultratree)
-p <- p + geom_tiplab(size=2.2)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
-rownames(f_cdata) = rownames(cdata)
-f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
-f_cdata[match(tree$tip.label,rownames(f_cdata)),]
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
 ```
@@
 
-######VARS data barplot code
-```{r vars_plot, echo=F, message=F, warning=F}
-#Get Raw Data
-raw_data = read.csv(file="character_coding/trim_siphdepths.tsv", sep = "\t")[,c(4,3,5:11)]
-raw_data = raw_data[order(raw_data$ConceptName),]
-
-#Clean Raw Data
-names(raw_data)[1] = "Species"
-raw_data$Species = as.character(raw_data$Species)
-raw_data$Species[which(raw_data$Species == "Kephyes")] = "Kephyes ovata"
-raw_data$Species[which(raw_data$Species == "Lilyopsis")] = "Lilyopsis fluoracantha"
-raw_data$Species[which(raw_data$Species == "Agalma")] = "Agalma elegans"
-raw_data$Species[which(raw_data$Species == "Frillagalma")] = "Frillagalma vityazi"
-raw_data$Species[which(raw_data$Species == "Stephalia")] = "Stephalia dilata"
-raw_data$Species[which(raw_data$Species == "Lychnagalma")] = "Lychnagalma utricularia"
-raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
-raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
-raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
-raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
-raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
-raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Thermopalia sp"
-raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Physonect sp"
-
-#Adjust Data to tree structure
-raw_data[which(raw_data$Species %in% ultratree$tip.label),] -> intree_data
-tree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))] %>% length() -> missSPP
-missing_data_proxy <- raw_data[1:missSPP,]
-missing_data_proxy$Species = ultratree$tip.label[which(!(ultratree$tip.label %in% raw_data$Species))]
-missing_data_proxy$Depth = rep(20, missSPP)
-parched_data <- rbind(missing_data_proxy, raw_data[which(raw_data$Species %in% ultratree$tip.label),])
-parched_data = parched_data[which(!is.na(parched_data$Depth)),]
-parched_data$Species <- factor(parched_data$Species, levels = ultratree$tip.label)
-
-#Plots for FIGURE
-ggplot(parched_data, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
-ggplot(parched_data, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
-
-#90% Cut-off of extrame values
-SPDepths = parched_data[,1:2]
-byspp_SPD = split(SPDepths, SPDepths$Species)
-for(i in 1:length(byspp_SPD)){
-  if(nrow(byspp_SPD[[i]])>200){
-#   byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 0.02*nrow(byspp_SPD[[i]]))
-    byspp_SPD[[i]] = sample_n(byspp_SPD[[i]], 200)
-  }
-}
-trimmed_SDdata = bind_rows(byspp_SPD)
-
-sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp", "Resomia ornicephala", "Frillagalma vityazi", "Physonect sp", "Marrus claudanielis", "Stephalia dilata", "Thermopalia sp", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Prayidae D27D2", "Prayidae D27SS7", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp", "Rhizophysa filiformis", "Physalia physalis")
-
-trimmed_SDdata$Species = as.character(trimmed_SDdata$Species)
-trimmed_SDdata = trimmed_SDdata[which(!is.na(trimmed_SDdata$Species)),]
-trimmed_SDdata$Species <- factor(trimmed_SDdata$Species, levels = sp_order)
-
-ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
-ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
-
-```
-
 ### Software versions

### 2017-11-21 manuscript.rmd
made supplemental phylogenetic signal table into xtable

@@
 ```{r phylosig_binaries, echo=F, message=F, warning=F}
-binaries_phylosig # make it xtable
+ # make it xtable
+xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
+print(xtable_PS, scalebox=0.7)
 ```

### 2017-11-22 manuscript.rmd
added dev.off() to pdf outputs

@@
 #add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(sex_sim) %>% pdf("density_sex_sim")
+pdf("density_sex_sim")
+densityMap(sex_sim)
+dev.off()
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(palpons_sim)%>% pdf("palpons_sim")
+pdf("palpons_sim")
+densityMap(palpons_sim)
+dev.off()
 ```
@@
 #ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
-ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic() %>% pdf("trimmed_violinVars.pdf")
 
+pdf("trimmed_violinVars.pdf")
+ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
+dev.off()
 ```
@@
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1) %>% pdf("overview_ggheatmap_chars.pdf")
+pdf("overview_ggheatmap_chars.pdf")
+gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
+dev.off()
 ```

### 2017-11-22 manuscript.rmd
fixed path to folder figures_for_AI and added par(ask=F) to avoid Hit Return <>

@@
 ```{r sex, echo=F, message=F, warning=F, include=F}
+par(ask=F)
 #Simmap SEX
@@
 #add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-pdf("density_sex_sim")
+pdf("figures_for_AI/density_sex_sim.pdf")
 densityMap(sex_sim)
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-pdf("palpons_sim")
+pdf("figures_for_AI/palpons_sim.pdf")
 densityMap(palpons_sim)
@@
 
-pdf("trimmed_violinVars.pdf")
+pdf("figures_for_AI/trimmed_violinVars.pdf")
 ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
@@
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
-pdf("overview_ggheatmap_chars.pdf")
+pdf("figures_for_AI/overview_ggheatmap_chars.pdf")
 gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)

### 2017-11-22 manuscript.rmd
updated figures. New knit

@@
 
-\* Corresponding author, casey_dunn@brown.edu
+\* Corresponding author, casey.dunn@yale.edu
 
@@
 Most siphonophores are planktonic, with the exception of one pleustonic species (*Physalia physalis*, Portuguese man of war) that lives at the interface of water and air, and a small clade of benthic siphonophores, the Rhodaliidae [@Totton1960;@pugh1983benthic].
+
+![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top, and the ventral side to the left.
+Adapted from http://commons.wikimedia.org/wiki/File:Nanomia_bijuga_whole_animal_and_growth_zones.svg, drawn by Freya Goetz.
+(*A*) Overview of the whole mature colony.
+(*B*) Inset of the nectosomal growth zone with pneumatophore.
+A series of buds gives rise to nectophores.
+(*C*) Inset of the siphosomal growth zone.
+Probuds subdivide to give rise to zooids in repeating-units (cormidia).
+Cormidial boundaries are marked by a gastrozooid.](figures/Figure2.pdf)
 

### 2017-11-22 manuscript.rmd
fixed typo, Casey's email

@@
 
-Catriona Munro^1^, Stefan Siebert^1,2^, Felipe Zapata^1,3^, Mark Howison^4^, Alejandro Damian Serrano^1,9^, Samuel H Church^1, 5^, Freya Goetz^1,6^, Phil Pugh^7^, Steven H.D.
+Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4^, Alejandro Damian Serrano^1,9^, Samuel H Church^1,5^, Freya Goetz^1,6^, Phil Pugh^7^, Steven H.D.
 Haddock^8^, Casey W.
-Dunn^1,9^*
+Dunn^1,9^**
 
@@
 
-\* Corresponding author, casey.dunn@yale.edu
+\* Authors contributed equally
+
+\*\* Corresponding author, casey.dunn@yale.edu
 

### 2017-11-22 manuscript.rmd
removed pie simaps, left only densityMaps in SuppInfo

@@
 #Simmap TENTILLA
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 densityMap(tentilla_sim)
@@
 #Simmap TENTILLA
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 densityMap(tentilla_sim2)
@@
 #Simmap NECTOSOME
-plotTree(ultratree, lwd = 4)
-colors = c("black", "red")
-names(colors) = c("Absent", "Present")
-nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
 densityMap(nectosome_sim)

### 2017-11-22 manuscript.rmd
added align function to phylosig xtable

@@
 xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
+align( xtable_PS ) <- c('l|',rep('p{0.5in}',ncol(binaries_phylosig)) )
 print(xtable_PS, scalebox=0.7)
+
 ```

### 2017-11-22 manuscript.rmd
placeholders for figures 4 and 5 in text, changed Figure reference numbers in text

@@
 
+Figure 4. Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
+
 In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters left with equivocal evolutionary histories, due to unresolved relationships between physonects.
@@
 Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA (most recent common ancestor) as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
-Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
+Figure 5a shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
 There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the *Marrus-Erenna* clade.
+
+Figure 5. Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
 
@@
 
-The ancestral siphonophore certainly had a pneumatophore (Supp Figure X).
+The ancestral siphonophore certainly had a pneumatophore, since both Cystonects and most Codonophorans possess one (Figure 4).
 This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
@@
 Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
-They were present in the MRCA of siphonophores (Supp Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
+They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.
@@
 This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Figure 5b).
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Supp Figure).
 
@@
 
-Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (ref , *Bargmannia sp* 3888m VARS unpublished).
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (ref , Figure 4).
 While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny suggests the siphonophore MRCA was planktonic, as most extant taxa are.

### 2017-11-22 manuscript.rmd
included mention to bract-less Physophora hydrostatica in text

@@
 As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
-The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and in *Gymnopraia lapislazula*.
+The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
+gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula*.
 Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy.
-Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony Bidigare, R.
-R., & Biggs, D.
-C.
-(1980).
+Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
 

### 2017-11-22 manuscript.rmd
included mention to loss of palpons a 3rd time with M. claudanielis in text

@@
 Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
-They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae.
+They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Bargmannia*, in Calycophorans, and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.

### 2017-11-22 manuscript.rmd
commmented out AI output lines so we dont have to overwrite them every time the code runs

@@
 #add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-pdf("figures_for_AI/density_sex_sim.pdf")
-densityMap(sex_sim)
-dev.off()
+# pdf("figures_for_AI/density_sex_sim.pdf")
+# densityMap(sex_sim)
+# dev.off()
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-pdf("figures_for_AI/palpons_sim.pdf")
-densityMap(palpons_sim)
-dev.off()
+# pdf("figures_for_AI/palpons_sim.pdf")
+# densityMap(palpons_sim)
+# dev.off()
 ```
@@
 
-pdf("figures_for_AI/trimmed_violinVars.pdf")
-ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
-dev.off()
+# pdf("figures_for_AI/trimmed_violinVars.pdf")
+# ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
+# dev.off()
 ```
@@
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
-pdf("figures_for_AI/overview_ggheatmap_chars.pdf")
-gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
-dev.off()
+# pdf("figures_for_AI/overview_ggheatmap_chars.pdf")
+# gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
+# dev.off()
 ```

### 2017-11-22 manuscript.rmd
changed Bargmannia to Pyrostephidae in Palpon loss text

@@
 Our ancestral reconstructions for this character (Supp figure) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae (sensu stricto) and the branch leading to the *Bargmannia* species.
+
 All Codonophora (with the exception of *Athorybia* species) have a nectosome, but the number and subtypes of nectophores present varies greatly between species.
@@
 Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
-They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Bargmannia*, in Calycophorans, and in *Marrus claudanielis*.
+They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.

### 2017-11-22 manuscript.rmd
adjusted table, added latex stylesheet

--- 2017-11-22 manuscript.rmd
@@
     fig_caption: yes
+    includes:
+            in_header: mystyles.sty
 ---
@@
 
-```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-
-xtable_summary<-xtable(summary_stats,auto=TRUE)
-align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
-print(xtable_summary, scalebox=0.7, include.rownames=FALSE)
-
-```
-Table 1 (at the end of file): Summary statistics for libraries.
-
 #####SIMMAP Presence of tentilla
@@
 binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
+names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
 ```
@@
 #####Phylogenetic signal in Binary Traits
-```{r phylosig_binaries, echo=F, message=F, warning=F}
+```{r phylosig_binaries, results='asis',echo=F, message=F, warning=F}
  # make it xtable
 xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
-align( xtable_PS ) <- c('l|',rep('p{0.5in}',ncol(binaries_phylosig)) )
-print(xtable_PS, scalebox=0.7)
+print(xtable_PS, scalebox=0.7,comment=F)
 
 ```
+```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
 
+xtable_summary<-xtable(summary_stats,auto=TRUE)
+align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
+print(xtable_summary, include.rownames=FALSE,comment=F, floating=FALSE, tabular.environment = "longtable")
+
+```
 ### Software versions

### 2017-11-22 manuscript.rmd
B. sibogae -> B. conifera

--- 2017-11-22 manuscript.rmd
@@
 Since *B.
-sibogae* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
+conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
 When we reconstrauct the evolution of this character on the current phylogeny, we find that 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia*.

### 2017-11-22 manuscript.rmd
minor aesthetic table updates

@@
 	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
-	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size standard deviation","Number of read pairs","Percent kept after rRNA removal","Percent kept after assembly","Adapter fails","Quality fails","Base Composition fails","Total assembled transcripts","Coding transcripts")
+	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
+fails","Total transcripts","Coding transcripts")
 	specimen_data = read.csv("supplementary_info/Specimen_data.csv")
@@
 xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
-print(xtable_PS, scalebox=0.7,comment=F)
+print(xtable_PS,comment=F)
 
 ```
+#####Library assembly summary
+
 ```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
 
-xtable_summary<-xtable(summary_stats,auto=TRUE)
-align( xtable_summary ) <- c('l|',rep('p{0.5in}',ncol(summary_stats)) )
-print(xtable_summary, include.rownames=FALSE,comment=F, floating=FALSE, tabular.environment = "longtable")
+xtable_summary<-xtable(summary_stats)
+align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
+print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
 

### 2017-11-22 manuscript.rmd
added F4 and F5 calls in manuscript

@@
 
-Figure 4. Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
+![Figure 4. Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
 
@@
 
-Figure 5. Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
+![Figure 5. Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.](figures/Figure5.pdf)
 

### 2017-11-22 manuscript.rmd
moved gene occupancy to supplementary info

--- 2017-11-22 manuscript.rmd
@@
 
+####£Agalma analysis
+
+![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
+
+```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
+
+xtable_summary<-xtable(summary_stats)
+align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
+print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
+
+```
+The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.
+
 #####SIMMAP Presence of tentilla
@@
 
-####Phylogenetic signal in the character data
-```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
-binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
-names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
-```
-
 #####Phylogenetic signal in Binary Traits
@@
  # make it xtable
+binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
+names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
 xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
@@
 ```
-#####Library assembly summary
+Supplementary table x.
+Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
 
-```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-
-xtable_summary<-xtable(summary_stats)
-align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
-print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
-
-```
 ### Software versions

### 2017-11-23 manuscript.rmd
changed csl to MPE
updated intro

@@
 bibliography: manuscript.bib
-csl: systematic-biology.csl
+csl: molecular-phylogenetics-and-evolution.csl
 output:
@@
 fails","Total transcripts","Coding transcripts")
-	specimen_data = read.csv("supplementary_info/Specimen_data.csv")
+	specimen_data <- read.csv("supplementary_info/Specimen_data.csv")
 names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
@@
 A siphonophore colony arises from a single embryo, which forms a protozooid and a growth zone from which other genetically identical zooids bud asexually [@carre1969etude;@carre1991complete;@carre1967;@carre1995ordre].
-Each zooid arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
+Each zooid is generated asexually, and arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
 2) [@dunn2006evolution].
+Siphonophores are found at all depths - most species are planktonic, with the exception of one pleustonic species (*Physalia physalis*, Portuguese man of war) that lives at the interface of water and air, and a small clade of benthic siphonophores, the Rhodaliidae [@carre1995ordre;@Totton1960;@pugh1983benthic].
+They are among the most abundant gelatinous predators in the open ocean, and play an important ecological role in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
 
-Most siphonophores are planktonic, with the exception of one pleustonic species (*Physalia physalis*, Portuguese man of war) that lives at the interface of water and air, and a small clade of benthic siphonophores, the Rhodaliidae [@Totton1960;@pugh1983benthic].
+Siphonophores are monophyletic and nested within the Hydroidolina, although relationships among the major lineages of the Hydroidolina remain difficult to resolve [@cartwright2010character;@Cartwright2008kl;@zapata2015phylogenomic].
+The most resolved siphonophore phylogeny to date focused on two genes (16S, 18S) from 52 siphonophore taxa, and resolved many long standing questions about siphonophore biology, including the relationships of the three historically recognised groups (Cystonectae, Physonectae, and Calycophorae) [@Dunn:2005dy].
+The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects (the name Codonophora was given to this clade)[@Dunn:2005dy].
+The Apolemiidae are sister to all other Codonophora, however there was little resolution of deep relationships within Codonophora.
+The history of colony structure and budding was investigated by mapping these traits to the tree, and suggesting a complex history of zooid gain and loss.
+Resolving deep relationships within Codonophora is key to resolving the evolution of several character traits, including sexual systems (monoecy vesus dioecy) or the gain and loss of particular zooids, such as palpons.
+
+Siphonophores are difficult to collect, in part because they are fragile (nets destroy all but the most robust parts of the colony), but also because many species are found in the deep sea [@youngbluth1984water;@Dunn:2005dy].
+Specimen sampling for the previous phylogeny, and this new work, is enabled by modern collection techniques, including blue-water scuba diving and remotely operated vehicles (ROVs).
+Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 34 siphonophore species and 9 outgroup species.
+Using 1071 genes shared across species, we are able to resolve deep relationships within the siphonophore phylogeny and reconstruct the evolutionary history of characters, including zooid type, life history traits, and habitat.
 

### 2017-11-28 manuscript.rmd
Fig. 1 added, New knit
updated results section

@@
 
-	#summary statistics for libraries
-
-	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
-	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
-fails","Total transcripts","Coding transcripts")
-	specimen_data <- read.csv("supplementary_info/Specimen_data.csv")
-names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
 ```
@@
 A siphonophore colony arises from a single embryo, which forms a protozooid and a growth zone from which other genetically identical zooids bud asexually [@carre1969etude;@carre1991complete;@carre1967;@carre1995ordre].
-Each zooid is generated asexually, and arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
+Each zooid is generated asexually, arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
 2) [@dunn2006evolution].
@@
 The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects (the name Codonophora was given to this clade)[@Dunn:2005dy].
+
+![Photographs of representatives of the major groups of siphonophores.
+(A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
+(B) *Bathyphysa conifera*, scale bar = XXX.
+(C) *Hippopodius hippopus*, scale bar = 5 mm.
+(D) *Physophora hydrostatica*, scale bar = 5 mm.
+(E) *Nanomia bijuga*, scale bar = 1 cm.
+(F) *Desmophyes haematogaster*, scale bar = 5 mm.
+(G) *Sphaeronectes christiansonae*, scale bar = 2 mm.
+(H) *Lychnagalma utricularia*, scale bar = 1 cm.
+(I) *Kephyes hiulcus*, scale bar = 2 mm.
+(J) *Praya dubia*, scale bar = 4 cm .
+(K) *Apolemia sp.*, scale bar = 1 cm.](figures/Figure1.jpg)
+
 The Apolemiidae are sister to all other Codonophora, however there was little resolution of deep relationships within Codonophora.
@@
 
-When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table 1).
+When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table x).
 mRNA was extracted directly from tissue using a variety of methods (supplementary table x): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
@@
 In the case of large tissue inputs, libraries were sequenced separately for each tissue, subsequently subsampled and pooled *in silico*.
-Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (supplementary table 1).
-Summary statistics for expression libaries are given in Table 1.
+Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (supplementary table x).
 
@@
 
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids.
 Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
@@
 
-Morphological character data were obtained from the literature, or from direct observation of available voucher material.
-
-Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.
+Morphological character data used in trait mapping were obtained from the literature, or from direct observation of available voucher material.
+We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169, @huelsenbeck2003stochastic].
+Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package.
+See Supplementary Information for R package version numbers.
 
 #### Hypothesis testing
+We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
+As the sexual system of *Rudjakovia sp* is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia sp* included as a monoecious species, and one without.
+We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
+version 0.39 (commit fd68ef5733c095c7000a4f92dc8c0daaddeec3b9) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
+For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (figure 3 x).
+We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
 
@@
 
-### Sample collecting and sequencing
+### Species phylogeny and hypothesis testing
 
-### Species phylogeny
+Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California (table 1).
+All sequence data have been deposited in the sequence read archive (SRA).
+The analyses presented here consider 34 siphonophore species and 9 outgroup species.
+This includes new data for 30 species.
+Summary statistics for expression libaries are given in supplementary table x.
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
+```{r specimen_summary, results='asis', echo=FALSE, comment=NA, warning=FALSE}
+specimen_data <- read.csv("supplementary_info/Specimen_data.csv")
+names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth (m)","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
+
+specimen_data_reduced <- specimen_data[,c(1,2,4,14,16)]
+
+xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="Table 1. A complete list of specimens collected for this work.")
+print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.75')
+
+```
+Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
+We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
+This left `r length( phylobayes_trees )` trees in the posterior.
+The phylobayes chains did not converge after a long CPU time, and uncertainty remains around the placement of *Erenna richardi*, and also *Nanomia bijuga*.
+Within the ML analysis, the placement of *Erenna richardi* is also unstable.
+Alternative topologies for these nodes are shown in figure 3 x.
 
@@
 
-(XX This paragraph on sampling through Agalma analyses) The analyses presented here consider XXX siphonophore species and 8 outgroup species.
-This includes new data for XXX species.
-Summary stats on assemblies XX (Table XX).
-Matrix has XX genes, XX sites, and occupancy is XX.
+These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
+Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.
+In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
+*Physophora gilmeri*, along with *Lychnagalma utricularia* (not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*.
+*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*.
+*Cordagalma cordiforme* was previously unresolved, while in this analysis *Cordagalma sp* falls within the Agalmatidae.
+Additionally, there is strong support for *Bargmannia* as sister to all non-Apolemid codonophorans.
 
-(XX This paragraph on summarizes phylogeny runs, apart from tree topology) Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
-We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
-This left `r length( phylobayes_trees )` trees in the posterior.
-XXConvergence...
-
-These findings are entirely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
-Cystonectae is the sister group to the remaining siphonophores, and Calycophorae is nested within the paraphyletic "Physonectae".
-In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
-These findings include XXX.
+We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakova sp.*) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakova sp.*) are monophyletic (figure 3 xx).
+In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03).
 
@@
 
-Siphonohores have evolved a fascinating diversity of morphological features, zooid types, life history traits, and habitats. Here we explore the evolutionary history of some of these features.
-
-![Figure 4. Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
+![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
 
@@
 
-####£Agalma analysis
+####Agalma analysis
 
-![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
+![Supplementary Figure 1.
+60% gene occupancy matrix for 41 species across 1,071 genes.
+Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
 
 ```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
+	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
+	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
+fails","Total transcripts","Coding transcripts")
 
-xtable_summary<-xtable(summary_stats)
+xtable_summary<-xtable(summary_stats, caption="Supplementary table 1.The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.")
 align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
@@
 ```
-The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.
 
@@
 names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
-xtable_PS<-xtable(binaries_phylosig,auto=TRUE)
+xtable_PS<-xtable(binaries_phylosig,auto=TRUE, caption="Supplementary table 3.
+Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
+")
 print(xtable_PS,comment=F)
@@
 ```
-Supplementary table x.
-Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
 

### 2017-11-28 manuscript.rmd
Figure 3 now inserted from file.

@@
 
-```{r species_phylogram, echo=FALSE, comment=NA, warning=FALSE, fig.width=6.5, fig.height=7, fig.cap="Phylogram of siphonophore relationships.
-Node labels indicate bootstrap support percent, unnumbered nodes have 100% support.
-The image was rendered with ggtree [@Yu:2016fo]" }
+```{r species_phylogram, echo=FALSE, comment=NA, warning=FALSE }
 
@@
 
+	pdf("figures_for_AI/figure_3_raw.pdf")
 	flip( p_phylogram, euphysonectae_node, calycophora_node )
+	dev.off()
 
 ```
+
+![Phylogram of siphonophore relationships. Node labels indicate bootstrap support percent, unnumbered nodes have 100% support. The image was rendered with ggtree [@Yu:2016fo].](figures/Figure3.pdf)
 

### 2017-11-28 manuscript.rmd
small changes to figures, methods, part of character mapping section

--- 2017-11-28 manuscript.rmd
@@
 
-## Results and Discussion
+## Results
 
@@
 specimen_data_reduced <- specimen_data[,c(1,2,4,14,16)]
+specimen_data_reduced$`New data`<-sub("^N"," ",specimen_data_reduced$`New data`)
 
-xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="Table 1. A complete list of specimens collected for this work.")
+xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="Table 1.
+A complete list of specimens collected for this work.
+New data indicated by Y, blank fields indicate that data are already published.")
 print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.75')
@@
 
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
+These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (figure 3) [@Dunn:2005dy].
 Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.
@@
 ## Character Evolution
+
+The morphological character matrix, and all the code used to generate the character mapping is available in the git repository and in the supplementary materials.
+A  number of traits were mapped to branches on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (fig 4, 5, supplementary figures 2-5).
 
@@
 
+##Discussion
+
+###Siphonophore Phylogeny
+
+The strong phylogenetic signal in the characters traditinally used for taxonomic diagnostics is a positive indicator of the applicability and unambiguity of these characters.
+
 ![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
@@
 
-![Figure 5. Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.](figures/Figure5.pdf)
+![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.](figures/Figure5.pdf)
 
@@
 
-##Discussion
-
-The strong phylogenetic signal in the characters traditinally used for taxonomic diagnostics is a positive indicator of the applicability and unambiguity of these characters.
-
 ## Conclusions
@@
 ## Supplementary Information
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\makeatletter 
+\renewcommand{\thefigure}{S\@arabic\c@figure}
+\renewcommand{\thetable}{S\@arabic\c@table}
+\makeatother
 
@@
 
-![Supplementary Figure 1.
-60% gene occupancy matrix for 41 species across 1,071 genes.
-Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
+![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
 
@@
 
-xtable_summary<-xtable(summary_stats, caption="Supplementary table 1.The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.")
-align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
-print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
+#xtable_summary<-xtable(summary_stats, caption="Supplementary table 1.The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.")
+#align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
+#print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
 
@@
 
-#####SIMMAP Presence of tentilla
-```{r tentilla, echo=F, message=F,warning=F}
+#####Stochastic Character maps
+```{r tentilla, echo=F, message=F,warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla"}
 #Simmap TENTILLA
@@
 
-#####SIMMAP Presence of tentilla - *Physalia* corrected
-```{r tentilla2, echo=F, message=F, warning=F}
+```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with Physalia included"}
 #Simmap TENTILLA
@@
 
-#####SIMMAP Presence of Nectosome
-```{r nectosome, echo=F, message=F, warning=F}
+```{r nectosome, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
 #Simmap NECTOSOME
@@
 ```
-Posterior probabilities of states (1 Present, 0 Absent).
 
-#####Median depth reconstruction:
-```{r median_cont, echo=F, message=F, warning=F}
+```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of median depth of species"}
 depth_value = depth_median
@@
 
-#####Phylogenetic signal in Binary Traits
 ```{r phylosig_binaries, results='asis',echo=F, message=F, warning=F}
@@
 names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
-xtable_PS<-xtable(binaries_phylosig,auto=TRUE, caption="Supplementary table 3.
-Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
+xtable_PS<-xtable(binaries_phylosig,auto=TRUE, caption="Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
 ")

### 2017-11-28 manuscript.rmd
fixed merge issue

@@
 
+	pdf("figures_for_AI/figure_3_raw.pdf")
 	flip( p_phylogram, euphysonectae_node, calycophora_node )
+	dev.off()
 
 ```
+These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (figure 3) [@Dunn:2005dy].
+Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.
+In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
+*Physophora gilmeri*, along with *Lychnagalma utricularia* (not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*.
+*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*.
+*Cordagalma cordiforme* was previously unresolved, while in this analysis *Cordagalma sp* falls within the Agalmatidae.
+Additionally, there is strong support for *Bargmannia* as sister to all non-Apolemid codonophorans.
 
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (figure 3) [@Dunn:2005dy].
+![Phylogram of siphonophore relationships. Node labels indicate bootstrap support percent, unnumbered nodes have 100% support. The image was rendered with ggtree [@Yu:2016fo].](figures/Figure3.pdf)
+
+These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
 Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.

### 2017-11-28 manuscript.rmd
figure 5 legend <<The color gradients shows the reconstructed probability estimate of the discrete character states along the branches.>>

@@
 
-![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.](figures/Figure5.pdf)
+![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
+The color gradients shows the reconstructed probability estimate of the discrete character states along the branches.](figures/Figure5.pdf)
 

### 2017-11-28 manuscript.rmd
shows->show
intermediate = uncertainty. Removed Siph_depth_pruned.csv

@@
 ![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
-The color gradients shows the reconstructed probability estimate of the discrete character states along the branches.](figures/Figure5.pdf)
+The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
+Intermediate values reflect uncertainty.](figures/Figure5.pdf)
 

### 2017-11-29 manuscript.rmd
merged results & discussion - expanded phylogeny discussion

--- 2017-11-28 manuscript.rmd
@@
 
-## Results
+## Results and Discussion
 
 ### Species phylogeny and hypothesis testing
-
 Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California (table 1).
@@
 These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (figure 3) [@Dunn:2005dy].
-Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.
-In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
-*Physophora gilmeri*, along with *Lychnagalma utricularia* (not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*.
-*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*.
-*Cordagalma cordiforme* was previously unresolved, while in this analysis *Cordagalma sp* falls within the Agalmatidae.
-Additionally, there is strong support for *Bargmannia* as sister to all non-Apolemid codonophorans.
+There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora.
+We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other Codonophora.
+Multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
+There is strong support for the Pyrostephidae as sister to all non-Apolemid codonophorans.
+Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae, which includes the remaining non-Apolemid, non-Pyrostephid historically recognized Physonectae.
+Given the relatively shallow sampling of this analysis, we define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to the Calycophorae.
+
+In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae,*Halistemma* and *Nanomia* [@Dunn:2005dy].
+In the rDNA study, *Physophora hydrostatica* was sister to the Forskaliidae with low support.
+*Cordagalma cordiforme* (= *Cordagalma ordinatum* [@pugh2016synopsis]) was previously unresolved, while in this analysis *Cordagalma* sp is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
+*Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic,@pugh1998re, @Dunn:2005dy].
+Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
+*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*
 
 ![Phylogram of siphonophore relationships. Node labels indicate bootstrap support percent, unnumbered nodes have 100% support. The image was rendered with ggtree [@Yu:2016fo].](figures/Figure3.pdf)
-
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
-Cystonectae is the sister group to the remaining siphonophores, Calycophorae is nested within the paraphyletic "Physonectae", and *Apolemia* is sister to all other Codonophora.
-In addition, multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
-*Physophora gilmeri*, along with *Lychnagalma utricularia* (not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*.
-*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*.
-*Cordagalma cordiforme* was previously unresolved, while in this analysis *Cordagalma sp* falls within the Agalmatidae.
-Additionally, there is strong support for *Bargmannia* as sister to all non-Apolemid codonophorans.
 
@@
 
-##Discussion
+![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
 
-###Siphonophore Phylogeny
+###Evolution of Monoecy
 
 The strong phylogenetic signal in the characters traditinally used for taxonomic diagnostics is a positive indicator of the applicability and unambiguity of these characters.
-
-![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
-
-In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters left with equivocal evolutionary histories, due to unresolved relationships between physonects.
-With our current cladistic resolution, we were able to answer some of the questions left open:
-
-###Evolution of Monoecy
 
@@
 ![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
-The color gradients shows the reconstructed probability estimate of the discrete character states along the branches.](figures/Figure5.pdf)
+The color gradients show the reconstructed probability estimate of the discrete character states along the branches.](figures/Figure5.pdf)
 

### 2017-11-29 manuscript.rmd

@@
 ![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
-The color gradients show the reconstructed probability estimate of the discrete character states along the branches.](figures/Figure5.pdf)
+The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
+Intermediate values reflect uncertainty.](figures/Figure5.pdf)
 

### 2017-11-29 manuscript.rmd
modified discussion

@@
 All sequence data have been deposited in the sequence read archive (SRA).
-The analyses presented here consider 34 siphonophore species and 9 outgroup species.
+The analyses presented here consider 33 siphonophore species and 8 outgroup species.
 This includes new data for 30 species.
@@
 There is strong support for the Pyrostephidae as sister to all non-Apolemid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae, which includes the remaining non-Apolemid, non-Pyrostephid historically recognized Physonectae.
+Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae, which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
 Given the relatively shallow sampling of this analysis, we define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to the Calycophorae.
@@
 *Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic,@pugh1998re, @Dunn:2005dy].
+
 Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
-*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus*
+Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1988siphonophore]).
+*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
+As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
 
@@
 
+###Evolution of Monoecy
+
+Siphonophores colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
+Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within the codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
+The current better-resolved tree model (Figure 5a), indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
+There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
+
+###The Evolution of Zooid Types
+
+One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
+For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores).
+Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
+Here we reconstruct the evolutionary origins of the different zooid types on the present transcriptome tree (fig 4).
+
+Nectophores are retained modified medusae that Codonophora use for coordinated colony-level swimming.
+The nectosome is the region of the colony that develops from the nectosomal growth zone (figure 1B).
+Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
+Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra).
+It is possible that the common ancestor of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
+We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora.
+It is suggested that the nectosome arose as a duplication of the siphosome, followed by functional specialization in propelling the colony [@dunn2006evolution].
+The nectosome has been lost within Codonophora in the genus *Athorybia*.
+
 ![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
 
-###Evolution of Monoecy
+Following the colony development orientation framework [@haddock_dunn_pugh_2005], the nectosome can be located in a dorsal or a ventral position.
+Our ancestral reconstructions for this character (S fig 4) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
-The strong phylogenetic signal in the characters traditinally used for taxonomic diagnostics is a positive indicator of the applicability and unambiguity of these characters.
+Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
+As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
+The MRCA of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
+gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula* (not included in present phylogeny).
+Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy.
+Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
 
-[Citation needed] noticed for the first time that some siphonophores were monoeicous and others were dioiecious.
-Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA (most recent common ancestor) as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy.
-Figure 5a shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato).
-There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the *Marrus-Erenna* clade.
+The ancestral siphonophore had a pneumatophore (fig 1B), since both Cystonects and most Codonophorans possess one (fig 4).
+This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
+Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
+The pneumatophore was lost in the Calycophorae and never gained again in that clade.
+Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.
+
+Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [cite].
+They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
+These taxa might have found other avenues to effectively circulate nutrients across the colony.
 
@@
 
-###The Evolution of Zooid Types
-
-One of the most striking aspects of siphonophore biology is their diversity of unique zooid types.
-Other colonial cnidarians (such as Hydractinia) and some bryozoans (example) have been found to have up to X different zooid types [Citation1 , Citation 2].
-The siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores).
-Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
-Here we reconstruct the evolutionary origins ofthe different zooid types on the present transcriptome tree.
-
-Nectophores are retained modified medusae that Codonophora use for coordinated colony-level swimming.
-The nectosome is the region of the colony that develops from the nectosomal growth zone.
-Unlike the siphosomal growth zone, the nectosome does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-If fact, with the exception of  *Physalia physalis* (which grows small nectophores near the gonodendra), siphonophore nectophores are exclusively found on the nectosome.
-It is possible that the MRCA of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
-We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora.
-The nectosome probably arose as a duplication of the siphosome, followed by functional specialization in propelling the colony.
-The nectosome has been lost within Codonophora in the genus *Athorybia*.
-
-Following the colony development orientation framework presented in (citation), the nectosome can be located in a dorsal or a ventral position.
-Our ancestral reconstructions for this character (Supp figure) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae (sensu stricto) and the branch leading to the *Bargmannia* species.
-
-All Codonophora (with the exception of *Athorybia* species) have a nectosome, but the number and subtypes of nectophores present varies greatly between species.
-As shown in Figure X, most Codonophora presents the ancestral nectosome with multiple nectophores of the same subtype.
-However, Calycophorans evolved a different system with just 2 nectophores of one type.
-This shift may be associated with the loss of the pneumatophore.
-Not all Calycophorans remained with this arrangement.
-The Hippopodidae returned to bearing multiple identical nectophores, many of which are inactive and serve functions of defense (like a shell to retract in) and buoyancy.
-As in the rest of Calycophorans, the Hippopodids only use 2 nectophores to propel the colony.
-Another interesting shift occurs in the branch leading to Diphyomorpha, where the 2 nectophores specialize into 2 subtypes, associated with a shift into a vertically aligned position and pointed bell shapes.
-The 2 types function together in a coupled hydrodynamic system that allows very fast escape responses (Mackie 1964).
-
-Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
-As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
-The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
-gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula*.
-Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy.
-Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
-
-The ancestral siphonophore certainly had a pneumatophore, since both Cystonects and most Codonophorans possess one (Figure 4).
-This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
-Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  (Church, 2013) suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis.
-Despite its multiple biological functions, it was lost in the Calycophorae and never gained again in that clade.
-Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.
-
-Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid.
-They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
-These taxa might have found other avenues to effectively circulate nutrients across the colony.
-
 ###The Gain and Loss of Tentilla
 
-The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla.
+The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla [fig 1A].
 Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
@@
 conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
-When we reconstrauct the evolution of this character on the current phylogeny, we find that 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia*.
-However, this leaves a 30% support for a simple-tentacled MRCA followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.
+When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (S Fig 2).
+However, this leaves a 30% support for a simple-tentacled MRCA followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *(Bargmannia, Diphyes)*.
 
@@
 This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Supp Figure).
-
-Siphonophore tentilla present an astounding diversity of sizes, shapes, colors, and nematocyst complements, and some have been observed to rapidly uncoil in contact with prey.
-Future research should explore the evolutionary history of these unique structures.
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (S Fig 3).
 

### 2017-11-29 manuscript.rmd
reworked intro, did not regenerate pdf

--- 2017-11-29 manuscript.rmd
@@
 
-Siphonophores are a group of 188 valid species within Hydrozoa (Cnidaria), the vast majority of which are members of the plankton (fig.
-1).
-Siphonophores are colonial, and are composed of zooids that are each homologous to solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1988siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from all other colonial hydrozoans in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
-A siphonophore colony arises from a single embryo, which forms a protozooid and a growth zone from which other genetically identical zooids bud asexually [@carre1969etude;@carre1991complete;@carre1967;@carre1995ordre].
-Each zooid is generated asexually, arises in the same repeating species-specific pattern, and is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
-2) [@dunn2006evolution].
-Siphonophores are found at all depths - most species are planktonic, with the exception of one pleustonic species (*Physalia physalis*, Portuguese man of war) that lives at the interface of water and air, and a small clade of benthic siphonophores, the Rhodaliidae [@carre1995ordre;@Totton1960;@pugh1983benthic].
-They are among the most abundant gelatinous predators in the open ocean, and play an important ecological role in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
+Siphonophores (Fig.
+1) are among the most abundant gelatinous predators in the open ocean, and play an important ecological role in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
+Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
+*Physalia physalis*, the Portuguese man of war, floats at the surface and, as a result, is the most familiar siphonophore species.
+Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length.
+There is also a small clade of benthic siphonophores, the Rhodaliidae [@carre1995ordre;@Totton1960;@pugh1983benthic], that are tethered to the bottom for part of their lives.
+In all, there are currently 188 valid described species.
 
-Siphonophores are monophyletic and nested within the Hydroidolina, although relationships among the major lineages of the Hydroidolina remain difficult to resolve [@cartwright2010character;@Cartwright2008kl;@zapata2015phylogenomic].
-The most resolved siphonophore phylogeny to date focused on two genes (16S, 18S) from 52 siphonophore taxa, and resolved many long standing questions about siphonophore biology, including the relationships of the three historically recognised groups (Cystonectae, Physonectae, and Calycophorae) [@Dunn:2005dy].
-The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects (the name Codonophora was given to this clade)[@Dunn:2005dy].
+Siphonophores remain poorly known, in large part because they are fragile and difficult to collect.
+They have, however, been of great interest for more than 150 years (XXX) due to their very unique structure and development.
+Like many other cnidarians, they are colonial -- they grow by incomplete asexual reproduction.
+Each siphonophore colony arises from a single embryo that forms the protozooid, the first body.
+A growth zone then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+These zooids are each homologous to a solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1988siphonophore; @dunn2006evolution].
+Siphonophores differ significantly from all other colonial animals in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
+Each zooid is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
+2) [@dunn2006evolution], and they are arranged in a repeating species-specific pattern.
 
@@
 
-The Apolemiidae are sister to all other Codonophora, however there was little resolution of deep relationships within Codonophora.
-The history of colony structure and budding was investigated by mapping these traits to the tree, and suggesting a complex history of zooid gain and loss.
+Understanding the unique ecology, morphology, and development of siphonophores requires an understanding of their phylogenetic relationships.
+The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
+A siphonophore phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced serveral long standing questions about siphonophore biology.
+These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
+The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects.
+The name Codonophora was given to this clade of physonects and calycophorans [@Dunn:2005dy].
+
+Major questions remained after this early work, though.
+There was, in particular, little support for important deep relationships within Codonophora.
 Resolving deep relationships within Codonophora is key to resolving the evolution of several character traits, including sexual systems (monoecy vesus dioecy) or the gain and loss of particular zooids, such as palpons.
-
-Siphonophores are difficult to collect, in part because they are fragile (nets destroy all but the most robust parts of the colony), but also because many species are found in the deep sea [@youngbluth1984water;@Dunn:2005dy].
-Specimen sampling for the previous phylogeny, and this new work, is enabled by modern collection techniques, including blue-water scuba diving and remotely operated vehicles (ROVs).
 Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 34 siphonophore species and 9 outgroup species.
-Using 1071 genes shared across species, we are able to resolve deep relationships within the siphonophore phylogeny and reconstruct the evolutionary history of characters, including zooid type, life history traits, and habitat.
+Using 1071 genes shared across species, we find strong support for imporant relationships within the siphonophore phylogeny.
+Using this phylogeny, we reconstruct the evolutionary history of characters central to their unique biology, including zooid type, life history traits, and habitat.
 

### 2017-11-29 manuscript.rmd
fixed wording

--- 2017-11-29 manuscript.rmd
@@
 Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1988siphonophore]).
-*Craseoa lathetica* and *Desmophyes haematogaster* are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
+*Craseoa lathetica* and *Desmophyes* sp are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
 As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
@@
 The morphological character matrix, and all the code used to generate the character mapping is available in the git repository and in the supplementary materials.
-A  number of traits were mapped to branches on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (fig 4, 5, supplementary figures 2-5).
+A  number of traits were mapped to branches on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (Fig.
+4, 5, S2-5).
 
@@
 Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within the codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
-The current better-resolved tree model (Figure 5a), indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
+The current better-resolved tree model (Fig.
+5a), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
 There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
@@
 Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
-Here we reconstruct the evolutionary origins of the different zooid types on the present transcriptome tree (fig 4).
+Here we reconstruct the evolutionary origins of the different zooid types on the present transcriptome tree (Fig.
+4).
 
 Nectophores are retained modified medusae that Codonophora use for coordinated colony-level swimming.
-The nectosome is the region of the colony that develops from the nectosomal growth zone (figure 1B).
+The nectosome is the region of the colony that develops from the nectosomal growth zone (Fig.
+1B).
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
@@
 Following the colony development orientation framework [@haddock_dunn_pugh_2005], the nectosome can be located in a dorsal or a ventral position.
-Our ancestral reconstructions for this character (S fig 4) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+Our ancestral reconstructions for this character (Fig.
+S4) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
@@
 
-The ancestral siphonophore had a pneumatophore (fig 1B), since both Cystonects and most Codonophorans possess one (fig 4).
+The ancestral siphonophore had a pneumatophore (Fig.
+1B), since both Cystonects and most Codonophorans possess one (Fig.
+4).
 This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
@@
 Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [cite].
-They were present in the MRCA of siphonophores (Figure 5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
+They were present in the MRCA of siphonophores (Fig.
+5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.
@@
 conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
-When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (S Fig 2).
+When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
+S2).
 However, this leaves a 30% support for a simple-tentacled MRCA followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *(Bargmannia, Diphyes)*.
@@
 This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (S Fig 3).
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
+S3).
 
@@
 
-Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (ref , Figure 4).
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figure 4, S5) [@mapstone2014global].
 While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny suggests the siphonophore MRCA was planktonic, as most extant taxa are.

### 2017-11-29 manuscript.rmd
incorporating Casey's comments from origin/casey

@@
 
-The Apolemiidae are sister to all other Codonophora, however there was little resolution of deep relationships within Codonophora.
-The history of colony structure and budding was investigated by mapping these traits to the tree, and suggesting a complex history of zooid gain and loss.
+Understanding the unique ecology, morphology, and development of siphonophores requires an understanding of their phylogenetic relationships.
+The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
+A siphonophore phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced serveral long standing questions about siphonophore biology.
+These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
+The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects.
+The name Codonophora was given to this clade of physonects and calycophorans [@Dunn:2005dy].
+
+Major questions remained after this early work, though.
+There was, in particular, little support for important deep relationships within Codonophora.
 Resolving deep relationships within Codonophora is key to resolving the evolution of several character traits, including sexual systems (monoecy vesus dioecy) or the gain and loss of particular zooids, such as palpons.
-
-Siphonophores are difficult to collect, in part because they are fragile (nets destroy all but the most robust parts of the colony), but also because many species are found in the deep sea [@youngbluth1984water;@Dunn:2005dy].
-Specimen sampling for the previous phylogeny, and this new work, is enabled by modern collection techniques, including blue-water scuba diving and remotely operated vehicles (ROVs).
-Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 34 siphonophore species and 9 outgroup species.
-Using 1071 genes shared across species, we are able to resolve deep relationships within the siphonophore phylogeny and reconstruct the evolutionary history of characters, including zooid type, life history traits, and habitat.
+Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were excluded).
+Using 1071 genes shared across species, we find strong support for imporant relationships within the siphonophore phylogeny.
+Using this phylogeny, we reconstruct the evolutionary history of characters central to their unique biology, including zooid type, life history traits, and habitat.
 

### 2017-11-30 manuscript.rmd
removed unnecessary columns from main_characters, corrected for column number shifts in manuscripts, removed unnecesary variable declarations

@@
 
-vrange = cdata$Vertical.range
-names(vrange) = rownames(cdata)
-
 #Depth pruned data
@@
 names(depth_median) = rownames(dpruned_data)
-
-depth_mean = dpruned_data$Depth.Mean
-names(depth_mean) = rownames(dpruned_data)
-
-depth_min = dpruned_data$Depth.Min
-names(depth_min) = rownames(dpruned_data)
-
-depth_max = dpruned_data$Depth.Max
-names(depth_max) = rownames(dpruned_data)
-
-depth_sd = dpruned_data$Depth.StdDev
-names(depth_sd) = rownames(dpruned_data)
-
-depth_range = depth_max - depth_min
-
-abundance = dpruned_data$Depth.Count
-names(abundance) = rownames(dpruned_data)
-
 ```
@@
 p <- p + geom_tiplab(size=2.2)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:8,12)]
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:9)]
 rownames(f_cdata) = rownames(cdata)

### 2017-11-30 manuscript.rmd
fleshed Evolution of Vertical habitat

@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figure 4, S5) [@mapstone2014global].
-While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny suggests the siphonophore MRCA was planktonic, as most extant taxa are.
-Some interesting questions arise from these facts, including 1) what was the bathymetric niche of the siphonophore MRCA, and 2) how did siphonophore's vertical habitat use of the water columns evolve along the phylogeny.
-Our results indicate a mesopelagic MRCA, with several convergent transition events to epipelagic and bathypelagic waters.
-There was only a single transition to benthic lifestyle on the stem of *Rhodaliidae*.
+While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny indicates that the siphonophore MRCA was probably planktonic, as most extant siphonophore taxa are.
+Some siphonophore species (such as *Bargmannia lata*) have been observed at abyssopelagic depths near 4000m.
+An interesting question that arises from these facts is what was the evolutionary history of siphonophore vertical habitat use of the water columns.
+Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Supplementary figure S5).
+This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
+This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelgic habitat.
+Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Figure 4).
+There was only a single transition to benthic lifestyle on the stem of *Rhodaliidae*, and a single transition to a pleustonic lifestyle in the stem of *Physalia physalis*.
+Further research should address whether these transitions are following evolutionary changes in feeding habits and prey availability.
 

### 2017-11-30 manuscript.rmd
added VARS ref n the methods
added VARS ref

@@
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-*Physalia physalis*, the Portuguese man of war, floats at the surface and, as a result, is the most familiar siphonophore species.
+*Physalia physalis*, the Portuguese man of war, floats at the surface and, as a result, is the most familiar siphonophore species[@Totton1960].
 Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length.
-There is also a small clade of benthic siphonophores, the Rhodaliidae [@carre1995ordre;@Totton1960;@pugh1983benthic], that are tethered to the bottom for part of their lives.
+There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
 In all, there are currently 188 valid described species.
@@
 Siphonophores remain poorly known, in large part because they are fragile and difficult to collect.
-They have, however, been of great interest for more than 150 years (XXX) due to their very unique structure and development.
+They have, however, been of great interest for more than 150 years due to their very unique structure and development [@mapstone2014global;].
 Like many other cnidarians, they are colonial -- they grow by incomplete asexual reproduction.
@@
 Morphological character data used in trait mapping were obtained from the literature, or from direct observation of available voucher material.
+Depth distribution data was queried from the MBARI VARS ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/, [@schlining2006mbari]) database (with assistance and permission from SHD Haddock), including records from several MBARI cruises.
 We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169, @huelsenbeck2003stochastic].
@@
 In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae,*Halistemma* and *Nanomia* [@Dunn:2005dy].
-In the rDNA study, *Physophora hydrostatica* was sister to the Forskaliidae with low support.
-*Cordagalma cordiforme* (= *Cordagalma ordinatum* [@pugh2016synopsis]) was previously unresolved, while in this analysis *Cordagalma* sp is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
+In the rDNA study, *Physophora hydrostatica* (sister species to *P.
+gilmeri*) was sister to the Forskaliidae with low support.
+*Cordagalma cordiforme* (= *Cordagalma ordinatum*)[@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
 *Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic,@pugh1998re, @Dunn:2005dy].

### 2017-11-30 manuscript.rmd
added more references

@@
 Siphonophores remain poorly known, in large part because they are fragile and difficult to collect.
-They have, however, been of great interest for more than 150 years due to their very unique structure and development [@mapstone2014global;].
+They have, however, been of great interest for more than 150 years due to their very unique structure and development [@mapstone2014global;@mackie1987siphonophore].
 Like many other cnidarians, they are colonial -- they grow by incomplete asexual reproduction.
@@
 A growth zone then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
-These zooids are each homologous to a solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1988siphonophore; @dunn2006evolution].
+These zooids are each homologous to a solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
 Siphonophores differ significantly from all other colonial animals in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
@@
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
-Accession numbers are given in supplementary table X.
-In cases where physical vouchers were unavailable we provide photographs to document species identity (table x).
+Accession numbers are given in supplementary table 1.
+In cases where physical vouchers were unavailable we provide photographs to document species identity (see git repository).
 
@@
 
-When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (supplementary table x).
-mRNA was extracted directly from tissue using a variety of methods (supplementary table x): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
+When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (see supplementary document).
+mRNA was extracted directly from tissue using a variety of methods (see supplementary document): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
 Extractions were performed according to the manufacturer's instruction.
@@
 In the case of large tissue inputs, libraries were sequenced separately for each tissue, subsequently subsampled and pooled *in silico*.
-Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (supplementary table x).
+Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (see supplementary document).
 
@@
 Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
-Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1988siphonophore]).
+Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1987siphonophore]).
 *Craseoa lathetica* and *Desmophyes* sp are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
@@
 
-We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakova sp.*) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakova sp.*) are monophyletic (figure 3 xx).
+We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakova* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakova* sp.) are monophyletic (figure 3 xx).
 In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03).
@@
 Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
-As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not.
-The MRCA of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
+As with the nectosome, we have ambiguity determining whether the common ancestor of siphonophores had bracts or not.
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
 gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula* (not included in present phylogeny).
@@
 4).
-This unique zooid fills itself with gas, which helps the colony float and maintain its orientation in the water column.
+It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
+The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column.
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
 The pneumatophore was lost in the Calycophorae and never gained again in that clade.
-Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.
+Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy [@mackie1974locomotion].
 
-Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [cite].
-They were present in the MRCA of siphonophores (Fig.
+Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+They were present in the common ancestor of siphonophores (Fig.
 5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
@@
 Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
-It is still an open question whether the MRCA of Siphonophora had simple or branched tentacles.
+It is still an open question whether the common ancestor of Siphonophora had simple or branched tentacles.
 The only siphonophores genera regarded as lacking tentilla are *Physalia physalis* and *Apolemia* spp., and *Bathyphysa conifera*.
@@
 conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
-When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an MRCA bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
+When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
 S2).
-However, this leaves a 30% support for a simple-tentacled MRCA followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *(Bargmannia, Diphyes)*.
+However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *(Bargmannia, Diphyes)*.
 
@@
 This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing MRCA followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
 S3).
@@
 
-Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figure 4, S5) [@mapstone2014global].
-While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny indicates that the siphonophore MRCA was probably planktonic, as most extant siphonophore taxa are.
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
+4, S5) [@mapstone2014global].
+While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny indicates that the siphonophore common ancestor was probably planktonic, as most extant siphonophore taxa are.
 Some siphonophore species (such as *Bargmannia lata*) have been observed at abyssopelagic depths near 4000m.
 An interesting question that arises from these facts is what was the evolutionary history of siphonophore vertical habitat use of the water columns.
-Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Supplementary figure S5).
+Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
+S5).
 This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
 This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelgic habitat.
-Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Figure 4).
+Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
+4).
 There was only a single transition to benthic lifestyle on the stem of *Rhodaliidae*, and a single transition to a pleustonic lifestyle in the stem of *Physalia physalis*.

### 2017-12-01 manuscript.rmd
added reference
spelling, ref Mackie year 1988)

@@
 *Physalia physalis*, the Portuguese man of war, floats at the surface and, as a result, is the most familiar siphonophore species[@Totton1960].
-Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length.
+Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length [@mackie1987siphonophore].
 There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
@@
 A growth zone then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
-These zooids are each homologous to a solitary animals, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
+These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
 Siphonophores differ significantly from all other colonial animals in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].

### 2017-12-01 manuscript.rmd
edits, formatting, spelling

@@
 Each zooid is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
-2) [@dunn2006evolution], and they are arranged in a repeating species-specific pattern.
+2) [@dunn2006evolution], and they are arranged in a repeating,  species-specific pattern.
 
@@
 (J) *Praya dubia*, scale bar = 4 cm .
-(K) *Apolemia sp.*, scale bar = 1 cm.](figures/Figure1.jpg)
+(K) *Apolemia* sp., scale bar = 1 cm.](figures/Figure1.jpg)
 
@@
 Resolving deep relationships within Codonophora is key to resolving the evolution of several character traits, including sexual systems (monoecy vesus dioecy) or the gain and loss of particular zooids, such as palpons.
-Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were excluded).
-Using 1071 genes shared across species, we find strong support for imporant relationships within the siphonophore phylogeny.
+Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded).
+Using 1,071 genes shared across species, we find strong support for several relationships within the siphonophore phylogeny.
 Using this phylogeny, we reconstruct the evolutionary history of characters central to their unique biology, including zooid type, life history traits, and habitat.
@@
 We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
-As the sexual system of *Rudjakovia sp* is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia sp* included as a monoecious species, and one without.
+As the sexual system of *Rudjakovia* sp.
+is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
+included as a monoecious species, and one without.
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
@@
 This includes new data for 30 species.
-Summary statistics for expression libaries are given in supplementary table x.
+Summary statistics for expression libraries are given in supplementary table x.
 In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
@@
 There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora.
-We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other Codonophora.
+We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans.
 Multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
-There is strong support for the Pyrostephidae as sister to all non-Apolemid codonophorans.
+There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
 Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae, which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
@@
 
-In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae,*Halistemma* and *Nanomia* [@Dunn:2005dy].
+In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy].
 In the rDNA study, *Physophora hydrostatica* (sister species to *P.
 gilmeri*) was sister to the Forskaliidae with low support.
-*Cordagalma cordiforme* (= *Cordagalma ordinatum*)[@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
-*Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic,@pugh1998re, @Dunn:2005dy].
+*Cordagalma cordiforme* (= *Cordagalma ordinatum*)[@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
+is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
+*Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic;@pugh1998re;@Dunn:2005dy].
 

### 2017-12-01 manuscript.rmd
minor edits

@@
 Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1987siphonophore]).
-*Craseoa lathetica* and *Desmophyes* sp are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
+*Craseoa lathetica* and *Desmophyes* sp.
+are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
 As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
@@
 
-Siphonophores colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
-Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within the codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
+Siphonophore colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
+Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within Codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
 The current better-resolved tree model (Fig.
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores).
-Diphyomorphs have more than 1 type of nectophore, while Cystonects have none.
+For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)(CITATION).
+Diphyomorphs have more than 1 type of propulsive nectophore, while cystonects have none.
 Here we reconstruct the evolutionary origins of the different zooid types on the present transcriptome tree (Fig.
@@
 
-Nectophores are retained modified medusae that Codonophora use for coordinated colony-level swimming.
+Nectophores are retained modified medusae that codonophorans use for coordinated colony-level swimming.
 The nectosome is the region of the colony that develops from the nectosomal growth zone (Fig.
-1B).
+2B).
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra).
-It is possible that the common ancestor of siphonophores had a nectosome, which has lost on the branch leading to Cystonects.
-We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora.
+Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra (CITATION)).
+It is possible that the common ancestor of siphonophores had a nectosome, which has lost on the branch leading to cystonects.
+We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to Codonophora.
 It is suggested that the nectosome arose as a duplication of the siphosome, followed by functional specialization in propelling the colony [@dunn2006evolution].
@@
 
-Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora.
-As with the nectosome, we have ambiguity determining whether the common ancestor of siphonophores had bracts or not.
+Bracts are highly reduced zooids unique to siphonophores, but they are only present in Codonophora.
+As with the nectosome, we have ambiguity when determining whether the common ancestor of siphonophores had bracts or not.
 The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
 gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula* (not included in present phylogeny).
-Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy.
-Some calycophorans are able to actively exclude suphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
+Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy (CITATION).
+Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
 
 The ancestral siphonophore had a pneumatophore (Fig.
-1B), since both Cystonects and most Codonophorans possess one (Fig.
+2B), since both cystonects and most codonophorans possess one (Fig.
 4).
 It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
-The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column.
+The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column (CITATION).
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
-The pneumatophore was lost in the Calycophorae and never gained again in that clade.
+The pneumatophore was lost in Calycophorae and never gained again in that clade.
 Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy [@mackie1974locomotion].
@@
 They were present in the common ancestor of siphonophores (Fig.
-5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in Calycophorans, and in *Marrus claudanielis*.
+5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.
@@
 A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture.
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture (CITATION).
 Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
 Hessinger & Ford 1988 (in the Biology of Nematocysts) described *Physalia*'s buttons as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm.
-This suggests they are not just ectodermal swellings, but probably are reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
+This suggests they are not just ectodermal swellings, but probably  reduced tentilla.
+When we code *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
 S3).
@@
 This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
-This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelgic habitat.
+This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
 Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
 4).
-There was only a single transition to benthic lifestyle on the stem of *Rhodaliidae*, and a single transition to a pleustonic lifestyle in the stem of *Physalia physalis*.
+There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
 Further research should address whether these transitions are following evolutionary changes in feeding habits and prey availability.

### 2017-12-01 manuscript.rmd
added mention to vertical migration and reference to Pugh1984 in bib
fixed citations, added conclusion

@@
 			)
+#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
 
-	pdf("figures_for_AI/figure_3_raw.pdf")
-	flip( p_phylogram, euphysonectae_node, calycophora_node )
-	dev.off()
+#	pdf("figures_for_AI/figure_3_raw.pdf")
+#	flip( p_phylogram, euphysonectae_node, calycophora_node )
+#	dev.off()
 
@@
 par(ask=F)
+
+#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
+
 #Simmap SEX
@@
 #Simmap PALPONS
+
+#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
+
 # plotTree(ultratree, lwd = 4)
@@
 
+#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
+
 #ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
@@
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
+
+#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
+
 # pdf("figures_for_AI/overview_ggheatmap_chars.pdf")
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)(CITATION).
+For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Diphyomorphs have more than 1 type of propulsive nectophore, while cystonects have none.
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra (CITATION)).
+Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960].
 It is possible that the common ancestor of siphonophores had a nectosome, which has lost on the branch leading to cystonects.
@@
 gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula* (not included in present phylogeny).
-Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy (CITATION).
-Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony (Bidigare & Biggs, 1980).
+Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen].
+Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
 
@@
 It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
-The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column (CITATION).
+The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column [@church2015histology;@mackie1974locomotion].
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
@@
 
-A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture (CITATION).
+How we define absence of tentilla, especially for *Physalia physalis*, is also important.
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@BARDI2007].
 Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
-Hessinger & Ford 1988 (in the Biology of Nematocysts) described *Physalia*'s buttons as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm.
+*Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@Hessinger1988].
 This suggests they are not just ectodermal swellings, but probably  reduced tentilla.
-When we code *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
+When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
 S3).
@@
 4, S5) [@mapstone2014global].
-While there are some pleustonic (*Physalia*) and benthic (*Rhodaliidae*) siphonophores, the phylogeny indicates that the siphonophore common ancestor was probably planktonic, as most extant siphonophore taxa are.
+However, the depth distribution of siphonophore populations is not always static.
+Many species are strong vertical migrators, and some species such as *Chuniphyes multidentata* have exhibited synchronous diel migration patterns [@pugh1984diel].
+The phylogeny indicates that the siphonophore common ancestor was probably planktonic, as most extant siphonophore taxa are.
 Some siphonophore species (such as *Bargmannia lata*) have been observed at abyssopelagic depths near 4000m.
-An interesting question that arises from these facts is what was the evolutionary history of siphonophore vertical habitat use of the water columns.
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
@@
 This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
+The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives.
 Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
@@
 There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
-Further research should address whether these transitions are following evolutionary changes in feeding habits and prey availability.
 
 ## Conclusions
+Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support.
+A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-Apolemid codonophorans.
+Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
+Future studies with the addition of key taxa within the Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
 
@@
 ```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
-	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
+
+####The supplementary table is too long. We decided to include as a supplementary datafile instead
+
+#	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
+#	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
 fails","Total transcripts","Coding transcripts")

### 2017-12-01 manuscript.rmd
Made Supplementary headers for phylogenetic signal and vertical evolution

@@
 
+#####Vertical Habitat Evolution
 ```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of median depth of species"}
@@
 ```
+
+#####Phylogenetic Signal
 

### 2017-12-01 manuscript.rmd
removed vrange_simmap making

@@
 #print("SIMMAP Habitat")
-make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim
-
 ```

### 2017-12-01 manuscript.rmd
made extended depth evolution contmap with BW diving spp at 20m, physalia at 0m

@@
 #####Vertical Habitat Evolution
-```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of median depth of species"}
+```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
 depth_value = depth_median
 depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}
+median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
+names(median_aggregate) = c("Spp","Depth")
+median_aggregate[33,2] = 0
+median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
+median_depth = median_aggregate$Depth
+names(median_depth) = median_aggregate$Spp
+depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
 plot(depthmap)

### 2017-12-01 manuscript.rmd
fixed species names in manuscript code calls

@@
 read.csv('character_coding/main_characters.csv', header = T, sep = ',') -> cdata
-cdata$Species = str_replace_all(cdata$Species, "sp ", "sp")
 rownames(cdata) = cdata$Species
@@
 raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
-raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
-raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Prayidae D27SS7"
-raw_data$Species[which(raw_data$Species == "Craseoa lathetica")] = "Prayidae D27D2"
-raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
-raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Thermopalia sp"
-raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Physonect sp"
+raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp."
+raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Desmophyes sp."
+raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp."
+raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Stephalia sp." ##We should do something about this not corresponding...
+raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Undescribed Sp. L" #Is this correct?
 
@@
 
-sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp", "Resomia ornicephala", "Frillagalma vityazi", "Physonect sp", "Marrus claudanielis", "Stephalia dilata", "Thermopalia sp", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Prayidae D27D2", "Prayidae D27SS7", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp", "Rhizophysa filiformis", "Physalia physalis")
+sp_order = c("Agalma elegans", "Athorybia rosacea", "Halistemma rubrum", "Nanomia bijuga", "Lychnagalma utricularia", "Physophora gilmeri", "Forskalia asymmetrica", "Cordagalma sp.", "Resomia ornicephala", "Frillagalma vityazi", "Undescribed Sp.
+L", "Marrus claudanielis", "Stephalia dilata", "Stephalia sp.", "Erenna richardi", "Abylopsis tetragona", "Diphyes dispar", "Chelophyes appendiculata", "Kephyes ovata", "Chuniphyes multidentata", "Lilyopsis fluoracantha", "Craseoa lathetica", "Desmophyes sp.", "Hippopodius hippopus", "Bargmannia amoena", "Bargmannia elongata", "Bargmannia lata", "Rudjakovia sp", "Apolemia rubriversa", "Apolemia lanosa", "Apolemia sp.", "Rhizophysa filiformis", "Physalia physalis")
 

### 2017-12-04 manuscript.rmd
updated tip names

@@
 
+  gsub2 <- function(pattern, replacement, x, ...) {
+  for(i in 1:length(pattern))
+  x <- gsub(pattern[i], replacement[i], x, ...)
+  x
+  }
+
+  correctnames =read.csv("tipcorrection.csv") #updating species IDs relative to agalma species IDs
+
   fix_names = function( tree ){
@@
     tree$tip.label = sub( 'sp_', 'sp', tree$tip.label ) # Fix trailing _ in Physonect sp_
+    tree$tip.label = gsub2(correctnames$Agalma.ID, correctnames$Tree.ID,tree$tip.label)
     return( tree )
@@
 raw_data$Species[which(raw_data$Species == "Craseoa")] = "Craseoa lathetica"
-raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp."
-raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Desmophyes sp."
-raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp."
-raw_data$Species[which(raw_data$Species == "Thermopalia")] = "Stephalia sp." ##We should do something about this not corresponding...
-raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Undescribed Sp. L" #Is this correct?
+raw_data$Species[which(raw_data$Species == "Cordagalma")] = "Cordagalma sp"
+raw_data$Species[which(raw_data$Species == "Desmophyes haematogaster")] = "Desmophyes sp"
+raw_data$Species[which(raw_data$Species == "Rudjakovia")] = "Rudjakovia sp"
+raw_data$Species[which(raw_data$Species == "Physonectae sp. 1")] = "Undescribed sp L"
 
@@
 ```
+
 ```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}

### 2017-12-04 manuscript.rmd
moved pneumatophore to own section
added nectosome position reconstruction code with subset of spp and 2 states Dorsal/Ventral

@@
 
-necto_pos = cdata$Nectosome.position
-names(necto_pos) = rownames(cdata)
-
-necto_types = cdata$Nectophores
-names(necto_types) = rownames(cdata)
-
 palpons = cdata$Palpons
@@
 #print("SIMMAP Nectosome Position")
-make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
-#print("SIMMAP Nectophore Types")
-make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
-#print("SIMMAP Bracts")
 make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
 #print("SIMMAP Habitat")
+
+#Nectophore position SIMMAP setup
+necto_pos_data = cdata[which(cdata$Nectosome.position != "None"),]
+necto_pos = necto_pos_data$Nectosome.position
+names(necto_pos) = rownames(necto_pos_data)
+NectoPosTree = drop.tree(ultratree, which(!(ultratree$tip.label %in% names(necto_pos))))
+make.simmap(NectPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+#print("SIMMAP Nectophore Types")
 ```
@@
 
+Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+They were present in the common ancestor of siphonophores (Fig.
+5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+These taxa might have found other avenues to effectively circulate nutrients across the colony.
+
+![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
+The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
+Intermediate values reflect uncertainty.](figures/Figure5.pdf)
+
+###The Gain and Loss of the Pneumatophore
+
 The ancestral siphonophore had a pneumatophore (Fig.
@@
 Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy [@mackie1974locomotion].
-
-Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
-They were present in the common ancestor of siphonophores (Fig.
-5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
-These taxa might have found other avenues to effectively circulate nutrients across the colony.
-
-![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
-The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
-Intermediate values reflect uncertainty.](figures/Figure5.pdf)
 
@@
 
+```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome"}
+#SIMMAP NectoPos
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Dorsal", "Ventral")
+# necto_pos_sim %>% plotSimmap(lwd = 4, add = T)
+# nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(necto_pos_sim)
+```
+
 #####Vertical Habitat Evolution

### 2017-12-04 manuscript.rmd
fixed typos, reorganized pneumatophore text order, added Bathyphysa loss to the last sentence of tentilla evolution

@@
 
-We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakova* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakova* sp.) are monophyletic (figure 3 xx).
+We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic (figure 3 xx).
 In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03).
@@
 
+The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column [@church2015histology;@mackie1974locomotion].
+Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
 The ancestral siphonophore had a pneumatophore (Fig.
@@
 It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
-The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column [@church2015histology;@mackie1974locomotion].
-Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
 The pneumatophore was lost in Calycophorae and never gained again in that clade.
@@
 This suggests they are not just ectodermal swellings, but probably  reduced tentilla.
-When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by a single loss of tentilla in the branch leading to *Apolemiidae* (Fig.
-S3).
+When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to *Apolemiidae* (Fig.
+S3), and in *Bathyphysa conifera*.
 

### 2017-12-04 manuscript.rmd
incorporated many changes from Felipe, Sam & Phil

--- 2017-12-04 manuscript.rmd
@@
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-*Physalia physalis*, the Portuguese man of war, floats at the surface and, as a result, is the most familiar siphonophore species[@Totton1960].
+The Portuguese man of war *Physalia physalis* floats at the surface and, as a result, is the most familiar siphonophore species [@Totton1960].
 Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length [@mackie1987siphonophore].
@@
 There was, in particular, little support for important deep relationships within Codonophora.
-Resolving deep relationships within Codonophora is key to resolving the evolution of several character traits, including sexual systems (monoecy vesus dioecy) or the gain and loss of particular zooids, such as palpons.
+Resolving deep relationships within Codonophora is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (fig 2).
 Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded).
@@
 New data were analysed in conjunction with 13 publically available datasets, with a total number of 43 species.
-Sequence assembly, annotation, Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma [@Dunn:2013kw], v.
-1.00, and Bayesian Inference (BI) analyses were conducted using Phylobayes[@lartillot2009phylobayes] v.
-1.7a-mpi.
+Sequence assembly, annotation, Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma v.
+1.00 [@Dunn:2013kw; @Guang202416], and Bayesian Inference (BI) analyses were conducted using Phylobayes v.
+1.7a-mpi [@lartillot2009phylobayes].
 Source code for all analysis steps, sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
@@
 Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The CAT-GTR and CAT-poisson models did not converge after a long CPU time, and only the results from the CAT-poisson model are included here.
+The CAT-GTR and CAT-Poisson models did not converge after a long CPU time, and only the results from the CAT-Poisson model are included here.
 
@@
 Depth distribution data was queried from the MBARI VARS ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/, [@schlining2006mbari]) database (with assistance and permission from SHD Haddock), including records from several MBARI cruises.
-We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169, @huelsenbeck2003stochastic].
+We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
 Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package.
@@
 We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans.
-Multiple nodes that were not resolved in the previous two-gene analysis do recieve strong support in this 1,071-gene transcriptome analysis.
+Multiple nodes that were not resolved in the previous two-gene analysis recieve strong support in this 1,071-gene transcriptome analysis.
 There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
@@
 
-In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy].
+In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *Physophora hydrostatica* (sister species to *P.
 gilmeri*) was sister to the Forskaliidae with low support.
-*Cordagalma cordiforme* (= *Cordagalma ordinatum*)[@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
+*Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
 is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
-*Cordagalma* has traditionally been placed within the Agalmatidae *sensu lato* [@totton1965synopsis], although the Agalmatidae is considered to be a catch-all family in need of revision [@pugh2006taxonomic;@pugh1998re;@Dunn:2005dy].
+Placement of *Cordagalma* outside Agalmatidae *sensu stricto* is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].
 
 Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
-Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after [@mackie1987siphonophore]).
+Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after @mackie1987siphonophore).
+As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
 *Craseoa lathetica* and *Desmophyes* sp.
 are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
-As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
 
@@
 The morphological character matrix, and all the code used to generate the character mapping is available in the git repository and in the supplementary materials.
-A  number of traits were mapped to branches on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (Fig.
+A  number of traits were mapped to tips on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (Fig.
 4, 5, S2-5).
@@
 Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within Codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
-The current better-resolved tree model (Fig.
+The current better-resolved tree (Fig.
 5a), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+For example, the siphonophore genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Diphyomorphs have more than 1 type of propulsive nectophore, while cystonects have none.
@@
 Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960].
-It is possible that the common ancestor of siphonophores had a nectosome, which has lost on the branch leading to cystonects.
+It is possible that the common ancestor of siphonophores had a nectosome, which was lost on the branch leading to cystonects.
 We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to Codonophora.
@@
 Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support.
-A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-Apolemid codonophorans.
+A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemid codonophorans.
 Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
@@
 Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island.
+SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of Harvard cluster.
 We also thank the MBARI crews and ROV pilots for collection of the specimens.

### 2017-12-04 manuscript.rmd

@@
 names(nectosome) = rownames(cdata)
-
-necto_pos = cdata$Nectosome.position
-names(necto_pos) = rownames(cdata)
-
-necto_types = cdata$Nectophores
-names(necto_types) = rownames(cdata)
 

### 2017-12-04 manuscript.rmd
fixed typo in code

@@
 names(necto_pos) = rownames(necto_pos_data)
-NectoPosTree = drop.tree(ultratree, which(!(ultratree$tip.label %in% names(necto_pos))))
-make.simmap(NectPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+NectoPosTree = drop.tip(ultratree, which(!(ultratree$tip.label %in% names(necto_pos))))
+make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
 #print("SIMMAP Nectophore Types")

### 2017-12-04 manuscript.rmd
fixed bib file

@@
 
-```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}
-median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
-names(median_aggregate) = c("Spp","Depth")
-median_aggregate[33,2] = 0
-median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
-median_depth = median_aggregate$Depth
-names(median_depth) = median_aggregate$Spp
-depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
-plot(depthmap)
-```
+# ```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}
+# median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
+# names(median_aggregate) = c("Spp","Depth")
+# median_aggregate[33,2] = 0
+# median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
+# median_depth = median_aggregate$Depth
+# names(median_depth) = median_aggregate$Spp
+# depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+# plot(depthmap)
+# ```
 

### 2017-12-05 manuscript.rmd
updated file name to figures raw

--- 2017-12-04 manuscript.rmd
@@
 			)
-#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
 
-#	pdf("figures_for_AI/figure_3_raw.pdf")
-#	flip( p_phylogram, euphysonectae_node, calycophora_node )
-#	dev.off()
+pdf("figures_raw/figure_3_raw.pdf")
+flip( p_phylogram, euphysonectae_node, calycophora_node )
+dev.off()
 
@@
 #Simmap SEX
-#plotTree(ultratree, lwd = 4)
-#colors = c("black", "red")
-#names(colors) = c("Dioecious", "Monoecious")
-#sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
-#nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-#add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-# pdf("figures_for_AI/density_sex_sim.pdf")
-# densityMap(sex_sim)
-# dev.off()
+plotTree(ultratree, lwd = 4)
+colors = c("black", "red")
+names(colors) = c("Dioecious", "Monoecious")
+sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
+nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+ pdf("figures_raw/figure_5_raw_sex.pdf")
+ densityMap(sex_sim)
+ dev.off()
 ```
@@
 
-#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
-
-# plotTree(ultratree, lwd = 4)
-# colors = c("black", "red")
-# names(colors) = c("Absent", "Present")
-# palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-# pdf("figures_for_AI/palpons_sim.pdf")
-# densityMap(palpons_sim)
-# dev.off()
+ plotTree(ultratree, lwd = 4)
+ colors = c("black", "red")
+ names(colors) = c("Absent", "Present")
+ palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+ nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+ add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+ pdf("figures_raw/figure_5_raw_palpons.pdf")
+ densityMap(palpons_sim)
+ dev.off()
 ```
@@
 
-#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
-
 #ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_boxplot() + theme_classic()
 
-# pdf("figures_for_AI/trimmed_violinVars.pdf")
-# ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
-# dev.off()
+ pdf("figures_raw/figure_4_raw_violin.pdf")
+ ggplot(trimmed_SDdata, aes(x = Species, y=-Depth)) + geom_violin() + theme_classic()
+ dev.off()
 ```
@@
 
-#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
-
-# pdf("figures_for_AI/overview_ggheatmap_chars.pdf")
-# gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
-# dev.off()
+ pdf("figures_raw/figure_4_raw_heat.pdf")
+ gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 1)
+ dev.off()
 ```
@@
  ```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}
-median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
-names(median_aggregate) = c("Spp","Depth")
-median_aggregate[33,2] = 0
-median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
-median_depth = median_aggregate$Depth
-names(median_depth) = median_aggregate$Spp
-depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
-plot(depthmap)
+# median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
+# names(median_aggregate) = c("Spp","Depth")
+# median_aggregate[33,2] = 0
+# median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
+# median_depth = median_aggregate$Depth
+# names(median_depth) = median_aggregate$Spp
+# depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+# plot(depthmap)
 ```

### 2017-12-05 manuscript.rmd
added mantle canal characters

@@
 
+mantle = cdata$Mantle.Canal
+names(mantle)=rownames(cdata)
+mantle[is.na(mantle)]=0
+
 #Depth pruned data
@@
 #print("SIMMAP Nectosome Position")
-make.simmap(ultratree, bracts, nsim = 100) -> bracts_sim
+make.simmap(ultratree, bracts, nsim = 100, message=FALSE) -> bracts_sim
 #print("SIMMAP Habitat")
+make.simmap(ultratree, mantle, nsim = 100, message=FALSE) -> mantle_sim
 
@@
 
+```{r mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+#Simmap NECTOSOME
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(mantle_sim)
+```
+
 ```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome"}

### 2017-12-05 manuscript.rmd

@@
 
-mantle = cdata$Mantle.Canal
-names(mantle)=rownames(cdata)
-mantle[is.na(mantle)]=0
-
 #Depth pruned data
@@
 #print("SIMMAP Habitat")
-make.simmap(ultratree, mantle, nsim = 100, message=FALSE) -> mantle_sim
 
@@
 make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
-#print("SIMMAP Nectophore Types")
+
+#Mantle canal SIMMAP Setup
+mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
+mantle = mantle_data$Mantle.Canal
+names(mantle)=rownames(mantle_data)
+MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
+make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
 ```

### 2017-12-05 manuscript.rmd
quick edits to first two paragraphs

@@
 Siphonophores (Fig.
-1) are among the most abundant gelatinous predators in the open ocean, and play an important ecological role in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
+1) are among the most abundant gelatinous predators in the open ocean, and play important ecological roles in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-The Portuguese man of war *Physalia physalis* floats at the surface and, as a result, is the most familiar siphonophore species [@Totton1960].
-Most siphonophores species are planktonic, living in the midwater where some grow to be longer than 30 meters in length [@mackie1987siphonophore].
+The most familiar siphonophore species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
+Most siphonophore species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
 There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
@@
 They have, however, been of great interest for more than 150 years due to their very unique structure and development [@mapstone2014global;@mackie1987siphonophore].
-Like many other cnidarians, they are colonial -- they grow by incomplete asexual reproduction.
+Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction.
 Each siphonophore colony arises from a single embryo that forms the protozooid, the first body.
@@
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from all other colonial animals in terms of colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
-Each zooid is functionally specialized for a particular task (e.g feeding, reproducing, swimming) (fig.
+Siphonophores differ significantly from all other colonial animals in colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
+Each zooid is functionally specialized for a particular task (*e.g.*, feeding, reproducing, swimming) (fig.
 2) [@dunn2006evolution], and they are arranged in a repeating,  species-specific pattern.

### 2017-12-06 manuscript.rmd
spelling, knit

@@
 Siphonophores differ significantly from all other colonial animals in colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
-Each zooid is functionally specialized for a particular task (*e.g.*, feeding, reproducing, swimming) (fig.
+Each zooid is functionally specialized for a particular task (*e.g.*, feeding, reproducing, swimming) (Fig.
 2) [@dunn2006evolution], and they are arranged in a repeating,  species-specific pattern.
@@
 The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
-A siphonophore phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced serveral long standing questions about siphonophore biology.
+A siphonophore phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced several long standing questions about siphonophore biology.
 These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
@@
 There was, in particular, little support for important deep relationships within Codonophora.
-Resolving deep relationships within Codonophora is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (fig 2).
+Resolving deep relationships within Codonophora is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (Fig.
+2).
 Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded).
@@
 #Mantle canal SIMMAP Setup
-mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
-mantle = mantle_data$Mantle.Canal
-names(mantle)=rownames(mantle_data)
-MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
-make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
+#mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
+#mantle = mantle_data$Mantle.Canal
+#names(mantle)=rownames(mantle_data)
+#MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
+#make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(mantle_sim)
+#densityMap(mantle_sim)
 ```

### 2017-12-06 manuscript.rmd
comments Felipe

@@
 ```
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (figure 3) [@Dunn:2005dy].
+These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
+3) [@Dunn:2005dy].
 There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora.
 We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans.
-Multiple nodes that were not resolved in the previous two-gene analysis recieve strong support in this 1,071-gene transcriptome analysis.
+Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis.
 There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-Siphonophore nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960].
+Nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960].
 It is possible that the common ancestor of siphonophores had a nectosome, which was lost on the branch leading to cystonects.
@@
 
-![Stochastic mapping reconstruction of the evolutionary history of A) palpon zooids, and B) distribution of sexual zooids on the colonies.
+![Stochastic mapping reconstruction of the evolutionary history of A) distribution of sexual zooids on the colonies, and B) palpon zooids.
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
@@
 It is still an open question whether the common ancestor of Siphonophora had simple or branched tentacles.
-The only siphonophores genera regarded as lacking tentilla are *Physalia physalis* and *Apolemia* spp., and *Bathyphysa conifera*.
+The only siphonophores species regarded as lacking tentilla are *Physalia physalis*, *Apolemia* spp., and *Bathyphysa conifera*.
 Since *B.

### 2017-12-06 manuscript.rmd
minor edits

@@
 version 0.39 (commit fd68ef5733c095c7000a4f92dc8c0daaddeec3b9) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
-For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (figure 3 x).
+For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
+3 x).
 We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
@@
 
-In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *Physophora hydrostatica* (sister species to *P.
@@
 
-We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic (figure 3 xx).
+We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic (Fig.
+3 xx).
 In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03).
@@
 The current better-resolved tree (Fig.
-5a), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
+5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
 There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-Nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960].
+Nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960]).
 It is possible that the common ancestor of siphonophores had a nectosome, which was lost on the branch leading to cystonects.
@@
 Our ancestral reconstructions for this character (Fig.
-S4) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the braches leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+S5) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
@@
 They were present in the common ancestor of siphonophores (Fig.
-5b), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.
@@
 
-The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla [fig 1A].
+The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla (Fig.
+2A).
 Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
@@
 The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@BARDI2007].
-Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
+Siphonophore tentilla are complete diverticular branchings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
 *Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@Hessinger1988].
@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
-4, S5) [@mapstone2014global].
+4, S6) [@mapstone2014global].
 However, the depth distribution of siphonophore populations is not always static.
@@
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
-S5).
+S6).
 This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
@@
 
-```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with Physalia included"}
+```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included"}
 #Simmap TENTILLA

### 2017-12-06 manuscript.rmd
minor edits

@@
 S2).
-However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *(Bargmannia, Diphyes)*.
+However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.
 

### 2017-12-06 manuscript.rmd
small ms changes, new knit

@@
 The current better-resolved tree (Fig.
-5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Agalmatids *sensu lato*.
+5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to [**placeholder name** Agalmatids *sensu lato*].
 There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the siphonophore genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+For example, the genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Diphyomorphs have more than 1 type of propulsive nectophore, while cystonects have none.
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-Nectophores are exclusively found on the nectosome -with the exception of  *Physalia physalis* (which has no nectosome, and grows small nectophores near the gonodendra [@Totton1960]).
 It is possible that the common ancestor of siphonophores had a nectosome, which was lost on the branch leading to cystonects.
@@
 Bracts are highly reduced zooids unique to siphonophores, but they are only present in Codonophora.
-As with the nectosome, we have ambiguity when determining whether the common ancestor of siphonophores had bracts or not.
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica* (however, they are present in its sister species, *P.
-gilmeri* included in the present phylogeny), and in *Gymnopraia lapislazula* (not included in present phylogeny).
 Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen].
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
+As with the nectosome, we cannot be sure whether the common ancestor of siphonophores had bracts or not.
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
+Phylogenetic reconstructions are sensitive to taxon sampling, and these patterns of loss are not captured by the present sampling [**We don't have the reconstruction in supp info - do we want to add it?**].
 
-Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+We do not disinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
 They were present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of Stephanophyes [@totton1965synopsis]), and in *Marrus claudanielis*.
 These taxa might have found other avenues to effectively circulate nutrients across the colony.
@@
 
-The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony float and maintain its orientation in the water column [@church2015histology;@mackie1974locomotion].
+The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
-4, S6) [@mapstone2014global].
+4, S6) [@mapstone2014global; @mackie1987siphonophore].
 However, the depth distribution of siphonophore populations is not always static.
@@
 Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island.
-SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of Harvard cluster.
+SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster.
 We also thank the MBARI crews and ROV pilots for collection of the specimens.

### 2017-12-07 manuscript.rmd
fixed typoes

@@
 
-Bracts are highly reduced zooids unique to siphonophores, but they are only present in Codonophora.
+Bracts are highly reduced zooids unique to siphonophores, but they are only present in Codonophora (Fig.
+4).
 Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen].
@@
 Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
-We do not disinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
+We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
 They were present in the common ancestor of siphonophores (Fig.

### 2017-12-07 manuscript.rmd
fixed extended median depth contMap

@@
 
- ```{r median_cont_BWincluded, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations."}
-# median_aggregate = aggregate(trimmed_SDdata$Depth,by=list(trimmed_SDdata$Species), FUN=median)
-# names(median_aggregate) = c("Spp","Depth")
-# median_aggregate[33,2] = 0
-# median_aggregate = median_aggregate[match(median_aggregate$Spp, ultratree$tip.label),]
-# median_depth = median_aggregate$Depth
-# names(median_depth) = median_aggregate$Spp
-# depthmap <- contMap(ultratree, median_depth, plot=F) %>% setMap(colors=c("cyan","blue","black"))
-# plot(depthmap)
+```{r median_depth_extended, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations." }
+
+median_extended = cdata$Median.Depth.Extended
+names(median_extended) = rownames(cdata)
+median_extended = median_extended[match(names(median_extended), ultratree$tip.label),]
+depthmap_ext <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap_ext)
+
 ```

### 2017-12-07 manuscript.rmd
fixed typos in code median_depth_contmap

@@
 #Simmap PALPONS
-
  plotTree(ultratree, lwd = 4)
@@
 
-median_extended = cdata$Median.Depth.Extended
+median_extended = cdata$Depth.Median.Extended
 names(median_extended) = rownames(cdata)
-median_extended = median_extended[match(names(median_extended), ultratree$tip.label),]
+median_extended = median_extended[match(names(median_extended), ultratree$tip.label)]
 depthmap_ext <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))

### 2017-12-07 manuscript.rmd
descending/ascending canal simmaps independent in the code

@@
 
-#Mantle canal SIMMAP Setup
-#mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
-#mantle = mantle_data$Mantle.Canal
-#names(mantle)=rownames(mantle_data)
-#MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
-#make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
+#Descending mantle Canal SIMMAP Setup
+desc_mantle_data = cdata[which(!is.na(cdata$Desc.Mantle.Canal)),]
+desc_mantle = mantle_data$Desc.Mantle.Canal
+names(desc_mantle)=rownames(desc_mantle_data)
+DescMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(desc_mantle))))
+make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
+
+#Ascending Mantle canal SIMMAP Setup
+asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
+asc_mantle = mantle_data$Asc.Mantle.Canal
+names(asc_mantle)=rownames(asc_mantle_data)
+AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
+make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
 ```
@@
 
-```{r mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
-#Simmap NECTOSOME
+```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+#Simmap Descending Mantle Canal
 # plotTree(ultratree, lwd = 4)
@@
 # names(colors) = c("Absent", "Present")
-# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-#densityMap(mantle_sim)
+densityMap(desc_mantle_sim)
+```
+
+```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+#Simmap Ascending mantle canal
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# asc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(ascribe.simmap(asc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+densityMap(asc_mantle_sim)
 ```

### 2017-12-07 manuscript.rmd
changed figure legends for canal simmaps

@@
 
-```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores"}
 #Simmap Descending Mantle Canal
@@
 
-```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of an ascending mantle canal in the nectophores"}
 #Simmap Ascending mantle canal

### 2017-12-07 manuscript.rmd
code for figure 4 raw,ggheatmap included nectophore characters in code

@@
 p <- p + geom_tiplab(size=2.2)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:9)]
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:9,17,18,3)]
 rownames(f_cdata) = rownames(cdata)
-f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
+f_cdata = f_cdata[,c(7,4,1,3,6,2,10,8,9,5)]
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]

### 2017-12-07 manuscript.rmd
fixed mantle canals code make.simmap

@@
 desc_mantle_data = cdata[which(!is.na(cdata$Desc.Mantle.Canal)),]
-desc_mantle = mantle_data$Desc.Mantle.Canal
+desc_mantle = des_mantle_data$Desc.Mantle.Canal
 names(desc_mantle)=rownames(desc_mantle_data)
@@
 asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
-asc_mantle = mantle_data$Asc.Mantle.Canal
+asc_mantle = asc_mantle_data$Asc.Mantle.Canal
 names(asc_mantle)=rownames(asc_mantle_data)

### 2017-12-07 manuscript.rmd
fixed typo in mantle  make.simmap

@@
 desc_mantle_data = cdata[which(!is.na(cdata$Desc.Mantle.Canal)),]
-desc_mantle = des_mantle_data$Desc.Mantle.Canal
+desc_mantle = desc_mantle_data$Desc.Mantle.Canal
 names(desc_mantle)=rownames(desc_mantle_data)

### 2017-12-07 manuscript.rmd
changed color palettes of densitymaps, added details to Fig4

@@
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+sex_obj <- densityMap(sex_sim)
+sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(sex_obj$cols))
  pdf("figures_raw/figure_5_raw_sex.pdf")
- densityMap(sex_sim)
+ plot(sex_obj)
  dev.off()
@@
  add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+palpons_obj <- densityMap(palpons_sim)
+palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(palpons_obj$cols))
  pdf("figures_raw/figure_5_raw_palpons.pdf")
- densityMap(palpons_sim)
+ plot(palpons_obj)
  dev.off()
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim)
+tentilla_obj <- densityMap(tentilla_sim)
+tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(tentilla_obj$cols))
+plot(tentilla_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim2)
+tentilla2_obj <- densityMap(tentilla_sim2)
+tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(tentilla2_obj$cols))
+plot(tentilla2_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(nectosome_sim)
+nectosome_obj <- densityMap(nectosome_sim)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(desc_mantle_sim)
+desc_mantle_obj <- densityMap(desc_mantle_sim)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(asc_mantle_sim)
+asc_mantle_obj <- densityMap(asc_mantle_sim)
+asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(asc_mantle_obj$cols))
+plot(asc_mantle_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(necto_pos_sim)
+necto_pos_obj <- densityMap(necto_pos_sim)
+necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(necto_pos_obj$cols))
+plot(necto_pos_obj)
 ```

### 2017-12-07 manuscript.rmd
changed colors in Figure5 pdf

@@
 par(ask=F)
-
-#These are commented out to avoid outputting the figure on every knit (remove comment to generate)
-
 #Simmap SEX

### 2017-12-07 manuscript.rmd
changed pure blue to matte blue in densityMaps

--- 2017-12-07 manuscript.rmd
@@
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+sex_obj <- densityMap(sex_sim)
+sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj$cols))
  pdf("figures_raw/figure_5_raw_sex.pdf")
- densityMap(sex_sim)
+ plot(sex_obj)
  dev.off()
@@
  add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+palpons_obj <- densityMap(palpons_sim)
+palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj$cols))
  pdf("figures_raw/figure_5_raw_palpons.pdf")
- densityMap(palpons_sim)
+ plot(palpons_obj)
  dev.off()
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim)
+tentilla_obj <- densityMap(tentilla_sim)
+tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
+plot(tentilla_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim2)
+tentilla2_obj <- densityMap(tentilla_sim2)
+tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
+plot(tentilla2_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(nectosome_sim)
+nectosome_obj <- densityMap(nectosome_sim)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(desc_mantle_sim)
+desc_mantle_obj <- densityMap(desc_mantle_sim)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(asc_mantle_sim)
+asc_mantle_obj <- densityMap(asc_mantle_sim)
+asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
+plot(asc_mantle_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(necto_pos_sim)
+necto_pos_obj <- densityMap(necto_pos_sim)
+necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
+plot(necto_pos_obj)
 ```

### 2017-12-07 manuscript.rmd
incorporated Casey's changes

--- 2017-12-07 manuscript.rmd
@@
 
-This manuscript is an executable document computed directly from the data, providing an explicit and reproducible description of all findings.
 All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017).
@@
 
-Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in supplementary table 1.
-Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water scuba dives.
-*Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 04/13/11.
+Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California.Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in Supplementary Information.
+Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water SCUBA dives.
+*Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
-Accession numbers are given in supplementary table 1.
+Accession numbers are given in Supplementary Information.
 In cases where physical vouchers were unavailable we provide photographs to document species identity (see git repository).
@@
 
-When possible specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (see supplementary document).
-mRNA was extracted directly from tissue using a variety of methods (see supplementary document): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006)-  in case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol directly into the Illumina TruSeq Stranded Library Kit.
+When possible, specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (see supplementary document).
+mRNA was extracted directly from tissue using a variety of methods (see supplementary document): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
+In case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol extraction of total RNA directly into the Illumina TruSeq Stranded Library Kit.
 Extractions were performed according to the manufacturer's instruction.
-Any resulting higher rRNA read counts were dealt with further downstream in the bioinformatics workflow.
+All samples were DNase treated (TURBO DNA-free, Invitrogen #AM1907;or on column DNase treatment with Zymo Quick RNA MicroPrep).
 Libraries were prepared for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep Master Mix Set (NEB, #E6110S).
@@
 In the case of large tissue inputs, libraries were sequenced separately for each tissue, subsequently subsampled and pooled *in silico*.
-Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms (see supplementary document).
+Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms.
 
@@
 New data were analysed in conjunction with 13 publically available datasets, with a total number of 43 species.
-Sequence assembly, annotation, Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma v.
+Sequence assembly, annotation, and Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma v.
 1.00 [@Dunn:2013kw; @Guang202416], and Bayesian Inference (BI) analyses were conducted using Phylobayes v.
 1.7a-mpi [@lartillot2009phylobayes].
-Source code for all analysis steps, sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
+Sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
+All sequence data have been deposited in the NCBI sequence read archive (SRA).
 
-Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
+Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
 ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
@@
 Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The CAT-GTR and CAT-Poisson models did not converge after a long CPU time, and only the results from the CAT-Poisson model are included here.
+The CAT-GTR and CAT-Poisson models did not converge, and only the results from the CAT-Poisson model are included here.
+The analyses presented here consider 33 siphonophore species and 8 outgroup species.
+This includes new data for 30 species.
+Summary statistics for expression libraries are given in Supplementary Information.
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
 
-Morphological character data used in trait mapping were obtained from the literature, or from direct observation of available voucher material.
-Depth distribution data was queried from the MBARI VARS ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/, [@schlining2006mbari]) database (with assistance and permission from SHD Haddock), including records from several MBARI cruises.
+Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
+We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
+This left `r length( phylobayes_trees )` trees in the posterior.
+The phylobayes chains did not converge, and uncertainty remains around the placement of *Erenna richardi*, and also *Nanomia bijuga*.
+Within the ML analysis, the placement of *Erenna richardi* is also unstable.
+Alternative topologies for these nodes are shown in figure 3.
+
+Morphological character data used in trait mapping were obtained from the literature or direct observation of available voucher material.
+Depth distribution data was queried from the MBARI VARS ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/, [@schlining2006mbari]) database.
 We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
@@
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
-version 0.39 (commit fd68ef5733c095c7000a4f92dc8c0daaddeec3b9) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
+version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
 For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
-3 x).
+3).
 We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
@@
 ### Species phylogeny and hypothesis testing
-Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California (table 1).
-All sequence data have been deposited in the sequence read archive (SRA).
-The analyses presented here consider 33 siphonophore species and 8 outgroup species.
-This includes new data for 30 species.
-Summary statistics for expression libraries are given in supplementary table x.
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
+
 ```{r specimen_summary, results='asis', echo=FALSE, comment=NA, warning=FALSE}
@@
 ```
-Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
-We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
-This left `r length( phylobayes_trees )` trees in the posterior.
-The phylobayes chains did not converge after a long CPU time, and uncertainty remains around the placement of *Erenna richardi*, and also *Nanomia bijuga*.
-Within the ML analysis, the placement of *Erenna richardi* is also unstable.
-Alternative topologies for these nodes are shown in figure 3 x.
 
@@
 There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae, which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
-Given the relatively shallow sampling of this analysis, we define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to the Calycophorae.
+Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig.
+3), which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
+We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
+Within Euphysonectae, we identify Clade A and Clade B (Fig.
+3).
+Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
+Clade B is a group of deep-sea siphonophores found within the bathypelagic.
+Within Clade B there is low support for the placement of *Erenna richardi*, and more taxon sampling will be required to determine the relationship of species within this clade.
 
-In this phylogeny, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *Physophora hydrostatica* (sister species to *P.
 gilmeri*) was sister to the Forskaliidae with low support.
-*Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
+The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
 is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
@@
 Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after @mackie1987siphonophore).
-As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic.
+As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic with respect to the diphyomorphs.
 *Craseoa lathetica* and *Desmophyes* sp.
@@
 ## Character Evolution
-
-The morphological character matrix, and all the code used to generate the character mapping is available in the git repository and in the supplementary materials.
-A  number of traits were mapped to tips on the phylogeny, including: the presence/absence of palpons, tentilla, bracts, pneumatophores, nectosome, and eudoxia, monoecy vs dioecy, the orientation of nectosome relative to siphosome, as well as the median vertical distribution (Fig.
-4, 5, S2-5).
-
 ```{r char_evo_setup, include=F, echo=F}
@@
 
-#Descending mantle Canal SIMMAP Setup
-desc_mantle_data = cdata[which(!is.na(cdata$Desc.Mantle.Canal)),]
-desc_mantle = desc_mantle_data$Desc.Mantle.Canal
-names(desc_mantle)=rownames(desc_mantle_data)
-DescMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(desc_mantle))))
-make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
-
-#Ascending Mantle canal SIMMAP Setup
-asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
-asc_mantle = asc_mantle_data$Asc.Mantle.Canal
-names(asc_mantle)=rownames(asc_mantle_data)
-AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
-make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
+#Mantle canal SIMMAP Setup
+#mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
+#mantle = mantle_data$Mantle.Canal
+#names(mantle)=rownames(mantle_data)
+#MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
+#make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
 ```
@@
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-sex_obj <- densityMap(sex_sim)
-sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(sex_obj$cols))
  pdf("figures_raw/figure_5_raw_sex.pdf")
- plot(sex_obj)
+ densityMap(sex_sim)
  dev.off()
@@
  add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-palpons_obj <- densityMap(palpons_sim)
-palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(palpons_obj$cols))
  pdf("figures_raw/figure_5_raw_palpons.pdf")
- plot(palpons_obj)
+ densityMap(palpons_sim)
  dev.off()
@@
 p <- p + geom_tiplab(size=2.2)
-f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:9,17,18,3)]
+f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(1,2,5:9)]
 rownames(f_cdata) = rownames(cdata)
-f_cdata = f_cdata[,c(7,4,1,3,6,2,10,8,9,5)]
+f_cdata = f_cdata[,c(4,7,1,3,6,2,5)]
 f_cdata[match(tree$tip.label,rownames(f_cdata)),]
@@
 Siphonophore colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
-Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora; however the lack of resolution of deep relationships within Codonophora made it impossible to rule out alternative evolutionary scenarios [@Dunn:2005dy].
 The current better-resolved tree (Fig.
-5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to [**placeholder name** Agalmatids *sensu lato*].
+5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Clade A.
 There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
+Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora [@Dunn:2005dy].
 
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
-Diphyomorphs have more than 1 type of propulsive nectophore, while cystonects have none.
-Here we reconstruct the evolutionary origins of the different zooid types on the present transcriptome tree (Fig.
+For example, the genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none.
+Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome tree (Fig.
 4).
 
-Nectophores are retained modified medusae that codonophorans use for coordinated colony-level swimming.
+Nectophores are non-reproductive propulsive medusae that codonophorans use for coordinated colony-level swimming (Fig.
+2).
+Nectophores are also present within the gonodendra (reproductive structures) of the cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@totton1960].
+It is not clear whether the nectophores found within the siphosome of the cystonects is homologous to the nectophores borne on the nectosome of the codonophora.
 The nectosome is the region of the colony that develops from the nectosomal growth zone (Fig.
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-It is possible that the common ancestor of siphonophores had a nectosome, which was lost on the branch leading to cystonects.
-We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to Codonophora.
-It is suggested that the nectosome arose as a duplication of the siphosome, followed by functional specialization in propelling the colony [@dunn2006evolution].
-The nectosome has been lost within Codonophora in the genus *Athorybia*.
+The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome.
 
@@
 
-Following the colony development orientation framework [@haddock_dunn_pugh_2005], the nectosome can be located in a dorsal or a ventral position.
+The nectosome can be located in a dorsal or a ventral position, following the orientation framework of @haddock_dunn_pugh_2005.
 Our ancestral reconstructions for this character (Fig.
-S5) show that a ventrally-oriented nectosome was the ancestral form in siphonophores, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+S5) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
-Bracts are highly reduced zooids unique to siphonophores, but they are only present in Codonophora (Fig.
+Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig.
 4).
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-As with the nectosome, we cannot be sure whether the common ancestor of siphonophores had bracts or not.
 The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
@@
 We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
-They were present in the common ancestor of siphonophores (Fig.
+We reconstruct them as present in the common ancestor of siphonophores (Fig.
 5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of Stephanophyes [@totton1965synopsis]), and in *Marrus claudanielis*.
-These taxa might have found other avenues to effectively circulate nutrients across the colony.
 
@@
 The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
+It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
@@
 4).
-It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
 The pneumatophore was lost in Calycophorae and never gained again in that clade.
@@
 2A).
-Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches).
-It is still an open question whether the common ancestor of Siphonophora had simple or branched tentacles.
+Most hydrozoans bear simple tentacles (tentacles with no side branches).
+It is still an open question whether the common ancestor of Siphonophora had tentilla.
 The only siphonophores species regarded as lacking tentilla are *Physalia physalis*, *Apolemia* spp., and *Bathyphysa conifera*.
@@
 Many species are strong vertical migrators, and some species such as *Chuniphyes multidentata* have exhibited synchronous diel migration patterns [@pugh1984diel].
-The phylogeny indicates that the siphonophore common ancestor was probably planktonic, as most extant siphonophore taxa are.
 Some siphonophore species (such as *Bargmannia lata*) have been observed at abyssopelagic depths near 4000m.
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-tentilla_obj <- densityMap(tentilla_sim)
-tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(tentilla_obj$cols))
-plot(tentilla_obj)
+densityMap(tentilla_sim)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-tentilla2_obj <- densityMap(tentilla_sim2)
-tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(tentilla2_obj$cols))
-plot(tentilla2_obj)
+densityMap(tentilla_sim2)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-nectosome_obj <- densityMap(nectosome_sim)
-nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(nectosome_obj$cols))
-plot(nectosome_obj)
+densityMap(nectosome_sim)
 ```
 
-```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores"}
-#Simmap Descending Mantle Canal
+```{r mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+#Simmap NECTOSOME
 # plotTree(ultratree, lwd = 4)
@@
 # names(colors) = c("Absent", "Present")
-# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-desc_mantle_obj <- densityMap(desc_mantle_sim)
-desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(desc_mantle_obj$cols))
-plot(desc_mantle_obj)
+#densityMap(mantle_sim)
 ```
 
-```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of an ascending mantle canal in the nectophores"}
-#Simmap Ascending mantle canal
-# plotTree(ultratree, lwd = 4)
-# colors = c("black", "red")
-# names(colors) = c("Absent", "Present")
-# asc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(ascribe.simmap(asc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-asc_mantle_obj <- densityMap(asc_mantle_sim)
-asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(asc_mantle_obj$cols))
-plot(asc_mantle_obj)
-```
-
-```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome"}
+```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.
+Cystonects were excluded as they do not have a nectosome."}
 #SIMMAP NectoPos
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-necto_pos_obj <- densityMap(necto_pos_sim)
-necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("blue", "orange"), space="Lab")(length(necto_pos_obj$cols))
-plot(necto_pos_obj)
+densityMap(necto_pos_sim)
 ```

### 2017-12-07 manuscript.rmd

@@
 
-#Mantle canal SIMMAP Setup
-#mantle_data = cdata[which(!is.na(cdata$Mantle.Canal)),]
-#mantle = mantle_data$Mantle.Canal
-#names(mantle)=rownames(mantle_data)
-#MantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(mantle))))
-#make.simmap(MantleTree, mantle, nsim = 100, message=FALSE) -> mantle_sim
+#Descending mantle Canal SIMMAP Setup
+desc_mantle_data = cdata[which(!is.na(cdata$Desc.Mantle.Canal)),]
+desc_mantle = desc_mantle_data$Desc.Mantle.Canal
+names(desc_mantle)=rownames(desc_mantle_data)
+DescMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(desc_mantle))))
+make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
+
+#Ascending Mantle canal SIMMAP Setup
+asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
+asc_mantle = asc_mantle_data$Asc.Mantle.Canal
+names(asc_mantle)=rownames(asc_mantle_data)
+AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
+make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(nectosome_sim)
+nectosome_obj <- densityMap(nectosome_sim)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
 ```
 
-```{r mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
-#Simmap NECTOSOME
+```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores"}
+#Simmap Descending Mantle Canal
 # plotTree(ultratree, lwd = 4)
@@
 # names(colors) = c("Absent", "Present")
-# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-#densityMap(mantle_sim)
+desc_mantle_obj <- densityMap(desc_mantle_sim)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
+```
+
+```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of an ascending mantle canal in the nectophores"}
+#Simmap Ascending mantle canal
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# asc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(ascribe.simmap(asc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+asc_mantle_obj <- densityMap(asc_mantle_sim)
+asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
+plot(asc_mantle_obj)
 ```

### 2017-12-08 manuscript.rmd
modified figures, fixed some code. New knit

--- 2017-12-07 manuscript.rmd
@@
 ```{r specimen_summary, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-specimen_data <- read.csv("supplementary_info/Specimen_data.csv")
+specimen_data <- read.csv("supplementary_info/Supplementary_data_1_Specimen_data.csv")
 names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth (m)","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
@@
 add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+sex_obj <- densityMap(sex_sim, plot=FALSE)
+sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj$cols))
  pdf("figures_raw/figure_5_raw_sex.pdf")
- densityMap(sex_sim)
+ plot(sex_obj)
  dev.off()
@@
  add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+palpons_obj <- densityMap(palpons_sim, plot=FALSE)
+palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj$cols))
  pdf("figures_raw/figure_5_raw_palpons.pdf")
- densityMap(palpons_sim)
+ plot(palpons_obj)
  dev.off()
@@
 2).
-Nectophores are also present within the gonodendra (reproductive structures) of the cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@totton1960].
+Nectophores are also present within the gonodendra (reproductive structures) of the cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
 It is not clear whether the nectophores found within the siphosome of the cystonects is homologous to the nectophores borne on the nectosome of the codonophora.
@@
 
+![Constrained topologies specified in SOWH testing.
+Test statistic and p-value for each tree estimated under constraint are given.](sowh/sowh_tests_supplemental_figure/sowh_tests_supplemental_figure.png)
+
 ```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim)
+tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
+tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
+plot(tentilla_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(tentilla_sim2)
+tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
+tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
+plot(tentilla2_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(nectosome_sim)
+nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
 ```
 
-```{r mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
-#Simmap NECTOSOME
+```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores.
+Cystonects and Athorybia were excluded as they do not have a nectosome.
+"}
+#Simmap Descending Mantle Canal
 # plotTree(ultratree, lwd = 4)
@@
 # names(colors) = c("Absent", "Present")
-# nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# desc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(describe.simmap(desc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-#densityMap(mantle_sim)
+desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
+```
+
+```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of an ascending mantle canal in the nectophores.
+Cystonects and Athorybia were excluded as they do not have a nectosome."}
+#Simmap Ascending mantle canal
+# plotTree(ultratree, lwd = 4)
+# colors = c("black", "red")
+# names(colors) = c("Absent", "Present")
+# asc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
+# nodelabels(pie=(ascribe.simmap(asc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
+# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
+asc_mantle_obj <- densityMap(asc_mantle_sim, plot=FALSE)
+asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
+plot(asc_mantle_obj)
 ```
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-densityMap(necto_pos_sim)
+necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
+necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
+plot(necto_pos_obj)
 ```
 
-#####Vertical Habitat Evolution
 ```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
@@
 ```
-
-#####Phylogenetic Signal
 

### 2017-12-09 manuscript.rmd
updated text, character mapping, figure 3

@@
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-The most familiar siphonophore species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
-Most siphonophore species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
+The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
+Most species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
 There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
@@
 Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction.
-Each siphonophore colony arises from a single embryo that forms the protozooid, the first body.
+Each colony arises from a single embryo that forms the protozooid, the first body.
 A growth zone then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
@@
 The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
-A siphonophore phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced several long standing questions about siphonophore biology.
+A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced several long standing questions about siphonophore biology.
 These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
@@
 
-Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California.Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in Supplementary Information.
+Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California.
+Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in Supplementary Information.
 Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water SCUBA dives.
 *Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
-Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
+Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
 Accession numbers are given in Supplementary Information.
@@
 Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
+This clade also shares the prescence of a descending mantle canal within the nectophores (Fig.
+S6).
 Clade B is a group of deep-sea siphonophores found within the bathypelagic.
@@
 3 xx).
-In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03).
+In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
+S2).
 
@@
 #Ascending Mantle canal SIMMAP Setup
-asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
-asc_mantle = asc_mantle_data$Asc.Mantle.Canal
-names(asc_mantle)=rownames(asc_mantle_data)
-AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
-make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
+#asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
+#asc_mantle = asc_mantle_data$Asc.Mantle.Canal
+#names(asc_mantle)=rownames(asc_mantle_data)
+#AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
+#make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
 ```
@@
 Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
-The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome.
+The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
+S5).
 
@@
 Our ancestral reconstructions for this character (Fig.
-S5) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
@@
 The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
-Phylogenetic reconstructions are sensitive to taxon sampling, and these patterns of loss are not captured by the present sampling [**We don't have the reconstruction in supp info - do we want to add it?**].
+These patterns of loss are not captured by the present sampling.
 
@@
 When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
-S2).
+S3).
 However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.
@@
 When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to *Apolemiidae* (Fig.
-S3), and in *Bathyphysa conifera*.
+S4), and in *Bathyphysa conifera*.
 
@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
-4, S6) [@mapstone2014global; @mackie1987siphonophore].
+4, S9) [@mapstone2014global; @mackie1987siphonophore].
 However, the depth distribution of siphonophore populations is not always static.
@@
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
-S6).
+S9).
 This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
@@
 4).
+We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving (Fig.
+S10), and still find support for a mesopelagic ancestor.
 There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
@@
 # add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-asc_mantle_obj <- densityMap(asc_mantle_sim, plot=FALSE)
-asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
-plot(asc_mantle_obj)
+#asc_mantle_obj <- densityMap(asc_mantle_sim, plot=FALSE)
+#asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
+#plot(asc_mantle_obj)
 ```

### 2017-12-09 manuscript.rmd
added Phil's comment, removed ascending mantle.

@@
 make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
-
-#Ascending Mantle canal SIMMAP Setup
-#asc_mantle_data = cdata[which(!is.na(cdata$Asc.Mantle.Canal)),]
-#asc_mantle = asc_mantle_data$Asc.Mantle.Canal
-#names(asc_mantle)=rownames(asc_mantle_data)
-#AscMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(asc_mantle))))
-#make.simmap(AscMantleTree, asc_mantle, nsim = 100, message=FALSE) -> asc_mantle_sim
 ```
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the genus *Forskalia* has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 10 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none.
@@
 2B).
-Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
+Unlike the siphosomal growth zone, the nectosomal growth zone does not b ud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
 The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
@@
 
-```{r asc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of an ascending mantle canal in the nectophores.
-Cystonects and Athorybia were excluded as they do not have a nectosome."}
-#Simmap Ascending mantle canal
-# plotTree(ultratree, lwd = 4)
-# colors = c("black", "red")
-# names(colors) = c("Absent", "Present")
-# asc_mantle_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
-# nodelabels(pie=(ascribe.simmap(asc_mantle_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
-# add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
-#asc_mantle_obj <- densityMap(asc_mantle_sim, plot=FALSE)
-#asc_mantle_obj$cols[1:length(asc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(asc_mantle_obj$cols))
-#plot(asc_mantle_obj)
-```
-
 ```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.

### 2017-12-10 manuscript.rmd
Fig. 1 legend update
text edits

@@
 
-![Photographs of representatives of the major groups of siphonophores.
+![Photographs of representatives of the three major groups of siphonophores.
+Cystonectae (A-B), Calycophorae (C-F) and Physonectae (G-K).
 (A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
-(B) *Bathyphysa conifera*, scale bar = XXX.
+(B) *Bathyphysa conifera*, scale bar = 2cm.
 (C) *Hippopodius hippopus*, scale bar = 5 mm.
-(D) *Physophora hydrostatica*, scale bar = 5 mm.
-(E) *Nanomia bijuga*, scale bar = 1 cm.
-(F) *Desmophyes haematogaster*, scale bar = 5 mm.
-(G) *Sphaeronectes christiansonae*, scale bar = 2 mm.
+(D) *Kephyes hiulcus*, scale bar = 2 mm.
+(E) *Desmophyes haematogaster*, scale bar = 5 mm.
+(F) *Sphaeronectes christiansonae*, scale bar = 2 mm.
+(G) *Nanomia bijuga*, scale bar = 1 cm.
 (H) *Lychnagalma utricularia*, scale bar = 1 cm.
-(I) *Kephyes hiulcus*, scale bar = 2 mm.
+(I) *Physophora hydrostatica*, scale bar = 5 mm.
 (J) *Praya dubia*, scale bar = 4 cm .
@@
 There was, in particular, little support for important deep relationships within Codonophora.
-Resolving deep relationships within Codonophora is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (Fig.
+Resolving these relationships is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (Fig.
 2).
 Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded).
-Using 1,071 genes shared across species, we find strong support for several relationships within the siphonophore phylogeny.
+Using 1,071 genes shared across species, we find strong support for several relationships within the group.
 Using this phylogeny, we reconstruct the evolutionary history of characters central to their unique biology, including zooid type, life history traits, and habitat.
@@
 3) [@Dunn:2005dy].
-There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora.
-We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans.
+There is strong support for Cystonectae as sister to all other siphonophores, the codonophorans.
+We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for Apolemidae as sister to all other codonophorans.
 Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis.
-There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig.
-3), which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
-We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
+There is strong support for Pyrostephidae as sister to all other non-Apolemid codonophorans.
+Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we call Euphysonectae (Fig.
+3), which includes the remaining non-Apolemid, non-Pyrostephid physonects.
+We define Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
 Within Euphysonectae, we identify Clade A and Clade B (Fig.
 3).
-Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
+Clade A comprises siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
 This clade also shares the prescence of a descending mantle canal within the nectophores (Fig.

### 2017-12-10 manuscript.rmd
edits, knit

@@
 ```
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
+The phylogenetic relationships recovered in this study are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
 3) [@Dunn:2005dy].
@@
 
-Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
-In the rDNA study, *Physophora hydrostatica* (sister species to *P.
-gilmeri*) was sister to the Forskaliidae with low support.
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+In the rRNA study, *Physophora hydrostatica* (sister species to *P.
+gilmeri*) was sister to Forskaliidae with low support.
 The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
-is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
+is in a clade with *Forskalia asymmetrica*, falling outside of Agalmatidae *sensu stricto*.
 Placement of *Cordagalma* outside Agalmatidae *sensu stricto* is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].
@@
 5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Clade A.
-There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of the Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
+There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
 Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora [@Dunn:2005dy].
@@
 Nectophores are also present within the gonodendra (reproductive structures) of the cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
-It is not clear whether the nectophores found within the siphosome of the cystonects is homologous to the nectophores borne on the nectosome of the codonophora.
+It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans.
 The nectosome is the region of the colony that develops from the nectosomal growth zone (Fig.
 2B).
-Unlike the siphosomal growth zone, the nectosomal growth zone does not b ud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
+Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
 The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
@@
 Our ancestral reconstructions for this character (Fig.
-S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
 
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
 These patterns of loss are not captured by the present sampling.
@@
 We reconstruct them as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of Stephanophyes [@totton1965synopsis]), and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of *Stephanophyes* [@totton1965synopsis]), and in *Marrus claudanielis*.
 
@@
 ## Conclusions
-Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support.
-A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemid codonophorans.
+Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
+A previous rRNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemid codonophorans.
 Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
-Future studies with the addition of key taxa within the Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
+Future studies with the addition of key taxa within Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
 

### 2017-12-11 manuscript.rmd
added Phil's changes

--- 2017-12-09 manuscript.rmd
@@
 Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis.
-There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans.
+There is strong support for the Pyrostephidae as sister to all other non-apolemiid codonophorans.
 Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig.
-3), which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae".
+3), which includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
 We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
@@
 This clade also shares the prescence of a descending mantle canal within the nectophores (Fig.
-S6).
+S6), and is also monoecious (Fig.
+5).
 Clade B is a group of deep-sea siphonophores found within the bathypelagic.
@@
 
-Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *Physophora hydrostatica* (sister species to *P.
@@
 The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
-is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*.
-Placement of *Cordagalma* outside Agalmatidae *sensu stricto* is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].
+is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae.
+Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].
 
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 10 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none.
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodiidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
 These patterns of loss are not captured by the present sampling.
@@
 We reconstruct them as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of Stephanophyes [@totton1965synopsis]), and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.
 
@@
 Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support.
-A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemid codonophorans.
+A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemiid codonophorans.
 Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.

### 2017-12-11 manuscript.rmd

@@
 ```
-These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
+
+The phylogenetic relationships recovered in this study are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
 3) [@Dunn:2005dy].
@@
 Our ancestral reconstructions for this character (Fig.
-S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to the Agalmatidae *sensu stricto* and the branch leading to the *Bargmannia* species.
+S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to Agalmatidae and the branch leading to the *Bargmannia* species.
 
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodiidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
 These patterns of loss are not captured by the present sampling.
@@
 ## Conclusions
-Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support.
-A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemiid codonophorans.
+Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
+A previous rRNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemiid codonophorans.
 Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
-Future studies with the addition of key taxa within the Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
+Future studies with the addition of key taxa within Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
 

### 2017-12-15 manuscript.rmd
added section on different types of monoecy

@@
 5).
-Clade B is a group of deep-sea siphonophores found within the bathypelagic.
 Within Clade B there is low support for the placement of *Erenna richardi*, and more taxon sampling will be required to determine the relationship of species within this clade.
@@
 
+In all siphonophores, each gonophoren(sexual medusa that produces gametes) is either male or female.
 Siphonophore colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
@@
 Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora [@Dunn:2005dy].
+
+The Calycophorae and Clade A each have very different arangements of monoecy, providing evidence for multiple gains of monoecy.
+In both the Calycophorae and Clade A, male and female gonophores are located within the same colony.
+In calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
+2) bears either male or female zooids.
+In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row.
+In either case, male and female zooids are found at the same corresponding locations within the cormidia.
+In Clade A, however, male and female zooids are found within the same cormidium.In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
 

### 2017-12-18 manuscript.rmd
photo credit Bathyphysa
updated specimen table

@@
 
-The Calycophorae and Clade A each have very different arangements of monoecy, providing evidence for multiple gains of monoecy.
+The Calycophorae and Clade A each have very different arrangements of male and female gonophores, providing evidence for multiple gains of monoecy.
 In both the Calycophorae and Clade A, male and female gonophores are located within the same colony.
@@
 SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster.
+We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph.
 We also thank the MBARI crews and ROV pilots for collection of the specimens.

### 2017-12-18 manuscript.rmd
fixed supplementary figs captions

@@
 #####Stochastic Character maps
-```{r tentilla, echo=F, message=F,warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla"}
+```{r tentilla, echo=F, message=F,warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla."}
 #Simmap TENTILLA
@@
 
-```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included"}
+```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla."}
 #Simmap TENTILLA
@@
 
-```{r nectosome, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome"}
+```{r nectosome, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
 #Simmap NECTOSOME

### 2017-12-18 manuscript.rmd
added model testing for median depth reconstruciton
added model testing for palpons

@@
 #Simmap PALPONS
+
+##MODEL test
+fitDiscrete(ultratree, palpons, model="ER")$opt$aicc #24.85
+fitDiscrete(ultratree, palpons, model="ARD")$opt$aicc #24.60
+
  plotTree(ultratree, lwd = 4)
@@
 
+fitContinuous(ultratree, median_extended, model="white")$opt$aicc
+fitContinuous(ultratree, median_extended, model="BM")$opt$aicc
+fitContinuous(ultratree, median_extended, model="OU")$opt$aicc
+
 ```

### 2017-12-29 manuscript.rmd
cited support for BM in vertical evolution text
added Casey edits

--- 2017-12-18 manuscript.rmd
@@
 Siphonophores (Fig.
-1) are among the most abundant gelatinous predators in the open ocean, and play important ecological roles in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous].
+1) are among the most abundant gelatinous predators in the open ocean, and play important ecological roles in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous; @Choy20172116].
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
+The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the ocean surface and can conspicuously wash ashore onto beaches in large numbers [@Totton1960].
 Most species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
@@
 mRNA was extracted directly from tissue using a variety of methods (see supplementary document): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
-In case of anticipated very small total RNA quantities, only a single round of bead purification was performed; or Trizol extraction of total RNA directly into the Illumina TruSeq Stranded Library Kit.
+In case of anticipated very small total RNA quantities, only a single round of bead purification was performed.
 Extractions were performed according to the manufacturer's instruction.
@@
 Sequence assembly, annotation, and Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma v.
-1.00 [@Dunn:2013kw; @Guang202416], and Bayesian Inference (BI) analyses were conducted using Phylobayes v.
+1.00 [@Dunn:2013kw; @Guang202416].
+This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @stamatakis2006raxml, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
+Bayesian Inference (BI) analyses were conducted using Phylobayes v.
 1.7a-mpi [@lartillot2009phylobayes].
@@
 All sequence data have been deposited in the NCBI sequence read archive (SRA).
+Summary statistics for expression libraries are given in Supplementary Information.
 
@@
 Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The CAT-GTR and CAT-Poisson models did not converge, and only the results from the CAT-Poisson model are included here.
+The CAT-GTR and CAT-Poisson models did not converge (CAT-Poisson: maxdiff= 1 &
+meandiff=0.0147679; CAT-GTR: maxdiff = 1 & meandiff= 0.0126582), and only the results from the CAT-Poisson model are included here.
 The analyses presented here consider 33 siphonophore species and 8 outgroup species.
 This includes new data for 30 species.
-Summary statistics for expression libraries are given in Supplementary Information.
 In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
@@
 #### Hypothesis testing
-We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
-As the sexual system of *Rudjakovia* sp.
-is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
-included as a monoecious species, and one without.
-We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
-version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
+We implemented the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] using SOWHAT [@doi:10.1093/sysbio/syv055] dev.
+version 0.39 (commit fd68ef57).
+The SOWH tests were carried out in parallel with the default options and an initial sample size of 100.
 For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
@@
 
-We tested the following three alternative phylogenetic hypotheses against the most likely tree topology: (i) physonect siphonophores are monophyletic, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic (Fig.
-3 xx).
+Using the SOWH test, we statistically tested two alternative phylogenetic hypotheses against the maximum-likelihood tree topology recovered in these analyses.
+We evaluated two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy] (Fig.
+3).
+As the sexual system of *Rudjakovia* sp.
+is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
+included as a monoecious species, and one without.
 In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
-S2).
+S2), supporting the present topology, where physonects and monoecious species are paraphyletic.
 
@@
 
-In all siphonophores, each gonophoren(sexual medusa that produces gametes) is either male or female.
+In all siphonophores, each gonophore (sexual medusa that produces gametes) is either male or female.
 Siphonophore colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny).
-These patterns of loss are not captured by the present sampling.
+The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
+These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.
 

### 2017-12-29 manuscript.rmd

--- 2017-12-18 manuscript.rmd
@@
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
-S9).
+S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
 This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.

### 2018-01-03 manuscript.rmd
removed pdf output into knitr document
edits throughout manuscript, did not rebuild the pdf

@@
 Siphonophores (Fig.
-1) are among the most abundant gelatinous predators in the open ocean, and play important ecological roles in oceanic waters [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous; @Choy20172116].
+1) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous] [XX cite Anela and Steve's new paper].
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
-The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the ocean surface and can conspicuously wash ashore onto beaches in large numbers [@Totton1960].
+The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
 Most species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
 There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
-In all, there are currently 188 valid described species.
+There are currently 188 valid described species.
 
@@
 Each colony arises from a single embryo that forms the protozooid, the first body.
-A growth zone then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+One or two growth zones then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from all other colonial animals in colony structure, development, and the degree to which they are functionally specialized [@Beklemishev1969;@cartwright2010character].
+Siphonophores differ significantly from all other colonial animals in colony structure, development, and the high degree to which their zooids are functionally specialized [@Beklemishev1969;@cartwright2010character].
 Each zooid is functionally specialized for a particular task (*e.g.*, feeding, reproducing, swimming) (Fig.
@@
 There was, in particular, little support for important deep relationships within Codonophora.
-Resolving these relationships is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) or the gain and loss of particular zooids, such as palpons (Fig.
+Resolving these relationships is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig.
 2).
@@
 Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms.
+All sequence data have been deposited in the NCBI sequence read archive (SRA) with accession number XXX.
 
@@
 
-New data were analysed in conjunction with 13 publically available datasets, with a total number of 43 species.
-Sequence assembly, annotation, and Maximum Likelihood (ML) phylogenetic analysis were conducted with the tool Agalma v.
+New data were analysed in conjunction with 13 publicly available datasets, with a total number of 43 species.
+Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.
 1.00 [@Dunn:2013kw; @Guang202416].
 This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @stamatakis2006raxml, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
-Bayesian Inference (BI) analyses were conducted using Phylobayes v.
+Maximum likelihood analyses of the supermatrix were conducted with XX.
+Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
 1.7a-mpi [@lartillot2009phylobayes].
 Sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
-All sequence data have been deposited in the NCBI sequence read archive (SRA).
-Summary statistics for expression libraries are given in Supplementary Information.
 
 Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
-ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates.
+The final analyses presented here consider 33 siphonophore species and 8 outgroup species.
+This includes new data for 30 species.
+Summary statistics for expression libraries are given in Supplementary Information [XX This sentence seems out of place - delete?
+Or rephrase and move up.].
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
+
+ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
 BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
 Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The CAT-GTR and CAT-Poisson models did not converge (CAT-Poisson: maxdiff= 1 &
-meandiff=0.0147679; CAT-GTR: maxdiff = 1 & meandiff= 0.0126582), and only the results from the CAT-Poisson model are included here.
-The analyses presented here consider 33 siphonophore species and 8 outgroup species.
-This includes new data for 30 species.
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
+The CAT-GTR and CAT-Poisson models did not converge [XX include the maxdiff in the text here], and only the results from the CAT-Poisson model are included here.
+Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
+This left `r length( phylobayes_trees )` trees in the posterior.
 
-Maximimum likelihood analyses had `r length(bootstrap_trees)` replicates.
-We ran `r length( input_files_phylobayes )` phylobayes chains, and visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all runs.
-This left `r length( phylobayes_trees )` trees in the posterior.
-The phylobayes chains did not converge, and uncertainty remains around the placement of *Erenna richardi*, and also *Nanomia bijuga*.
-Within the ML analysis, the placement of *Erenna richardi* is also unstable.
-Alternative topologies for these nodes are shown in figure 3.
+We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
+As the sexual system of *Rudjakovia* sp.
+is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
+included as a monoecious species, and one without.
+We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
+version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
+For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
+3).
+We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
 
@@
 See Supplementary Information for R package version numbers.
-
-#### Hypothesis testing
-We implemented the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] using SOWHAT [@doi:10.1093/sysbio/syv055] dev.
-version 0.39 (commit fd68ef57).
-The SOWH tests were carried out in parallel with the default options and an initial sample size of 100.
-For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
-3).
-We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
 
@@
 
-```{r species_phylogram, echo=FALSE, comment=NA, warning=FALSE }
+```{r species_phylogram, verbose= FALSE, include=F, message = FALSE, echo=FALSE, comment=NA, warning=FALSE }
 
@@
 
-The phylogenetic relationships recovered in this study are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig.
-3) [@Dunn:2005dy].
-There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora.
-We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans.
-Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis.
+Most relationships received strong support across analysis methods (Figure 3A), with a couple localized exceptions (Figure 3B).
+The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
+Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
+Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses.
 There is strong support for the Pyrostephidae as sister to all other non-apolemiid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig.
-3), which includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
+Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we here name Euphysonectae (Fig.
+3).
+It includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
 We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
-Within Euphysonectae, we identify Clade A and Clade B (Fig.
-3).
+
+Euphysonectae consists of two reciprocally monophyletic groups we here provisionally refer to as Clade A and Clade B (Fig.
+3A).
 Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
@@
 5).
-Within Clade B there is low support for the placement of *Erenna richardi*, and more taxon sampling will be required to determine the relationship of species within this clade.
+Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses, and more taxon sampling will be required to determine the relationship of species within this clade.
 
 Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
-In the rDNA study, *Physophora hydrostatica* (sister species to *P.
+In the rDNA study, *Physophora hydrostatica* (the presumed sister species to *P.
 gilmeri*) was sister to the Forskaliidae with low support.
@@
 Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
-Calycophorans have in the past been split into convenient groupings of prayomorph and diphyomorph based on morphology (after @mackie1987siphonophore).
-As in the previous study, this phylogeny indicates that the prayomorphs are paraphyletic with respect to the diphyomorphs.
+Calycophorans have in the past been split into two groups, prayomorph and diphyomorph, based on morphology (after @mackie1987siphonophore).
+As in the previous study, the results presented here indicate that the prayomorphs are paraphyletic with respect to the diphyomorphs.
 *Craseoa lathetica* and *Desmophyes* sp.
@@
 
-Using the SOWH test, we statistically tested two alternative phylogenetic hypotheses against the maximum-likelihood tree topology recovered in these analyses.
-We evaluated two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy] (Fig.
-3).
-As the sexual system of *Rudjakovia* sp.
-is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
-included as a monoecious species, and one without.
-In all three cases the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
-S2), supporting the present topology, where physonects and monoecious species are paraphyletic.
+Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
+3C): (i) monophyletic Physonectae, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic.
+We performed tests (ii) and (iii) because it is not known if *Rudjakovia* sp.
+is monoecious or dioecious.
+In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
+S2).
 
@@
 In all siphonophores, each gonophore (sexual medusa that produces gametes) is either male or female.
-Siphonophore colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
+Within each siphonophore species, colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
 The current better-resolved tree (Fig.
@@
 
-The Calycophorae and Clade A each have very different arrangements of male and female gonophores, providing evidence for multiple gains of monoecy.
-In both the Calycophorae and Clade A, male and female gonophores are located within the same colony.
+Calycophorae and Clade A each have very different arrangements of male and female gonophores, consistent with the multiple gains of monoecy supported here.
 In calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
@@
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-In Clade A, however, male and female zooids are found within the same cormidium.In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
+In Clade A, however, male and female zooids are found within the same cormidium.
+In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
 
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
-For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
-Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none.
+For example, members of *Forskalia* have 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
 Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome tree (Fig.
@@
 It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans.
-The nectosome is the region of the colony that develops from the nectosomal growth zone (Fig.
+The nectosome is the region of the codonophoran colony that develops from the nectosomal growth zone (Fig.
 2B).
-Unlike the siphosomal growth zone, the nectosomal growth zone does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons).
+The nectosomal growth zone buds nectophores (and in the case of *Apolemia*, also palpons).
 The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
@@
 Our ancestral reconstructions for this character (Fig.
-S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to Agalmatidae and the branch leading to the *Bargmannia* species.
+S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to Agalmatidae and in *Bargmannia*.
 
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
+The common ancestor of Codonophora had only one bract subtype [XX provide citation, indicate which analyses support this, or remove].
+Bracts were lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
 These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.
@@
 The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
-It is not considered a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
+It is not a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
 Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
@@
 S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
-This model indicates a mesopelagic common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
+This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
 This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.

### 2018-01-03 manuscript.rmd
manuscript edits through line 306, did not rebuild pdf

@@
 
-# Siphonophore Phylogeny
+# Phylogenetic analyses of trait evolution in Siphonophora (Cnidaria)
 
@@
 Siphonophores (Fig.
-1) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical;@purcell1981dietary;@pages2001gelatinous] [XX cite Anela and Steve's new paper].
+1 and 2) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical; @purcell1981dietary; @pages2001gelatinous;  @Choy20172116].
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
 The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
-Most species are planktonic, living in the water column, where some grow to be up to 30 meters in length [@mackie1987siphonophore].
+Most species are planktonic, living in the water column, where some grow to be more than 30 meters in length [@mackie1987siphonophore].
 There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
-There are currently 188 valid described species.
+There are currently 188 valid described siphonophore species.
 
 Siphonophores remain poorly known, in large part because they are fragile and difficult to collect.
-They have, however, been of great interest for more than 150 years due to their very unique structure and development [@mapstone2014global;@mackie1987siphonophore].
+They have, however, been of great interest for more than 150 years in part due to their unique structure and development [@mapstone2014global;@mackie1987siphonophore].
 Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction.
 Each colony arises from a single embryo that forms the protozooid, the first body.
-One or two growth zones then arises that asexually produces other genetically identical zooids [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+One or two growth zones (Fig.
+2) then arises that asexually produces other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from all other colonial animals in colony structure, development, and the high degree to which their zooids are functionally specialized [@Beklemishev1969;@cartwright2010character].
-Each zooid is functionally specialized for a particular task (*e.g.*, feeding, reproducing, swimming) (Fig.
-2) [@dunn2006evolution], and they are arranged in a repeating, species-specific pattern.
+Siphonophores differ significantly from other colonial animals in colony structure, development, and the high degree to which their zooids are functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character].
+The functions that zooids are specialized for include feeding, reproducing, or swimming (Fig.
+2) [@dunn2006evolution].
 
-![Photographs of representatives of the three major groups of siphonophores.
-Cystonectae (A-B), Calycophorae (C-F) and Physonectae (G-K).
+![Photographs of living siphonophores.
+They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-F, J), Clade A within Euphysonectae (G-I) and Apolemidae (K).
 (A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
@@
 
-Understanding the unique ecology, morphology, and development of siphonophores requires an understanding of their phylogenetic relationships.
+Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group.
 The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
@@
 These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
-The cystonects were found to be sister to all other siphonophores, while the calycophorans were nested within physonects.
-The name Codonophora was given to this clade of physonects and calycophorans [@Dunn:2005dy].
+The Cystonectae was found to be sister to all other siphonophores, while the Calycophorae were nested within "Physonectae".
+The name Codonophora was given to this clade of "Physonectae" and "Calycophorae" [@Dunn:2005dy].
 
@@
 There was, in particular, little support for important deep relationships within Codonophora.
-Resolving these relationships is key to resolving the evolution of several traits, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig.
+Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig.
 2).
-Here we present a broadly sampled phylogenomic analysis of Siphonophora, assessing transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded).
-Using 1,071 genes shared across species, we find strong support for several relationships within the group.
-Using this phylogeny, we reconstruct the evolutionary history of characters central to their unique biology, including zooid type, life history traits, and habitat.
+Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling).
+Using 1,071 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study.
+Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.
 
-![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top, and the ventral side to the left.
+![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top of the page, and the ventral side to the left.
 Adapted from http://commons.wikimedia.org/wiki/File:Nanomia_bijuga_whole_animal_and_growth_zones.svg, drawn by Freya Goetz.
-(*A*) Overview of the whole mature colony.
-(*B*) Inset of the nectosomal growth zone with pneumatophore.
+(A) Overview of the whole mature colony.
+(B) Inset of the pneumatophore and nectosomal growth zone.
 A series of buds gives rise to nectophores.
-(*C*) Inset of the siphosomal growth zone.
+(C) Inset of the siphosomal growth zone.
 Probuds subdivide to give rise to zooids in repeating-units (cormidia).
-Cormidial boundaries are marked by a gastrozooid.](figures/Figure2.pdf)
+The gastrozooid is the posterior-most zooid within each cormidium.](figures/Figure2.pdf)
 
@@
 Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California.
-Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods and sequencing details are summarized in Supplementary Information.
+Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods, and sequencing details are summarized in the file Supplementary data 1.
 Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water SCUBA dives.
@@
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
-Accession numbers are given in Supplementary Information.
+Accession numbers are given in Supplementary data 1.
 In cases where physical vouchers were unavailable we provide photographs to document species identity (see git repository).
@@
 
-When possible, specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time point of specimen collection (see supplementary document).
-mRNA was extracted directly from tissue using a variety of methods (see supplementary document): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
+When possible, specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time of specimen collection.
+mRNA was extracted directly from tissue using a variety of methods (Supplementary data 1): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
 In case of anticipated very small total RNA quantities, only a single round of bead purification was performed.
@@
 Libraries were sequenced on the HiSeq 2000, 2500, and 3000 sequencing platforms.
+Summary statistics for each library are given in the file Supplementary data 2.
 All sequence data have been deposited in the NCBI sequence read archive (SRA) with accession number XXX.
@@
 This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @stamatakis2006raxml, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
-Maximum likelihood analyses of the supermatrix were conducted with XX.
+Maximum likelihood analyses of the supermatrix were conducted with [XX add raxml version and citation].
 Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
 1.7a-mpi [@lartillot2009phylobayes].
-Sequence alignments, sampled and consensus trees, and voucher information are available in a git repository https://github.com/caseywdunn/siphonophore_phylogeny_2017.
+Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository.
 
@@
 This includes new data for 30 species.
-Summary statistics for expression libraries are given in Supplementary Information [XX This sentence seems out of place - delete?
-Or rephrase and move up.].
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (gene occupancy matrix - supplementary figure 1).
+In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (Fig.
+S1).
 
@@
 Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The CAT-GTR and CAT-Poisson models did not converge [XX include the maxdiff in the text here], and only the results from the CAT-Poisson model are included here.
-Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
+The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0127).
+Only the results from the CAT-Poisson model are presented here.Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
 This left `r length( phylobayes_trees )` trees in the posterior.
 
-We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses: (i) physonects are monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
+We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses (Fig.
+3C, S2): (i) "Physonectae" is monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
 As the sexual system of *Rudjakovia* sp.
 is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
-included as a monoecious species, and one without.
+included as a monoecious species, and one without (Fig.
+S2).
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
-version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (source code can be found in the git repository).
+version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (analysis code can be found in the git repository).
 For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
@@
 Morphological character data used in trait mapping were obtained from the literature or direct observation of available voucher material.
-Depth distribution data was queried from the MBARI VARS ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/, [@schlining2006mbari]) database.
+Depth distribution data was queried from the MBARI VARS database ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/) [@schlining2006mbari].
 We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
@@
 
-Most relationships received strong support across analysis methods (Figure 3A), with a couple localized exceptions (Figure 3B).
+Most relationships received strong support across analysis methods (Fig.
+3A), with a couple localized exceptions (Fig.
+3B).
 The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we here name Euphysonectae (Fig.
-3).
+3A).
 It includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
@@
 5).
-Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses, and more taxon sampling will be required to determine the relationship of species within this clade.
+Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses (Fig.
+3B).
+More taxon sampling will be required to determine the relationship of species within this clade.
 
@@
 is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae.
-Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].
+Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of morphological and molecular data [@pugh2006taxonomic; @Dunn:2005dy].
 
-Within the calycophorans, taxon sampling is shallower, however there is broad agreement with the previous analysis.
-Calycophorans have in the past been split into two groups, prayomorph and diphyomorph, based on morphology (after @mackie1987siphonophore).
+Within Calycophorae, taxon sampling is shallower here than in the previous study.
+The calycophoran relationships that can be investigated, however, are in broad agreement with the previous analysis.
+Calycophorans have in the past been split into two groups, prayomorphs and diphyomorphs, based on morphology (after @mackie1987siphonophore).
 As in the previous study, the results presented here indicate that the prayomorphs are paraphyletic with respect to the diphyomorphs.

### 2018-01-04 manuscript.rmd
finished pass at edits in main text

@@
 Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository.
+Tree figures were rendered with ggtree [@Yu:2016fo].
 
@@
 
-![Phylogram of siphonophore relationships. Node labels indicate bootstrap support percent, unnumbered nodes have 100% support. The image was rendered with ggtree [@Yu:2016fo].](figures/Figure3.pdf)
+![(A) Maximum likelihood (ML) phylogram with bipartation frequencies from the ML bootstraps and the Bayesian posterior distribution of trees.
+Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
+(B) The topologies found in the psoterior distribution of trees that conflict with the ML tree.
+(C) The topologies evaluated by the SOWH tests.
+For more details on the SOWH topologies refer to Fig.
+S2.](figures/Figure3.pdf)
 
 Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) monoecious siphonophores (not including *Rudjakovia* sp.) are monophyletic, and (iii) monoecious siphonophores (including *Rudjakovia* sp.) are monophyletic.
-We performed tests (ii) and (iii) because it is not known if *Rudjakovia* sp.
-is monoecious or dioecious.
+3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores (both with and without *Rudjakovia* sp.).
 In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
@@
 Within each siphonophore species, colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
-The current better-resolved tree (Fig.
-5A), indicates that monoecy in siphonophores evolved twice from a dioecious ancestor, in the branch leading to Calycophorae and in the branch leading to Clade A.
-There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Codonophora, with a subsequent shift to dioecy in the *Marrus-Erenna* clade.
-Previous analyses suggested that the common ancestor of siphonophores was dioecious, and there is a single gain of monoecy within the Codonophora [@Dunn:2005dy].
+Previous analyses suggested that the common ancestor of siphonophores was dioecious, and was consistent with a single gain of monoecy within the Codonophora and no secondary losses [@Dunn:2005dy].
+The better-resolved tree in the current analyses indicates that the evolution of monoecy is more complicated than this.
+The two clades of monophyloetic siphonophores, Calycophorae and Clade A (Fig.
+3A), do not form a monophyletic group.
+This is because Clade B, which contains dioecious species, is also desceneded from their most recent common ancestor.
+SOWH tests strongly reject the placement of the monecious clades Calycophorae and Clade A as a group that excludes Clade B (Figs.
+3C and S2).
+The positions of the only two taxa from Clade B in the previous analysies [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
+This difference in conclusions regarding trait evolution therefore does not reflect a contradication between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
 
-Calycophorae and Clade A each have very different arrangements of male and female gonophores, consistent with the multiple gains of monoecy supported here.
+The distribution of monoecy is consistent with two scenarios.
+In the first, there is a single shift from dioecy to monoecy along the branch that give rise to the most recont common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
+In the second, monoecy arose twice -- once along the branch that gave rise to Clade A and again along the branch that gave rise to Calycophorae.
+
+Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig.
+5A), once in Calycophorae and once in Clade A.
+This is consistent with differences in the arrangements of male and female gonophores in the two clades.
 In calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
@@
 
-One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character].
+One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969; @cartwright2010character].
 For example, members of *Forskalia* have 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
@@
 
-Nectophores are non-reproductive propulsive medusae that codonophorans use for coordinated colony-level swimming (Fig.
-2).
-Nectophores are also present within the gonodendra (reproductive structures) of the cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
+Nectophores (Fig.
+2) are non-reproductive propulsive medusae.
+In Codonophora, the nectophores are localized to a region known as the nectosome (Fig.
+2B), which has its own growth zone, and are used for coordinated colony-level swimming.
+Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
 It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans.
-The nectosome is the region of the codonophoran colony that develops from the nectosomal growth zone (Fig.
-2B).
-The nectosomal growth zone buds nectophores (and in the case of *Apolemia*, also palpons).
-The present tree, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
+The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
 S5).
@@
 
-The nectosome can be located in a dorsal or a ventral position, following the orientation framework of @haddock_dunn_pugh_2005.
+Within the nectosome, the nectophores can be attached along the dorsal or ventral side of the stem, following the orientation framework of @haddock_dunn_pugh_2005.
 Our ancestral reconstructions for this character (Fig.
-S8) show that a ventrally-oriented nectosome was the ancestral form in Codonophora, and that a dorsal nectosome has evolved twice independently, in the branch leading to Agalmatidae and in *Bargmannia*.
+S8) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment hasindependently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
 
@@
 
-The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla (Fig.
+The most complex nematocyst batteries of Cnidaria are found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla (Fig.
 2A).
@@
 
-Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
-4, S9) [@mapstone2014global; @mackie1987siphonophore].
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs.
+4 and S9) [@mapstone2014global; @mackie1987siphonophore].
 However, the depth distribution of siphonophore populations is not always static.

### 2018-01-10 manuscript.rmd
incorporated Phil's changes

@@
 ![Photographs of living siphonophores.
-They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-F, J), Clade A within Euphysonectae (G-I) and Apolemidae (K).
+They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-F, J), Clade A within Euphysonectae (G-I) and Apolemiidae (K).
 (A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
@@
 
-![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top of the page, and the ventral side to the left.
+![Schematic of the siphonophore *Nanomia bijuga*, orientated with the anterior of the colony at the top of the page, and the ventral side to the left.
 Adapted from http://commons.wikimedia.org/wiki/File:Nanomia_bijuga_whole_animal_and_growth_zones.svg, drawn by Freya Goetz.
@@
 The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
-Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
+Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
 Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses.
@@
 3A).
-Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis].
+Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the cnidoband [@totton1965synopsis].
 This clade also shares the prescence of a descending mantle canal within the nectophores (Fig.
@@
 The calycophoran relationships that can be investigated, however, are in broad agreement with the previous analysis.
-Calycophorans have in the past been split into two groups, prayomorphs and diphyomorphs, based on morphology (after @mackie1987siphonophore).
+Calycophorans have in the past been split into two groups, prayomorphs and diphyomorphs, based on morphology after @mackie1987siphonophore.
 As in the previous study, the results presented here indicate that the prayomorphs are paraphyletic with respect to the diphyomorphs.
 *Craseoa lathetica* and *Desmophyes* sp.
-are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *Craseoa lathetica* and the clade including *Hippopodius hippopus* was unresolved.
+are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *C.
+lathetica* and the clade including *H.
+hippopus* was unresolved.
 
@@
 This is consistent with differences in the arrangements of male and female gonophores in the two clades.
-In calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
+In Clade A, male and female zooids are found within the same cormidium.
+In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
+Meanwhile in calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
 2) bears either male or female zooids.
@@
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-In Clade A, however, male and female zooids are found within the same cormidium.
-In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
+However, there are exceptions -- in *Sulculeolaria*, colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
 
@@
 It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans.
+Similarly, the homology of the special nectophore associated with gonophores of the calycophoran *Stephanophyes superba* is also unclear [@chun1891canarischen].
+As such, we only consider the evolution of the nectosome in this study, and not the presence/absence of nectophores.
 The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
@@
 Our ancestral reconstructions for this character (Fig.
-S8) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment hasindependently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
+S8) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
 
@@
 It is not a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
-Recent evidence of neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
+Recent descriptions of the neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
 The ancestral siphonophore had a pneumatophore (Fig.
-2B), since both cystonects and most codonophorans possess one (Fig.
+2B), since both cystonects and all "physonects" possess one (Fig.
 4).
@@
 
-The most complex nematocyst batteries of Cnidaria are found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla (Fig.
+The most complex nematocyst batteries of Cnidaria are found among the siphonophores.
+In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig.
 2A).
-Most hydrozoans bear simple tentacles (tentacles with no side branches).
+Outside of Siphonophora, most hydrozoans bear simple tentacles without side branches.
 It is still an open question whether the common ancestor of Siphonophora had tentilla.
@@
 Since *B.
-conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we can safely assume this is a case of secondary loss.
+conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we assume this is a case of secondary loss.
 When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
@@
 4 and S9) [@mapstone2014global; @mackie1987siphonophore].
-However, the depth distribution of siphonophore populations is not always static.
-Many species are strong vertical migrators, and some species such as *Chuniphyes multidentata* have exhibited synchronous diel migration patterns [@pugh1984diel].
-Some siphonophore species (such as *Bargmannia lata*) have been observed at abyssopelagic depths near 4000m.
+The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel].
+Some species such as *Nanomia bijuga* exhibit synchronous diel migration patterns [@barham1966deep].
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
@@
 Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
-A previous rRNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemiid codonophorans.
-Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
+A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans.
+Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
 Future studies with the addition of key taxa within Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.

### 2018-01-11 manuscript.rmd
expanded broader perspectives; conclusions

@@
 1.00 [@Dunn:2013kw; @Guang202416].
-This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @stamatakis2006raxml, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
-Maximum likelihood analyses of the supermatrix were conducted with [XX add raxml version and citation].
+This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575,, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
+Maximum likelihood analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
 Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
@@
 Most relationships received strong support across analysis methods (Fig.
-3A), with a couple localized exceptions (Fig.
+3A), with a couple of localized exceptions (Fig.
 3B).
@@
 Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses.
-There is strong support for the Pyrostephidae as sister to all other non-apolemiid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we here name Euphysonectae (Fig.
+There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans.
+Within the clade that is sister to the Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig.
 3A).
 It includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
-We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
+We define Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.
 
@@
 Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
-In the rDNA study, *Physophora hydrostatica* (the presumed sister species to *P.
+In the rDNA study, *P.
+hydrostatica* (the presumed sister species to *P.
 gilmeri*) was sister to the Forskaliidae with low support.
@@
 S2).
+
+The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae.
+This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focussed.
+Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig.
+3A, 3B).
+Expanding sampling of species that probably fall in Clade B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Clade B synapomorphies.
+Similarly, within Calycophorae, increased taxon sampling is needed.
+This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons.
+In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis flyoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
+Expanded taxon sampling, particularly *P.
+dubia* or a nectopyramid, but also extensive sampling across the major prayomorph and diphyomorph groups, will expand our understanding of relationships within Calycophorae.
+This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig.
+4), or the arrangement of male and female zooids along the stem (see below).
 
@@
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-However, there are exceptions -- in *Sulculeolaria*, colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+However, there are exceptions -- in *Sulculeolaria* (not included in this phylogeny), colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
 
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-The common ancestor of Codonophora had only one bract subtype [XX provide citation, indicate which analyses support this, or remove].
 Bracts were lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
@@
 This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
-This reconstruction only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
+There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
+Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis, suggesting that depth may be more labile than it appears.
+Phylogenetic conservatism is also observed for the mesopelagic Apolemiidae and Pyrostephidae clades, but not for the Calycophorae, which show a diversified evolutionary trajectory for depth.
+The present sampling, however, is not sufficient to capture significant variation in depth distributions between closely related species.
+Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical].
+
+This reconstruction (Fig.
+S9) only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
 The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives.
@@
 4).
-We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving (Fig.
+We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or swimming (Fig.
 S10), and still find support for a mesopelagic ancestor.
-There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
+It is important to note, however, that *H.
+hippopus* and *C.
+appendiculata* were both collected in the bay of  Villefrance-sur-mer, France, where an upwelling is known to bring deeper species closer to the surface [@nival1976evolution].
 
@@
 A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans.
+We find support for two monophyletic groups within the clade that is sister to Pyrostephidae -- Calycophorae and a new clade we name Euphysonectae.
+Within Euphysonectae, we find support for two monophyletic groups dubbed Clade A and Clade B.
 Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
-Future studies with the addition of key taxa within Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
+These results are consistent with a hypothesis that monoecy evolved twice, once in Clade A and once in Calycophorae.
+This hypothesis is consistent with differences in the arrangement of male and female zooids in the two monoecious groups.
+Future studies with the addition of key taxa within Euphysonectae will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution, including the evolution of zooid types, that cannot be fully assessed here.
 

### 2018-01-12 manuscript.rmd
fixed typo

@@
 This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons.
-In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis flyoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
+In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
 Expanded taxon sampling, particularly *P.

### 2018-01-12 manuscript.rmd
fixed typos, updated sup figures references in text

@@
 
-# Phylogenetic analyses of trait evolution in Siphonophora (Cnidaria)
+# Improved phylogenetic resolution within Siphonophora (Cnidaria) with implications for trait evolution
 
@@
 One or two growth zones (Fig.
-2) then arises that asexually produces other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+2) then arise that asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
@@
 Extractions were performed according to the manufacturer's instruction.
-All samples were DNase treated (TURBO DNA-free, Invitrogen #AM1907;or on column DNase treatment with Zymo Quick RNA MicroPrep).
+All samples were DNase treated (TURBO DNA-free, Invitrogen #AM1907; or on column DNase treatment with Zymo Quick RNA MicroPrep).
 Libraries were prepared for sequencing using the Illumina TruSeq RNA Sample Prep Kit (Illumina, #FC-122-1001, #FC-122-1002), the Illumina TruSeq Stranded Library Prep Kit (Illumina, #RS-122-2101) or the NEBNext RNA Sample Prep Master Mix Set (NEB, #E6110S).
@@
 1.00 [@Dunn:2013kw; @Guang202416].
-This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575,, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
+This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
 Maximum likelihood analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
@@
 included as a monoecious species, and one without (Fig.
-S2).
+3C, S2).
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
@@
 
-xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="Table 1.
-A complete list of specimens collected for this work.
+xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work.
 New data indicated by Y, blank fields indicate that data are already published.")
@@
 
-Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *P.
@@
 Our ancestral reconstructions for this character (Fig.
-S8) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
+S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
 
@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs.
-4 and S9) [@mapstone2014global; @mackie1987siphonophore].
+4 and S8) [@mapstone2014global; @mackie1987siphonophore].
 The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel].
@@
 This reconstruction (Fig.
-S9) only included species recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
+S8) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
 The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives.
@@
 4).
-We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or swimming (Fig.
-S10), and still find support for a mesopelagic ancestor.
+We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or via a shallow trawl (Fig.
+S9), and still find support for a mesopelagic ancestor.
 It is important to note, however, that *H.

### 2018-01-12 manuscript.rmd
numbered sections, added abstract

@@
 
-## Introduction
+Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface of the ocean, like the familiar Portuguese man of war, to the deep sea.
+Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators in the open ocean.
+Previous work with a few molecular markers provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
+Here, we present a new siphonophore phylogeny, obtained by collecting new transcriptome data from 30 siphonophore species, and analyzing them in combination with 13 publicly available genomic and transcriptomic datasets.
+We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including gain and loss of zooid types, life history traits, and habitat.
+Sequence assembly, annotation, homology evaluation, and supermatrix construction were all conducted using the Agalma pipeline, and the supermatrix was subsequently analysed using Maximum Likelihood and Bayesian approaches.
+The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but finds strong support for new clades within Codonophora that were previously unresolved.
+These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose twice.
+
+## 1. Introduction
 
@@
 
-## Methods
+## 2. Material and methods
 
@@
 
-#### Collecting
+#### 2.1 Collecting
 
@@
 
-#### Sequencing
+#### 2.2 Sequencing
 
@@
 
-#### Analysis
+#### 2.3 Analysis
 
@@
 
-## Results and Discussion
+## 3. Results and Discussion
 
-### Species phylogeny and hypothesis testing
+### 3.1 Species phylogeny and hypothesis testing
 
@@
 This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig.
-4), or the arrangement of male and female zooids along the stem (see below).
+4), or the arrangement of male and female zooids along the stem (see section 3.2 below).
 
-## Character Evolution
+## 3.2 Character Evolution
 ```{r char_evo_setup, include=F, echo=F}
@@
 
-###Evolution of Monoecy
+### 3.3 Evolution of Monoecy
 
@@
 
-###The Evolution of Zooid Types
+### 3.4 The Evolution of Zooid Types
 
@@
 
-###The Gain and Loss of the Pneumatophore
+### 3.5 The Gain and Loss of the Pneumatophore
 
@@
 
-###The Gain and Loss of Tentilla
+### 3.6 The Gain and Loss of Tentilla
 
@@
 
-###The Evolution of Vertical Habitat Use
+### 3.7 The Evolution of Vertical Habitat Use
 
@@
 
-## Conclusions
+## 4. Conclusions
 Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
@@
 We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph.
+We also thank Zack Lewis for help with RNA extractions.
 We also thank the MBARI crews and ROV pilots for collection of the specimens.

### 2018-01-15 manuscript.rmd
updated subsample script

@@
 
-Euphysonectae consists of two reciprocally monophyletic groups we here provisionally refer to as Clade A and Clade B (Fig.
+Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Clade B (Fig.
 3A).
@@
 Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores (both with and without *Rudjakovia* sp.).
+3C): (i) monophyletic Physonectae, (ii) and (iii) monophyletic monoecious siphonophores,with and without *Rudjakovia* sp.
+respectively.
 In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
@@
 The positions of the only two taxa from Clade B in the previous analysies [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
-This difference in conclusions regarding trait evolution therefore does not reflect a contradication between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
+This difference in conclusions regarding trait evolution, therefore, does not reflect a contradication between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
 

### 2018-01-15 agalma/SRA/subsample.sh

@@ -1,45 +1,20 @@
-!/bin/bash
+#!/bin/bash
 set -e
 mkdir -p subsampled
 cd subsampled
-
-BARGMANNIA_IDS=(
-HWI-ST625-51-C02UNACXX-8-CAGATC
-HWI-ST625-51-C02UNACXX-8-CGATGT
-HWI-ST625-51-C02UNACXX-8-GCCAAT
-HWI-ST625-51-C02UNACXX-8-GGCTAC
-HWI-ST625-51-C02UNACXX-8-TAGCTT
-)
-
-BARGMANNIA_FRACTIONS=(
-0.000 0.124 0.249 0.373 0.497
-0.000 0.100 0.201 0.301 0.402
-0.000 0.099 0.197 0.296 0.394
-0.359 0.270 0.180 0.090 0.000
-1.262 0.946 0.631 0.315 0.000
-)
-
-for i in {0..4}
+for id in HWI-ST625-51-C02UNACXX-8-CAGATC HWI-ST625-51-C02UNACXX-8-CGATGT HWI-ST625-51-C02UNACXX-8-GCCAAT HWI-ST625-51-C02UNACXX-8-GGCTAC HWI-ST625-51-C02UNACXX-8-TAGCTT
 do
-	id=${BARGMANNIA_IDS[i]}
-	for j in {0..4}
-	do
-		bl-filter-illumina -i ../filtered/${id}_1.fastq -i ../filtered/${id}_2.fastq -o ${id}_N${j}_1.fastq -o ${id}_N${j}_2.fastq -F ${BARGMANNIA_FRACTIONS[5*i+j]} -S 5850687106356126 -q 0
-	done
+    bl-filter-illumina -i ../filtered/${id}_1.fastq -i ../filtered/${id}_2.fastq -o ${id}_1.fastq -o ${id}_2.fastq -F 0.242 -S 5850687106356126 -q 0
 done
-
 for id in HWI-ST625-51-C02UNACXX-6-ATCACG HWI-ST625-51-C02UNACXX-6-CGATGT HWI-ST625-51-C02UNACXX-6-TGACCA
 do
-	bl-filter-illumina -i ../filtered/${id}_1.fastq -i ../filtered/${id}_2.fastq -o ${id}_1.fastq -o ${id}_2.fastq -F 0.570 -S 7587381619978254 -q 0
[…]

### 2018-01-15 manuscript.rmd
added Phil's changes

@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-Bracts were lost in Hippopodidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
+Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
 These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.
@@
 We reconstruct them as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to Pyrostephidae (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
 Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.
@@
 This suggests they are not just ectodermal swellings, but probably reduced tentilla.
-When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to *Apolemiidae* (Fig.
+When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig.
 S4), and in *Bathyphysa conifera*.
@@
 This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
-There was only a single transition to benthic lifestyle on the branch of *Rhodaliidae*, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
+There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
 Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis, suggesting that depth may be more labile than it appears.

### 2018-01-16 manuscript.rmd
updated figure 1, figure caption

@@
 ![Photographs of living siphonophores.
-They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-F, J), Clade A within Euphysonectae (G-I) and Apolemiidae (K).
+They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-G), Apolemiidae (H), and Clade A within Euphysonectae (I-K).
 (A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
@@
 (F) *Sphaeronectes christiansonae*, scale bar = 2 mm.
-(G) *Nanomia bijuga*, scale bar = 1 cm.
-(H) *Lychnagalma utricularia*, scale bar = 1 cm.
-(I) *Physophora hydrostatica*, scale bar = 5 mm.
-(J) *Praya dubia*, scale bar = 4 cm .
-(K) *Apolemia* sp., scale bar = 1 cm.](figures/Figure1.jpg)
+(G) *Praya dubia*, scale bar = 4 cm .
+(H) *Apolemia* sp., scale bar = 1 cm.
+(I) *Lychnagalma utricularia*, scale bar = 1 cm.
+(J) *Nanomia* sp., scale bar = 1 cm.
+(K) *Physophora hydrostatica*, scale bar = 5 mm.
+Photo credits: S.
+Siebert (C,H,I,K), S.
+Haddock (A,D,E,F), R.
+Sherlock (B), MBARI (G), C.
+Munro (J)](figures/Figure1.jpg)
 

### 2018-01-17 manuscript.rmd
added Bioproject accession
light edits throughout, did not rebuild pdf

@@
 Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators in the open ocean.
-Previous work with a few molecular markers provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
-Here, we present a new siphonophore phylogeny, obtained by collecting new transcriptome data from 30 siphonophore species, and analyzing them in combination with 13 publicly available genomic and transcriptomic datasets.
-We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including gain and loss of zooid types, life history traits, and habitat.
-Sequence assembly, annotation, homology evaluation, and supermatrix construction were all conducted using the Agalma pipeline, and the supermatrix was subsequently analysed using Maximum Likelihood and Bayesian approaches.
+A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
+Here, we present a new siphonophore phylogeny based on new transcriptome data from 30 siphonophore species analyzed in combination with 13 publicly available genomic and transcriptomic datasets.
+We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including sexual system (monoecy vs.
+dioecy), gain and loss of zooid types, life history traits, and habitat.
 The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but finds strong support for new clades within Codonophora that were previously unresolved.
@@
 Most species are planktonic, living in the water column, where some grow to be more than 30 meters in length [@mackie1987siphonophore].
-There is also a small clade of benthic siphonophores, the Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
+There is also a small clade of benthic siphonophores, Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
 There are currently 188 valid described siphonophore species.
@@
 These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
-The Cystonectae was found to be sister to all other siphonophores, while the Calycophorae were nested within "Physonectae".
+Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae".
 The name Codonophora was given to this clade of "Physonectae" and "Calycophorae" [@Dunn:2005dy].
@@
 There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans.
-Within the clade that is sister to the Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig.
+Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig.
 3A).
@@
 3A).
-Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the cnidoband [@totton1965synopsis].
-This clade also shares the prescence of a descending mantle canal within the nectophores (Fig.
-S6), and is also monoecious (Fig.
+The presence of an involucrum, a fold around the base of the cnidoband [@totton1965synopsis], is a potential synapomorphy for Clade A.
+Species of Clade A also have a descending mantle canal within the nectophores (Fig.
+S6), a structure that is also present in some calycophorans.
+Members of Clade A are also monoecious (Fig.
 5).
+There is not a clear synapomorphy for Clade B.
 Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses (Fig.
@@
 
-Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *P.
 hydrostatica* (the presumed sister species to *P.
-gilmeri*) was sister to the Forskaliidae with low support.
+gilmeri*) was sister to Forskaliidae with low support.
 The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
-is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae.
+is in a clade with *Forskalia asymmetrica*, falling outside of Agalmatidae.
 Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of morphological and molecular data [@pugh2006taxonomic; @Dunn:2005dy].
@@
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
-(B) The topologies found in the psoterior distribution of trees that conflict with the ML tree.
+(B) The topologies found in the posterior distribution of trees that conflict with the ML tree.
 (C) The topologies evaluated by the SOWH tests.
@@
 Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) and (iii) monophyletic monoecious siphonophores,with and without *Rudjakovia* sp.
+3C): (i) monophyletic Physonectae, (ii) and (iii) monophyletic monoecious siphonophores, with and without *Rudjakovia* sp.
 respectively.
@@
 Similarly, within Calycophorae, increased taxon sampling is needed.
-This study, and the previous study, suggest that the prayomorphs are paraphyletic, but for slightly different reasons.
+This study, and the previous phylogenetic study [@Dunn:2005dy], suggest that the prayomorphs are paraphyletic, but for slightly different reasons given the different sampling of the analyses.
 In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
@@
 Within each siphonophore species, colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
-Previous analyses suggested that the common ancestor of siphonophores was dioecious, and was consistent with a single gain of monoecy within the Codonophora and no secondary losses [@Dunn:2005dy].
+Previous analyses suggested that the common ancestor of siphonophores was dioecious, and was consistent with a single gain of monoecy within Codonophora and no secondary losses [@Dunn:2005dy].
 The better-resolved tree in the current analyses indicates that the evolution of monoecy is more complicated than this.
@@
 Our ancestral reconstructions for this character (Fig.
-S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the Agalmatidae stem and once along the Pyrostephidae stem.
+S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the stem of Agalmatidae and once along the stem of Pyrostephidae.
 
@@
 Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis, suggesting that depth may be more labile than it appears.
-Phylogenetic conservatism is also observed for the mesopelagic Apolemiidae and Pyrostephidae clades, but not for the Calycophorae, which show a diversified evolutionary trajectory for depth.
+Phylogenetic conservatism is also observed for the mesopelagic Apolemiidae and Pyrostephidae clades, but not for Calycophorae, which show a diversified evolutionary trajectory for depth.
 The present sampling, however, is not sufficient to capture significant variation in depth distributions between closely related species.

### 2018-01-17 manuscript.rmd
light edits, chelophyes carre2000, knit

@@
 Summary statistics for each library are given in the file Supplementary data 2.
-All sequence data have been deposited in the NCBI sequence read archive (SRA) with accession number XXX.
+All sequence data have been deposited in the NCBI sequence read archive (SRA) with Bioproject accession number PRJNA255132.
 
@@
 1.00 [@Dunn:2013kw; @Guang202416].
-This workflow integrates a variety of existing tools [@grabherr2011full, @altschul1990basic, @doi:10.1093/nar/30.7.1575, @sukumaran2010dendropy, @langmead2012fast, @li2011rsem, @talavera2007improvement, @li2009sequence, @katoh2013mafft] and new methods.
+This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods.
 Maximum likelihood analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
@@
 The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0127).
-Only the results from the CAT-Poisson model are presented here.Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
+Only the results from the CAT-Poisson model are presented here.
+Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
 This left `r length( phylobayes_trees )` trees in the posterior.
@@
 
-![(A) Maximum likelihood (ML) phylogram with bipartation frequencies from the ML bootstraps and the Bayesian posterior distribution of trees.
+![(A) Maximum likelihood (ML) phylogram with bipartition frequencies from the ML bootstraps and the Bayesian posterior distribution of trees.
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
@@
 The better-resolved tree in the current analyses indicates that the evolution of monoecy is more complicated than this.
-The two clades of monophyloetic siphonophores, Calycophorae and Clade A (Fig.
+The two clades of monophyletic siphonophores, Calycophorae and Clade A (Fig.
 3A), do not form a monophyletic group.
-This is because Clade B, which contains dioecious species, is also desceneded from their most recent common ancestor.
+This is because Clade B, which contains dioecious species, is also descended from their most recent common ancestor.
 SOWH tests strongly reject the placement of the monecious clades Calycophorae and Clade A as a group that excludes Clade B (Figs.
 3C and S2).
-The positions of the only two taxa from Clade B in the previous analysies [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
+The positions of the only two taxa from Clade B in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
 This difference in conclusions regarding trait evolution, therefore, does not reflect a contradication between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
@@
 The distribution of monoecy is consistent with two scenarios.
-In the first, there is a single shift from dioecy to monoecy along the branch that give rise to the most recont common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
+In the first, there is a single shift from dioecy to monoecy along the branch that give rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
 In the second, monoecy arose twice -- once along the branch that gave rise to Clade A and again along the branch that gave rise to Calycophorae.
@@
 However, there are exceptions -- in *Sulculeolaria* (not included in this phylogeny), colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+Environmental influences may play a role in determining the expressed sex.
+Colonies of the calycophoran *Chelophyes appendiculata* collected in the field always bear both male and female gonophores whereas when kept in culture only gonophores of one sex are maintained [@Carre:2000hb].
+This suggests a high plasticity of the sexual state in some taxa and underlines the need for caution when evaluation the state of this character in rarely collected species.
 
@@
 conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we assume this is a case of secondary loss.
-When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support an common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
+When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support a common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
 S3).

### 2018-01-18 manuscript.rmd
added Felipe's comments, updated conclusion

@@
 They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-G), Apolemiidae (H), and Clade A within Euphysonectae (I-K).
-(A) *Rhisophysa eysenhardtii*, scale bar = 1 cm.
+(A) *Rhizophysa eysenhardtii*, scale bar = 1 cm.
 (B) *Bathyphysa conifera*, scale bar = 2cm.
@@
 Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water SCUBA dives.
-*Chelophyes appendiculata* and *Hippopodius hippopus* specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
+*Chelophyes appendiculata* and *Hippopodius hippopus* (Fig.
+1C) specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
@@
 
-Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (Fig.
+1I) (both not included in the previous phylogeny) are sister to Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *P.
@@
 This study, and the previous phylogenetic study [@Dunn:2005dy], suggest that the prayomorphs are paraphyletic, but for slightly different reasons given the different sampling of the analyses.
-In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia*, *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
+In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia* (Fig.
+1G), *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
 Expanded taxon sampling, particularly *P.
@@
 The better-resolved tree in the current analyses indicates that the evolution of monoecy is more complicated than this.
-The two clades of monophyletic siphonophores, Calycophorae and Clade A (Fig.
+The two clades of monoecious siphonophores, Calycophorae and Clade A (Fig.
 3A), do not form a monophyletic group.
@@
 
-The distribution of monoecy is consistent with two scenarios.
+The distribution of monoecy is consistent with two scenarios (Fig.
+4).
 In the first, there is a single shift from dioecy to monoecy along the branch that give rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
@@
 This is consistent with differences in the arrangements of male and female gonophores in the two clades.
-In Clade A, male and female zooids are found within the same cormidium.
+In Clade A, male and female zooids are found within the same cormidium (a single reiterated sequence of zooids along the stem, see Fig.
+2).
 In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
-Meanwhile in calycophorans, each cormidium (a single reiterated sequence of zooids along the stem, see Fig.
-2) bears either male or female zooids.
+Meanwhile in calycophorans, each cormidium bears either male or female zooids.
 In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row.
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-However, there are exceptions -- in *Sulculeolaria* (not included in this phylogeny), colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
-Environmental influences may play a role in determining the expressed sex.
+In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
+
+There are exceptions, in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+Environmental influences may also play a role in determining the expressed sex.
 Colonies of the calycophoran *Chelophyes appendiculata* collected in the field always bear both male and female gonophores whereas when kept in culture only gonophores of one sex are maintained [@Carre:2000hb].
-This suggests a high plasticity of the sexual state in some taxa and underlines the need for caution when evaluation the state of this character in rarely collected species.
+This suggests a high plasticity of the sexual state in some calycophoran taxa and underlines the need for caution when evaluating the state of this character in rarely collected species.
 
@@
 2B), which has its own growth zone, and are used for coordinated colony-level swimming.
+Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
+1A) instead move through the water column using repeated contraction and relaxation of the stem, and use modified flattened gastrozooids called ptera to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the stem of Agalmatidae and once along the stem of Pyrostephidae.
+The functional role of dorsal vs.
+ventral attachment is not clear.
 
@@
 Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
-Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica*, and in *Gymnopraia lapislazula*.
-These patterns of loss are not captured by the present sampling, as most of these species are not included in the present phylogeny.
+Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica* (Fig.
+1K), and in *Gymnopraia lapislazula*.
+These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny.
+In species without bracts, a diversity of zooids appear to fulfil the roles of neutral buoyancy and protection.
+In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
+While *Hippopodius hippopus* have up to twelve nectophores and retract the stem within the nectophores -- the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
+1C shows the blanching of nectophores)[@bassot1978bioluminescence].
 
-Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+Palpons are modified reduced gastrozooids (specialized feeding polyps) used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
 We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
@@
 
-The most complex nematocyst batteries of Cnidaria are found among the siphonophores.
+Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture).
+As in other cnidarians, nematocyst stinging capsules, arranged in dense nematocyst batteries, play a critical role in prey capture.
 In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig.
@@
 It is still an open question whether the common ancestor of Siphonophora had tentilla.
-The only siphonophores species regarded as lacking tentilla are *Physalia physalis*, *Apolemia* spp., and *Bathyphysa conifera*.
+The only siphonophores species regarded as lacking tentilla are *Physalia physalis*, *Apolemia* spp.
+(Fig.
+1H), and *Bathyphysa conifera* (Fig.
+1B).
 Since *B.
@@
 S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
-This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic (light blue) and bathypelagic (dark blue) waters.
+This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters.
 There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
-Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis, suggesting that depth may be more labile than it appears.
-Phylogenetic conservatism is also observed for the mesopelagic Apolemiidae and Pyrostephidae clades, but not for Calycophorae, which show a diversified evolutionary trajectory for depth.
-The present sampling, however, is not sufficient to capture significant variation in depth distributions between closely related species.
+There is a strong signal of "phylogenetic depth structure", whereby clades are largely restricted to particular depth ranges, with the exception of Calycophorae which have diversified across the water column (Figs.
+S8, S9).
+However, the present sampling is not sufficient to capture significant variation in depth distributions between closely related species.
 Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical].
+This suggests that vertical habitat use is more labile than it appears and may be an important mechanism in siphonophore ecology.
 
@@
 ## 4. Conclusions
+
 Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
-A previous rRNA study was unable to resolve deep relationships within Codonophora, while our study suggests that Pyrostephidae are sister to all other non-apolemiid codonophorans.
-We find support for two monophyletic groups within the clade that is sister to Pyrostephidae -- Calycophorae and a new clade we name Euphysonectae.
-Within Euphysonectae, we find support for two monophyletic groups dubbed Clade A and Clade B.
-Our study supports the rejection of the monophyly of Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores.
-These results are consistent with a hypothesis that monoecy evolved twice, once in Clade A and once in Calycophorae.
-This hypothesis is consistent with differences in the arrangement of male and female zooids in the two monoecious groups.
-Future studies with the addition of key taxa within Euphysonectae will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution, including the evolution of zooid types, that cannot be fully assessed here.
+Using 33 siphonophore species, sampled across the phylogeny, we were able to gain more resolution of the major clades within Siphonophora.
+We suggest that monoecy likely evolved twice and siphonophores evolved very different ways to be monoecious, however we are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora.
+This phylogeny suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae and Calycophorae to further resolve the internal relationships within these clades.
+The vast majority of valid described siphonophore species likely fall within these two groups (172 species), and a large proportion of known siphonophore trait diversity is found within these two groups.
 

### 2018-01-18 manuscript.rmd
added hook sentence at the end of tentilla section
added comments

@@
 There is also a small clade of benthic siphonophores, Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
-There are currently 188 valid described siphonophore species.
+There are currently 187 valid described siphonophore species.
 
@@
 Probuds subdivide to give rise to zooids in repeating-units (cormidia).
-The gastrozooid is the posterior-most zooid within each cormidium.](figures/Figure2.pdf)
+The gastrozooid (specialized feeding polyp) is the posterior-most zooid within each cormidium.](figures/Figure2.pdf)
 
@@
 In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
-Meanwhile in calycophorans, each cormidium bears either male or female zooids.
+Meanwhile in calycophorans, each cormidium bears either male or female gonophores.
 In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row.
@@
 Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
-1A) instead move through the water column using repeated contraction and relaxation of the stem, and use modified flattened gastrozooids called ptera to increase surface area and prevent colony sinking [@biggs1976siphonophore].
+1A) instead move through the water column using repeated contraction and relaxation of the stem, and use modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the stem of Agalmatidae and once along the stem of Pyrostephidae.
-The functional role of dorsal vs.
+The functional implication of dorsal vs.
 ventral attachment is not clear.
@@
 
-Palpons are modified reduced gastrozooids (specialized feeding polyps) used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
+Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
 We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
@@
 There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
-There is a strong signal of "phylogenetic depth structure", whereby clades are largely restricted to particular depth ranges, with the exception of Calycophorae which have diversified across the water column (Figs.
+There is evidence that habitat depth is conserved in within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
 S8, S9).
-However, the present sampling is not sufficient to capture significant variation in depth distributions between closely related species.
+Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis.
+The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species.
 Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical].
@@
 This phylogeny suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae and Calycophorae to further resolve the internal relationships within these clades.
-The vast majority of valid described siphonophore species likely fall within these two groups (172 species), and a large proportion of known siphonophore trait diversity is found within these two groups.
+The vast majority of valid described siphonophore species likely fall within these two groups, with 109 species in Calycophorae and 62 species in Euphysonectae.
 

### 2018-01-18 manuscript.rmd
modified conclusions, legends

@@
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
+The numbers of valid described species estimated to be based in each clade based on taxonomy are shown below beach clade name on the right.
 (B) The topologies found in the posterior distribution of trees that conflict with the ML tree.
@@
 
-![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.](figures/Figure4.pdf)
+![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
+Bottom: siphonophore phylogeny, colored by clade.
+Middle panel: diagram showing the presence/absence of traits across Siphonophora, with the physical location of the trait shown on a schematic of *Nanomia bijuga* (schematic by Freya Goetz).
+Top: Bathymetric distribution of siphonophore species.
+Physalia illustration by Noah Schlottman, taken from http://phylopic.org/](figures/Figure4.pdf)
 
@@
 S4), and in *Bathyphysa conifera*.
+The application of phylogenetic methods to the evolution of tentillum morphology would be a crucial step towards understanding the evolution of these structures, and their relationship with the feeding ecology of siphonophores.
 
@@
 Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
-Using 33 siphonophore species, sampled across the phylogeny, we were able to gain more resolution of the major clades within Siphonophora.
-We suggest that monoecy likely evolved twice and siphonophores evolved very different ways to be monoecious, however we are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora.
-This phylogeny suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae and Calycophorae to further resolve the internal relationships within these clades.
-The vast majority of valid described siphonophore species likely fall within these two groups, with 109 species in Calycophorae and 62 species in Euphysonectae.
+Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae.
+Our results suggest that monoecy arose twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.
+We are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora, which will require greater taxon sampling and improved morphological understanding of many poorly known species.
+The improved resolution presented in this study suggests that an important next step in understanding siphonophore evolution will be targeting molecular sampling within Euphysonectae (where we sampled 13 of 62 valid described species that likely belong to the group) and Calycophorae (where we sampled 9 species in a clade of 109 valid described species) to further resolve the internal relationships within these clades.
 

### 2018-01-18 manuscript.rmd
added Mark's comments

@@
 
-Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4^, Alejandro Damian Serrano^1,9^, Samuel H Church^1,5^, Freya Goetz^1,6^, Phil Pugh^7^, Steven H.D.
-Haddock^8^, Casey W.
-Dunn^1,9^**
+Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,6^, Alejandro Damian Serrano^1,10^, Samuel H Church^1,6^, Freya Goetz^1,7^, Phil Pugh^8^, Steven H.D.
+Haddock^9^, Casey W.
+Dunn^1,10^**
 
@@
 
-^5^ Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, MA 02138, USA
+^5^ Current address: Watson Institute for International and Public Affairs, Brown University, Providence, RI 02912, USA
 
-^6^ Current address: Smithsonian Institution, National Museum of Natural History, Washington, DC 20560, USA
+^6^ Current address: Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, MA 02138, USA
 
-^7^ National Oceanography Centre, Southampton, SO14 3ZH, UK
+^7^ Current address: Smithsonian Institution, National Museum of Natural History, Washington, DC 20560, USA
 
-^8^ Monterey Bay Aquarium Research Institute, Moss Landing, CA 95039, USA
+^8^ National Oceanography Centre, Southampton, SO14 3ZH, UK
 
-^9^ Current address: Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA
+^9^ Monterey Bay Aquarium Research Institute, Moss Landing, CA 95039, USA
+
+^10^ Current address: Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA
 
@@
 Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.
-1.00 [@Dunn:2013kw; @Guang202416].
+1.0.0 [@Dunn:2013kw; @Guang202416].
 This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods.

### 2018-01-19 manuscript.rmd
edited affilations, names

@@
 
-Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,6^, Alejandro Damian Serrano^1,10^, Samuel H Church^1,6^, Freya Goetz^1,7^, Phil Pugh^8^, Steven H.D.
+Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
+Church^1,6^, Freya Goetz^1,7^, Philip R.
+Pugh^8^, Steven H.D.
 Haddock^9^, Casey W.

### 2018-01-19 manuscript.rmd
light edits abstract

@@
 
-^2^ Current address: Department of Molecular & Cellular Biology, University of California at Davis, Davis, CA 95616, USA
+^2^ Current address: Department of Molecular & Cellular Biology, University of California Davis, Davis, CA 95616, USA
 
@@
 
-Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface of the ocean, like the familiar Portuguese man of war, to the deep sea.
-Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators in the open ocean.
+Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea.
+Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators.
 A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
@@
 dioecy), gain and loss of zooid types, life history traits, and habitat.
-The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but finds strong support for new clades within Codonophora that were previously unresolved.
+The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but we find strong support for new clades within Codonophora that were previously unresolved.
 These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose twice.

### 2018-01-19 manuscript.rmd
fixed typos

--- 2018-01-19 manuscript.rmd
@@
 The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
-A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa advanced several long standing questions about siphonophore biology.
-These include the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
+A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology.
+Including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
 Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae".
-The name Codonophora was given to this clade of "Physonectae" and "Calycophorae" [@Dunn:2005dy].
+The name Codonophora was given to this clade of "Physonectae" and Calycophorae [@Dunn:2005dy].
 
@@
 
-Specimens were collected in the eastern Pacific Ocean, Mediterranean, and the Gulf of California.
-Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods, and sequencing details are summarized in the file Supplementary data 1.
+Specimens were collected in the northeastern Pacific Ocean, Mediterranean, and the Gulf of California.
+Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods, and sequencing details are summarized in the file Supplementary data 1 (also found in the git repository).
 Monterey Bay and Gulf of California specimens were collected by remotely operated underwater vehicle (ROV) or during blue-water SCUBA dives.
@@
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
-The numbers of valid described species estimated to be based in each clade based on taxonomy are shown below beach clade name on the right.
+The numbers of valid described species estimated to be based in each clade based on taxonomy are shown below each clade name on the right.
 (B) The topologies found in the posterior distribution of trees that conflict with the ML tree.
@@
 The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae.
-This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focussed.
+This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focused.
 Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig.

### 2018-01-19 manuscript.rmd
minor edits

@@
 Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
-Church^1,6^, Freya Goetz^1,7^, Philip R.
+Church^1,6^, Freya E.
+Goetz^1,7^, Philip R.
 Pugh^8^, Steven H.D.
@@
 xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work.
-New data indicated by Y, blank fields indicate that data are already published.")
+New data indicated by Y, blank fields indicate that data were already published.")
 print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.75')
@@
 In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
-While *Hippopodius hippopus* have up to twelve nectophores and retract the stem within the nectophores -- the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
+While *Hippopodius hippopus* have up to twelve nectophores and can retract the stem within the nectophores -- the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
 1C shows the blanching of nectophores)[@bassot1978bioluminescence].
@@
 There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
-There is evidence that habitat depth is conserved in within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
+There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
 S8, S9).

### 2018-01-19 manuscript.rmd
switched places of citations in Physalia tentilla paragraph

@@
 How we define absence of tentilla, especially for *Physalia physalis*, is also important.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@BARDI2007].
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@Hessinger1988].
 Siphonophore tentilla are complete diverticular branchings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).
-*Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@Hessinger1988].
+*Physalia*'s buttons enclose individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm [@BARDI2007].
 This suggests they are not just ectodermal swellings, but probably reduced tentilla.

### 2018-01-19 manuscript.rmd
adjusted affiliations

--- 2018-01-19 manuscript.rmd
@@
 
-Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
-Church^1,6^, Freya Goetz^1,7^, Philip R.
+Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
+Church^1,6^, Freya E.
+Goetz^1,7^, Philip R.
 Pugh^8^, Steven H.D.
 Haddock^9^, Casey W.
-Dunn^1,10^**
+Dunn^10^**
 
@@
 
-^3^ Current address: Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA
+^3^ Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA
 
@@
 
-^10^ Current address: Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA
+^10^ Department of Ecology and Evolutionary Biology, Yale University, New Haven, CT 06520, USA
 

### 2018-01-19 manuscript.rmd
fixed

--- 2018-01-19 manuscript.rmd
@@
 
-Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^1,3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
+Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
 Church^1,6^, Freya E.
@@
 Haddock^9^, Casey W.
-Dunn^1,10^**
+Dunn^10^**
 

### 2018-01-19 manuscript.rmd
added note about on depth and sexual mode

@@
 
-Catriona Munro^1^\*, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
+Catriona Munro^1^\*^,^**, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
 Church^1,6^, Freya E.
@@
 Haddock^9^, Casey W.
-Dunn^10^**
+Dunn^10^
 
@@
 
-\*\* Corresponding author, casey.dunn@yale.edu
+\*\* Corresponding author, catriona_munro@brown.edu
 
@@
 In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
+Both Calycophorae and Clade A both have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode.
+Similar independent transitions from gonochorism (separate sex) to hermaphrodism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated].
+To test this hypothesis, a more extensive taxon sampling of the Calycophorae is needed.
 
-There are exceptions, in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+Within Calycophorae there are some variations in sexual mode -- in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
 Environmental influences may also play a role in determining the expressed sex.
@@
 How we define absence of tentilla, especially for *Physalia physalis*, is also important.
-The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain on their ectoderm all active and functionally arranged nematocysts used by the organism for prey capture [@Totton1960;@Hessinger1988].
+The tentacles of this species, when uncoiled, show very prominent, evenly spaced, bulging buttons which contain in the ectoderm all functional nematocytes (carrying mature nematocysts) used by the organism for prey capture [@Totton1960;@Hessinger1988].
 Siphonophore tentilla are complete diverticular branchings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm).

### 2018-01-20 manuscript.rmd
fixed typos, formatting

@@
 Siphonophores remain poorly known, in large part because they are fragile and difficult to collect.
-They have, however, been of great interest for more than 150 years in part due to their unique structure and development [@mapstone2014global;@mackie1987siphonophore].
+They have, however, been of great interest for more than 150 years due to their unique structure and development [@mapstone2014global;@mackie1987siphonophore].
 Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction.
@@
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from other colonial animals in colony structure, development, and the high degree to which their zooids are functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character].
+Siphonophores differ significantly from other colonial animals in terms of colony structure and development -- their zooids are highly functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character].
 The functions that zooids are specialized for include feeding, reproducing, or swimming (Fig.
@@
 (B) Inset of the pneumatophore and nectosomal growth zone.
-A series of buds gives rise to nectophores.
+A series of buds give rise to nectophores.
 (C) Inset of the siphosomal growth zone.
@@
 ## 3.2 Character Evolution
-```{r char_evo_setup, include=F, echo=F}
+```{r char_evo_setup, include=F, echo=FALSE}
 #Load data
@@
 
-```{r simmap_making, echo=F, message=F, warning=F, include=F}
+```{r simmap_making, echo=FALSE, message=FALSE, warning=FALSE, include=F}
 #Stochastic trait mapping:
@@
 
-```{r sex, echo=F, message=F, warning=F, include=F}
+```{r sex, echo=FALSE, message=F, warning=FALSE, include=F}
 par(ask=F)
@@
 
-```{r palpons, echo=F, message=F, warning=F, include=F}
+```{r palpons, echo=FALSE, message=F, warning=FALSE, include=F}
 #Simmap PALPONS
@@
 
-```{r vars_plot, echo=F, message=F, warning=F, include=F}
+```{r vars_plot, echo=FALSE, message=F, warning=FALSE, include=F}
 #Get Raw Data
@@
 
-```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, include=F}
+```{r ggtree_bin_heatmap, echo=FALSE, message=F, warning=FALSE, include=F}
 p <- ggtree(ultratree)
@@
 In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
-Both Calycophorae and Clade A both have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode.
+Both Calycophorae and Clade A have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode.
 Similar independent transitions from gonochorism (separate sex) to hermaphrodism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated].
@@
 
-![Stochastic mapping reconstruction of the evolutionary history of A) distribution of sexual zooids on the colonies, and B) palpon zooids.
+![Stochastic mapping reconstruction of the evolutionary history of A) sexual mode, whether a colony is monoecious or dioecious and B) presence/absence of palpons (modified reduced gastrozooids).
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
@@
 
+\pagebreak
+
+####SOWH analysis
+
 ![Constrained topologies specified in SOWH testing.
@@
 
-```{r summary_table, results='asis', echo=FALSE, comment=NA, warning=FALSE}
-
-####The supplementary table is too long. We decided to include as a supplementary datafile instead
-
-#	summary_stats = read.table("supplementary_info/Library_summary.csv", header=TRUE, sep=",")
-#	names(summary_stats)<-c("Species","Mean insert size (bp)","Insert size sd","# read pairs","% kept after rRNA removal","% kept after assembly","Adapter fails","Quality fails","Base Comp.
-fails","Total transcripts","Coding transcripts")
-
-#xtable_summary<-xtable(summary_stats, caption="Supplementary table 1.The assembly summary statistics for all included taxa, showing mean insert size, and the percentage of reads kept after pipeline steps.")
-#align( xtable_summary ) <- c('l|','p{0.8in}','p{0.3in}','p{0.3in}','p{0.5in}','p{0.3in}','p{0.3in}','p{0.4in}','p{0.5in}','p{0.5in}','p{0.5in}','p{0.5in}')
-#print(xtable_summary,include.rownames=FALSE,comment=F, floating =FALSE, tabular.environment = "longtable")
-
-```
+\pagebreak
 
 #####Stochastic Character maps
-```{r tentilla, echo=F, message=F,warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla."}
+
+```{r tentilla, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla."}
 #Simmap TENTILLA
@@
 
-```{r tentilla2, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla."}
+```{r tentilla2, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla."}
 #Simmap TENTILLA
@@
 
-```{r nectosome, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
+```{r nectosome, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
 #Simmap NECTOSOME
@@
 
-```{r desc_mantle, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores.
+```{r desc_mantle, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores.
 Cystonects and Athorybia were excluded as they do not have a nectosome.
@@
 
-```{r nectosome_pos, echo=F, message=F, warning=F,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.
+```{r nectosome_pos, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.
 Cystonects were excluded as they do not have a nectosome."}
@@
 
-```{r median_cont, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
+```{r median_cont, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
 depth_value = depth_median
@@
 
-```{r median_depth_extended, echo=F, message=F, warning=F, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations." }
+```{r median_depth_extended, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations." }
 
@@
 
-fitContinuous(ultratree, median_extended, model="white")$opt$aicc
-fitContinuous(ultratree, median_extended, model="BM")$opt$aicc
-fitContinuous(ultratree, median_extended, model="OU")$opt$aicc
+#below are the model fit details
+
+modelfit_white <- fitContinuous(ultratree, median_extended, model="white")$opt$aicc
+modelfit_BM <- fitContinuous(ultratree, median_extended, model="BM")$opt$aicc
+modelfit_OU <- fitContinuous(ultratree, median_extended, model="OU")$opt$aicc
 
@@
 
-```{r phylosig_binaries, results='asis',echo=F, message=F, warning=F}
- # make it xtable
-binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5:8)]), ultratree)
-names(binaries_phylosig)<-c("K","PICvar obs","PICvar rnd","P-value","Z-score")
-xtable_PS<-xtable(binaries_phylosig,auto=TRUE, caption="Phylogenetic signal in the binary traits, including Blomberg's K statistic, the mean observed PIC variance, the random variance of PICs, p-value of the comparison of observed and random variance, and the z-score.
-")
-print(xtable_PS,comment=F)
-
-```
+\pagebreak
 

### 2018-01-20 manuscript.rmd
spelling errors

@@
 
-Specimens were collected in the northeastern Pacific Ocean, Mediterranean, and the Gulf of California.
+Specimens were collected in the north-eastern Pacific Ocean, Mediterranean, and the Gulf of California.
 Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods, and sequencing details are summarized in the file Supplementary data 1 (also found in the git repository).
@@
 The positions of the only two taxa from Clade B in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
-This difference in conclusions regarding trait evolution, therefore, does not reflect a contradication between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
+This difference in conclusions regarding trait evolution, therefore, does not reflect a contradiction between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
 
@@
 Both Calycophorae and Clade A have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode.
-Similar independent transitions from gonochorism (separate sex) to hermaphrodism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated].
+Similar independent transitions from gonochorism (separate sex) to hermaphroditism (both sexes in the same individual) have been identified in shallow-water scleractinian corals (Anthozoa, Cnidaria) [@kerr2011correlated].
 To test this hypothesis, a more extensive taxon sampling of the Calycophorae is needed.
 
-Within Calycophorae there are some variations in sexual mode -- in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protoandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+Within Calycophorae there are some variations in sexual mode -- in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protandrous, with female gonophores developing after the release of male gonophores [@carre1979].
 Environmental influences may also play a role in determining the expressed sex.
@@
 S8) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
-The ROV observations are reliable below 200m, and no quantative measurements were made on SCUBA dives.
+The ROV observations are reliable below 200m, and no quantitative measurements were made on SCUBA dives.
 Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.

### 2018-01-23 manuscript.rmd
changed Rudjakovia Sex and DMC to NA in data, and adjusted code to render them as Dioecious and 0

@@
 names(sex) = rownames(cdata)
+sex[which(names(sex)=="Rudjakovia sp")] <- "Dioecious" #Modified for graphing purposes
 
@@
 names(desc_mantle)=rownames(desc_mantle_data)
+desc_mantle[which(names(desc_mantle)=="Rudjakovia sp")] <- 0 #Modified for graphing purposes
 DescMantleTree =  drop.tip(ultratree, which(!(ultratree$tip.label %in% names(desc_mantle))))

### 2018-01-26 manuscript.rmd
added details about Rudjakovia & SRA numbers

@@
 The phylogenetic relationships in this study are largely consistent with the previous phylogeny, but we find strong support for new clades within Codonophora that were previously unresolved.
-These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose twice.
+These results have important implications for trait evolution within Siphonophora, including favoring the hypothesis that monoecy arose at least twice.
 
@@
 
+specimen_data_reduced<-as.data.frame(specimen_data_reduced)
 xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work.
 New data indicated by Y, blank fields indicate that data were already published.")
-print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.75')
+print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.7')
 
@@
 In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
+The sexual system and cormidial arrangement of *Rudjakovia* is undescribed, although preliminary observations suggest that this species may be monoecious and that monoecy arose a third time in the Pyrostephidae.
+A detailed redescription of *Rudjakovia* would help clarify this.
+
 Both Calycophorae and Clade A have a large proportion of shallow water species (see section 3.7), suggesting that there may be an association between habitat depth and sexual mode.
@@
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
-Intermediate values reflect uncertainty.](figures/Figure5.pdf)
+Intermediate values reflect uncertainty.
+The grey dashed branch leading to *Rudjakovia* sp.
+indicates that the sexual mode of this species is unknown.](figures/Figure5.pdf)
 
@@
 Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae.
-Our results suggest that monoecy arose twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.
+Our results suggest that monoecy arose at least twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.
 We are unable to fully capture some of the complex patterns of zooid gain and loss within Codonophora, which will require greater taxon sampling and improved morphological understanding of many poorly known species.

### 2018-01-26 manuscript.rmd
oriented/orientated debate

@@
 output:
+  word_document: default
   pdf_document:
@@
 
-![Schematic of the siphonophore *Nanomia bijuga*, orientated with the anterior of the colony at the top of the page, and the ventral side to the left.
+![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top of the page, and the ventral side to the left.
 Adapted from http://commons.wikimedia.org/wiki/File:Nanomia_bijuga_whole_animal_and_growth_zones.svg, drawn by Freya Goetz.

### 2018-01-26 manuscript.rmd
Edits throughout the text.

@@
 
-Catriona Munro^1^\*^,^**, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
+Catriona Munro^1^\*^,^‡, Stefan Siebert^1,2^\*, Felipe Zapata^3^, Mark Howison^4,5^, Alejandro Damian Serrano^1,10^, Samuel H.
 Church^1,6^, Freya E.
@@
 
-\*\* Corresponding author, catriona_munro@brown.edu
+\‡ Corresponding author, catriona_munro@brown.edu
 
@@
 Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea.
-Siphonophores play an important role in ocean ecosystems, and are among the most abundant gelatinous predators.
+Siphonophores play important roles in ocean ecosystems, and are among the most abundant gelatinous predators.
 A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
@@
 One or two growth zones (Fig.
-2) then arise that asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+2) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
-Siphonophores differ significantly from other colonial animals in terms of colony structure and development -- their zooids are highly functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character].
-The functions that zooids are specialized for include feeding, reproducing, or swimming (Fig.
+Siphonophores differ significantly from other colonial animals in their colony structure and development -- their zooids are highly functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character].
+Zooids are specialized for a range of functions, including feeding, reproduction, or swimming (Fig.
 2) [@dunn2006evolution].
@@
 ![Photographs of living siphonophores.
-They correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-G), Apolemiidae (H), and Clade A within Euphysonectae (I-K).
+Colored circles correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-G), Apolemiidae (H), and Clade A within Euphysonectae (I-K).
 (A) *Rhizophysa eysenhardtii*, scale bar = 1 cm.
@@
 Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group.
-The relationship of siphonophores to other hydrozoans has been difficult to resolve [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
-A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology.
-Including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
+The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
+A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
 Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae".
@@
 Major questions remained after this early work, though.
-There was, in particular, little support for important deep relationships within Codonophora.
+In particular, there was little support for important deep relationships within Codonophora.
 Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig.
@@
 (C) Inset of the siphosomal growth zone.
-Probuds subdivide to give rise to zooids in repeating-units (cormidia).
+Probuds subdivide to give rise to zooids in repeating units (cormidia).
 The gastrozooid (specialized feeding polyp) is the posterior-most zooid within each cormidium.](figures/Figure2.pdf)
@@
 All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017).
-The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")`.
+The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag XXXXXXXXXX.
 
@@
 When possible, specimens were starved overnight in filtered seawater at temperatures close to ambient water temperatures at the time of specimen collection.
-mRNA was extracted directly from tissue using a variety of methods (Supplementary data 1): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
+Subsequently, mRNA was extracted directly from tissue using a variety of methods (Supplementary data 1): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
 In case of anticipated very small total RNA quantities, only a single round of bead purification was performed.
@@
 is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
-included as a monoecious species, and one without (Fig.
+included as a monoecious species, and one with it considered dioecious (Fig.
 3C, S2).
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
-version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel with the default options and an initial sample size of 100 (analysis code can be found in the git repository).
+version 0.39 (commit fd68ef57) to carry out the SOWH tests in parallel, using the default options and an initial sample size of 100 (analysis code can be found in the git repository).
 For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
@@
 
-Most relationships received strong support across analysis methods (Fig.
+The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig.
 3A), with a couple of localized exceptions (Fig.
 3B).
-The phylogenetic relationships recovered in this study are largely consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
-Relationships that receive strong support in both studies include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
+Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
+Relationships that receive strong support in both include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
 Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses.
@@
 is in a clade with *Forskalia asymmetrica*, falling outside of Agalmatidae.
-Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of morphological and molecular data [@pugh2006taxonomic; @Dunn:2005dy].
+Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses of molecular and morphological data [@pugh2006taxonomic; @Dunn:2005dy].
 
-Within Calycophorae, taxon sampling is shallower here than in the previous study.
+Within Calycophorae, taxon sampling is less comprehensive here than in the previous study.
 The calycophoran relationships that can be investigated, however, are in broad agreement with the previous analysis.
@@
 Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) and (iii) monophyletic monoecious siphonophores, with and without *Rudjakovia* sp.
-respectively.
+3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores including *Rudjakovia* sp.
+and (iii) monophyletic monoecious with dioecious *Rudjakovia* sp.
+<!--  XXX This needs to be fixed to coincide with the two hypotheses listed in section 2.3 XXX -->.
 In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
@@
 
-The broad sampling approach of this phylogeny provides new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae.
+The broad taxon sampling and more extensive gene sampling of this phylogeny provide new evidence for the relationships between major siphonophore clades within Codonophora, specifically between Pyrostephidae, Calycophorae, and the newly named Euphysonectae.
 This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focused.
-Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, however there is weaker support for Clade B (Fig.
+Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, although there is weaker support for Clade B (Fig.
 3A, 3B).
@@
 1G), *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
-Expanded taxon sampling, particularly *P.
+Expanded transcriptome sequencing, particularly *P.
 dubia* or a nectopyramid, but also extensive sampling across the major prayomorph and diphyomorph groups, will expand our understanding of relationships within Calycophorae.
@@
 Within each siphonophore species, colonies are either monoecious (male and female gonophores are on the same colony) or dioecious (male and female gonophores are on different colonies).
-Previous analyses suggested that the common ancestor of siphonophores was dioecious, and was consistent with a single gain of monoecy within Codonophora and no secondary losses [@Dunn:2005dy].
+Previous analysis suggested that the common ancestor of siphonophores was dioecious, and was consistent with a single gain of monoecy within Codonophora and no secondary losses [@Dunn:2005dy].
 The better-resolved tree in the current analyses indicates that the evolution of monoecy is more complicated than this.
@@
 3C and S2).
-The positions of the only two taxa from Clade B in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that previous study.
+The positions of the only two taxa from Clade B that were included in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that study.
 This difference in conclusions regarding trait evolution, therefore, does not reflect a contradiction between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
 
-The distribution of monoecy is consistent with two scenarios (Fig.
+The distribution of monoecy is consistent with two potential scenarios (Fig.
 4).
-In the first, there is a single shift from dioecy to monoecy along the branch that give rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
-In the second, monoecy arose twice -- once along the branch that gave rise to Clade A and again along the branch that gave rise to Calycophorae.
+In the first, there is a single shift from dioecy to monoecy along the branch that gave rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
+In the second, monoecy arose twice: once along the branch that gave rise to Clade A and again along the branch that gave rise to Calycophorae.
 
@@
 2).
-In these species, the male and female zooids are placed at very different but well defined locations within the cormidium.
+In these species, the male and female zooids are placed at different but well defined locations within the cormidium.
 Meanwhile in calycophorans, each cormidium bears either male or female gonophores.
-In this form of monoecy, the male and female cormidia can either be in an alternating pattern, or there can be several male or female cormidia in a row.
+In this form of monoecy, the male and female cormidia can either occur in an alternating pattern, or there can be several male or female cormidia in a row.
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-In sum, homoplasy in sexual system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
+In sum, homoplasy in sexual-system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
 The sexual system and cormidial arrangement of *Rudjakovia* is undescribed, although preliminary observations suggest that this species may be monoecious and that monoecy arose a third time in the Pyrostephidae.
@@
 
-Within Calycophorae there are some variations in sexual mode -- in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex, however they are monoecious and protandrous, with female gonophores developing after the release of male gonophores [@carre1979].
+Within Calycophorae there are additional variations of the sexual mode: in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex at a time.
+However they are monoecious and protandrous, with female gonophores developing after the release of male gonophores [@carre1979].
 Environmental influences may also play a role in determining the expressed sex.
-Colonies of the calycophoran *Chelophyes appendiculata* collected in the field always bear both male and female gonophores whereas when kept in culture only gonophores of one sex are maintained [@Carre:2000hb].
+Colonies of the calycophoran *Chelophyes appendiculata* collected in the field always bear both male and female gonophores, whereas when kept in culture only gonophores of one sex are maintained [@Carre:2000hb].
 This suggests a high plasticity of the sexual state in some calycophoran taxa and underlines the need for caution when evaluating the state of this character in rarely collected species.
@@
 One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969; @cartwright2010character].
-For example, members of *Forskalia* have 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
-Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome tree (Fig.
+For example, *Forskalia* and other physonects have at least 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, there can be nine zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision].
+Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome-based tree (Fig.
 4).
@@
 In Codonophora, the nectophores are localized to a region known as the nectosome (Fig.
-2B), which has its own growth zone, and are used for coordinated colony-level swimming.
+2B), which has its own growth zone, and they are used for coordinated colony-level swimming.
 Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
-1A) instead move through the water column using repeated contraction and relaxation of the stem, and use modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
+1A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 Similarly, the homology of the special nectophore associated with gonophores of the calycophoran *Stephanophyes superba* is also unclear [@chun1891canarischen].
-As such, we only consider the evolution of the nectosome in this study, and not the presence/absence of nectophores.
+In this study, we only consider the evolution of the nectosome, and not the presence/absence of nectophores.
 The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
@@
 Top: Bathymetric distribution of siphonophore species.
-Physalia illustration by Noah Schlottman, taken from http://phylopic.org/](figures/Figure4.pdf)
+*Physalia* illustration by Noah Schlottman, taken from http://phylopic.org/](figures/Figure4.pdf)
 
 Within the nectosome, the nectophores can be attached along the dorsal or ventral side of the stem, following the orientation framework of @haddock_dunn_pugh_2005.
+The apparent placement of the nectophores on opposite sides of the nectosome occurs through twisting of the stem during development.
 Our ancestral reconstructions for this character (Fig.
@@
 
-Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig.
+Bracts are greatly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig.
 4).
@@
 These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny.
-In species without bracts, a diversity of zooids appear to fulfil the roles of neutral buoyancy and protection.
+In species without bracts, other zooids appear to fulfill the roles of neutral buoyancy and protection.
 In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
-While *Hippopodius hippopus* have up to twelve nectophores and can retract the stem within the nectophores -- the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
+*Hippopodius hippopus* has up to twelve nectophores and can retract the stem within the nectophores.
+The nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
 1C shows the blanching of nectophores)[@bassot1978bioluminescence].
+<!-- XXX This hippo section seems like a total non-sequitur in the bract paragraph, even considering the potantial role in buoyancy XXX -->
 
@@
 
-![Stochastic mapping reconstruction of the evolutionary history of A) sexual mode, whether a colony is monoecious or dioecious and B) presence/absence of palpons (modified reduced gastrozooids).
+![Stochastic mapping reconstruction of the evolutionary history of (A) sexual mode, whether a colony is monoecious or dioecious and (B) presence/absence of palpons (modified reduced gastrozooids).
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
@@
 
-The pneumatophore is a gas filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
+The pneumatophore is a gas-filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
 It is not a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
-Recent descriptions of the neural arrangement in the pneumatophore of *Nanomia bijuga*  suggests it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
+Recent descriptions of the neural arrangement in the pneumatophore of *Nanomia bijuga*  suggest it could also gather information on relative pressure changes (and thus depth changes), helping regulate geotaxis [@church2015histology].
 The ancestral siphonophore had a pneumatophore (Fig.
@@
 Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture).
-As in other cnidarians, nematocyst stinging capsules, arranged in dense nematocyst batteries, play a critical role in prey capture.
+As in other cnidarians, stinging capsules, arranged in dense batteries of nematocysts, play a critical role in prey capture.
 In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig.
@@
 S8, S9).
-Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow living species) and Clade B (deep dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis.
+Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Clade B (deep-dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis.
 The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species.
@@
 The ROV observations are reliable below 200m, and no quantitative measurements were made on SCUBA dives.
-Species such as *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
+Species such as *Nanomia bijuga*, *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
 4).
@@
 S9), and still find support for a mesopelagic ancestor.
-It is important to note, however, that *H.
+It should be noted, however, that *H.
 hippopus* and *C.
@@
 Using phylogenomic tools we were able to resolve deep relationships within Siphonophora with strong support.
-Among other relationships, we identify the clade Euphysonectae as the sister group to Calycophorae.
+We identify the clade Euphysonectae as the sister group to Calycophorae.
 Our results suggest that monoecy arose at least twice, based both on phylogenetic reconstruction and differences in the way monoecy is realized in different clades.

### 2018-01-27 manuscript.rmd
incorporating final comments

@@
 output:
-  word_document: default
   pdf_document:
@@
       in_header: mystyles.sty
+  word_document: default
 ---
@@
 All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017).
-The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag XXXXXXXXXX.
+The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v1".
 
@@
 
-Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following three alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores including *Rudjakovia* sp.
-and (iii) monophyletic monoecious with dioecious *Rudjakovia* sp.
-<!--  XXX This needs to be fixed to coincide with the two hypotheses listed in section 2.3 XXX -->.
+Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following two alternative phylogenetic hypotheses against the most likely tree topology (Fig.
+3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores.
+As the sexual mode of  *Rudjakovia* is undescribed, we carried out two tests: with one test including *Rudjakovia* sp.
+as a monoecious species, and the other with dioecious *Rudjakovia* sp.
 In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
@@
 In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
-*Hippopodius hippopus* has up to twelve nectophores and can retract the stem within the nectophores.
-The nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
+While in *Hippopodius hippopus* the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
 1C shows the blanching of nectophores)[@bassot1978bioluminescence].
-<!-- XXX This hippo section seems like a total non-sequitur in the bract paragraph, even considering the potantial role in buoyancy XXX -->
 
@@
 
-![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/gene_occupancy.png)
+![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
 
@@
 
-![Constrained topologies specified in SOWH testing.
-Test statistic and p-value for each tree estimated under constraint are given.](sowh/sowh_tests_supplemental_figure/sowh_tests_supplemental_figure.png)
+![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](supplementary_info/Supplementary_Figure2.png)
 
@@
 plot(tentilla_obj)
+
+pdf("supplementary_info/Supplementary_Figure3.pdf")
+plot(tentilla_obj)
+dev.off()
+
 ```
@@
 plot(tentilla2_obj)
+
+pdf("supplementary_info/Supplementary_Figure4.pdf")
+plot(tentilla2_obj)
+dev.off()
 ```
@@
 plot(nectosome_obj)
+
+pdf("supplementary_info/Supplementary_Figure5.pdf")
+plot(nectosome_obj)
+dev.off()
 ```
@@
 plot(desc_mantle_obj)
+
+pdf("supplementary_info/Supplementary_Figure6.pdf")
+plot(desc_mantle_obj)
+dev.off()
 ```
@@
 plot(necto_pos_obj)
+
+pdf("supplementary_info/Supplementary_Figure7.pdf")
+plot(necto_pos_obj)
+dev.off()
 ```
@@
 plot(depthmap)
+
+pdf("supplementary_info/Supplementary_Figure8.pdf")
+plot(depthmap)
+dev.off()
 ```
@@
 plot(depthmap_ext)
+
+pdf("supplementary_info/Supplementary_Figure9.pdf")
+plot(depthmap_ext)
+dev.off()
 

### 2018-01-27 supplementary_info/Supplementary_figures.Rmd
incorporating final comments

---
output:
  pdf_document:
    fig_caption: yes
---

## Supplementary Figures
\setcounter{figure}{0}
\setcounter{table}{0}
\makeatletter
\renewcommand{\thefigure}{S\@arabic\c@figure}
\renewcommand{\thetable}{S\@arabic\c@table}
\makeatother

####Agalma analysis

![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](Supplementary_Figure1.png)

\pagebreak

####SOWH analysis

![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](Supplementary_Figure2.png)

\pagebreak

#####Stochastic Character maps

![Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla.](Supplementary_Figure3.pdf)

![Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla.](Supplementary_Figure4.pdf)

![Stochastic character map of presence of nectosome.](Supplementary_Figure5.pdf)

![Stochastic character map of presence of a descending mantle canal in the nectophores. Cystonects and Athorybia were excluded as they do not have a nectosome.](Supplementary_Figure6.pdf)

![Stochastic character map for the evolution of the position of the nectosome. Cystonects were excluded as they do not have a nectosome.](Supplementary_Figure7.pdf)

![Brownian Motion character map of median depth of species observed with an MBARI ROV](Supplementary_Figure8.pdf)

![Brownian Motion character map of median depth of species including blue water diving observations.](Supplementary_Figure9.pdf)

### 2018-01-28 manuscript.rmd
added reference to table, updated figure 4
minor changes to manuscript and figure 4

@@
 
-Specimens were collected in the north-eastern Pacific Ocean, Mediterranean, and the Gulf of California.
+Specimens were collected in the north-eastern Pacific Ocean, Mediterranean, and the Gulf of California (see table 1).
 Collection data on all examined specimens, a description of the tissue that was sampled from the colony, collection mode, sample processing details, mRNA extraction methods, sequencing library preparation methods, and sequencing details are summarized in the file Supplementary data 1 (also found in the git repository).
@@
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
-The numbers of valid described species estimated to be based in each clade based on taxonomy are shown below each clade name on the right.
+The numbers of valid described species estimated to be in each clade based on taxonomy are shown below each clade name on the right.
 (B) The topologies found in the posterior distribution of trees that conflict with the ML tree.

### 2018-02-28 manuscript.rmd
updated numbers

@@
 A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
-Here, we present a new siphonophore phylogeny based on new transcriptome data from 30 siphonophore species analyzed in combination with 13 publicly available genomic and transcriptomic datasets.
+Here, we present a new siphonophore phylogeny based on new transcriptome data from 29 siphonophore species analyzed in combination with 14 publicly available genomic and transcriptomic datasets.
 We use this new phylogeny to reconstruct several traits that are central to siphonophore biology, including sexual system (monoecy vs.
@@
 
-New data were analysed in conjunction with 13 publicly available datasets, with a total number of 43 species.
+New data were analysed in conjunction with 14 publicly available datasets, with a total number of 43 species.
 Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.

### 2018-03-01 manuscript.rmd
updated supp. info, bib file

@@
 Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group.
-The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic], but there has been progress on their internal relationships.
+The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2015phylogenetic], but there has been progress on their internal relationships.
 A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
@@
 1C) specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
-Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the United States National Museum (Smithsonian Institution), Washington, DC.
+Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the Smithsonian National Museum of Natural History, Washington, DC.
 Accession numbers are given in Supplementary data 1.
-In cases where physical vouchers were unavailable we provide photographs to document species identity (see git repository).
+In cases where physical vouchers were unavailable we provide photographs to document species identity (see git repository: https://github.com/caseywdunn/siphonophore_phylogeny_2017/tree/master/supplementary_info/photographic_vouchers).
 
@@
 
-New data were analysed in conjunction with 14 publicly available datasets, with a total number of 43 species.
+New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic, @Dunn:2013kw, @lehnert2012developing, @philippe2009phylogenomics, @fidler2014unique, @chapman2010dynamic, @sanders2014differential, @putnam2007sea, @sanders2015interspecific], with a total number of 43 species.
 Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.
@@
 specimen_data <- read.csv("supplementary_info/Supplementary_data_1_Specimen_data.csv")
-names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Collection date","Specimen id","Depth (m)","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
-
-specimen_data_reduced <- specimen_data[,c(1,2,4,14,16)]
+names(specimen_data)<-c("New data","Species","ID","SRA Number", "Library ID", "Library type","Tissue","Sequencer", "Sequencing center", "Sample prep","Collected by","Identified by","Collection date","Specimen id","Depth (m)","Geo loc name","Lat Lon","Voucher Type","Voucher location","Voucher Cat No.","Collection mode")
+
+specimen_data_reduced <- specimen_data[,c(1,2,15,17,4)]
 specimen_data_reduced$`New data`<-sub("^N"," ",specimen_data_reduced$`New data`)
@@
 specimen_data_reduced<-as.data.frame(specimen_data_reduced)
-xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens collected for this work.
-New data indicated by Y, blank fields indicate that data were already published.")
+xtable_specimen<-xtable(specimen_data_reduced,auto=TRUE, caption="A complete list of specimens used in this work, information from already published datasets added where available.
+New data indicated by Y, blank fields indicate that data were already published.
+For the species not on SRA, a link to the data is included in supplementary file 1.")
 print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.7')

### 2018-03-02 manuscript.rmd
changes throughout text & to bib file

@@
 
-Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at all depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea.
+Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at most depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea.
 Siphonophores play important roles in ocean ecosystems, and are among the most abundant gelatinous predators.
@@
 1 and 2) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical; @purcell1981dietary; @pages2001gelatinous;  @Choy20172116].
-Siphonophores, which belong to Hydrozoa (Cnidaria), are found at all depths in the ocean.
+Siphonophores, which belong to Hydrozoa (Cnidaria), are found at most depths in the ocean, with the deepest recorded species found around 4300m [@lindsay2005planktonic].
 The most familiar species is the Portuguese man of war *Physalia physalis*, which floats at the surface and can wash up conspicuously onto beaches [@Totton1960].
 Most species are planktonic, living in the water column, where some grow to be more than 30 meters in length [@mackie1987siphonophore].
-There is also a small clade of benthic siphonophores, Rhodaliidae [@pugh1983benthic], that are tethered to the bottom for part of their lives.
-There are currently 187 valid described siphonophore species.
+There is also a small clade of benthic siphonophores, Rhodaliidae, that are tethered to the bottom for part of their lives [@pugh1983benthic].
+There are currently 187 valid described siphonophore species [@worms].
 
@@
 One or two growth zones (Fig.
-2) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1995ordre].
+2) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
@@
 There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans.
+We provisionally refer here to Pyrostephidae as the clade including *Rudjakovia* sp., although sampling of *Pyrostephos vanhoeffeni* is needed in order to determine if *Rudjakovia* sp.
+falls within Pyrostephidae or is sister to it.
 Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig.
@@
 
-## 3.2 Character Evolution
 ```{r char_evo_setup, include=F, echo=FALSE}
@@
 
-### 3.3 Evolution of Monoecy
+### 3.2 Evolution of Monoecy
 
@@
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
+One exception to this is Abylids, where both male or female gonophores may be found within the same eudoxid [@carre1967].
 In sum, homoplasy in sexual-system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
@@
 
-### 3.4 The Evolution of Zooid Types
+### 3.3 The Evolution of Zooid Types
 
@@
 We reconstruct them as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to Pyrostephidae (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*.
+5B), retained in most species, but lost three times independently in the branches leading to Pyrostephidae (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
+L.
 Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.
@@
 
-### 3.5 The Gain and Loss of the Pneumatophore
-
-The pneumatophore is a gas-filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
+### 3.4 The Gain and Loss of the Pneumatophore
+
+The pneumatophore (Fig.
+2A) is a gas-filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
 It is not a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
@@
 
-### 3.6 The Gain and Loss of Tentilla
+### 3.5 The Gain and Loss of Tentilla
 
@@
 
-### 3.7 The Evolution of Vertical Habitat Use
+### 3.6 The Evolution of Vertical Habitat Use
 
@@
 We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph.
-We also thank Zack Lewis for help with RNA extractions.
-We also thank the MBARI crews and ROV pilots for collection of the specimens.
+We also thank Zack Lewis for help with RNA extractions and the MBARI crews and ROV pilots for collection of the specimens.
 

### 2018-03-02 manuscript.rmd
typo in bib reference

@@
 
-New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic, @Dunn:2013kw, @lehnert2012developing, @philippe2009phylogenomics, @fidler2014unique, @chapman2010dynamic, @sanders2014differential, @putnam2007sea, @sanders2015interspecific], with a total number of 43 species.
+New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw, @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species.
 Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.

### 2018-03-02 manuscript.rmd
missing comma

@@
 
-New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw, @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species.
+New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw; @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species.
 Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.

### 2018-03-02 manuscript.rmd
added line addressing Grossmann 2015 R2 comment

@@
 This suggests that vertical habitat use is more labile than it appears and may be an important mechanism in siphonophore ecology.
+The observed variation in depth distribution could be attributed to any of the correlated environmental variables (i.e.
+temperature, chlorophyll, oxygen…).
+Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores (Grossman et al 2015).
+Since our oceanographic data is restricted to the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.
 

### 2018-03-15 manuscript.rmd
added citation to Grossmann 2015 to bib
added note about supplementary trees and robustness of model selection

@@
 This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods.
-Maximum likelihood analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
+Maximum likelihood (ML) analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
+We also used IQTree to implement other ML analyses using different models of substitution to test the robustness of the results to model selection [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
 Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
@@
 3B).
+The ML results were robust to model selection (Fig.
+S10).
 Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 temperature, chlorophyll, oxygen…).
-Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores (Grossman et al 2015).
+Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores [@grossmann2015diversity].
 Since our oceanographic data is restricted to the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.
@@
 
+```{r supplementary trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution.
+Topologies are the same.
+A.
+GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
+B.
+JTT + Empirically counted frequencies from alignment + FreeRate model with 6 categories.
+C.
+WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
+"}
+
+fix_trees<-function(tree){
+tree<-fix_names(tree)
+tree =
+	  ape::root(
+	    tree,
+	    c( "Nematostella vectensis", "Aiptasia pallida" ),
+	    resolve.root=TRUE
+	   )
+return(tree)
+}
+
+GTRtree<-ape::read.tree("supplementary_info/Siphonophore_GTR_IQtree.tre")
+JTTtree<-ape::read.tree("supplementary_info/Siphonophore_JTT_IGtree.tre")
+WAGtree<-ape::read.tree("supplementary_info/Siphonophore_WAG_IGtree.tre")
+
+trees<-c(GTRtree,JTTtree,WAGtree)
+trees<-lapply(trees,fix_trees)
+class(trees)<-"multiPhylo"
+names(trees)<-c("A. GTR20+FO+R6 ","B. JTT+F+R6","C. WAG+FO+R6")
+ggtree(trees) + facet_wrap(~.id, scale="free", ncol=3) +geom_tiplab(size=2) + xlim(0, 2)
+
+```
+
 \pagebreak

### 2018-03-15 manuscript.rmd
added likelihood values for all but 1 supp. tree
added likelihood value for JTT tree
corrected
actually the correct values.
added bootstrap support

@@
 ```{r supplementary trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution.
-Topologies are the same.
+Unlabeled nodes have support >0.99.
 A.
 GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
+Likelihood: -5964623.1967; AIC score: 11929462.3933.
 B.
 JTT + Empirically counted frequencies from alignment + FreeRate model with 6 categories.
+Likelihood:-5964771.2358; AIC: 11929758.4715.
 C.
 WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
-"}
+Likelihood:-5985124.7530; AIC score: 11970465.5060 "}
 
@@
 class(trees)<-"multiPhylo"
-names(trees)<-c("A. GTR20+FO+R6 ","B. JTT+F+R6","C. WAG+FO+R6")
-ggtree(trees) + facet_wrap(~.id, scale="free", ncol=3) +geom_tiplab(size=2) + xlim(0, 2)
+
+support = as.numeric( trees[[1]]$node.label )
+support[ support > 99 ] = NA
+support1 = c( rep( NA, length( trees[[1]]$tip.label ) ), support )
+
+support = as.numeric( trees[[2]]$node.label )
+support[ support > 99 ] = NA
+support2 = c( rep( NA, length( trees[[2]]$tip.label ) ), support )
+
+support = as.numeric( trees[[3]]$node.label )
+support[ support > 99 ] = NA
+support3 = c( rep( NA, length( trees[[3]]$tip.label ) ), support )
+
+p1<-ggtree(trees[[1]])+geom_tiplab(size=2)+geom_text(aes(label=support1),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("A. GTR20+FO+R6")
+
+p2<-ggtree(trees[[2]])+geom_tiplab(size=2)+geom_text(aes(label=support2),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("B. JTT+F+R6")
+
+p3<-ggtree(trees[[3]])+geom_tiplab(size=2)+geom_text(aes(label=support3),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("C. WAG+FO+R6")
+
+multiplot(p1, p2, p3, ncol=3)
 

### 2018-03-15 manuscript.rmd
added clarification on palpons

@@
 
-Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore].
-We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis].
-We reconstruct them as present in the common ancestor of siphonophores (Fig.
+Palpons are typically defined as modified reduced gastrozooids [@mackie1987siphonophore].
+Across different species and clades, palpons vary in their function, morphology, and location in the colony.
+In many species palpons are thought to play a role in digestion and circulation of the gastrovascular fluid, while other species may use them for defense (e.g _Physophora_) or sensory functions [@totton1965synopsis].
+Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle); palpons found along the stem of the siphosome are termed palpons (typically have a tentacle, called a palpacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in _Apolemia_) [@totton1965synopsis;@siebert2013re].
+It is not clear how structure and function differs among different palpon subtypes, and a detailed histological investigation of the palpons found at different locations within species is needed.
+For this reason, here we only assess the presence or absence of palpons as a category, without assessing subtypes of palpons.
+This presumes that palpons located at different regions in the colony are derived from other palpons rather than each arising de novo by independent modification of gastrozooids, a hypothesis that itself could be challenged upon closer histological examination of palpon diversity.
+
+We reconstruct palpons as present in the common ancestor of siphonophores (Fig.
 5B), retained in most species, but lost three times independently in the branches leading to Pyrostephidae (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis* and Undescribed sp.

### 2018-03-16 manuscript.rmd
updated palpons section

@@
 Palpons are typically defined as modified reduced gastrozooids [@mackie1987siphonophore].
-Across different species and clades, palpons vary in their function, morphology, and location in the colony.
 In many species palpons are thought to play a role in digestion and circulation of the gastrovascular fluid, while other species may use them for defense (e.g _Physophora_) or sensory functions [@totton1965synopsis].
 Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle); palpons found along the stem of the siphosome are termed palpons (typically have a tentacle, called a palpacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in _Apolemia_) [@totton1965synopsis;@siebert2013re].
-It is not clear how structure and function differs among different palpon subtypes, and a detailed histological investigation of the palpons found at different locations within species is needed.
+It is not clear how structure and function differs among different palpon subtypes, and a detailed histological investigation of palpons found at different locations within species is needed.
 For this reason, here we only assess the presence or absence of palpons as a category, without assessing subtypes of palpons.

### 2018-03-16 manuscript.rmd
added comment about oleocysts

@@
 We reconstruct palpons as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to Pyrostephidae (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
+5B), retained in most species, but lost three times independently in the branches leading to *Bargmannia* and *Rudjakovia* sp., in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
 L.
+The pyrostephid *Pyrostephos vanhoeffeni* (not sampled) has modified tentacle-less palpons (termed oleocysts), but the relationship between this species and *Rudjakovia* sp.
+is not known, so the exact patterns of loss within *Pyrostephidae* remain unclear.
 Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.

### 2018-03-16 manuscript.rmd
updated support/grants. Caveats about depth data

@@
 The observed variation in depth distribution could be attributed to any of the correlated environmental variables (i.e.
-temperature, chlorophyll, oxygen…).
+temperature, chlorophyll, oxygen).
 Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores [@grossmann2015diversity].
-Since our oceanographic data is restricted to the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.
+Since we only have oceanographic data from the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.
 
@@
 appendiculata* were both collected in the bay of Villefrance-sur-mer, France, where an upwelling is known to bring deeper species closer to the surface [@nival1976evolution].
+Additionally, while we are confident about many of the species IDs in the VARS dataset, it is difficult to distinguish *Kephyes ovata* and *K.
+hiulcus* from images alone and the distribution likely includes data points from both species.
+*Halistemma rubrum* distributions were obtained from cruises in the Gulf of California, where the only *Halistemma* species collected by ROV is *H.
+rubrum*.
+Where we could not be certain of species identifications in the VARS dataset, we only included a few data points from specimens that were collected and identified.
 
@@
 This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award).
+CM was also supported in part by a RI-EPSCoR Fellowship, NSF EPS-1004057.
 Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057.
@@
 SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster.
+Analyses were also conducted on the Hoffman2 cluster of the Institute for Digital Research and Education of the University of California, Los Angeles.
 We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph.

### 2018-04-25 manuscript.rmd
updated tree stats

@@
 Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling).
-Using 1,071 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study.
+Using 1,423 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study.
 Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.
@@
 New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw; @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species.
-Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v.
-1.0.0 [@Dunn:2013kw; @Guang202416].
+Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma (commit 6bd9988, running BioLite commit 784edc6) [@Dunn:2013kw; @Guang202416].
 This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods.
@@
 
-Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes).
+Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 17.0% and 17.3% respectively in a 60% occupancy matrix with 3,379 genes).
 The final analyses presented here consider 33 siphonophore species and 8 outgroup species.
 This includes new data for 30 species.
-In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (Fig.
+In the final analyses, we sampled 1,423 genes to generate a supermatrix with 80% occupancy and a length of 395,699 amino acids (Fig.
 S1).
@@
 BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
-Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model.
-The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0127).
+Two independent MCMC chains were run under the CAT-GTR model, and eight independent MCMC chains were run under the CAT-Poisson model.
+The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0221709).
 Only the results from the CAT-Poisson model are presented here.
+Two chains (chain1 and chain3) did not burn in after 1405 and 4695 iterations and were excluded from the analyses.
 Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
@@
 Relationships that receive strong support in both include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae".
-Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses.
+Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,423-gene transcriptome analyses.
 There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans.
@@
 
-![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
+![60% gene occupancy matrix for 41 species across 1,423 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
 
@@
 GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
-Likelihood: -5964623.1967; AIC score: 11929462.3933.
+Likelihood: -8133157.335; AIC:16266530.7277.
 B.
-JTT + Empirically counted frequencies from alignment + FreeRate model with 6 categories.
-Likelihood:-5964771.2358; AIC: 11929758.4715.
+JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories.
+Likelihood:-8113694.922; AIC: .
 C.
 WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
-Likelihood:-5985124.7530; AIC score: 11970465.5060 "}
+Likelihood: -8156043.772; AIC score: 16312303.6120 "}
 

### 2018-05-01 manuscript.rmd
updated poisson chains and stats about runs

@@
 	 )
+
+	#remove two chains that are stuck in local maxima
+	input_files_phylobayes= input_files_phylobayes[-which(input_files_phylobayes %in% c("chain1_trimmed60_poisson.treelist","chain3_trimmed60_poisson.treelist"))]
 
@@
 Two independent MCMC chains were run under the CAT-GTR model, and eight independent MCMC chains were run under the CAT-Poisson model.
-The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0221709).
+The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0185032; CAT-GTR: maxdiff=1, meandiff=0.045415).
 Only the results from the CAT-Poisson model are presented here.
-Two chains (chain1 and chain3) did not burn in after 1405 and 4695 iterations and were excluded from the analyses.
+Two chains (chain1 and chain3) were stuck in local maxima after 1405 and 4695 iterations and were excluded from the analyses.
 Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
@@
 JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories.
-Likelihood:-8113694.922; AIC: .
+Likelihood:-8113694.922; AIC:16227609.9565 .
 C.

### 2018-05-04 manuscript.rmd
mapped traits to BI topology

@@
 
-Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Clade B (Fig.
+In ML analyses Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Clade B (Fig.
 3A).
+In BI analyses, clade B is graded (Fig.
+3B).
 The presence of an involucrum, a fold around the base of the cnidoband [@totton1965synopsis], is a potential synapomorphy for Clade A.
@@
 There is not a clear synapomorphy for Clade B.
-Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses (Fig.
+Within Clade B there is high support for the placement of *Erenna richardi* in ML analyses, but it is placed as sister to Clade A in BI analyses (Fig.
 3B).
@@
 As in the previous study, the results presented here indicate that the prayomorphs are paraphyletic with respect to the diphyomorphs.
-*Craseoa lathetica* and *Desmophyes* sp.
-are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *C.
+In the previous study, the relationship between *C.
 lathetica* and the clade including *H.
 hippopus* was unresolved.
+In this study, *Craseoa lathetica* and *Desmophyes* sp.
+are sister to *Hippopodius hippopus* in ML analyses with high support, while in BI analyses, *H.
+hippopus* is sister to *Lilyopsis fluoracantha* and the diphyomorphs (Fig.
+3B).
 
@@
 
+```{r BI tree ancestral trait reconstructions}
+
+BIspeciestree<-read.tree(file="agalma/Outfiles/combined_trimmed60_poisson.con.tre")
+BIspeciestree = ape::ladderize( BIspeciestree )
+
+BIspeciestree = fix_names( BIspeciestree )
+
+BIspeciestree =
+	  ape::root(
+	    BIspeciestree,
+	    c( "Nematostella vectensis", "Aiptasia pallida" ),
+	    resolve.root=TRUE
+	   )
+
+tree<-BIspeciestree
+
+tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
+
+ultratree <- chronos(tree)
+class( ultratree  ) = "phylo"
+
+dpruned_tree <- drop.tip(ultratree, which(!(ultratree$tip.label %in% rownames(dpruned_data))))
+dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
+
+depth_median = dpruned_data$Depth.Median
+names(depth_median) = rownames(dpruned_data)
+
+#Stochastic trait mapping:
+#Make simmap objects
+#print("SIMMAP Sex Distribution")
+make.simmap(ultratree , sex, nsim = 100, message=FALSE) -> sex_sim
+#print("SIMMAP Palpons")
+make.simmap(ultratree , palpons, nsim = 100, message=FALSE) -> palpons_sim
+#print("SIMMAP Tentilla")
+make.simmap(ultratree , tentilla, nsim = 100, message=FALSE) -> tentilla_sim
+#print("SIMMAP Tentilla2")
+make.simmap(ultratree , tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
+#print("SIMMAP Nectosome")
+make.simmap(ultratree , nectosome, nsim = 100, message=FALSE) -> nectosome_sim
+#print("SIMMAP Nectosome Position")
+make.simmap(ultratree , bracts, nsim = 100, message=FALSE) -> bracts_sim
+#print("SIMMAP Habitat")
+
+#Nectophore position SIMMAP setup
+NectoPosTree = drop.tip(ultratree , which(!(ultratree $tip.label %in% names(necto_pos))))
+make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+
+#Descending mantle Canal SIMMAP Setup
+DescMantleTree =  drop.tip(ultratree , which(!(ultratree $tip.label %in% names(desc_mantle))))
+make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
+```
+
+```{r BI mono, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of sexual characters on the consensus Bayesian tree topology." }
+#Simmap SEX
+sex_obj <- densityMap(sex_sim, plot=FALSE)
+sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj$cols))
+ plot(sex_obj)
+```
+
+```{r BI palp, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of palpon presence/absence on the consensus Bayesian tree topology." }
+#Simmap PALPONS
+palpons_obj <- densityMap(palpons_sim, plot=FALSE)
+palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj$cols))
+
+plot(palpons_obj)
+```
+
+```{r tentilla BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla, mapped on the consensus Bayesian tree topology."}
+tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
+tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
+plot(tentilla_obj)
+```
+
+```{r tentilla ext BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla, mapped on the consensus Bayesian tree topology."}
+tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
+tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
+plot(tentilla2_obj)
+```
+
+```{r nectosome BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome, mapped on the consensus Bayesian tree topology."}
+nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
+```
+
+```{r mantle BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores, mapped on the consensus Bayesian tree topology."}
+desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
+```
+
+```{r necto BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome, mapped on the consensus Bayesian tree topology."}
+necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
+necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
+plot(necto_pos_obj)
+```
+
+```{r depth BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV, mapped on the consensus Bayesian tree topology."}
+depth_value = depth_median
+depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+
+```{r depth BI extended, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations, mapped on the consensus Bayesian tree topology."}
+median_extended = cdata$Depth.Median.Extended
+names(median_extended) = rownames(cdata)
+median_extended = median_extended[match(names(median_extended), ultratree$tip.label)]
+depthmap_ext <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap_ext)
+
+```
+
 \pagebreak

### 2018-05-04 manuscript.rmd
updated text to include new supp figures

@@
 Subsequently, mRNA was extracted directly from tissue using a variety of methods (Supplementary data 1): Magnetic mRNA Isolation Kit (NEB, #S1550S), Invitrogen Dynabeads mRNA Direct Kit (Ambion, #61011), Zymo Quick RNA MicroPrep (Zymo #R1050), or from total RNA after Trizol (Ambion, #15596026) extraction and through purification using Dynabeads mRNA Purification Kit (Ambion, #61006).
-In case of anticipated very small total RNA quantities, only a single round of bead purification was performed.
+In case of small total RNA quantities, only a single round of bead purification was performed.
 Extractions were performed according to the manufacturer's instruction.
@@
 Species of Clade A also have a descending mantle canal within the nectophores (Fig.
-S6), a structure that is also present in some calycophorans.
+S6, S16), a structure that is also present in some calycophorans.
 Members of Clade A are also monoecious (Fig.
@@
 Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig.
-5A), once in Calycophorae and once in Clade A.
+5A, S11), once in Calycophorae and once in Clade A.
 This is consistent with differences in the arrangements of male and female gonophores in the two clades.
@@
 The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
-S5).
-
-![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
-Bottom: siphonophore phylogeny, colored by clade.
+S5, S15).
+
+![Siphonophore ML phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
+Bottom: siphonophore ML phylogeny, colored by clade.
 Middle panel: diagram showing the presence/absence of traits across Siphonophora, with the physical location of the trait shown on a schematic of *Nanomia bijuga* (schematic by Freya Goetz).
@@
 We reconstruct palpons as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Bargmannia* and *Rudjakovia* sp., in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
+5B, S12), retained in most species, but lost three times independently in the branches leading to *Bargmannia* and *Rudjakovia* sp., in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
 L.
@@
 
-![Stochastic mapping reconstruction of the evolutionary history of (A) sexual mode, whether a colony is monoecious or dioecious and (B) presence/absence of palpons (modified reduced gastrozooids).
+![Stochastic mapping reconstruction on the ML tree of the evolutionary history of (A) sexual mode, whether a colony is monoecious or dioecious and (B) presence/absence of palpons (modified reduced gastrozooids).
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
@@
 When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support a common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
-S3).
+S3, S13).
 However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.
@@
 When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig.
-S4), and in *Bathyphysa conifera*.
+S4, S14), and in *Bathyphysa conifera*.
 The application of phylogenetic methods to the evolution of tentillum morphology would be a crucial step towards understanding the evolution of these structures, and their relationship with the feeding ecology of siphonophores.
@@
 There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
-S8, S9).
+S8, S9, S18, S19).
 Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Clade B (deep-dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis.

### 2018-05-10 manuscript.rmd
updated to bookdown

--- 2018-05-01 manuscript.rmd
@@
 csl: molecular-phylogenetics-and-evolution.csl
+site: "bookdown::bookdown_site"
 output:
-  pdf_document:
-    fig_caption: yes
-    includes:
-      in_header: mystyles.sty
-  word_document: default
+  bookdown::pdf_document2:
+    toc: FALSE
 ---
@@
 Siphonophores (Fig.
-1 and 2) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical; @purcell1981dietary; @pages2001gelatinous;  @Choy20172116].
+\@ref(siphonophores) and \@ref(colony)) are among the most abundant gelatinous predators in the open ocean, and have a large impact on ocean ecosystems [@pugh1984diel;@pugh1997vertical; @williams1981vertical; @purcell1981dietary; @pages2001gelatinous;  @Choy20172116].
 Siphonophores, which belong to Hydrozoa (Cnidaria), are found at most depths in the ocean, with the deepest recorded species found around 4300m [@lindsay2005planktonic].
@@
 One or two growth zones (Fig.
-2) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre].
+\@ref(colony)) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
@@
 Zooids are specialized for a range of functions, including feeding, reproduction, or swimming (Fig.
-2) [@dunn2006evolution].
-
-![Photographs of living siphonophores.
+\@ref(colony)) [@dunn2006evolution].
+
+![\label{siphonophores}Photographs of living siphonophores.
 Colored circles correspond to the clades shown in Figure 3 as follows: Cystonectae (A-B), Calycophorae (C-G), Apolemiidae (H), and Clade A within Euphysonectae (I-K).
@@
 Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig.
-2).
+\@ref(colony)).
 Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling).
@@
 
-![Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top of the page, and the ventral side to the left.
+![\label{colony}Schematic of the siphonophore *Nanomia bijuga*, oriented with the anterior of the colony at the top of the page, and the ventral side to the left.
 Adapted from http://commons.wikimedia.org/wiki/File:Nanomia_bijuga_whole_animal_and_growth_zones.svg, drawn by Freya Goetz.
@@
 *Chelophyes appendiculata* and *Hippopodius hippopus* (Fig.
-1C) specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
+\@ref(siphonophores)C) specimens were collected in the bay of Villefranche-sur-Mer, France, during a plankton trawl on 13 April 2011.
 Available physical vouchers have been deposited at the Museum of Comparative Zoology (Harvard University), Cambridge, MA, the Peabody Museum of Natural History (Yale University), New Haven, CT, or had been previously deposited at the Smithsonian National Museum of Natural History, Washington, DC.
@@
 In the final analyses, we sampled 1,423 genes to generate a supermatrix with 80% occupancy and a length of 395,699 amino acids (Fig.
-S1).
+\@ref(gene-occupancy)).
 
@@
 We used the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic] to evaluate two hypotheses (Fig.
-3C, S2): (i) "Physonectae" is monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
+\@ref(phylogeny)C, S2): (i) "Physonectae" is monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
 As the sexual system of *Rudjakovia* sp.
@@
 included as a monoecious species, and one with it considered dioecious (Fig.
-3C, S2).
+\@ref(phylogeny)C, S2).
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
@@
 For each hypothesis we defined a topology with a single constrained node that was inconsistent with the most likely topology (Fig.
-3).
+\@ref(phylogeny)).
 We used a threshold for significance of 0.05 and following the initial 100 samples, we evaluated the confidence interval around the p-value to determine if more samples were necessary.
@@
 The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig.
-3A), with a couple of localized exceptions (Fig.
-3B).
+\@ref(phylogeny)A), with a couple of localized exceptions (Fig.
+\@ref(phylogeny)B).
 The ML results were robust to model selection (Fig.
-S10).
+\@ref(fig:supplementary-trees)).
 Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig.
-3A).
+\@ref(phylogeny)A).
 It includes the remaining non-apolemiid, non-pyrostephid "Physonectae".
@@
 
-Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Clade B (Fig.
-3A).
+In ML analyses Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Group B (Fig.
+\@ref(phylogeny)A).
+In BI analyses, Group B is graded (Fig.
+\@ref(phylogeny)B).
 The presence of an involucrum, a fold around the base of the cnidoband [@totton1965synopsis], is a potential synapomorphy for Clade A.
 Species of Clade A also have a descending mantle canal within the nectophores (Fig.
-S6), a structure that is also present in some calycophorans.
+\@ref(fig:desc-mantle), \@ref(fig:mantle-BI)), a structure that is also present in some calycophorans.
 Members of Clade A are also monoecious (Fig.
-5).
-There is not a clear synapomorphy for Clade B.
-Within Clade B there is low support for the placement of *Erenna richardi*, which is placed as sister to Clade A in some analyses (Fig.
-3B).
+\@ref(reconstruct)).
+There is not a clear synapomorphy for Group B.
+Within Group B there is high support for the placement of *Erenna richardi* in ML analyses, but it is placed as sister to Clade A in BI analyses (Fig.
+\@ref(phylogeny)B).
 More taxon sampling will be required to determine the relationship of species within this clade.
@@
 Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (Fig.
-1I) (both not included in the previous phylogeny) are sister to Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
+\@ref(siphonophores)I) (both not included in the previous phylogeny) are sister to Agalmatidae, a clade restricted to *Agalma*, *Athorybia*, *Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic].
 In the rDNA study, *P.
@@
 As in the previous study, the results presented here indicate that the prayomorphs are paraphyletic with respect to the diphyomorphs.
-*Craseoa lathetica* and *Desmophyes* sp.
-are sister to *Hippopodius hippopus* in this study, while in the previous study, the relationship between *C.
+In the previous study, the relationship between *C.
 lathetica* and the clade including *H.
 hippopus* was unresolved.
-
-![(A) Maximum likelihood (ML) phylogram with bipartition frequencies from the ML bootstraps and the Bayesian posterior distribution of trees.
+In this study, *Craseoa lathetica* and *Desmophyes* sp.
+are sister to *Hippopodius hippopus* in ML analyses with high support, while in BI analyses, *H.
+hippopus* is sister to *Lilyopsis fluoracantha* and the diphyomorphs (Fig.
+\@ref(phylogeny)B).
+
+![\label{phylogeny}(A) Maximum likelihood (ML) phylogram with bipartition frequencies from the ML bootstraps and the Bayesian posterior distribution of trees.
 Unlabeled nodes have support >0.99 for both bootstraps and posteriors.
@@
 For more details on the SOWH topologies refer to Fig.
-S2.](figures/Figure3.pdf)
+\@ref(sowh).](figures/Figure3.pdf)
 
 Using the Swofford-Olsen-Waddell-Hillis (SOWH) test [@swofford1996phylogenetic], we evaluated the following two alternative phylogenetic hypotheses against the most likely tree topology (Fig.
-3C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores.
+\@ref(phylogeny)C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores.
 As the sexual mode of  *Rudjakovia* is undescribed, we carried out two tests: with one test including *Rudjakovia* sp.
@@
 In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
-S2).
+\@ref(sowh)).
 
@@
 This opens up new questions about key relationships within both Calycophorae and Euphysonectae -- where future transcriptome sampling efforts should be focused.
-Within Euphysonectae, two clades (Clade A and Clade B) are hypothesized, although there is weaker support for Clade B (Fig.
-3A, 3B).
-Expanding sampling of species that probably fall in Clade B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Clade B synapomorphies.
+Within Euphysonectae, two clades (Clade A and Group B) are hypothesized, although there is weaker support for Group B (Fig.
+\@ref(phylogeny)A, 3B).
+Expanding sampling of species that probably fall in Group B, including other *Erenna* species, rhodaliids, and relatives of Undescribed sp L, will greatly expand our understanding of these two groups and perhaps provide evidence of Group B synapomorphies.
 Similarly, within Calycophorae, increased taxon sampling is needed.
@@
 In @Dunn:2005dy, a clade of prayomorphs including *Praya dubia* (Fig.
-1G), *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
+\@ref(siphonophores)G), *Nectadamas diomedeae*, and *Nectopyramis natans* (not included in this study) were found to be sister to all other calycophorans, while in this study, the prayomorph *Lilyopsis fluoracantha* (not included in the previous study) is found in a clade including diphyomorph calycophorans that is sister to all other prayomorphs.
 Expanded transcriptome sequencing, particularly *P.
@@
 This will be especially important for understanding trait evolution within Calycophorae,  for example, the release of eudoxids (Fig.
-4), or the arrangement of male and female zooids along the stem (see section 3.2 below).
+\@ref(traitdist)), or the arrangement of male and female zooids along the stem (see section 3.2 below).
 
@@
 The two clades of monoecious siphonophores, Calycophorae and Clade A (Fig.
-3A), do not form a monophyletic group.
-This is because Clade B, which contains dioecious species, is also descended from their most recent common ancestor.
-SOWH tests strongly reject the placement of the monecious clades Calycophorae and Clade A as a group that excludes Clade B (Figs.
+\@ref(phylogeny)A), do not form a monophyletic group.
+This is because Group B, which contains dioecious species, is also descended from their most recent common ancestor.
+SOWH tests strongly reject the placement of the monecious clades Calycophorae and Clade A as a group that excludes Group B (Figs.
 3C and S2).
-The positions of the only two taxa from Clade B that were included in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that study.
+The positions of the only two taxa from Group B that were included in the previous analysis [@Dunn:2005dy], *Erenna* and *Stephalia*, were unresolved in that study.
 This difference in conclusions regarding trait evolution, therefore, does not reflect a contradiction between alternative strongly supported results, but the resolution of earlier polytomies in a way that indicates there has been homoplasy in the evolution of monoecy.
@@
 The distribution of monoecy is consistent with two potential scenarios (Fig.
-4).
-In the first, there is a single shift from dioecy to monoecy along the branch that gave rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Clade B.
+\@ref(traitdist)).
+In the first, there is a single shift from dioecy to monoecy along the branch that gave rise to the most recent common ancestor of Calycophorae and Euphysonectae, followed by a shift back to dioecy along the branch that gave rise to Group B.
 In the second, monoecy arose twice: once along the branch that gave rise to Clade A and again along the branch that gave rise to Calycophorae.
@@
 Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig.
-5A), once in Calycophorae and once in Clade A.
+\@ref(reconstruct)A, \@ref(fig:BI-mono)), once in Calycophorae and once in Clade A.
 This is consistent with differences in the arrangements of male and female gonophores in the two clades.
 In Clade A, male and female zooids are found within the same cormidium (a single reiterated sequence of zooids along the stem, see Fig.
-2).
+\@ref(colony)).
 In these species, the male and female zooids are placed at different but well defined locations within the cormidium.
@@
 Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome-based tree (Fig.
-4).
+\@ref(traitdist)).
 
 Nectophores (Fig.
-2) are non-reproductive propulsive medusae.
+\@ref(colony)) are non-reproductive propulsive medusae.
 In Codonophora, the nectophores are localized to a region known as the nectosome (Fig.
-2B), which has its own growth zone, and they are used for coordinated colony-level swimming.
+\@ref(colony)B), which has its own growth zone, and they are used for coordinated colony-level swimming.
 Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
-1A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
+\@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig.
-S5).
-
-![Siphonophore phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
-Bottom: siphonophore phylogeny, colored by clade.
+\@ref(fig:nectosome-origin), \@ref(fig:nectosome-BI)).
+
+![\label{traitdist}Siphonophore ML phylogeny showing the distribution of the main anatomical characters and the bathymetric distributions of the different species.
+Bottom: siphonophore ML phylogeny, colored by clade.
 Middle panel: diagram showing the presence/absence of traits across Siphonophora, with the physical location of the trait shown on a schematic of *Nanomia bijuga* (schematic by Freya Goetz).
@@
 Our ancestral reconstructions for this character (Fig.
-S7) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the stem of Agalmatidae and once along the stem of Pyrostephidae.
+\@ref(fig:nectosome-pos), \@ref(fig:necto-BI)) suggest that ventral attachment of nectophores was the ancestral state in Codonophora, and that dorsal attachment has independently evolved twice -- once along the stem of Agalmatidae and once along the stem of Pyrostephidae.
 The functional implication of dorsal vs.
@@
 Bracts are greatly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig.
-4).
+\@ref(traitdist)).
 Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen].
@@
 Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica* (Fig.
-1K), and in *Gymnopraia lapislazula*.
+\@ref(siphonophores)K), and in *Gymnopraia lapislazula*.
 These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny.
@@
 While in *Hippopodius hippopus* the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
-1C shows the blanching of nectophores)[@bassot1978bioluminescence].
+\@ref(siphonophores)C shows the blanching of nectophores)[@bassot1978bioluminescence].
 
@@
 We reconstruct palpons as present in the common ancestor of siphonophores (Fig.
-5B), retained in most species, but lost three times independently in the branches leading to *Bargmannia* and *Rudjakovia* sp., in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
+\@ref(reconstruct)B, \@ref(fig:BI-palp)), retained in most species, but lost three times independently in the branches leading to *Bargmannia* and *Rudjakovia* sp., in calycophorans, and in *Marrus claudanielis* and Undescribed sp.
 L.
@@
 
-![Stochastic mapping reconstruction of the evolutionary history of (A) sexual mode, whether a colony is monoecious or dioecious and (B) presence/absence of palpons (modified reduced gastrozooids).
+![\label{reconstruct}Stochastic mapping reconstruction on the ML tree of the evolutionary history of (A) sexual mode, whether a colony is monoecious or dioecious and (B) presence/absence of palpons (modified reduced gastrozooids).
 The color gradients show the reconstructed probability estimate of the discrete character states along the branches.
@@
 The pneumatophore (Fig.
-2A) is a gas-filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
+\@ref(colony)A) is a gas-filled float located at the anterior end of the colony, which helps the colony maintain its orientation in the water column, and plays a role in flotation in the case of the cystonects [@church2015histology;@mackie1974locomotion; @totton1965synopsis].
 It is not a zooid, as it is not formed by budding but by invagination at the aboral end of the planula during early development [@garstang1946morphology; @leloup1935siphonophores; @carre1969etude].
@@
 The ancestral siphonophore had a pneumatophore (Fig.
-2B), since both cystonects and all "physonects" possess one (Fig.
-4).
+\@ref(colony)B), since both cystonects and all "physonects" possess one (Fig.
+\@ref(traitdist)).
 The pneumatophore was lost in Calycophorae and never gained again in that clade.
@@
 In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig.
-2A).
+\@ref(colony)A).
 Outside of Siphonophora, most hydrozoans bear simple tentacles without side branches.
@@
 (Fig.
-1H), and *Bathyphysa conifera* (Fig.
-1B).
+\@ref(siphonophores)H), and *Bathyphysa conifera* (Fig.
+\@ref(siphonophores)B).
 Since *B.
@@
 When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support a common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig.
-S3).
+\@ref(fig:tentilla), \@ref(fig:tentilla-BI)).
 However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.
@@
 When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig.
-S4), and in *Bathyphysa conifera*.
+\@ref(fig:tentilla2), \@ref(fig:tentilla-ext-BI)), and in *Bathyphysa conifera*.
 The application of phylogenetic methods to the evolution of tentillum morphology would be a crucial step towards understanding the evolution of these structures, and their relationship with the feeding ecology of siphonophores.
@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs.
-4 and S8) [@mapstone2014global; @mackie1987siphonophore].
+\@ref(traitdist) and \@ref(fig:median-cont)) [@mapstone2014global; @mackie1987siphonophore].
 The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel].
@@
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
-S9), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
+\@ref(fig:median-depth-extended)), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
 This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters.
@@
 There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
-S8, S9).
-Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Clade B (deep-dwelling species); however several shallow-living species that likely belong in Clade B were not included in this analysis.
+\@ref(fig:median-cont), \@ref(fig:median-depth-extended), \@ref(fig:depth-BI), \@ref(fig:depth-BI-extended)).
+Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species); however several shallow-living species that likely belong in Group B were not included in this analysis.
 The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species.
@@
 This reconstruction (Fig.
-S8) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
+\@ref(fig:median-cont)) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
 The ROV observations are reliable below 200m, and no quantitative measurements were made on SCUBA dives.
 Species such as *Nanomia bijuga*, *Hippopodius hippopus*, *Athorybia rosacea*, *Diphyes dispar*, and *Chelophyes appendiculata* are often encountered blue water diving less than 20m from the surface (Fig.
-4).
+\@ref(traitdist)).
 We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or via a shallow trawl (Fig.
-S9), and still find support for a mesopelagic ancestor.
+\@ref(fig:median-depth-extended)), and still find support for a mesopelagic ancestor.
 It should be noted, however, that *H.
@@
 
-![60% gene occupancy matrix for 41 species across 1,423 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
+![\label{gene-occupancy}60% gene occupancy matrix for 41 species across 1,423 genes.
+Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
 
@@
 
-![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](supplementary_info/Supplementary_Figure2.png)
+![\label{sowh}Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](supplementary_info/Supplementary_Figure2.png)
 
@@
 
-```{r nectosome, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
+```{r nectosome-origin, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,  fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome."}
 #Simmap NECTOSOME
@@
 
-```{r desc_mantle, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores.
+```{r desc-mantle, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores.
 Cystonects and Athorybia were excluded as they do not have a nectosome.
@@
 
-```{r nectosome_pos, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.
+```{r nectosome-pos, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome.
 Cystonects were excluded as they do not have a nectosome."}
@@
 
-```{r median_cont, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
+```{r median-cont, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV."}
 depth_value = depth_median
@@
 
-```{r median_depth_extended, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations." }
+```{r median-depth-extended, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations." }
 
@@
 
-```{r supplementary trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution.
+```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution.
 Unlabeled nodes have support >0.99.
@@
 
-p1<-ggtree(trees[[1]])+geom_tiplab(size=2)+geom_text(aes(label=support1),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("A. GTR20+FO+R6")
-
-p2<-ggtree(trees[[2]])+geom_tiplab(size=2)+geom_text(aes(label=support2),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("B. JTT+F+R6")
-
-p3<-ggtree(trees[[3]])+geom_tiplab(size=2)+geom_text(aes(label=support3),size=3, hjust=-0.1, col="red") + xlim(0, 2)+ ggtitle("C. WAG+FO+R6")
+p1<-ggtree(trees[[1]])+geom_tiplab(size=2)+geom_text(aes(label=support1),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("A. GTR20+FO+R6")
+
+p2<-ggtree(trees[[2]])+geom_tiplab(size=2)+geom_text(aes(label=support2),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("B. JTT+F+R7")
+
+p3<-ggtree(trees[[3]])+geom_tiplab(size=2)+geom_text(aes(label=support3),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("C. WAG+FO+R6")
 
 multiplot(p1, p2, p3, ncol=3)
+
+```
+
+```{r BI tree ancestral trait reconstructions}
+
+BIspeciestree<-read.tree(file="agalma/Outfiles/combined_trimmed60_poisson.con.tre")
+BIspeciestree = ape::ladderize( BIspeciestree )
+
+BIspeciestree = fix_names( BIspeciestree )
+
+BIspeciestree =
+	  ape::root(
+	    BIspeciestree,
+	    c( "Nematostella vectensis", "Aiptasia pallida" ),
+	    resolve.root=TRUE
+	   )
+
+tree<-BIspeciestree
+
+tree <- drop.tip(tree, which(!(tree$tip.label %in% rownames(cdata))))
+
+ultratree <- chronos(tree)
+class( ultratree  ) = "phylo"
+
+dpruned_tree <- drop.tip(ultratree, which(!(ultratree$tip.label %in% rownames(dpruned_data))))
+dpruned_data <- dpruned_data[match(dpruned_tree$tip.label, rownames(dpruned_data)),]
+
+depth_median = dpruned_data$Depth.Median
+names(depth_median) = rownames(dpruned_data)
+
+#Stochastic trait mapping:
+#Make simmap objects
+#print("SIMMAP Sex Distribution")
+make.simmap(ultratree , sex, nsim = 100, message=FALSE) -> sex_sim
+#print("SIMMAP Palpons")
+make.simmap(ultratree , palpons, nsim = 100, message=FALSE) -> palpons_sim
+#print("SIMMAP Tentilla")
+make.simmap(ultratree , tentilla, nsim = 100, message=FALSE) -> tentilla_sim
+#print("SIMMAP Tentilla2")
+make.simmap(ultratree , tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
+#print("SIMMAP Nectosome")
+make.simmap(ultratree , nectosome, nsim = 100, message=FALSE) -> nectosome_sim
+#print("SIMMAP Nectosome Position")
+make.simmap(ultratree , bracts, nsim = 100, message=FALSE) -> bracts_sim
+#print("SIMMAP Habitat")
+
+#Nectophore position SIMMAP setup
+NectoPosTree = drop.tip(ultratree , which(!(ultratree $tip.label %in% names(necto_pos))))
+make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+
+#Descending mantle Canal SIMMAP Setup
+DescMantleTree =  drop.tip(ultratree , which(!(ultratree $tip.label %in% names(desc_mantle))))
+make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
+```
+
+```{r BI-mono, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of sexual characters on the consensus Bayesian tree topology." }
+#Simmap SEX
+sex_obj <- densityMap(sex_sim, plot=FALSE)
+sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj$cols))
+ plot(sex_obj)
+```
+
+```{r BI-palp, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of palpon presence/absence on the consensus Bayesian tree topology." }
+#Simmap PALPONS
+palpons_obj <- densityMap(palpons_sim, plot=FALSE)
+palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj$cols))
+
+plot(palpons_obj)
+```
+
+```{r tentilla-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla, mapped on the consensus Bayesian tree topology."}
+tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
+tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
+plot(tentilla_obj)
+```
+
+```{r tentilla-ext-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla, mapped on the consensus Bayesian tree topology."}
+tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
+tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
+plot(tentilla2_obj)
+```
+
+```{r nectosome-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome, mapped on the consensus Bayesian tree topology."}
+nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
+nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
+plot(nectosome_obj)
+```
+
+```{r mantle-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores, mapped on the consensus Bayesian tree topology."}
+desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
+desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
+plot(desc_mantle_obj)
+```
+
+```{r necto-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome, mapped on the consensus Bayesian tree topology."}
+necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
+necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
+plot(necto_pos_obj)
+```
+
+```{r depth-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species observed with an MBARI ROV, mapped on the consensus Bayesian tree topology."}
+depth_value = depth_median
+depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap)
+```
+
+```{r depth-BI-extended, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Brownian Motion character map of median depth of species including blue water diving observations, mapped on the consensus Bayesian tree topology."}
+median_extended = cdata$Depth.Median.Extended
+names(median_extended) = rownames(cdata)
+median_extended = median_extended[match(names(median_extended), ultratree$tip.label)]
+depthmap_ext <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap_ext)
 

### 2018-05-10 manuscript.rmd
added code for bayesian tree

@@
 
+siphonophora_node_BI = ape::mrca( BIspeciestree )[ "Physalia physalis", "Agalma elegans" ]
+cystonectae_node_BI = ape::mrca( BIspeciestree )[ "Physalia physalis", "Rhizophysa filiformis" ]
+calycophora_node_BI = ape::mrca( BIspeciestree )[ "Kephyes ovata", "Craseoa lathetica" ]
+euphysonectae_node_BI = ape::mrca( BIspeciestree )[ "Agalma elegans", "Undescribed sp L" ]
+
+	support = as.numeric( BIspeciestree$node.label )
+	support[ support > 99 ] = NA
+
+	# pad tip values
+	support = c( rep( NA, length( BIspeciestree$tip.label ) ), support )
+
+	p_phylogram = ggtree( BIspeciestree ) +
+      geom_text(aes(label=get_node_support(BIspeciestree, phylobayes_trees)), vjust=1.5, hjust=1, size=2.5, col="blue") +
+			geom_tiplab( size=3 ) +
+			xlim( 0, 12 ) +
+			geom_cladelabel( node=cystonectae_node_BI, label="Cystonectae", align=FALSE, offset=1, color='red' ) +
+			geom_cladelabel( node=calycophora_node_BI, label="Calycophora", align=FALSE, offset=1, color='blue' ) +
+			geom_strip(
+				which( species_tree$tip.label=="Agalma elegans" ),
+				which( species_tree$tip.label=="Apolemia white" ),
+				barsize=0.5,
+				color='orange',
+				offset=1,
+				label='"Physonectae"'
+			) +
+			geom_strip(
+				which( species_tree$tip.label=="Hydractinia symbiolongicarpus" ),
+				which( species_tree$tip.label=="Nematostella vectensis" ),
+				barsize=0.5,
+				color='gray',
+				offset=1,
+				label='"Outgroups"'
+			)
+
+pdf("figures_raw/figure_S11_raw.pdf")
+flip(p_phylogram, euphysonectae_node_BI, calycophora_node_BI)
+dev.off()
+
 tree<-BIspeciestree

### 2018-05-10 manuscript.rmd
fixed node support issue

@@
     clades = prop.clades(tree, treeset) / length(treeset)
+    clades[is.na(clades)]<-0
     nodes = c(rep(NA,length(tree$tip.label)), round(clades*100)) # Offset with NA values for tips
@@
 
-	support = as.numeric( BIspeciestree$node.label )
-	support[ support > 99 ] = NA
-
-	# pad tip values
-	support = c( rep( NA, length( BIspeciestree$tip.label ) ), support )
-
 	p_phylogram = ggtree( BIspeciestree ) +

### 2018-05-11 manuscript.rmd
added figure 11 and updated figure 3/text

@@
 \@ref(phylogeny)A), with a couple of localized exceptions (Fig.
-\@ref(phylogeny)B).
+\@ref(phylogeny)B and \@ref(bayesian)).
 The ML results were robust to model selection (Fig.
 \@ref(fig:supplementary-trees)).
+BI results were not robust to model selection: BI CAT models produced strongly supported topologies that differed from the ML topology (Fig.
+\@ref(phylogeny)B and \@ref(bayesian)), while the BI WAG+$\Gamma$ topology was consistent with the ML topology (Fig.
+\@ref(fig:BI-WAG)).
 Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 \@ref(phylogeny)A).
-In BI analyses, Group B is graded (Fig.
+In BI CAT-Poisson analyses, Group B is graded (Fig.
 \@ref(phylogeny)B).
@@
 There is not a clear synapomorphy for Group B.
-Within Group B there is high support for the placement of *Erenna richardi* in ML analyses, but it is placed as sister to Clade A in BI analyses (Fig.
+Within Group B there is high support for the placement of *Erenna richardi* in ML analyses, but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig.
 \@ref(phylogeny)B).
@@
 In this study, *Craseoa lathetica* and *Desmophyes* sp.
-are sister to *Hippopodius hippopus* in ML analyses with high support, while in BI analyses, *H.
+are sister to *Hippopodius hippopus* in ML analyses with high support, while in BI CAT-Poisson analyses, *H.
 hippopus* is sister to *Lilyopsis fluoracantha* and the diphyomorphs (Fig.
@@
 
-	p_phylogram = ggtree( BIspeciestree ) +
+	p_phylogram_BI = ggtree( BIspeciestree ) +
       geom_text(aes(label=get_node_support(BIspeciestree, phylobayes_trees)), vjust=1.5, hjust=1, size=2.5, col="blue") +
@@
 pdf("figures_raw/figure_S11_raw.pdf")
-flip(p_phylogram, euphysonectae_node_BI, calycophora_node_BI)
+flip(p_phylogram_BI, euphysonectae_node_BI, calycophora_node_BI)
 dev.off()
@@
 ```
+![\label{bayesian}Bayesian (BI) CAT poisson phylogram with bipartition frequencies from the Bayesian posterior distribution of trees.
+Unlabeled nodes have support >0.99.
+The numbers of valid described species estimated to be in each clade based on taxonomy are shown below each clade name on the right.](supplementary_info/Supplementary_Figure11.pdf)
+
+```{r BI-WAG, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Bayesian (BI) WAG and Gamma phylogram with bipartition frequencies from the Bayesian posterior distribution of trees.
+Unlabeled nodes have support >0.99." }
+
+WAG_BItree<-ape::read.tree("agalma/Outfiles/combined_trimmed60_WAG.con.tre")
+WAG_BItree<-fix_trees(WAG_BItree)
+
+support = as.numeric( WAG_BItree$node.label )
+support=support*100
+support[ support > 99 ] = NA
+support = c( rep( NA, length( WAG_BItree$tip.label ) ), support )
+
+calycophora_node = ape::mrca( WAG_BItree )[ "Kephyes ovata", "Hippopodius hippopus" ]
+euphysonectae_node = ape::mrca( WAG_BItree )[ "Agalma elegans", "Erenna richardi" ]
+
+p<-ggtree(WAG_BItree)+geom_tiplab(size=2)+geom_text(aes(label=support),size=2, hjust=-0.1, col="blue") + xlim(0, 3.5)
+
+flip(p,calycophora_node,euphysonectae_node)
+```
 

### 2018-05-11 manuscript.rmd
updated text, figures, formatting

@@
 
-# Improved phylogenetic resolution within Siphonophora (Cnidaria) with implications for trait evolution
+# Improved phylogenetic resolution within Siphonophora (Cnidaria) with implications for trait evolution {-}
 
@@
 
-## Abstract
+## Abstract {-}
 
 Siphonophores are a diverse group of hydrozoans (Cnidaria) that are found at most depths of the ocean - from the surface, like the familiar Portuguese man of war, to the deep sea.
-Siphonophores play important roles in ocean ecosystems, and are among the most abundant gelatinous predators.
+They play important roles in ocean ecosystems, and are among the most abundant gelatinous predators.
 A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
@@
 
-## 1. Introduction
+## 1. Introduction {-}
 
@@
 
-## 2. Material and methods
+## 2. Material and methods {-}
 
@@
 
-#### 2.1 Collecting
+### 2.1 Collecting{-}
 
@@
 
-#### 2.2 Sequencing
+### 2.2 Sequencing{-}
 
@@
 
-#### 2.3 Analysis
+### 2.3 Analysis{-}
 
@@
 Maximum likelihood (ML) analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
-We also used IQTree to implement other ML analyses using different models of substitution to test the robustness of the results to model selection [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
 Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
@@
 This left `r length( phylobayes_trees )` trees in the posterior.
+
+To test the robustness of the results to model selection, we also used IQTree to implement other ML analyses using different models of substitution [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
+We also evaluated effects of method (ML vs.
+BI), model, and program by running a fixed-state WAG+$\Gamma$ model in Phylobayes and an ML approximation of CAT models in IQTree.
 
@@
 
-## 3. Results and Discussion
-
-### 3.1 Species phylogeny and hypothesis testing
+## 3. Results and Discussion{-}
+
+### 3.1 Species phylogeny and hypothesis testing{-}
 
@@
 gilmeri*) was sister to Forskaliidae with low support.
-The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
+The position of *Cordagalma cordiforme* (= *C.
+ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp.
 is in a clade with *Forskalia asymmetrica*, falling outside of Agalmatidae.
@@
 
-### 3.2 Evolution of Monoecy
+### 3.2 Evolution of Monoecy {-}
 
@@
 
-### 3.3 The Evolution of Zooid Types
+### 3.3 The Evolution of Zooid Types {-}
 
@@
 Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
-\@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
+\@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and, in the case of *B.
+sibogae*, modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 In species without bracts, other zooids appear to fulfill the roles of neutral buoyancy and protection.
-In *Physophora hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
+In *P.
+hydrostatica*, enlarged palpons surround all other siphosomal zooids and move in a coordinated manner to inflict a powerful sting [@totton1965synopsis].
 While in *Hippopodius hippopus* the nectophores play a role in maintaining neutral buoyancy and possibly also in defense, by bioluminescing and blanching in response to stimuli (Fig.
@@
 Palpons are typically defined as modified reduced gastrozooids [@mackie1987siphonophore].
-In many species palpons are thought to play a role in digestion and circulation of the gastrovascular fluid, while other species may use them for defense (e.g _Physophora_) or sensory functions [@totton1965synopsis].
-Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle); palpons found along the stem of the siphosome are termed palpons (typically have a tentacle, called a palpacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in _Apolemia_) [@totton1965synopsis;@siebert2013re].
+In many species palpons are thought to play a role in digestion and circulation of the gastrovascular fluid, while other species may use them for defense (e.g *Physophora*) or sensory functions [@totton1965synopsis].
+Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle, called a palpacle); palpons found along the stem of the siphosome are termed palpons (typically have a tentacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in *Apolemia*) [@totton1965synopsis;@siebert2013re].
 It is not clear how structure and function differs among different palpon subtypes, and a detailed histological investigation of palpons found at different locations within species is needed.
 For this reason, here we only assess the presence or absence of palpons as a category, without assessing subtypes of palpons.
-This presumes that palpons located at different regions in the colony are derived from other palpons rather than each arising de novo by independent modification of gastrozooids, a hypothesis that itself could be challenged upon closer histological examination of palpon diversity.
+This presumes that palpons located at different regions in the colony are derived from other palpons rather than each arising *de novo* by independent modification of gastrozooids, a hypothesis that itself could be challenged upon closer histological examination of palpon diversity.
 
@@
 
-### 3.4 The Gain and Loss of the Pneumatophore
+### 3.4 The Gain and Loss of the Pneumatophore {-}
 
@@
 
-### 3.5 The Gain and Loss of Tentilla
+### 3.5 The Gain and Loss of Tentilla {-}
 
@@
 It is still an open question whether the common ancestor of Siphonophora had tentilla.
-The only siphonophores species regarded as lacking tentilla are *Physalia physalis*, *Apolemia* spp.
+The only siphonophores species regarded as lacking tentilla are *P.
+physalis*, *Apolemia* spp.
 (Fig.
@@
 This suggests they are not just ectodermal swellings, but probably reduced tentilla.
-When we define *Physalia physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig.
+When we define *P.
+physalis* as tentilla bearing, the results for the character reconstruction lead to a more robust support for a tentilla-bearing common ancestor followed by independent losses of tentilla in the branch leading to Apolemiidae (Fig.
 \@ref(fig:tentilla2), \@ref(fig:tentilla-ext-BI)), and in *Bathyphysa conifera*.
@@
 
-### 3.6 The Evolution of Vertical Habitat Use
+### 3.6 The Evolution of Vertical Habitat Use {-}
 
@@
 This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters.
-There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *Physalia physalis*.
+There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *P.
+physalis*.
 There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
@@
 
-## 4. Conclusions
+## 4. Conclusions {-}
 
@@
 
-## Acknowledgements
+## Acknowledgements {-}
 
@@
 
-## Supplementary Information
+## Supplementary Information {-}
 \setcounter{figure}{0}
@@
 
-####Agalma analysis
-
-![\label{gene-occupancy}60% gene occupancy matrix for 41 species across 1,423 genes.
+###Agalma analysis {-}
+
+![\label{gene-occupancy}80% gene occupancy matrix for 41 species across 1,423 genes.
 Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
@@
 
-####SOWH analysis
+###SOWH analysis {-}
 
@@
 
-#####Stochastic Character maps
+###Stochastic Character maps {-}
 
@@
 ```
+
+###Bayesian models & model fit {-}
 
@@
 
-### Software versions
+### Software versions {-}
 
@@
 
-## References
+## References {-}

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.10.sh
pushing results for mon20 and phy22

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.10 -n ml.0.10 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.14.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.14 -n ml.0.14 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.17.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.17 -n ml.0.17 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.3.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.3 -n ml.0.3 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.32.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.32 -n ml.0.32 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.47.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.47 -n ml.0.47 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.51.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.51 -n ml.0.51 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.55.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.55 -n ml.0.55 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.57.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.57 -n ml.0.57 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.65.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.65 -n ml.0.65 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.75.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.75 -n ml.0.75 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.81.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.81 -n ml.0.81 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.85.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.85 -n ml.0.85 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.ml.0.94.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.94 -n ml.0.94 >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.23.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.23 -n t1.0.23 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.25.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.25 -n t1.0.25 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.35.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.35 -n t1.0.35 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.48.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.48 -n t1.0.48 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.54.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.54 -n t1.0.54 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.61.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.61 -n t1.0.61 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.63.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.63 -n t1.0.63 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.71.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.71 -n t1.0.71 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.77.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.77 -n t1.0.77 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.8.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.8 -n t1.0.8 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.85.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.85 -n t1.0.85 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.88.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.88 -n t1.0.88 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/good_trees/monoecious_19/sowhat_scratch/tree_scripts/raxml_script.t1.0.98.sh

@@ -1,1 +1,0 @@
-raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/ -m PROTGAMMAWAG -s /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowhat_scratch/aln.phy.0.98 -n t1.0.98 -g monoecious_19.tre >> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stdout_monoecious_19.txt 2>> /n/home03/shchurch/sowhat/good_trees/monoecious_19/sowh_stderr_monoecious_19.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.12.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.12 -n ml.0.12 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.19.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.19 -n ml.0.19 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.21.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.21 -n ml.0.21 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.29.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.29 -n ml.0.29 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.3.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.3 -n ml.0.3 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.30.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.30 -n ml.0.30 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.36.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.36 -n ml.0.36 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.38.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.38 -n ml.0.38 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.39.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.39 -n ml.0.39 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.43.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.43 -n ml.0.43 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.52.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.52 -n ml.0.52 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.6.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.6 -n ml.0.6 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.62.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.62 -n ml.0.62 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.71.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.71 -n ml.0.71 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.72.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.72 -n ml.0.72 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.81.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.81 -n ml.0.81 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.92.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.92 -n ml.0.92 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.93.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.93 -n ml.0.93 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.ml.0.97.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.97 -n ml.0.97 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.0.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.0 -n t1.0.0 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.1.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.1 -n t1.0.1 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.19.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.19 -n t1.0.19 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.21.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.21 -n t1.0.21 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.38.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.38 -n t1.0.38 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.40.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.40 -n t1.0.40 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.48.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.48 -n t1.0.48 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.57.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.57 -n t1.0.57 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.6.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.6 -n t1.0.6 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.74.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.74 -n t1.0.74 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.83.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.83 -n t1.0.83 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.89.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.89 -n t1.0.89 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/monoecious_20/sowhat_scratch/tree_scripts/raxml_script.t1.0.97.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowhat_scratch/aln.phy.0.97 -n t1.0.97 -g monoecious_20.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stdout_monoecious_20.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/monoecious_20/sowh_stderr_monoecious_20.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.32.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.32 -n ml.0.32 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.54.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.54 -n ml.0.54 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.57.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.57 -n ml.0.57 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.6.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.6 -n ml.0.6 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.60.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.60 -n ml.0.60 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.61.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.61 -n ml.0.61 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.7.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.7 -n ml.0.7 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.74.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.74 -n ml.0.74 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.76.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.76 -n ml.0.76 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.77.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.77 -n ml.0.77 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.84.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.84 -n ml.0.84 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.85.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.85 -n ml.0.85 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.89.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.89 -n ml.0.89 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.90.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.90 -n ml.0.90 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.95.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.95 -n ml.0.95 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.ml.0.98.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.98 -n ml.0.98 >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.10.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.10 -n t1.0.10 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.13.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.13 -n t1.0.13 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.19.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.19 -n t1.0.19 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.27.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.27 -n t1.0.27 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.3.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.3 -n t1.0.3 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.32.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.32 -n t1.0.32 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.43.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.43 -n t1.0.43 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.48.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.48 -n t1.0.48 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.5.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.5 -n t1.0.5 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.51.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.51 -n t1.0.51 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.53.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.53 -n t1.0.53 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.60.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.60 -n t1.0.60 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.66.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.66 -n t1.0.66 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.67.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.67 -n t1.0.67 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.71.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.71 -n t1.0.71 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.72.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.72 -n t1.0.72 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.77.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.77 -n t1.0.77 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.83.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.83 -n t1.0.83 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.84.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.84 -n t1.0.84 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.85.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.85 -n t1.0.85 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.94.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.94 -n t1.0.94 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 sowh/sowhat_results_files/physonect_22/sowhat_scratch/tree_scripts/raxml_script.t1.0.95.sh

@@ -1,0 +1,1 @@
+raxmlHPC-PTHREADS -T 8 -p 1234 --no-bfgs -w /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/ -m PROTGAMMAWAG -s /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowhat_scratch/aln.phy.0.95 -n t1.0.95 -g physonect_22.tre >> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stdout_physonect_22.txt 2>> /n/regal/extavour_lab/shchurch/sowhat_Apr6/physonect_22/sowh_stderr_physonect_22.txt

### 2018-05-16 manuscript.rmd

@@
 Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs.
-\@ref(traitdist) and \@ref(fig:median-cont)) [@mapstone2014global; @mackie1987siphonophore].
+\@ref(traitdist),\@ref(fig:median-cont),\@ref(fig:depth-BI)) [@mapstone2014global; @mackie1987siphonophore].
 The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel].
@@
 Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig.
-\@ref(fig:median-depth-extended)), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
+\@ref(fig:median-cont) and \@ref(fig:depth-BI)), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck).
 This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters.
@@
 physalis*.
-There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Figs.
-\@ref(fig:median-cont), \@ref(fig:median-depth-extended), \@ref(fig:depth-BI), \@ref(fig:depth-BI-extended)).
-Depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species); however several shallow-living species that likely belong in Group B were not included in this analysis.
+There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Fig.
+\@ref(fig:median-cont), and \@ref(fig:depth-BI)).
+Under the ML topology, depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species) (Fig.
+\@ref(fig:median-cont)), while under the BI-CAT topology, a mesopelagic common ancestor is predicted, with a transition to epipelagic waters in Clade A (Fig.
+\@ref(fig:depth-BI)); however several shallow-living species that likely belong in Group B were not included in this analysis.
 The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species.
@@
 We also reconstructed the median depth changes along the phylogeny using median depths of 20m for all species collected by SCUBA diving or via a shallow trawl (Fig.
-\@ref(fig:median-depth-extended)), and still find support for a mesopelagic ancestor.
+\@ref(fig:median-depth-extended) and \@ref(fig:depth-BI-extended)), and still find support for a mesopelagic ancestor.
 It should be noted, however, that *H.

### 2018-05-16 manuscript.rmd
adjusted wording

@@
 \@ref(phylogeny)C, S2): (i) "Physonectae" is monophyletic [@totton1965synopsis]; (ii) monoecious species are monophyletic [@Dunn:2005dy].
-As the sexual system of *Rudjakovia* sp.
-is unclear, we carried out two tests of the monophyly of monoecy, one with *Rudjakovia* sp.
-included as a monoecious species, and one with it considered dioecious (Fig.
-\@ref(phylogeny)C, S2).
+The sexual mode of  *Rudjakovia* is undescribed, but preliminary observations suggest that they are monoecious, so we include *Rudjakovia* as a monoecious species in this test.
 We used SOWHAT [@doi:10.1093/sysbio/syv055] dev.
@@
 \@ref(phylogeny)C): (i) monophyletic Physonectae, (ii) monophyletic monoecious siphonophores.
-As the sexual mode of  *Rudjakovia* is undescribed, we carried out two tests: with one test including *Rudjakovia* sp.
-as a monoecious species, and the other with dioecious *Rudjakovia* sp.
-In all three tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
+In both tests the alternative hypothesis was rejected (p-value <0.01, confidence interval: <0.001 - 0.03, Fig.
 \@ref(sowh)).
@@
 This is because Group B, which contains dioecious species, is also descended from their most recent common ancestor.
-SOWH tests strongly reject the placement of the monecious clades Calycophorae and Clade A as a group that excludes Group B (Figs.
+The SOWH test strongly rejects the placement of the monecious clades Calycophorae and Clade A as a group that excludes Group B (Figs.
 3C and S2).

### 2018-05-17 manuscript.rmd
adjusted wording

@@
 In either case, male and female zooids are found at the same corresponding locations within the cormidia.
-One exception to this is Abylids, where both male or female gonophores may be found within the same eudoxid [@carre1967].
+One known exception to this can be found in abylid calycophorans, where both male or female gonophores may be found within the same eudoxid [@carre1967].
 In sum, homoplasy in sexual-system evolution along with variation in the spatial arrangement of gonophores within a colony suggest that siphonophores have evolved different ways to be monoecious.
@@
 These patterns of loss are not captured in this study, as most of these species are not included in the present phylogeny.
-In species without bracts, other zooids appear to fulfill the roles of neutral buoyancy and protection.
+In species without bracts, other zooids appear to fulfill the roles of buoyancy control and protection.
 In *P.

### 2018-05-17 manuscript.rmd
update Figure references

@@
 The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0185032; CAT-GTR: maxdiff=1, meandiff=0.045415).
-Only the results from the CAT-Poisson model are presented here.
+Only the results from the CAT-Poisson model are presented here (Fig.
+S11).
 Two chains (chain1 and chain3) were stuck in local maxima after 1405 and 4695 iterations and were excluded from the analyses.
@@
 We also evaluated effects of method (ML vs.
-BI), model, and program by running a fixed-state WAG+$\Gamma$ model in Phylobayes and an ML approximation of CAT models in IQTree.
+BI), model, and program by running a fixed-state WAG+$\Gamma$ model in Phylobayes (Fig.
+S12) and an ML approximation of CAT models in IQTree (Fig.
+S10).
 

### 2018-05-17 manuscript.rmd
update discussion palpon ancestral character reconstruction

@@
 L.
+It remains to be clarified if small buds associated with nectophores within the nectosome of *Bargmannia* species (Dunn 2005) actually correspond to reduced palpons.
 The pyrostephid *Pyrostephos vanhoeffeni* (not sampled) has modified tentacle-less palpons (termed oleocysts), but the relationship between this species and *Rudjakovia* sp.
-is not known, so the exact patterns of loss within *Pyrostephidae* remain unclear.
+is not known, so the exact patterns of loss within *Pyrostephidae* (here provisionally including *Rudjakovia* sp.) remain unclear.
 Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.

### 2018-05-18 manuscript.rmd
update wording

@@
 New data indicated by Y, blank fields indicate that data were already published.
-For the species not on SRA, a link to the data is included in supplementary file 1.")
+For the species not on SRA, a link to the data is included in supplementary data 1.")
 print(xtable_specimen,comment=F, include.rownames=FALSE, scalebox='0.7')
@@
 L.
-It remains to be clarified if small buds associated with nectophores within the nectosome of *Bargmannia* species (Dunn 2005) actually correspond to reduced palpons.
+It remains to be clarified if small buds associated with nectophores within the nectosome of *Bargmannia* species [@Dunn:2005fv] actually correspond to reduced palpons.
 The pyrostephid *Pyrostephos vanhoeffeni* (not sampled) has modified tentacle-less palpons (termed oleocysts), but the relationship between this species and *Rudjakovia* sp.

### 2018-05-18 manuscript.rmd
added kayal reference

@@
 Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group.
-The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2015phylogenetic], but there has been progress on their internal relationships.
+The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships.
 A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.

### 2018-05-18 manuscript.rmd
update Figure reference

@@
 
-Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Figs.
-\@ref(traitdist),\@ref(fig:median-cont),\@ref(fig:depth-BI)) [@mapstone2014global; @mackie1987siphonophore].
+Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig.
+\@ref(traitdist), \@ref(fig:median-cont), \@ref(fig:depth-BI)) [@mapstone2014global; @mackie1987siphonophore].
 The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel].

### 2018-05-18 manuscript.rmd
update Figure reference
fixed warning
updated code, caption for S10

@@
 There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Fig.
-\@ref(fig:median-cont), and \@ref(fig:depth-BI)).
+\@ref(fig:median-cont), \@ref(fig:depth-BI)).
 Under the ML topology, depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species) (Fig.
@@
 A.
+JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories (best model fit).
+Likelihood:-8113694.922; AIC:16227609.9565 .
+B.
 GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
 Likelihood: -8133157.335; AIC:16266530.7277.
-B.
-JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories.
-Likelihood:-8113694.922; AIC:16227609.9565 .
 C.
@@
 
-p1<-ggtree(trees[[1]])+geom_tiplab(size=2)+geom_text(aes(label=support1),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("A. GTR20+FO+R6")
-
-p2<-ggtree(trees[[2]])+geom_tiplab(size=2)+geom_text(aes(label=support2),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("B. JTT+F+R7")
+p1<-ggtree(trees[[2]])+geom_tiplab(size=2)+geom_text(aes(label=support2),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("A. JTT+F+R7")
+
+p2<-ggtree(trees[[1]])+geom_tiplab(size=2)+geom_text(aes(label=support1),size=2, hjust=-0.1, col="red") + xlim(0, 3.5)+ ggtitle("B. GTR20+FO+R6")
 
@@
 
-```{r BI tree ancestral trait reconstructions}
+```{r BI tree ancestral trait reconstructions, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,}
 

### 2018-05-18 manuscript.rmd
updated text and S10 figure re: IQTree runs/model selection

@@
 
-To test the robustness of the results to model selection, we also used IQTree to implement other ML analyses using different models of substitution [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
-We also evaluated effects of method (ML vs.
+To test the robustness of the results to model selection, we also used IQTree to implement other ML analyses using different models of substitution (GTR, JTT)  [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
+We used ModelFinder to assess model fit [@kalyaanamoorthy2017modelfinder].We also evaluated effects of method (ML vs.
 BI), model, and program by running a fixed-state WAG+$\Gamma$ model in Phylobayes (Fig.
-S12) and an ML approximation of CAT models in IQTree (Fig.
-S10).
+S12).
 
@@
 
-![\label{gene-occupancy}80% gene occupancy matrix for 41 species across 1,423 genes.
-Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)
+![\label{gene-occupancy}80% gene occupancy matrix for 41 species across 1,423 genes.](supplementary_info/Supplementary_Figure1.png)
 
@@
 
-```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution.
+```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree to test the robustness of the results to model selection.
+We used ModelFinder [@kalyaanamoorthy2017modelfinder] to assess model fit and to test the robustness of the results to model selection, we also used the WAG and GTR model.
 Unlabeled nodes have support >0.99.

### 2018-05-18 manuscript.rmd
edits through end of 3.1, did not rebuild pdf

@@
 They play important roles in ocean ecosystems, and are among the most abundant gelatinous predators.
-A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups, however there was little support for many deep relationships within the clade Codonophora.
+A previous phylogenetic study based on two ribosomal RNA genes provided insight into the internal relationships between major siphonophore groups.
+There was, however, little support for many deep relationships within the clade Codonophora.
 Here, we present a new siphonophore phylogeny based on new transcriptome data from 29 siphonophore species analyzed in combination with 14 publicly available genomic and transcriptomic datasets.
@@
 This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods.
-Maximum likelihood (ML) analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma.
-Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v.
-1.7a-mpi [@lartillot2009phylobayes].
-Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository.
-Tree figures were rendered with ggtree [@Yu:2016fo].
 
@@
 
-ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
-BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112].
-Two independent MCMC chains were run under the CAT-GTR model, and eight independent MCMC chains were run under the CAT-Poisson model.
-The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0185032; CAT-GTR: maxdiff=1, meandiff=0.045415).
-Only the results from the CAT-Poisson model are presented here (Fig.
-S11).
-Two chains (chain1 and chain3) were stuck in local maxima after 1405 and 4695 iterations and were excluded from the analyses.
+We used ModelFinder [@kalyaanamoorthy2017modelfinder], as implemented in IQTree v1.6.3 [@doi:10.1093/molbev/msu300], to assess relative model fit.
+ModelFinder selected XX based on XX.
+To assess the robustness of our results, we conducted phylogenetic analyses using multiple software programs, methods (Maximum likelihood (ML) and Bayesian Inference (BI)), and models (including the model selected by ModelFinder and several other commonly used models).
+Maximum likelihood (ML) analyses were conducted with RAxML v8.2.0 [@stamatakis2006raxml] and IQTree v1.6.3 [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
+Bayesian Inference (BI) were conducted with Phylobayes v.
+1.7a-mpi [@lartillot2009phylobayes].
+Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository.
+Tree figures were rendered with ggtree [@Yu:2016fo].
+
+RAxML ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution (Figure XX).
+RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
+IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories (Figure XX).
+
+BI was conducted in phylobayes using two different models: fixed-state WAG+$\Gamma$ (Fig.
+S12) and CAT-Poisson (Figure S11).
+Eight chains were run under the CAT-Poisson model.
+The WAG+$\Gamma$ runs (XX provide the same detail as is available for CAT poisson).
+The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0185032).
+Closer inspection revealed that chain1 and chain3 were stuck in local maxima with low likelihood relative to other chains after 1405 and 4695 iterations.
+These two chains were excluded from the analyses, and the results pressented here are based on the remaining 6 CAT-Poisson chains.
 Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
-This left `r length( phylobayes_trees )` trees in the posterior.
-
-To test the robustness of the results to model selection, we also used IQTree to implement other ML analyses using different models of substitution (GTR, JTT)  [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281].
-We used ModelFinder to assess model fit [@kalyaanamoorthy2017modelfinder].We also evaluated effects of method (ML vs.
-BI), model, and program by running a fixed-state WAG+$\Gamma$ model in Phylobayes (Fig.
-S12).
+This left `r length( phylobayes_trees )` trees in the CAT-Poisson posterior.
 
@@
 \@ref(phylogeny)B and \@ref(bayesian)).
-The ML results were robust to model selection (Fig.
+All of the ML analyses were congruent with each other, regardless of model and software used (Fig.
 \@ref(fig:supplementary-trees)).
-BI results were not robust to model selection: BI CAT models produced strongly supported topologies that differed from the ML topology (Fig.
-\@ref(phylogeny)B and \@ref(bayesian)), while the BI WAG+$\Gamma$ topology was consistent with the ML topology (Fig.
+These ML results were also congruent with the Phylobayes BI WAG+$\Gamma$ analyses (Fig.
 \@ref(fig:BI-WAG)).
+The Phylobayes BI CAT-Poisson result (Fig.
+\@ref(bayesian)), however, had a strongly supported topology that differed (Fig.
+\@ref(phylogeny)B) from the ML topology in a couple localized regions as described below.
+The fact that the Phylobayes BI WAG+$\Gamma$ is consistent with the WAG (and other) ML analyses suggests that the different topology recovered in the Phylobayes BI CAT-Poisson analyses is due to the different model rather than different software or methods.
+Here we take the conservative approach of considering relationships that differ between the Phylobayes BI CAT-Poisson analyses and other analyses to be unresolved.
+
 Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy].
@@
 
-In ML analyses Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Group B (Fig.
+In ML analyses and BI WAG analyses, Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Group B (Fig.
 \@ref(phylogeny)A).
-In BI CAT-Poisson analyses, Group B is graded (Fig.
+In BI CAT-Poisson analyses, Group B is paraphyletic (Fig.
 \@ref(phylogeny)B).
@@
 There is not a clear synapomorphy for Group B.
-Within Group B there is high support for the placement of *Erenna richardi* in ML analyses, but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig.
+Within Group B there is high support for the placement of *Erenna richardi* in ML analyses (XX and BI WAG??), but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig.
 \@ref(phylogeny)B).
-More taxon sampling will be required to determine the relationship of species within this clade.
+More taxon sampling will be required to determine the relationship of species within this group.
 
@@
 
-```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree to test the robustness of the results to model selection.
-We used ModelFinder [@kalyaanamoorthy2017modelfinder] to assess model fit and to test the robustness of the results to model selection, we also used the WAG and GTR model.
+```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree.
+We used ModelFinder [@kalyaanamoorthy2017modelfinder] to identify which examined model as the best relative model fit.
+We also present results for the commonly used WAG and GTR models.
 Unlabeled nodes have support >0.99.
 A.
-JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories (best model fit).
-Likelihood:-8113694.922; AIC:16227609.9565 .
+JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories (the model identified by ModelFinder as having the best fit).
+Log likelihood:-8113694.922; AIC:16227609.9565 .
 B.
 GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
-Likelihood: -8133157.335; AIC:16266530.7277.
+Log likelihood: -8133157.335; AIC:16266530.7277.
 C.
 WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories.
-Likelihood: -8156043.772; AIC score: 16312303.6120 "}
+Log likelihood: -8156043.772; AIC score: 16312303.6120 "}
 

### 2018-05-18 manuscript.rmd
incorporated author comments, fixed typos, updated references

@@
 \@ref(colony)) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre].
+In some species additional zooids are added outside the growth zone along the siphosomal stem [@siebert2013re].
 These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution].
@@
 Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group.
-The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships.
+The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2013cnidarian; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships.
 A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
@@
 Using 1,423 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study.
-Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.
+Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and vertical habitat use.
 
@@
 All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017).
-The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v1".
+The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v2".
 
@@
 We used ModelFinder [@kalyaanamoorthy2017modelfinder], as implemented in IQTree v1.6.3 [@doi:10.1093/molbev/msu300], to assess relative model fit.
-ModelFinder selected XX based on XX.
+ModelFinder selected JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories based on the Bayesian Information Criterion.
 To assess the robustness of our results, we conducted phylogenetic analyses using multiple software programs, methods (Maximum likelihood (ML) and Bayesian Inference (BI)), and models (including the model selected by ModelFinder and several other commonly used models).
@@
 
-RAxML ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution (Figure XX).
+RAxML ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution (Fig.
+\@ref(phylogeny)A).
 RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
-IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories (Figure XX).
+IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories ((Fig.
+\@ref(fig:supplementary-trees))).
 
 BI was conducted in phylobayes using two different models: fixed-state WAG+$\Gamma$ (Fig.
-S12) and CAT-Poisson (Figure S11).
+\@ref(fig:BI-WAG)) and CAT-Poisson (Fig.
+\@ref(bayesian)).
 Eight chains were run under the CAT-Poisson model.
-The WAG+$\Gamma$ runs (XX provide the same detail as is available for CAT poisson).
-The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0185032).
+Four chains were run udner WAG+$\Gamma$, these runs did not converge (maxdiff=1, meandiff=0.0130273).
+The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0565898).
 Closer inspection revealed that chain1 and chain3 were stuck in local maxima with low likelihood relative to other chains after 1405 and 4695 iterations.
-These two chains were excluded from the analyses, and the results pressented here are based on the remaining 6 CAT-Poisson chains.
+These two chains were excluded from the analyses, and the results presented here are based on the remaining 6 CAT-Poisson chains (maxdiff=1, meandiff=0.0185032).
 Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs.
@@
 Depth distribution data was queried from the MBARI VARS database ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/) [@schlining2006mbari].
-We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
+We used stochastic character mapping to infer the most probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
 Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package.
@@
 \@ref(bayesian)), however, had a strongly supported topology that differed (Fig.
-\@ref(phylogeny)B) from the ML topology in a couple localized regions as described below.
+\@ref(phylogeny)B) from the ML topology in localized regions as described below.
 The fact that the Phylobayes BI WAG+$\Gamma$ is consistent with the WAG (and other) ML analyses suggests that the different topology recovered in the Phylobayes BI CAT-Poisson analyses is due to the different model rather than different software or methods.
@@
 There is not a clear synapomorphy for Group B.
-Within Group B there is high support for the placement of *Erenna richardi* in ML analyses (XX and BI WAG??), but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig.
+Within Group B there is high support for the placement of *Erenna richardi* in ML analyses and BI WAG (Fig.
+\@ref(phylogeny) and \@ref(fig:BI-WAG)), but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig.
 \@ref(phylogeny)B).
@@
 Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig.
-\@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and, in the case of *B.
-sibogae*, modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
+\@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and in the case of *B.
+sibogae*, use modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].
 Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960].
@@
 
-Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture).
+Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture, and rhodalids, where some tentacles are used to anchor to the substrate and do not participate in feeding).
 As in other cnidarians, stinging capsules, arranged in dense batteries of nematocysts, play a critical role in prey capture.
@@
 Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores [@grossmann2015diversity].
-Since we only have oceanographic data from the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.
+Since most of our specimens were sampled only in the Monterey Bay region, our analyses of the local oceanographic and depth distribution data cannot disentangle the effects of these different variables on the vertical distributions.
 
@@
 
-This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award).
+This work was supported by the National Science Foundation (DEB-1256695, the Waterman Award, and DEB-1701272).
 CM was also supported in part by a RI-EPSCoR Fellowship, NSF EPS-1004057.
@@
 ```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree.
-We used ModelFinder [@kalyaanamoorthy2017modelfinder] to identify which examined model as the best relative model fit.
+We used ModelFinder [@kalyaanamoorthy2017modelfinder] to identify the model with the best relative model fit.
 We also present results for the commonly used WAG and GTR models.

### 2018-05-18 manuscript.rmd
new knit with supp. info pdf output fix

@@
 All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017).
-The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v2".
+The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "") %>% strtrim(8)` with tag "paper_v2".
 
@@
 RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
-IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories ((Fig.
+IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used models GTR + Optimized base frequencies + Free rate model with 6 categories and WAG + Optimized base frequencies + Free rate model with 6 categories ((Fig.
 \@ref(fig:supplementary-trees))).
@@
 We used stochastic character mapping to infer the most probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic].
+For continuous character traits, model fit was tested using
+`fitContinuous` in the `geiger` R package.
 Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package.
@@
 Ancestral character state reconstructions favor the hypothesis that monoecy arose twice (Fig.
-\@ref(reconstruct)A, \@ref(fig:BI-mono)), once in Calycophorae and once in Clade A.
+\@ref(reconstruct)A and \@ref(fig:BI-mono)), once in Calycophorae and once in Clade A.
 This is consistent with differences in the arrangements of male and female gonophores in the two clades.
@@
 
-Within Calycophorae there are additional variations of the sexual mode: in *Sulculeolaria* (not included in this phylogeny) colonies appear to consist of only one sex at a time.
+Within Calycophorae there are additional variations of the sexual mode: in *Sulculeolaria* (not included in this phylogeny) colonies appear to present a single sex at a time.
 However they are monoecious and protandrous, with female gonophores developing after the release of male gonophores [@carre1979].
@@
 In many species palpons are thought to play a role in digestion and circulation of the gastrovascular fluid, while other species may use them for defense (e.g *Physophora*) or sensory functions [@totton1965synopsis].
-Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle, called a palpacle); palpons found along the stem of the siphosome are termed palpons (typically have a tentacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in *Apolemia*) [@totton1965synopsis;@siebert2013re].
+Palpons are subcategorised based on their location - palpons that are associated with gonodendra are termed gonopalpons (typically with a reduced tentacle, called a palpacle); palpons found along the stem of the siphosome are termed palpons (typically have a palpacle); and palpons found along the stem of the nectosome are termed nectosomal palpons (as in *Apolemia*) [@totton1965synopsis;@siebert2013re].
 It is not clear how structure and function differs among different palpon subtypes, and a detailed histological investigation of palpons found at different locations within species is needed.
@@
 We also thank Zack Lewis for help with RNA extractions and the MBARI crews and ROV pilots for collection of the specimens.
+\clearpage
 
@@
 
-pdf("supplementary_info/Supplementary_Figure3.pdf")
-plot(tentilla_obj)
-dev.off()
-
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure4.pdf")
-plot(tentilla2_obj)
-dev.off()
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure5.pdf")
-plot(nectosome_obj)
-dev.off()
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure6.pdf")
-plot(desc_mantle_obj)
-dev.off()
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure7.pdf")
-plot(necto_pos_obj)
-dev.off()
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure8.pdf")
-plot(depthmap)
-dev.off()
 ```
@@
 
-pdf("supplementary_info/Supplementary_Figure9.pdf")
-plot(depthmap_ext)
-dev.off()
-
 #below are the model fit details
@@
 ```
-
-###Bayesian models & model fit {-}
 
@@
 
-```{r BI tree ancestral trait reconstructions, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE,}
+```{r BI-tree-trait-reconstructions, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, include=F}
 
@@
 #print("SIMMAP Sex Distribution")
-make.simmap(ultratree , sex, nsim = 100, message=FALSE) -> sex_sim
+make.simmap(ultratree , sex, nsim = 100, message=FALSE) -> sex_sim_BI
 #print("SIMMAP Palpons")
-make.simmap(ultratree , palpons, nsim = 100, message=FALSE) -> palpons_sim
+make.simmap(ultratree , palpons, nsim = 100, message=FALSE) -> palpons_sim_BI
 #print("SIMMAP Tentilla")
-make.simmap(ultratree , tentilla, nsim = 100, message=FALSE) -> tentilla_sim
+make.simmap(ultratree , tentilla, nsim = 100, message=FALSE) -> tentilla_sim_BI
 #print("SIMMAP Tentilla2")
-make.simmap(ultratree , tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
+make.simmap(ultratree , tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2_BI
 #print("SIMMAP Nectosome")
-make.simmap(ultratree , nectosome, nsim = 100, message=FALSE) -> nectosome_sim
+make.simmap(ultratree , nectosome, nsim = 100, message=FALSE) -> nectosome_sim_BI
 #print("SIMMAP Nectosome Position")
-make.simmap(ultratree , bracts, nsim = 100, message=FALSE) -> bracts_sim
+make.simmap(ultratree , bracts, nsim = 100, message=FALSE) -> bracts_sim_BI
 #print("SIMMAP Habitat")
@@
 NectoPosTree = drop.tip(ultratree , which(!(ultratree $tip.label %in% names(necto_pos))))
-make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
+make.simmap(NectoPosTree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim_BI
 
@@
 DescMantleTree =  drop.tip(ultratree , which(!(ultratree $tip.label %in% names(desc_mantle))))
-make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim
-```
+make.simmap(DescMantleTree, desc_mantle, nsim = 100, message=FALSE) -> desc_mantle_sim_BI
+```
+\clearpage
+
 ![\label{bayesian}Bayesian (BI) CAT poisson phylogram with bipartition frequencies from the Bayesian posterior distribution of trees.
@@
 #Simmap SEX
-sex_obj <- densityMap(sex_sim, plot=FALSE)
-sex_obj$cols[1:length(sex_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj$cols))
- plot(sex_obj)
+sex_obj_BI <- densityMap(sex_sim_BI, plot=FALSE)
+sex_obj_BI$cols[1:length(sex_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(sex_obj_BI$cols))
+ plot(sex_obj_BI)
 ```
@@
 #Simmap PALPONS
-palpons_obj <- densityMap(palpons_sim, plot=FALSE)
-palpons_obj$cols[1:length(palpons_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj$cols))
-
-plot(palpons_obj)
+palpons_obj_BI <- densityMap(palpons_sim_BI, plot=FALSE)
+palpons_obj_BI$cols[1:length(palpons_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(palpons_obj_BI$cols))
+
+plot(palpons_obj_BI)
 ```
@@
 ```{r tentilla-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla, mapped on the consensus Bayesian tree topology."}
-tentilla_obj <- densityMap(tentilla_sim,plot=FALSE)
-tentilla_obj$cols[1:length(tentilla_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj$cols))
-plot(tentilla_obj)
+tentilla_obj_BI <- densityMap(tentilla_sim_BI,plot=FALSE)
+tentilla_obj_BI$cols[1:length(tentilla_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla_obj_BI$cols))
+plot(tentilla_obj_BI)
 ```
@@
 ```{r tentilla-ext-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla, mapped on the consensus Bayesian tree topology."}
-tentilla2_obj <- densityMap(tentilla_sim2, plot=FALSE)
-tentilla2_obj$cols[1:length(tentilla2_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj$cols))
-plot(tentilla2_obj)
+tentilla2_obj_BI <- densityMap(tentilla_sim2_BI, plot=FALSE)
+tentilla2_obj_BI$cols[1:length(tentilla2_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(tentilla2_obj_BI$cols))
+plot(tentilla2_obj_BI)
 ```
@@
 ```{r nectosome-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of nectosome, mapped on the consensus Bayesian tree topology."}
-nectosome_obj <- densityMap(nectosome_sim, plot=FALSE)
-nectosome_obj$cols[1:length(nectosome_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj$cols))
-plot(nectosome_obj)
+nectosome_obj_BI <- densityMap(nectosome_sim_BI, plot=FALSE)
+nectosome_obj_BI$cols[1:length(nectosome_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(nectosome_obj_BI$cols))
+plot(nectosome_obj_BI)
 ```
@@
 ```{r mantle-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map of presence of a descending mantle canal in the nectophores, mapped on the consensus Bayesian tree topology."}
-desc_mantle_obj <- densityMap(desc_mantle_sim, plot=FALSE)
-desc_mantle_obj$cols[1:length(desc_mantle_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj$cols))
-plot(desc_mantle_obj)
+desc_mantle_obj_BI <- densityMap(desc_mantle_sim_BI, plot=FALSE)
+desc_mantle_obj_BI$cols[1:length(desc_mantle_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(desc_mantle_obj_BI$cols))
+plot(desc_mantle_obj_BI)
 ```
@@
 ```{r necto-BI, echo=FALSE, message=F, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Stochastic character map for the evolution of the position of the nectosome, mapped on the consensus Bayesian tree topology."}
-necto_pos_obj <- densityMap(necto_pos_sim, plot=FALSE)
-necto_pos_obj$cols[1:length(necto_pos_obj$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj$cols))
-plot(necto_pos_obj)
+necto_pos_obj_BI <- densityMap(necto_pos_sim_BI, plot=FALSE)
+necto_pos_obj_BI$cols[1:length(necto_pos_obj_BI$cols)]<-colorRampPalette(colors=c("#4A4ACC", "orange"), space="Lab")(length(necto_pos_obj_BI$cols))
+plot(necto_pos_obj_BI)
 ```
@@
 depth_value = depth_median
-depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
-plot(depthmap)
+depthmap_BI <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap_BI)
 ```
@@
 median_extended = median_extended[match(names(median_extended), ultratree$tip.label)]
-depthmap_ext <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+depthmap_ext_BI <- contMap(ultratree, median_extended, plot=F) %>% setMap(colors=c("cyan","blue","black"))
+plot(depthmap_ext_BI)
+```
+
+\clearpage
+
+```{r pdf-supplemental, verbose= FALSE, include=F, message = FALSE, echo=FALSE, comment=NA, warning=FALSE }
+pdf("supplementary_info/Supplementary_Figure3.pdf")
+plot(tentilla_obj)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure4.pdf")
+plot(tentilla2_obj)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure5.pdf")
+plot(nectosome_obj)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure6.pdf")
+plot(desc_mantle_obj)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure7.pdf")
+plot(necto_pos_obj)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure8.pdf")
+plot(depthmap)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure9.pdf")
 plot(depthmap_ext)
-
-```
-
-\pagebreak
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure10.pdf")
+multiplot(p1, p2, p3, ncol=3)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure12.pdf")
+flip(p,calycophora_node,euphysonectae_node)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure13.pdf")
+plot(sex_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure14.pdf")
+plot(palpons_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure15.pdf")
+plot(tentilla_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure16.pdf")
+plot(tentilla2_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure17.pdf")
+plot(nectosome_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure18.pdf")
+plot(desc_mantle_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure19.pdf")
+plot(necto_pos_obj_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure20.pdf")
+plot(depthmap_BI)
+dev.off()
+
+pdf("supplementary_info/Supplementary_Figure21.pdf")
+plot(depthmap_ext_BI)
+dev.off()
+
+```
 

### 2018-05-18 manuscript.rmd
minor typos and references fixed

@@
 The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2013cnidarian; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships.
-A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae.
+A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognized groups, Cystonectae, Physonectae, and Calycophorae.
 Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae".
@@
 RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates.
-IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used models GTR + Optimized base frequencies + Free rate model with 6 categories and WAG + Optimized base frequencies + Free rate model with 6 categories ((Fig.
-\@ref(fig:supplementary-trees))).
+IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used models GTR + Optimized base frequencies + Free rate model with 6 categories and WAG + Optimized base frequencies + Free rate model with 6 categories (Fig.
+\@ref(fig:supplementary-trees)).
 
@@
 Eight chains were run under the CAT-Poisson model.
-Four chains were run udner WAG+$\Gamma$, these runs did not converge (maxdiff=1, meandiff=0.0130273).
+Four chains were run under WAG+$\Gamma$, these runs did not converge (maxdiff=1, meandiff=0.0130273).
 The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0565898).
@@
 In this study, *Craseoa lathetica* and *Desmophyes* sp.
-are sister to *Hippopodius hippopus* in ML analyses with high support, while in BI CAT-Poisson analyses, *H.
+are sister to *Hippopodius hippopus* in ML and BI-WAG analyses with high support, while in BI CAT-Poisson analyses, *H.
 hippopus* is sister to *Lilyopsis fluoracantha* and the diphyomorphs (Fig.
-\@ref(phylogeny)B).
+\@ref(phylogeny)B and \@ref(bayesian)).
 
@@
 Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen].
-Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
+Some calycophorans are able to actively exclude sulfate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role].
 Bracts were lost in Hippopodiidae, some clausophyids, *Physophora hydrostatica* (Fig.
@@
 There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Fig.
-\@ref(fig:median-cont), \@ref(fig:depth-BI)).
+\@ref(fig:median-cont) and \@ref(fig:depth-BI)).
 Under the ML topology, depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species) (Fig.
@@
 This reconstruction (Fig.
-\@ref(fig:median-cont)) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
+\@ref(fig:median-cont) and \@ref(fig:depth-BI)) only included depths recorded using an ROV, thus it excludes many other independent colonizations of the epipelagic habitat.
 The ROV observations are reliable below 200m, and no quantitative measurements were made on SCUBA dives.
